### 2020-12-02 paper/paper.tex
initial commit from leuven paper

\documentclass[envcountsame]{llncs}

\usepackage[table]{xcolor}
\usepackage[utf8x]{inputenc}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{amsmath}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage[noend]{algpseudocode}
\usepackage{stmaryrd}
\usepackage{listings}
\usepackage{algorithm}
\usepackage{array}
\usepackage[colon,round,sort]{natbib}
\usepackage{textcomp}
\usepackage{pifont}
% \usepackage{bbding}
\usepackage{comment}
\usepackage{graphicx}
\usepackage[svgpath=figures/]{svg}
\usepackage[labelfont=bf]{caption}
\captionsetup[table]{skip=5pt}
\usepackage{tikz}
\usetikzlibrary{patterns}
\usetikzlibrary{calc}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{shapes.geometric}
% \usepackage{tikz}
% \usetikzlibrary{patterns}
% \usetikzlibrary{calc}
% \usetikzlibrary{decorations.pathreplacing}
% \usetikzlibrary{shapes,arrows}
% \usetikzlibrary{shapes.geometric}
% \usetikzlibrary{arrows}

\usepackage{ntheorem}
\newtheorem*{definition*}{Definition}

\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{relsize}
\tikzset{fontscale/.style = {font=\relsize{#1}}}
\usepackage{marvosym}
\usepackage{paper}

% \usepackage{academicons}
% \usepackage{cite}
\usepackage[hidelinks,
plainpages=false,
pdftitle={Reinforcement Learning Methods for Operations Research Applications -- Reloaded},
pdfauthor={Manuel Schneckenreither and Stefan Haeussler},
pdfsubject={Reinforcement Learning Methods for Operations Research Applications -- Reloaded},
pdfkeywords={production planning, order release, machine learning, reinforcement learning,
  operations research},
]{hyperref}
\usepackage{url}
\pagestyle{plain}

\newcommand{\SH}[1]{\begin{center} \textcolor{green!50!black}{#1} \end{center}}
\newcommand\MS[2][r]{\ifx t#1 \textcolor{blue}{[\textbf{MS:} #2]}
  \else \begin{center}\textcolor{blue}{\textbf{MS:} #2} \end{center} \fi}

% \newcolumntype{x}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}
\title{
  Reinforcement Learning Methods for Operations Research Applications -- Reloaded
}
\author{Manuel Schneckenreither% \inst{1}
  % \orcidID{0000-0002-4812-4665}
  \and Stefan Haeussler% \inst{2}
  % \orcidID{0000-0003-2589-1367}
}

% \authorrunning{Schneckenreither and Haeussler}
\institute{
  Department of Information Systems,
  Production and Logistics Management,
  University of Innsbruck, Austria\\
  email: \email{manuel.schneckenreither@uibk.ac.at,stefan.haeussler@uibk.ac.at}
}

\begin{document}

\maketitle

\begin{abstract}
  % teilw. von IJPR paper uebernommen
  An essential task in manufacturing planning and control is to determine when to release orders to
  the shop floor. One key parameter is the lead time which is the planned time that elapses between
  the release of an order and its completion. Lead times are normally determined based on the actual
  duration orders previously took to traverse through the production system (flow times).
  Traditional order release models assume static lead times, although it has been shown that they
  should be set dynamically to react the dynamic operational characteristics of the system.
  Therefore, we present an order release model which sets lead times dynamically by using an
  reinforcement learning approach. Therefore, we provide an in-depth analysis of reinforcement
  learning to show that average reward reinforcement learning is a better approach for operations
  research applications as discounted reinforcement learning. Additionally we present an average
  reward reinforcement learning algorithm which infers
  (Blackwell-)optimal\footnote{Blackwell-optimality is not yet fully proven! A first version of such
    a proof is attached in the Appendix.} policies.
  % \SH{Add RL stuff}
  % Therefore, we present a flow time estimation procedure to set lead times dynamically
  % by using an artificial neural network. Additionally, we implement a safety lead time
  % to incorporate the underlying cost ratio in the order release model. We test our
  % proposed approach using a simulation model of a three-stage make-to-order flow-
  % shop and compare the forecast accuracy and the accumulated costs to other forecast-
  % based order release models from literature. We show that our proposed model using
  We use a simulation model of a two-stage flow-shop to compare the algorithm
  to % and show that our approach outperforms
  well-known order release mechanisms. We show that in the current version our proposed model using
  reinforcement learning outperforms some, but not all other tested approaches.
  % EVTL: especially for scenarios with...
  \keywords{operations research, production planning, order release, machine learning, reinforcement
    learning}
\end{abstract}

\section{Introduction}
\label{sec:introduction} An important goal in Manufacturing Planning and Control (MPC) systems is to
achieve short and predictable flow times, especially where high flexibility in meeting customer
demand is required. Besides achieving short flow times, one should also maintain high output and
due-date performance while keeping the work-in-process level low. One approach to address this
problem is to collect all incoming orders in an order-pool and periodically decide which orders to
release to the shop floor. Once orders are released, costs start to accumulate as planned orders
materialise as actual jobs in the production system. The main challenge is to find a good compromise
of balancing the shop floor and timely completion of jobs. Although the performance of such systems
can be measured manifold the most overarching objective is to maximise profits by adequately
assigning holding and lateness costs.

One of the key modeling parameters for order release mechanisms is the \textit{lead time} which
refers to the planned time that elapses between the release of an order and its arrival in the
finished goods inventory. This planning parameter is often based on the observable time an order
needs to traverse through the production system, which we denote as \textit{flow time}. Flow times
consist of processing, setup, control, transport, and waiting times, whereas the latter is the
governing factor. Waiting times are a result from queuing (e.g. jobs queue before and after
processing) which depend heavily on the amount of jobs in the system (WIP) and thus, waiting times
are relatively difficult to estimate which makes the setting of favorable lead times so difficult
(e.g., \citealt{Tatsiopoulos1983, Wiendahl1995}). Most state-of-the-art order release mechanisms use
static or fixed lead times to address the order release problem, and thus neglect the nonlinear
relationship between resource utilisation and flow times which is well known from practice and
queuing theory (\citealt{Pahl2007}).

One way to address this nonlinear interaction effects is to set lead times dynamically. Thus,
intuitively the order release problem is solved by perfectly matching the lead times to the flow
times, but one faces ``sampling issues'' meaning that the flow times depend on the lead times. An
extreme scenario of this problem is the so called ``lead time syndrome'' which describes a vicious
cycle where increasing flow times perpetually inflate the lead times which leads to worse
performance (see e.g., \citealt{Mather1978, knollmann2013control, Selcuk2006}. Therefore,
\cite{SchneckenreitherHaeussler2019} propose to use reinforcement learning as approach to
incorporate these effects in this optimisation problem. Reinforcement learning is an optimisation
technique that stems from dynamic programming. The goal in reinforcement learning is to find the
best stationary policy for a given problem. This policy is usually provided by assessing the states
(or state-action pairs) of an underlying Markov Decision Process (MDP). The advantage of
reinforcement learning over dynamic programming is that (i) the problem space is explored by an
agent and thus only expectantly interesting parts of the problem space need to be assessed and (ii)
the knowledge (acquisition) of transition probabilities becomes unnecessary as the states are
evaluated by consecutively observed states solely.

Over the past decades reinforcement learning has been applied to various problems, for which
astonishing results have been reported. E.g. only recently \cite{mnih2015human} have presented a novel
value-iteration reinforcement learning agent which exceeds human-level abilities in playing many
classic Atari 2600 games~\citep{bellemare2012investigating}. \cite{mnih2016asynchronous} presents
improved results with asynchronous actor-critic reinforcement learning. Also games like
Go~\citep{silver2016mastering} and Chess~\citep{silver2017mastering} have been mastered with
superhuman performance by \textit{tabula rasa} reinforcement learning agents. Furthermore, the
method has also been applied in the setting of manufacturing system, e.g. to improve the ramp-up
process~\citep{doltsinis2012reinforcement}, in locally selecting appropriate dispatching rules
\citep{zhang1995reinforcement,wang2005application} or scheduling
\citep{zhang1995reinforcement, waschneck2018optimization}.
%
However, all these applications use discounted reinforcement learning and are either designed to
investigate a rather simple MDP, e.g. by selecting heuristics instead of optimising the underlying
problem itself, or by mapping the actual objective in a reward function that is approximately \(0\)
on average over time.
%
This is due to the fact that state value is largely composed of a term defined by the policy's
average reward value \citep{MillerVeinott1969,Blackwell62} which would otherwise dilute the state
values and thus decrease the solution quality, as for instance in
\cite{SchneckenreitherHaeussler2019} and \cite{gijsbrechts2018can}.
%

Therefore, most applications that incorporate and directly reflect costs or profit in the reward
function use average reward reinforcement learning. For instance \cite{aydin2000dynamic} use it in
the setting of scheduling, while in a series of papers Mahadevan et al. investigated several problem
domains starting with simple MDPs
\citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults,Mahadevan96_OptimalityCriteriaInReinforcementLearning}.
After these foundational works for average reward reinforcement learning they introduced a
continuous time average reward reinforcement learning algorithm named SMART
\citep{Mahadevan97_SelfimprovingFactorySimulationUsingContinuoustimeAveragerewardReinforcementLearning}.
Applications of SMART reach from the optimisation of queuing systems
\citep{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies,Mahadevan96_SensitiveDiscountOptimalityUnifyingDiscountedAndAverageRewardReinforcementLearning},
maintenance of an inventory system
\citep{Das99_SolvingSemiMarkovDecisionProblemsUsingAverageRewardReinforcementLearning} to optimising
transfer line in terms of maximizing demand, while keeping inventory levels of unfinished product as
low as possible \citep{Mahadevan98_OptimizingProductionManufacturingUsingReinforcementLearning}.

Like discounted reinforcement learning also average reward reinforcement learning is based on an
oracle function, in our case the accumulated profit of a period, to assess the decisions taken by
the agent. By repeatedly choosing different actions the agent examines the problem space and rates
possible actions for any observed state. The advantage of average reward reinforcement learning over
the widely applied discounted reinforcement learning framework is that the underlying optimisation
technique is able to find better policies. This yields from the fact that in the latter method the
states are assessed by a single value, while in the former the values are split up by the average
reward accumulated over time and a bias value which optimises the number of steps to reach the best
possible states of the examined process
\citep{Howard64,Puterman94,Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults}.
Thus, while in average reward reinforcement learning these terms are learned separately in the
discounted framework one value consisting of the addition of these values plus an additional error
term is estimated. This incautious combining of different kinds of state values as done in
discounted reinforcement learning leads to the problems that (i) the average reward, which is equal
for all states of usually investigated unichain MDPs, is dominating and thus diluting the bias
values, and (ii) the state values are deteriorated by the error term that is only imposed due to the
discounting technique.

Thus, as opposed to the commonly applied discounted reinforcement learning algorithm and as proposed
as future work by \cite{SchneckenreitherHaeussler2019} we use an average reward reinforcement
learning algorithm to adaptively release orders based on the assessed state values of the production
system. In contrast to the aforementioned works on average reward reinforcement learning our
algorithm incorporates the optimisation of not only the average reward over time, but also the bias
values. Only the work by
\cite{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}
integrates this second-level refinement optimisation. However, their algorithm requires the
selection of a reference state to prevent an infinite increase of state values. This results from
the lack of feedback in the iterative process of optimising the different refinement optimisation
levels.

Therefore, we propose a novel average reward reinforcement learning algorithm, which is the first
that solves the occurring cyclic constraint problem in the setting of average reward reinforcement
learning, to assess orders for their release to the shop floor. These cyclic constraint problem
emerges as the underlying constraint structure is based on the Laurent series expansion of the state
values as shown by \cite{MillerVeinott1969}, and thus easily imposes an infinite number of
interconnected constraints when handled unwisely.
%
The agent assesses the expected costs for the possible releases imposed by adapting the lead time.
According to the agent's estimates it sets a planned lead time and with that releases orders into
the production system.
% upon releasing an order for the current and future periods
% and, based on its estimates decides whether to release the order or not.
%
% \MS[t]{TODO} We provide ample evidence of the viability of the approach and compare it to well-known
% order release mechanisms. Furthermore, we show that the advantage of average reward reinforcement
% learning for the problem structures encountered in the context of production and logistics by
% comparing the results to the discounted reinforcement learning framework. The preliminary results
% show that our approach performs better than the other methods. With regard to practical implications
% we are confident that decision support tools based on average reward reinforcement learning increase
% the decision quality of human planners.

\paragraph{Structure.} The rest of the paper is structured as follows. The next section introduced
average reward reinforcement learning and presents the used algorithm (see also
\citealt{Schneckenreither2019}) including a proof-of-concept.
Section~\ref{sec:Experimental_Evaluation} describes the simulation model we use to evaluate the
approach. Then~\ref{sec:Preliminary} provides preliminary result data, wheras
Section~\ref{sec:conclusion} concludes this working paper.

\section{Average Reward Reinforcement Learning}
\label{sec:Average_Reward_Reinforcement_Learning}

This section introduces average reward reinforcement learning in comparison to the discounted
framework, elaborates on optimiality criteria and provides insights of the underlying algorithm.

Like \cite{MillerVeinott1969} we are considering problems that are observed in a sequence of points
in time labeled \(1,2,\ldots\) and can be modelled using a finite set of states \(\States\),
labelled \(1,2,\ldots,\size{\States}\), where the size \(\size{\States}\) is the number of elements
in \(\States\). At each point $t$ in time the system is in a state \(s_{t} \in \States\). Further,
by choosing an action $a_{t}$ of a finite set of possible actions \(A_{s}\) the system returns a
reward $r_{t} = r(s_{t}, a_{t})$ and transitions to another state \(s_{t+1} \in \States\) at time
\(t+1\) with conditional probability \(p(s_{t+1}, r_{t} \mid s_{t}, a_{t})\). That is we assume that
reaching state \(s_{t+1}\) from state \(s_{t}\) with reward \(r_{t}\) depends solely on the previous
state \(s_{t}\) and chosen action \(a_{t}\). In other words, we expect the system to possess the
Markov property \citep[p.63]{sutton1998introduction}. Reinforcement learning processes that possess
the Markov property are referred to as Markov decision processes (MDPs)
\citep[p.66]{sutton1998introduction}.
% \MS{does our system possesses the Markov property at all? :-o }

Thus, the action space is defined as \(F = \times_{s=1}^{\size{\States}} A_{s}\), where \(A_{s}\) is a
finite set of possible actions. A \emph{policy} is a sequence \(\pol = (f_{1},f_{2},\ldots)\) of
elements \(f_{t} \in F\). Using the policy \(\pol\) means that if the system is in state \(s\) at
time \(t\) the action \(f_{t}(s)\), i.e.% that is
the \(s\)-th component of \(f_{t}\), is chosen. A stationary policy \(\pol = (f,f,\ldots)\) does not
depend on time. In the sequel we are concerned with stationary policies only. Thus the goal in
reinforcement learning (RL) is to find the best stationary policy \(\pol\) for a given problem,
where the phrase ``best'' will be elaborated in detail below.

\subsection{Discounted Reinforcement Learning}
\label{subsec:Discounted_Reinforcement_Learning}

In the discounted framework the value of a state \(V_{\gamma}^{\pol}(s)\) is defined as the expected
discounted sum of rewards under the stationary policy \(\pol\) when starting in state \(s\). That is
\begin{align*}
  V_{\gamma}^{\pol}(s)=\lim_{N \to \infty} E[\sum_{t=0}^{N-1} \gamma^{t} R_{t}^{\pol}(s)]\tcom
\end{align*}

where \(0 \leqslant \gamma < 1\) is the discount factor and
\(R_{t}^{\pol}(s) = E_{\pol}[ r(s_{t},a_{t}) \mid s_{t} = s, a_{t} = a]\) % \MS[t]{check if correct!}
the reward received at time \(t\) upon starting in state \(s\) by following policy \(\pol\)
\citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults}.

The aim in the discounted framework is to find an optimal policy \(\polopt\), which when followed,
maximises the state value for all states \(s\) as compared to any other policy \(\pol\):
\(V_{\gamma}^{\polopt} - V_{\gamma}^{\pol} \geqslant 0\). This criteria is usually referred to as
\(\gamma\)-optimality as the discount factor \(\gamma\) is fixed. However, this also means that the
actual value set for \(\gamma\) defines the policy that is optimised for. For instance, as can be
seen in Figure~\ref{fig:printer} setting \(\gamma\) value to \(<0.8027\) the printer-loop is
preferred over the mail-loop by the agent, although the mail-loop accumulates more reward over time,
i.e. selecting the mail-loop is a sub-optimal choice.

\begin{figure}[t!]
  \centering
  \begin{tikzpicture}[thin, scale=0.75]
    % Nodes
    \foreach \x in {1,...,5}{
      \draw ({cos(\x*72-72)*2.5},{sin(\x*72-72)*2.5}) node(\x) [circle,draw,minimum size=25] {\footnotesize \(\x\)};
    };
    \foreach[evaluate={
      \y=int(\x+5);
    }]  \x in {2,...,3}{
      \draw ({-cos(\x*72-72)*2.5+5},{sin(\x*72-72)*2.5}) node(\y) [circle,draw,minimum size=25]
      {\footnotesize \(\x'\)};
    };
    \foreach [evaluate={
      \y=int(\x+5);
      \z=int(\x+5);
    }] \x in {4,...,5}{
      \draw ({-cos(\x*72-72)*2.5+5},{sin(\x*72-72)*2.5}) node(\z) [circle,draw,minimum size=25]
      {\footnotesize \(\y'\)};
    };

    % Edges
    \path[thin, ->, bend right, >=stealth] (1) edge[below left] node {\footnotesize\(\)} (2);
    \path[thin, ->, bend right, >=stealth] (2) edge[above]      node {\footnotesize\(\)} (3);
    \path[thin, ->, bend right, >=stealth] (3) edge[left ]      node {\footnotesize\(\)} (4);
    \path[thin, ->, bend right, >=stealth] (4) edge[above]      node {\footnotesize\(\)} (5);
    \path[thin, ->, bend right, >=stealth] (5) edge[above left] node {\footnotesize\(5\)} (1);

    \path[thin, ->, bend left, >=stealth]         (1)  edge[below right] node {\footnotesize\(\)} (7);
    \path[thin, ->, bend left, >=stealth]         (7)  edge[above right] node {\footnotesize\(\)} (8);
    \path[thin, ->, dashed, bend left, >=stealth] (8)  edge[right]       node {\(\ldots\)} (9) ;
    \path[thin, ->, bend left, >=stealth]         (9)  edge[above]       node {\footnotesize\(\)} (10);
    \path[thin, ->, bend left, >=stealth]         (10) edge[above right] node {\footnotesize\(20\)} (1);

    \draw (0,0) node[] { Printer };
    \draw (5,0) node[] { Mail };

  \end{tikzpicture}
  \caption{\label{fig:printer} A MDP with two different determinstic policies. The only action choice is in
    state \(1\), in which the agent can choose between doing the printer-loop or the mail-loop.
    Observe that the average reward received per step equals \(1\) for the printer-loop and \(2\)
    for the mail-loop. Thus, the (Blackwell-)optimal policy is to choose the mail-loop. However,
    if \(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted reinforcement
    learning chooses the printer loop.
    %
    Adapted from \cite{Mahadevan96_OptimalityCriteriaInReinforcementLearning}.}
\end{figure}

The main idea behind the \(\gamma\)-parameter of the discounted framework is to be able to balance
short-term rewards (low \(\gamma\)-values) and long-term rewards (high \(\gamma\)-values). However,
what seems to be an advantage rather becomes a disadvantage for most applications. The issue is that
the average reward value is non-linearly increased when \(\gamma\) approaches one. However, in
almost all cases the aim is to perform well over time which in terms of reward means to first
maximise for a the policy with highest average reward before more selectively choosing actions. That
is, if several processes exist, we are searching for the ones with highest average reward before
considering other criteria.
%
Therefore, in almost all RL studies the discount factor is set to a value very close to \(1\), for
instance to \(0.99\)~\cite[e.g.]{mnih2015human,mnih2016asynchronous,Lillicrap15}.
%
By doing so the above mentioned non-linear relationship leads to the fact that the state value
consists almost solely of the up-scaled average reward term, whereas the bias values can be
neglected. This leads to diluted state values and thus difficulties in distinguishing actions which
impose policies that possess the same average reward.
%
In contrast to that, average reward RL separately assesses these values and according to these
selectively chooses the best action, comparing one after the other.
%
To illustrate the idea reconsider Figure~\ref{fig:printer}. First the agent picks the policy
according to the highest average reward. Thus, all actions are assessed by an average reward of
\(2\) which is inferred from the Mail-loop. This however, makes choosing the Printer-loop
unattractive; i.e. the policy converges to the optimal policy of choosing the Mail loop.

\subsection{Average Reward Reinforcement Learning}
\label{subsec:Average_Reward_Reinforcement_Learning}

Due to \cite{Howard64} the average reward \(\avgrew^{\pol}(s)\) of a policy \(\pol\) and a starting
state \(s\) is defined as
\begin{align*}
  \avgrew^{\pol}(s) = \lim_{N \to \infty} \frac{\E [\sum_{t=0}^{N-1}R_{t}^{\pol}(s)]}{N}\tpkt
\end{align*}

In the common case of unichain MDPs, in which only a single set of recurrent states exists, the
average reward \(\avgrew^{\pol}(s)\) is equal for all states \(s\)
\citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults,Puterman94}.
For simplicity we concentrate on unichain MDPs in this work and thus may simply refer to it as
\(\avgrew^{\pol}\). A policy that maximises the average reward in every state is called
gain-optimal. Gain optimality is the least selective criteria an average reward reinforcement
learning aims for.

Further, for an unichain aperiodic\footnote{In the periodic case the Cesaro limit of degree \(1\)
  is required to ensure stationary state transition probabilities and thus stationary bias
  values \citep{Puterman94}. Therefore to ease readability we concentrate on unichain
  and aperiodic MDPs.} MDP problem, such as ergodic MDPs are, the average adjusted sum of rewards or
bias value is defined as
\begin{align*}
  V^{\pol}(s) = \lim_{N \to \infty}{ \E [ \sum_{t=0}^{N-1}(R_{t}^{\pol}(s) - \avgrew^{\pol} )]}\tcom
\end{align*}

where again \(R_{t}^{\pol}(s)\) is the reward received at time \(t\), starting in state \(s\) and
following policy \(\pol\). Note that the bias values are bounded due to the subtraction of the
average reward. Thus the bias value can be seen as the rewards that additionally sum up in case the
process starts in state \(s\). A policy that is gain-optimal is also bias-optimal if it maximises
the bias values in every state and compared to every other policy.

Especially for highly probabilistic systems bias-optimality is important. To clarify this consider
Figure~\ref{fig:three-states} which again consists of two possible deterministic policies with the
only choice in state \(1\). Both policies have the same average reward of \(1\). However, only
selecting the A-loop is bias-optimal, as the actions with non-zero rewards are selected earlier.
E.g.~consider starting in state 1. Under the policy \(\pol_{A}\) which takes the A-loop the reward
sequence is \((2,0,2,0,\ldots)\), while for the other policy \(\pol_{B}\) it is
\((0,2,0,2,\ldots)\).

\begin{figure}[t!]
  \centering
  \begin{tikzpicture}[thin, scale=0.75]
    % Nodes
    \draw (-3,0) node(0) [circle,draw,minimum size=25] {\footnotesize \(0\)};
    \draw (0,0)  node(1) [circle,draw,minimum size=25] {\footnotesize \(1\)};
    \draw (3,0)  node(2) [circle,draw,minimum size=25] {\footnotesize \(2\)};

    % Edges
    \path[thin, ->, bend right, >=stealth] (1) edge[above] node {\footnotesize\(2\)} (0);
    \path[thin, ->, bend right, >=stealth] (2) edge[above] node {\footnotesize\(2\)} (1);
    \path[thin, ->, bend right, >=stealth] (0) edge[below] node {\footnotesize\(0\)} (1);
    \path[thin, ->, bend right, >=stealth] (1) edge[below] node {\footnotesize\(0\)} (2);

    \draw (-1.5,0) node[] { A };
    \draw (1.5,0) node[] { B };

  \end{tikzpicture}
  \caption{\label{fig:three-states} A task with two different MDPs, A (going left in 1) and B
    (going right). As in the previous example the only action choice is in state \(1\). Observe
    that the average reward received per step equals \(1\) for both policies. However, only taking
    the A loop is bias-optimal, as its policy \(\pol_{A}\) leads to bias values
    \(V^{\pol_{A}}(1)=0.5\), \(V^{\pol_{A}}(0)=-0.5\) and \(V^{\pol_{A}}(2)=1.5\), while policy
    \(\pol_{B}\) which selects the B loop generates bias values \(V^{\pol_{B}}(1)=-0.5\),
    \(V^{\pol_{B}}(0)=-1.5\) and \(V^{\pol_{B}}(2)=0.5\). Adapted from
    \cite{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults}.}
\end{figure}

\subsection{More Refined Optimality Criteria}
\label{subsec:More_Refined_Optimality_Criteria}

The aforementioned and informally introduced notions of gain-optimality and bias-optimality can be
generalised to \(n\)-discount-optimality and will now be defined formally.

\begin{definition}
  Due to \cite{Veinott69} for MDPs a policy \(\polopt\) is \emph{\(n\)-discount-optimal} for
  \(n=-1,0,1,\ldots\) for all states \(s \in \States\) with discount factor
  \(\gamma\) % and \(V_{\gamma}^{\pol}(s)\) being the value function as defined above
  %
  if and only if
  \begin{align*}
    \lim_{\gamma \to 1}(1-\gamma)^{-n}\ (V_{\gamma}^{\polopt}(s) - V_{\gamma}^{\pol}(s)) \geqslant 0 \tpkt
  \end{align*}
\end{definition}

For the case of \(n=-1\) this lead to gain-optimality
\(\avgrew^{\polopt}(s) - \avgrew^{\pol}(s) \geqslant 0\) for all policies \(\pol\) and states \(s\)
\citep{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}. In
the case of \(n=0\) it describes bias-optimality with \(V^{\polopt}(s) - V^{\pol}(s) \geqslant 0\)
forall policies \(\pol\) and states \(s\)
\citep{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}.
%
Only if a policy is \(n\)-discount-optimal for all \(n < m\) it can be \(m\)-discount
optimal \citep{Puterman94,Veinott69}.
%
If a policy is \(\infty\)-discount-optimal then it is said to be
Blackwell-optimal \citep{Blackwell62}.
% weakening the argument!!!
I.e.~for Blackwell-optimal policies \(\polopt\) there exists a discount factor \(\gammaopt < 1\)
such that \(V_{\gamma}^{\polopt}(s) \geqslant V_{\gamma}^{\pol}(s)\) for all
\(\gamma \geqslant \gammaopt\) and under all policies \(\pol\)
\citep{Mahadevan96_SensitiveDiscountOptimalityUnifyingDiscountedAndAverageRewardReinforcementLearning,Blackwell62}.
Informally that means there exists a discount factor \(<1\) which finds Blackwell-optimal policies.
However, (i) in more complex, i.e. real world, MDPs this value can be arbitrary close to \(1\) and
(ii) the difference in state values may be very small, which (iii) due to the need of state value
function approximation likely causes errors when choosing among different actions in the discounted
framework.
% Real world applications underlie much more complex MDPs as the ones of Figures~\ref{fig:printer}
% and~\ref{fig:three-states}. Thus \(\gamma\)-values even very close to \(1\) can lead to sub-optimal
% policies. To investigate these differences we will now present a more elective optimality criteria.
%
%

In the sequel we present the Laurent series expansion which not only links discounted reinforcement
learning with average reward reinforcement learning, but also connects to \(n\)-discount-optimality.

% Furthermore, for the rest of the paper we occasionally write drop the \(V(s)\) and $V_{\gamma}$ instead of
% \(V^{\pol}(s)\) and \(V^{\pol}_{\gamma}\) when the policy is clear from the context.

\subsection{The Laurent Series Expansion of Discounted State Values}
\label{subsec:The_Laurent_Series_Expansion_of_Discounted_State_Values}

\cite{MillerVeinott1969} established the link between discounted RL state values
\(V_{\gamma}^{\pol}(s)\) and average reward RL values \(\avgrew^{\pol}(s)\) and \(V^{\pol}(s)\)
using the Laurent series expansion as
\begin{align*}
  V_{\gamma}^{\pol}(s) = \frac{\avgrew^{\pol}(s)}{1-\gamma} + V^{\pol}(s) + e_{\gamma}^{\pol}(s) \tcom
\end{align*}

where \cite{Puterman94} shows that \(\lim_{\gamma \to 1} e_{\gamma}^{\pol}(s) = 0\). Note how the
first term depending on the average reward \(\avgrew^{\pol}(s)\) converges to infinity as \(\gamma\)
increases.

However, an important insight is the connection between \(n\)-discount-optimality and the Laurent
series expansion of \(V_{\gamma}^{\pol}(s)\). Each addend corresponds to one step in
\(n\)-discount-optimality. That is for \((-1)\)-discount-optimality the average reward must be
maximised.
%
Then for \(0\)-discount-optimality the agent has to choose the highest bias values.
%
Finally \(n\)-discount-optimality for \(n \geqslant 1\) requires to maximise the term
\(e_{\gamma}^{\pol}(s)\) which only exists as \(\gamma\) is strictly less than \(1\). Thus this term
incorporates the number of steps its expected rewards on the path which are required for reaching
the highest rewards. That is, it optimises the expected reward according to the occurrence on the
paths, where shorter paths and those which collect the rewards sooner are preferred.

Note how the Laurent series expansion dissipates these values in single terms and therefore defines
a divide and conquer methodological approach to reinforcement learning. Thus in average reward
reinforcement learning these terms are learned separately, while in the discounted framework the
combined value for a fixed \(\gamma\)-value is learned.

The addends, where \(n=-1,0,\ldots\) denote the coefficients of the Laurent series expansion,
can be reformulated to following constraint problem with expected reward
\(R^{\pol}(s) = E_{\pol}[ r(s, a) ] = \sum_{a \in \mathcal{A}(s)}p(a | s) r(s, a)\) as shown by
\cite{MillerVeingsott1969} and \citet[p.346]{Puterman94}:

\begin{align}
  \label{eq:constr1}\avgrew^{\pol}(s) - E[\avgrew^{\pol}(s)] & = 0 &  & \text{for } n = -1 \\
  \label{eq:constr2}\avgrew^{\pol}(s) + V^{\pol}(s) - E[V^{\pol}(s)] & = R^{\pol}(s) && \text{for } n = 0 \\
  \label{eq:constr3}W^{\pol}_{n-1}(s) + W^{\pol}_{n}(s) - E[W^{\pol}_{n}(s)] & = 0 && \text{for } n \geqslant 1, \text{where } W_{0}^{\pol}(s) = V^{\pol}(s)
\end{align}

\citet[p.343ff]{Puterman94} shows that due to the given degree of freedom if \(n=-1,0,\ldots,M\)
constraints are satisfying the above conditions for all states \(s\), then only
\(\avgrew^{\pol}(s), V^{\pol}(s), W^{\pol}_{1}, \ldots, W^{\pol}_{M-1}\) are unique, whereas
\(W^{\pol}_{M}\) is offset by the vector \(u\) where for the transition probability matrix \(P\) the
vector \(u\) is characterised by \((I-P)u = 0\). Note that \(u\) is determined by the number of
closed irreducible classes of \(P\), that is for ergodic MDPs \(u\) is determined by a single
constant. Our algorithm is based on the above formulation.

\subsection{Algorithm}
\label{subsec:Algorithm}

\begin{algorithm}[t!]
  \begin{algorithmic}[1]
    \State{}Initialize values to \(0\) \While{the stopping criterion is not fulfilled}
    \State{}\begin{minipage}[t]{0.9\textwidth} With probability \(p_{exp}\) choose a random action
      and probability \(1-p_{exp}\) one that maximizes \MS[t]{max V$_{\gamma_{1}}(s,a)$ if more than
      one e?}
      \(\lex(\avgrew^{\pol}(s,a),\V^{\pol}(s,a),\Delta e^\pol_{\gamma_{1},\gamma_{0}}(s,a))\) at the
      current state \(s\) where
      \(\Delta e_{\gamma_{1},\gamma_{0}}^{\pol}(s,a) = V_{\gamma_{1}}^{\pol}(s,a) -
      V_{\gamma_{0}}^{\pol}(s,a) - \avgrew^{\pol}(s,a) (\frac{1}{1-\gamma_{1}} -
      \frac{1}{1-\gamma_{0}})\)
      % \(\Delta e^\pol_\gamma_{1},\gamma_{0}(s,a) = V^\pol_{\gamma_0}(s,a) - V^\pol_{\gamma_1}(s,a)\)
      for all actions for which
      % \(\frac{\avgrew^{\pol}(s,a)}{1-\gamma_{0}} + \V^{\pol}(s,a) < V^\pol_{\gamma_0}(s,a)\)
      \(\Delta e_{\gamma_{1},\gamma_{0}}^{\pol}(s,a) < 0\) holds or if no such actions exists, from
      all actions. Set \(a^{\pol} = 0\) if a random action was chosen,
      \(a^{\pol}=1\) otherwise.
    \end{minipage}
    \State{}Carry out action \(a\), observe reward \(r\) and resulting state \(s'\)
    \State{}Update values using exponential smoothing
      \MS[t]{why only random/non-random for rho and enforcing? very inconsistent!!!}
    \begin{align*}
      \qquad a) & & \avgrew^{\pol}(s,a)  & \gets (1- a^{\pol} \alpha) \avgrew^{\pol}(s,a) + a^{\pol} \alpha [r + \max_{a}\V^{\pol}(s',a) - \V^{\pol}(s,a)]     \\
      b) & & \V^{\pol}(s,a)        & \gets (1-\beta) \V^{\pol}(s,a) + \beta  \cdot \\
         &        &                      & [r - \avgrew^{\pol}(s,a) + \max_{a} \V^{\pol}(s',a) +
                                           a^{\pol} (\Psi_{V}^{\pol}(s,a) + \zeta \Psi_{W}^{\pol}(s,a))] \\
      c) & & W^{\pol}(s,a)        & \gets (1-\delta) W^{\pol}(s,a) + \delta [-\V^{\pol}(s,a) +
                                    \max_{a} W^{\pol}(s',a)]              \\
      d) & & \Psi_{V}^{\pol}(s,a) & \gets (1-\xi \beta) \Psi_{V}^{\pol}(s,a) + \xi \beta [r - \avgrew^{\pol}(s,a) - V(s,a) + \max_{a}V^{\pol}(s',a)]\\
      e) & & \Psi_{W}^{\pol}(s,a) & \gets (1-\xi \delta) \Psi_{W}^{\pol}(s,a) + \xi \delta [- V^{\pol}(s,a) - W^{\pol}(s,a) + \max_{a} W^{\pol}(s',a)]
    \end{align*}
    \State{Update discounted state-values}
    \begin{align*}
      V^\pol_{\gamma_0}(s,a)    & \gets (1-\gamma) V^\pol_{\gamma_0}(s,a) + \gamma (r + \gamma_{0} \max_{a} V^\pol_{\gamma_0}(s',a)) \\
      V^\pol_{\gamma_1}(s,a)    & \gets (1-\gamma) V^\pol_{\gamma_1}(s,a) + \gamma (r + \gamma_{1} \max_{a} V^\pol_{\gamma_1}(s',a))
    \end{align*}
    \State{}Set \(s \gets s'\), \(t \gets t+1\) and decay parameters accordingly
    \EndWhile{}
  \end{algorithmic}
  \caption{\label{alg}Model-free tabular version of the Blackwell-optimal RL algorithm.}
\end{algorithm}

% \MS{
%   Optimisations:
%   \begin{enumerate}
%   \item Use Lemma 3 of Miller \& Veinott to get rid of W2
%   \item Use Reference state as in Mahadevan to get rid of W
%   \item Average reward freed reinforcement learning
%   \end{enumerate}
% }

The reinforcement learning algorithm is depicted in Algorithm~\ref{alg}. After initialising all
values the agent enters the loop in which the first task is to choose an action (step 3). In this
action selection process we utilise an \(\epsilon\)-sensitive lexicographic order
\(a=(a_{1},\ldots,a_{n}) \lex (b_{1},\ldots,b_{n})=b\) defined as \(a \lex b\) if and only if
\(| a_{j} - b_{j} | \leqslant \epsilon\) for all \(j < i\) and \(|a_{i} - b_{i}| > \epsilon\). Note
that the resulting sets of actions may not be disjoint. Although this is an unusual order in
programming, taking the maximum as in our algorithm is straight-forward and thus cheap to compute.
%
The first action selection criteria maximises the average reward. In case of unichain and thus
ergodic MDPs only one scalar average reward value over all states is learned and thus the problem
simplified by abolishing the comparison in the first component. Then the set of actions which
maximise the bias values are selected. Finally, \(\Delta V^\pol_\gamma(s,a)\) is used to estimate
the slope of the error term \(e_{\gamma}^{\pol}(s,a)\), which can be either positive or negative.
Therefore, there is the need of a distinction of which actions to maximise according to the
discounted state value difference
\(\Delta V^{\pol}_{\gamma}(s,a) = V^\pol_{\gamma_1}(s,a) - V^\pol_{\gamma_0}(s,a)\). As the aim is
to collect rewards as soon as possible, a Blackwell-optimal reinforcement learning agents maximises
the error term, thus any actions converging to
\(\frac{\avgrew^{\pol}(s,a)}{1-\gamma_{0}} + \V^{\pol}(s,a)\) from above are preferred over those
that converge from below. Within these sets the action which maximises
\(\Delta V^{\pol}_{\gamma}(s,a)\) is selected. If multiple such actions the agent can choose
randomly among these.

As usual in reinforcement learning the equations are integrated by exponentially smoothing the
values utilising parameters \(\alpha, \beta, \gamma, \delta\).
% Further, as in this paper we concentrate on unichain MDPs the average reward
% \(\avgrew^{\pol}(s,a)\) can be learned as a scalar value \(\avgrew^{\pol}\), whereas the algorithm
% is supports multichain MDPs, cf. step \(5a\).
Step \(5a\) uses the Bellman equation for average reward reinforcement learning
(cf.~\citealt{Howard64, tadepalli1998model}) to find the average reward. Then step \(5b\), \(5c\)
and \(5d\) depict Equations (\ref{eq:constr2}) and (\ref{eq:constr3}) for \(n=0\), \(n=1\) and
\(n=2\). The need of learning values for \(n=2\) stems from the fact that the actual values
corresponding to the highest \(n\) are offset by vector \(u\).

% For \(n=0\) (step \(5b\)) the bias values are learned and further to ensure a unique solution of
% \(V^{\pol}\) step \(5c\) solves the constraints for \(n=1\). Thus, the first three assignments of
% step \(5\) correspond to the above constraint system for \(n=-1,0,1\). The last line of step \(5\)
% evaluates the deviation of \(W^{\pol}(s,a)\), which ought to be \(0\) once the system is correctly
% solved and if an action of the optimal policy \(\polopt\) was taken.
Step \(6\) infers the discounted state values using learning rate \(\gamma\) and two discount rates
\(0.5 \leqslant \gamma_{0} < \gamma_{1} < 1\) to estimate the state value difference
\(\Delta V^{\pol}_{\gamma}(s,a)\).

Steps \(7\) and \(8\) are crucial, but were not considered in average reward reinforcement learning
before. They feed back the deviations \(\Psi_{\cdot}(s,a)\) into the constraint problem creating a
recursive system of equations. Depending on the underlying MDP when this constraint is omitted one
likely observes steadily increasing deviations and therefore a wrong solution. This feedback loop
however is only activated in case a non-random action was chosen, as otherwise the deviations
\(\Psi_{\cdot}(s,a) > 0\) \citep{MillerVeinott1969}.

%
%
% Steps \(5d\) and \(5e\) calculate the deviations of actual values and expected values. In case the
% agent performs a non-random action
%
Recall that the algorithm evaluates the slope of \(e_{\gamma}^{\pol}(s)\) by learning the state
values using two different \(\gamma\) settings and discounted reinforcement learning. The slope of
\(e_{\gamma}^{\pol}(s)\) combined with the observation that for Blackwell-optimal policies
\(\polopt\) there exists a discount factor \(\gammaopt < 1\) such that
\(V_{\gamma}^{\polopt}(s) \geqslant V_{\gamma}^{\pol}(s)\) for all \(\gamma \geqslant \gammaopt\)
and under all policies \(\pol\)
\citep{Mahadevan96_SensitiveDiscountOptimalityUnifyingDiscountedAndAverageRewardReinforcementLearning,Blackwell62}
provides the most selective criteria for our algorithm, which theoretically can find
Blackwell-optimal policies\footnote{This claim is only not fully proven yet. See the appendix for a
  first version of the proof. } (see Section~\ref{sec:methods} in Appendix).

\subsection{Proof-of-Concept}
\label{subsec:Proof-of-Concept}

Reconsider the task depicted in Figure~\ref{fig:three-states}. The Blackwell-optimal MDP
\(\pol_{A}\), that is the one that chooses the A-loop, is shown on the left side of
Figure~\ref{fig:three-states2}. Recall that the bias values for \(\pol_{A}\) are
\(V^{\pol_{A}}(0)=-0.5\), \(V^{\pol_{A}}(1)=0.5\) and \(V^{\pol_{A}}(2)=1.5\). Our model-based
implementation infers the values \(\avgrew^{\pol_{A}} = 1\), \(V^{\pol_{A}}(0)=-0.5\), \(V^{\pol_{A}}(1)=0.5\)
and \(V^{\pol_{A}}(2)=1.4391\) fully automatically in \(200k\) steps. Furthermore, the addition of
\(V^{\pol_{A}}(2) + \Psi^{\pol_{A}}_{V}(2) = 1.4998\), where \(\Psi^{\pol_{A}}_{V}(2)\) is the exponentially
smoothed deviation for state \(2\). Thus the model-based algorithm is able to correctly approximate
the mathematical solution.

The right side of Figure~\ref{fig:three-states2} shows the same MDP for the model-free version. In
model-free reinforcement learning state-action pairs as opposed to state values are estimated. Thus
each state is replicated as often, as actions are available in that state. Therefore, state \(1\) is
replaced by states-action pairs \((1,l)\) and \((1,r)\) for moving to the left (taking the A-loop)
and to the right respectively. The dashed line indicates that these states are connected and thus
the agent has to choose among them when reaching either. In this case the inferred bias values are
\(V^{\pol_{A}}(0,r) = -0.5\), \(V^{\pol_{A}}(1,l)=V^{\pol_{A}}(1,r)=0.5\), and \(V^{\pol_{A}}(2,l)=1.5\). The above given tabular model-free
average reward reinforcement learning algorithm is able to exactly infer these values in less than
100k steps. This provides evidence that the algorithm in general infers correct solutions in regard to
the underlying constraints.

% \MS{Occasionally we write \(V^{\pol_{A}}(s)\) and $V_{\gamma}$ instead of \(V^{\pol}(s)\) and
% \(V^{\pol}_{\gamma}\) when the policy is clear from the context.}

Furthermore, in the later case \(V^{\pol_{A}}(1,l)=V^{\pol_{A}}(1,r)=0.5\), that is if the agent
resides in state \(1\) it has to choose among the both available actions. This cannot be done on
basis of the average reward, nor on the bias values. Therefore, it infers the slopes of the error
terms using the difference of the discounted state values \(V^{\pol_{A}}_{\gamma_{1}}(1,l) = 5.56\)
and \(V^{\pol_{A}}_{\gamma_{0}}(1,l) = 2.67\) for state-action pair \((1,l)\) and
\(V^{\pol_{A}}_{\gamma_{1}}(1,r) = 5.16\) and \(V^{\pol_{A}}_{\gamma_{0}}(1,r) = 1.67\) for
\((1,r)\), yielding \(\Delta V^{\pol_{A}}_\gamma(1,l) = -2.89\) and
\(\Delta V^{\pol_{A}}_\gamma(1,r) = -3.49\), where \(\gamma_{0} = 0.5\) and \(\gamma_{1}=0.8\).

The idea is that rewards are collected as soon as possible, thus the error term has to be maximised
when \(\gamma \to 1\). As we have
\(2.5 = \frac{\avgrew^{\pol_{A}}}{1-\gamma_{0}} + V^{\pol_{A}}(1,a) < V^{\pol_{A}}_{\gamma_{0}}(1,l)
= 2.67\), while \(2.5 \geqslant 1.67\), we choose action \(l\) as it is the only action with
positive error term.

% Current state:                               A
% Period:                                      5000
% Alpha:                                       0.0050
% Beta:                                        0.0700
% Delta:                                       0.0700
% Gamma:                                       0.0700
% Epsilon:                                     1.2913
% Exploration:                                 0.8609
% Algorithm:                                   BORL with gammas  (0.5,0.8); state values for rho; Deciding on V
% Zeta (for forcing V instead of W):           1.0000
% Xi (ratio of W error forcing to V):          0.2000
% Scaling (V,W,R0,R1) by V Config:             Tabular representation (no scaling needed)
% Psi Rho/Psi V/Psi W:                         (0.0000,0.0000,0.0000)
% Rho:                                         1.0000
% V                                       W
% [+0.0000]right: -0.5000                 [+0.0000]right: 0.2061
% [+1.0000]left : 0.5000                  [+1.0000]left : -0.2939
% [+1.0000]right: 0.5000                  [+1.0000]right: -2.2939
% [+2.0000]left : 1.5000                  [+2.0000]left : -1.7939
% R0                                      R1
% [+0.0000]right: 1.3333                  [+0.0000]right: 4.4444
% [+1.0000]left : 2.6667                  [+1.0000]left : 5.5556
% [+1.0000]right: 1.6667                  [+1.0000]right: 5.1556
% [+2.0000]left : 3.3333                  [+2.0000]left : 6.4444
% PsiV
% [+0.0000]right: -0.0000
% [+1.0000]left : -0.0000
% [+1.0000]right: -0.0000
% [+2.0000]left : -0.0000

\begin{figure}[t!]
  \centering
  \begin{tikzpicture}[thin, scale=0.75]
    % Nodes
    \draw (-3,0) node(0) [circle,draw,minimum size=25] {\footnotesize \(0\)};
    \draw (0,0)  node(1) [circle,draw,minimum size=25] {\footnotesize \(1\)};
    \draw (3,0)  node(2) [circle,draw,minimum size=25] {\footnotesize \(2\)};

    % Edges
    \path[thin, ->, bend right, >=stealth] (1) edge[above] node {\footnotesize\(2\)} (0);
    \path[thin, ->, bend right, >=stealth] (2) edge[above] node {\footnotesize\(2\)} (1);
    \path[thin, ->, bend right, >=stealth] (0) edge[below] node {\footnotesize\(0\)} (1);
    % \path[thin, ->, bend right, >=stealth] (1) edge[below] node {\footnotesize\(0\)} (2);

    \draw (-1.5,0) node[] { A };
    \draw (1.5,0) node[] { B };

    \draw (6,0) node(B0) [circle,draw,minimum size=25] {\footnotesize \((0,r)\)};
    \draw (9,1.0)  node(B1l) [circle,draw,minimum size=25] {\footnotesize \((1,l)\)};
    \draw (9,-1.0)  node(B1r) [circle,draw,minimum size=25] {\footnotesize \((1,r)\)};
    \draw (12,0)  node(B2) [circle,draw,minimum size=25] {\footnotesize \((2,l)\)};

    \path[thin, ->, bend right, >=stealth] (B1l) edge[above] node {\footnotesize\(2\)} (B0);
    \path[thin, ->, bend right, >=stealth] (B2) edge[above] node {\footnotesize\(2\)} (B1l);
    \path[thin, ->, bend right, >=stealth] (B0) edge[below] node {\footnotesize\(0\)} (B1l);
    \path[thin, ->, bend right, >=stealth] (B1r) edge[below] node {\footnotesize\(0\)} (B2);

    \path[thin, dashed] (B1r.west) edge[below] node {} (B1l.west);
    \path[thin, dashed] (B1r.east) edge[below] node {} (B1l.east);

  \end{tikzpicture}
  \caption{\label{fig:three-states2} The left side depicts the Blackwell-optimal MDP of
    Figure~\ref{fig:three-states}, that is the MDP that takes the A loop. On the right side is the
    model-free version of the same MDP with policy \(\pol_{A}\). As in the model-free scenario
    state-action pairs are evaluated state \(1\) is split into a state that takes the left loop
    and one that traverses to the right. They are linked with dashed lines to indicate that these
    states are connected and the agent has to decide for one among them. The bias values for the
    model-free version are \(V^{\pol_{A}}(0,r) = -0.5\),
    \(V^{\pol_{A}}(1,l)=V^{\pol_{A}}(1,r)=0.5\), and \(V^{\pol_{A}}(2,l)=1.5\).}
\end{figure}

\section{Experimental Evaluation}
\label{sec:Experimental_Evaluation}

This section introduces the simulation model and the parameterisation of the algorithm. To be able
to compare the performance of the algorithm we chose to implement the same setup as has been used by
\cite{SchneckenreitherHaeussler2019}.

\subsection{Simulation Model}
\label{subsec:Simulation_Model}

\begin{figure*}[t!]
  \centering
  \input{productionsystem}                      % Figure production system
  \caption{Production system of the simulation model with routing, processing time distributions and
    demand.} \label{fig:ps}
\end{figure*}

Therefore, we use the hypothetical flow shop make-to-order manufacturing system similar to the
system analyzed by \cite{Lee1997} and later by \cite{Schneckenreither2019}. The simulation model
consists of three work centres; each consists of a single machine and can process only one order at
a time. Figure~\ref{fig:ps} depicts the make-to-order production system setup.
%
The number of orders arriving is uniformly distributed between 3 and 15 order per period and
thus with a mean of 9 order per period.
%
% Orders arrive
% uniformly between 3 and 15 orders per period and thus a mean of 9 orders per period, or
In other words, on average every $0.11$ periods one order arrives in the order pool.
%
Incoming orders are queued at the order pool until released. The due date slack ($\DDS$), that is
the periods until incoming orders are due, is set to 7 periods.
%
Once orders are released they are queued at each work center and wait until being processed by the
machine. The queuing priority is first-come-first-serve. No preemptions are allowed. Order
routings are deterministic and embrace two production stages with diverging shape and no return
visits. This results in 2 different products. The operation times of the work centres are
uniformly distributed, cf. Table~\ref{tbl:costs}. These characteristics lead to a utilization rate
of 90\% for the bottleneck work center (WC3) in steady state. % Planning periods were set to 960
% minutes (16 hours).

% \begin{figure}[b!]
%   \centering
%   \def\arraystretch{1.2}          %  1 is the default, change whatever you need
%   \begin{table}[t!]
%     \centering
%     \begin{tabular}{|l||c|c|c|}
        %         \hline
        %     %         BEGIN RECEIVE ORGTBL optime
        %         Work Center & WC1 & WC2 & WC3\\
        %         \hline
        %         Operation time & \(\mathcal{U}(70,130)\) & \(\mathcal{U}(130,170)\) & \(\mathcal{U}(180,200)\)\\
        %     %         END RECEIVE ORGTBL optime
        %         \hline
        %       \end{tabular}
        %   %       \egroup

        %         \caption{\label{tab:OTs}Operation times.}
        %   %       \end{figure}
        %         \end{table}
\begin{table}[t]
  \centering
  \begin{tabular}{|l||c|c|c|}
    % BEGIN RECEIVE ORGTBL values
    \hline
    Work Center & WC1 & WC2 & WC3\\
    \hline
    Operation time & \(\mathcal{U}(70,130)\) & \(\mathcal{U}(130,170)\) & \(\mathcal{U}(180,200)\)\\
    % END RECEIVE ORGTBL optime
    \hline
    \multicolumn{4}{c}{}\\
    \hline
    Cost & Wc p. Order/Period & Fc p. Order/Period & Bc p. Order/Period \\
    \hline
    Value & 3 & 10 & 20 \\
    \hline
    % END RECEIVE ORGTBL values
  \end{tabular}

  \caption{\label{tbl:costs}Operation times (upper half) and costs setup (in monetary unit).}
\end{table}

        %         \begin{figure}[b!]
        %         \centering
        %         \def\arraystretch{1.2}          %  1 is the default, change whatever you need
        %         \setlength\tabcolsep{1.7pt}
        %         \begin{tabular}{|l||c|c|c|c|c|c|}

        %       \end{tabular}
        %         \egroup
        %         \caption{Operation times.}
        %         \label{tab:OTs}
        %         \end{figure}

To evaluate the performance of the different order release models we define
following performance measures similar to literature
(e.g.,~\citealt{Baykasoglu2011,Thuerer2012}):
\begin{itemize}
\item \textbf{Cost related measures:} average total holding costs for  WIP and finished goods inventory,
  costs for backorders. There are no earnings,
  thus the algorithms minimizes the costs. The setup for the actual values of the costs is given in
  Table~\ref{tbl:costs} and shows that late deliveries are especially expensive. All costs are per
  order and period, and are measured and reported at the end of each period.

  % \begin{table}[t]
  %   \centering
  %   \begin{tabular}{|l||c|c|c|}
  %     %     BEGIN RECEIVE ORGTBL values
        %         \hline
        %         Work Center & WC1 & WC2 & WC3\\
        %         \hline
        %         Operation time & \(\mathcal{U}(70,130)\) & \(\mathcal{U}(130,170)\) & \(\mathcal{U}(180,200)\)\\
        %     %         END RECEIVE ORGTBL optime
        %         \hline
        %         \multicolumn{4}{c}{}\\
        %         \hline
        %         Cost & Wc p. Order/Period & Fc p. Order/Period & Bc p. Order/Period \\
        %         \hline
        %         Value & 3 & 10 & 20 \\
        %         \hline
        %     %         END RECEIVE ORGTBL values
        %       \end{tabular}

        %         \caption{\label{tbl:costs}Operation times and costs setup (in monetary unit).}
        %         \end{table}

        %         \SH{The costs were set as follows: $Wc:Fc:Bc = 1:1.5:2.5$?}
\item \textbf{Delivery related measures:} % mean service level -- percentage of
        %         orders on time (\%SL),
  mean tardiness of late orders (TA), standard deviation of
  lateness ($\sigma$TA).
\item \textbf{Flow time related measures:} mean shop floor throughput time (time
  duration from release until entry of finished goods inventory; SFTT).
\end{itemize}

The length of each simulation run to evaluate the performance was 6000 periods including a warm-up
period of 1000 periods. Welch’s procedure was applied to approximate the length of the warm-up
period (see~\citealt{law:simulationc}).

        %         Each order release method was evaluated on 25 predefined demand
        %         streams.

        %         \MS{TODO}
\subsection{Conventional Order Release Rules}

As external benchmark for comparison we use
        %         use two different order release rules. First, we use immediate release which releases orders as
        %         soon as they arrive at the order pool. And second, we
different parameterized backward infinite loading (\BIL{}) techniques:

\begin{equation*}
  RD_{j} = DD_{j} - LT_{j} \tcom
  \label{eq:BIL}
\end{equation*}

where the release date \(RD_{j}\) of product type \(j\) is calculated by the difference of the due
date \(DD_{j}\) and the lead time \(LT_{j}\). Note that we set lead times for each product type and
not for every order as done by \cite{Ackerman1963}.

        %         The
        %         variants are named according to the used lead time $LT$. For example in case of $\BILOne$ we have
        %         $LT=1$. \MS{@Equation 1: Use LT$_{j}$ or LT?}

        %         \MS[t]{do we?}

\subsection{Algorithm Setup}
\label{subsec:Algorithm_Setup}

\paragraph{Markov Decision Process.} The underlying MDP is unichain and looks as follows.
Both, state space $\mathcal{S}$ and action space $\mathcal{A}$ are discrete. Any state
$s \in \mathcal{S}$ of the \textit{state space} is composed of the following information for each
product $p$:
        %         The following information is presented to the algorithm for both Product 1 and Product 2 as follows.
\begin{itemize}

\item The currently set lead time $LT_{p} \in \{1,2,\ldots,\DDS\}$ (Recall:
  $\text{Due Date Period} - \text{Lead Time} = \text{Release Period}$). Note that we bound the
  maximum lead time with due date slack $\DDS$, which is $7$ in our setup.

\item Counters $OP_{p,d}\in \N$ for the number of orders in the order pool divided in time buckets
  with $d \in \{1,2,\ldots,\DDS \}$, which stands for the number of periods until the due date.

  % \item Flag for each machine whether it is idle or not.

\item Counters $Q_{i}\in \N$ standing for the number of orders for each queue $i$.

\item Counters $FGI_{p,d} \in \N$ of orders in the finished goods inventory divided in time
  buckets with $d \in \{-5,-4,-3,\ldots, \DDS \}$, which stands for the number of periods until the
  due date. Orders with a due date with more than 5 periods ago are listed in the counter
  $FGI_{-5}$.

\item Counters $S_{p,d} \in \N$ of shipped orders from the last period divided in time buckets
  with $d \in\{-5,-4,-3,\ldots, \DDS \}$, which stands for the number of periods until the due date.
  Orders with a due date with more than 5 periods ago are listed in the counter $S_{-5}$.

\end{itemize}

\noindent The algorithm implicitly learns a function which maps the current state of the
production system to a release decision. In an optimal situation it does so by multiply exploring
every action for each state and assessing its economic viability.

Orders are released once the due date is within the interval $[t,t+\text{lead time}]$, whereas $t$
is the current period. Clearly, this yields bulk releases (either all or no orders with the same
due date and product type are released). % Note that we do not present past state
        %         information to the neural network.

The \textit{actions space} is composed of two independent decisions. These are the relative changes
of the lead times to the currently set lead times $LT_{p} \in \{1,2,\ldots,\DDS\}$ for each product
$p$. Recall that $p=2$.
        %
Furthermore, we restrict the action space for each state $s_{t+1}$ according to the last set lead
time $LT_{p}$ from state $s_{t}$ by restricting the change of the lead time for consecutive
periods to a maximum of 1. Thus, if $LT_{p}$ is the current lead time for product $p$ the action
space for this product is given by $\{1,2,\ldots,\DDS\} \cap \{LT_{p}-1,LT_{p},LT_{p}+1\}$. Put
differently the algorithm can increase or decrease the lead time by 1 or leave it as it is, as
long as it acts within the discrete action space given by the set $\{1,2,\ldots,\DDS\}$ for each
type of product. Thus the action space over all products is given by the full enumeration of
available actions of the individual products.

        %         We allow the agent a higher increase
        %         of the lead time, compared to the decrease, to give it the chance to react quickly to high
        %         number of
        %         incoming demand.

\paragraph*{Reward.}
In each period the agent chooses an action which generates a reward while traversing to the next
period by simulating the production system. The rewards are the accumulated costs at the end of the
period. These costs are consist of the number of backorders, the current WIP level and the number of
orders in the inventory.
        %
        %         The costs are divided by the normalizing parameter $\eta$ (see Table~\ref{tbl:variants}) and then
        %         clipped (cut off) to the interval $[-0.5,0.5]$.

\paragraph*{Adapted Algorithm.}
The algorithm depicted above is designed as a tabular version. We have adapted it to use neural
networks for the approximation of the different kind of state-action pair value functions. This is
required due to the exponential growth of the state space. The implementation collects the observed
information for each state-action function in a replay memory and randomly selects \(128\) data
points out of a collection of \(30k\) memories in each period and trains these on the neural
network. This technique is called experience replay memory
\citep{lin1993reinforcement,mnih2015human} and ought to overcome we instabilities. Nonetheless, this
approximation results in the fact that the algorithm can be unstable or may even diverge
\citep{tsitsiklis1997analysis}. Additionally we use a target and a worker network, whereas the
target network is overwritten every \(10k\) steps by the worker network. This has been proposed by
\cite{mnih2015human} and should further stabilize the algorithm.

\paragraph*{Neural Network Setup.}
        %
We use the same setup for the neural network as done by \cite{Schneckenreither2019}. However as they
use a actor-critic agent, whereas we approximate state-action-values the output of the network is a
one dimensional vector as opposed to a matrix. Therefore we use a single three layer fully-connected
network where the number of nodes and activation functions are 41-ReLU\footnote{ReLU stands for
  rectified linear unit.}-89-ReLU-20-ReLU-9, with the output activations being Tanh (hyperbolic
tangent function).
        %
The output for both networks consist of $3^{2}=9$ nodes. This is due to the fact that all
combinations of increase, decrease and no change of the lead time for each product type have to be
represented.
        %
For back-propagation we use the Adam optimiser with learning rate \(0.001\), and beta-values
\(\beta_{1}=0.9\), and \(\beta_{2}=0.999\).

\section{Preliminary Results}
\label{sec:Preliminary}

This section provides preliminary results for the described setup and gives an overview of the
current performance of the algorithm compared to the conventional static lead time setting
algorithms.

Currently we evaluate two different kinds of reward reporting as they have been proposed by
\cite{Schneckenreither2019}. They find that it is beneficial to overcome the time offset imposed by
the production system. That is, although orders are immediately released to the production system
and thus generate a response in terms of a reward to the agent, the full impact of the chosen action
takes several periods to materialise. For instance, by deciding not to release an order, which later
becomes a backorder, it is beneficial to reward the actions actually responsible for the backorder
instead of the last action (only). Thus they propose to keep track of the orders currently in the
order pool until all fully traversed through the production system and reward the actions according
to all orders in the order pool.

The parameter setup is given in Table~\ref{tbl:params}. The parameters are exponentially decayed
with rate \(0.05\) and \(350k\) steps\footnote{The decay is defined as \(0.05 ^{(t/350k)}\), where
  \(t\) is the current period}. We implemented the deep Q-network algorithm proposed by
\cite{mnih2015human}, which is parameterised by learning rate \(\delta\) and a discount factor of
\(0.99\).

\begin{table}[t]
  \centering
        %         \def\arraystretch{1.2}          %  1 is the default, change whatever you need

  \begin{tabular}{|l||c|c|c|c|c|c|c|c|c|c|}
    \hline
    Parameter & $\alpha$ & $\beta$ & $\delta$ & $\gamma$ & $\epsilon$ & $p_{exp}$ & $\xi$ \\
    \hline
    Value & 0.5 & 0.15 & 0.15 & 0.15 & 0.5 & 1.0 & 0.15 \\
    \hline
  \end{tabular}

        %         \egroup

  \caption{\label{tbl:params}Parameter Setup.}
        %         \end{figure}
\end{table}

Table~\ref{tbl:res} shows the results, where we evaluate the average reward and discounted
reinforcement learning algorithm by providing either direct reward feedback, that is the reward
accumulated in the next period, without keeping track of any orders, as \(\ActShipped\) and
\(\ActShippedDQN\) respectively. Further both algorithms are evaluated with keeping track of the
order pool orders, such that each action is responsible for all orders in the order pool. These are
labelled \(\ActOrdPool\) and \(\ActOrdPoolDQN\). For comparisons reasons we provide
\(\BILOne-\BILSix\) and immediate release (interval release).

The results show that at the current time we are facing issues with scaling the algorithm to the
order release problem. Likely this is due to parameterisation issues, which we are investigating
currently. However, it can be seen that i) the technique of keeping track of the order pool orders
leads to better policies and ii) that despite the parameterisation issues average reward
reinforcement learning finds better solutions than its discounted counterpart. However, the least
costs are generated by \(\BILThree\), which shows that there is room for improvement for the
reinforcement learning policies.

Interesting is also to see, that all reinforcement learning variants are releasing rather late,
leading to low finished goods inventory costs (FGIC), but high backorder costs (BOC). Furthermore,
the agents based on the order-pool orders, namely \(\ActOrdPool\) and \(\ActOrdPoolDQN\), are able
achieve lower work-in-process costs (WIPC) and also lower tardiness, than the other the agents
rewarded directly on the shipped orders.

\begin{table}[t!]
  \centering
  \def\arraystretch{1.2}          %  1 is the default, change whatever you need
  \setlength\tabcolsep{4pt}
  \begin{tabular}{|l||c|c|c|c|c|c|c|c|c|}
    % BEGIN RECEIVE ORGTBL res2
    \hline
    Algorithm/KPI & SUM & BOC & FGIC & WIPC & % \%SL &
                                                       TARD & \(\sigma\)TARD & SFTT\\
    \hline
    % Evaluation 1 & \multicolumn{11}{c|}{} \\
    \ActShipped    & 1777.74 & 1322.04 & 94.49 & 361.20       & 2.24 & 1.52 & 3.44\\
    \ActOrdPool    & 1593.03 & 1318.65 &14.01 &260.36         & 2.11 & 1.42 & 2.68\\
    \ActShippedDQN & 1940.62 & 1570.40 &6.30 &363.92          & 2.35 & 1.59 & 3.95 \\
    \ActOrdPoolDQN & 1618.88 & 1306.20 & 40.68& 272.00        & 2.08 & 1.38 & 2.58\\
    \BILOne        & 1736.98 & 1510.42  &  0.00  & 226.56     & 2.04 & 1.26 & 2.16 \\
    \BILTwo        & 1081.79 & 769.86 & 86.89   & 225.03      & 1.87 & 1.18 & 2.15\\
    \BILThree      &  \textbf{922.36} & 362.59 &  334.47  & 225.30     & 1.74 & 1.07 & 2.15\\
    \BILFour       & 1062.68 & 155.72  & 681.03   & 225.93    & 1.63 & 0.99 & 2.15\\
    \BILFive       & 1380.04 & 65.45  & 1088.94    & 225.64   & 1.67 & 1.02 & 2.15\\
    \BILSix        & 1777.77 & 25.04 &  1529.77  & 222.96     & 1.61 & 0.85 & 2.13\\
    \imre          & 2209.97 & 12.08 &1974.72 & 223.16        & 1.58 & 0.63 & 2.14 \\
    \hline
    % END RECEIVE ORGTBL res2
  \end{tabular}
        %         \egroup

  \caption{\label{tbl:res}Evaluation results where the monetary units are in \(k\)-values.}

        %         \end{figure}
\end{table}

\section{Conclusion}
\label{sec:conclusion}

This paper introduces deep theoretical insights in reinfocement learning and provides first evidence
that average reward reinfocement learning is better applicable to operations research problems than
the discounted framework.
        %         adds to the growing body of evidence that machine learning algorithms can contribute
        %         positively to a company's performance.
The paper describes an % successful
application of an order release model based on reinforcement learning. The performance is tested on
a multi-product, two-stage hypothetical flow shop and is measured by cost, delivery and lead time
related measures. We show that in the current version the machine learning approach is not able to
outperform all other tested order release
approaches. % By yielding lower \textit{UPDATE: total costs,
        %         less mean and standard deviation of tardiness and a shorter shop floor throughput time (SFTT)}.
        %         Despite the good performance of our proposed adaptive order release model based on \textit{UPDATE:
        %         reinforcement learning} we are aware of its limitations. Firstly, the results are limited to the
        %         simulated case and the validity of the results for other production systems must be assessed in
        %         future studies. Secondly, adding further experimental factors would be beneficial like including
        %         scenarios with machine failures and different scheduling rules. In conclusion we found that our
        %         proposed adaptive order release approach harbours enormous potential.
In the future we expect to be able to yield better results by investigating the parameterisation and
resolving current issues when scaling average reinforcement learning to complex scenarios.

\bibliographystyle{plainnat}
\bibliography{references}

\vfill
\pagebreak
\appendix

\section{N-Discount-Optimality}
\label{sec:methods}

This part proves (or rather ought to prove, as this is only the first version of such a proof)
Blackwell-optimality of the average reward reinforcement learning algorithm shown in
Algorithm~\ref{alg}. The idea is to prove \(n\)-discount optimality for \(n=-1\), \(n=0\), and
finally for \(n \geqslant 1\). But first recall the definition of \(n\)-discount-optimality.

\begin{definition*}
  Due to \cite{Veinott69} for MDPs a policy \(\polopt\) is \emph{\(n\)-discount-optimal} for
  \(n=-1,0,1,\ldots\) for all states \(s \in \States\) with discount factor
  \(\gamma\) % and \(V_{\gamma}^{\pol}(s)\) being the value function as defined above
        %
  if and only if
  \begin{align*}
    \lim_{\gamma \to 1}(1-\gamma)^{-n}\ (V_{\gamma}^{\polopt}(s) - V_{\gamma}^{\pol}(s)) \geqslant 0 \tpkt
  \end{align*}
\end{definition*}

\subsection{\(\mathbf{(-1)}\)-Discount-Optimality\cite{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}.}

Here we provide insights in the algorithm and
\(n\)-discount-optimality, where we concentrate on \(n=-1\). Recall that the discounted state value
can be expanded by the Laurent series expansion:
\(V_{\gamma}^{\pol} = \frac{\avgrew^{\pol}(s)}{1-\gamma} + V^{\pol} + e_{\gamma}^{\pol}(s)\), where
\(\lim_{\gamma \to 1} e_{\gamma}^{\pol}(s) = 0\)

\begin{align*}
  \lim_{\gamma \to 1} (1-\gamma)^{1} \cdot (V_{\gamma}^{\polopt}(s) - V_{\gamma}^{\pol}(s)) \geqslant 0\\
  \lim_{\gamma \to 1} (1-\gamma) \cdot (\frac{\avgrew^{\polopt}(s) - \avgrew^{\pol}(s)}{1-\gamma} + V^{\polopt}(s) - V^{\pol}(s) + e_{\gamma}^{\polopt}(s) - e_{\gamma}^{\pol}(s)) \geqslant 0\\
  \lim_{\gamma \to 1} (\avgrew^{\polopt}(s) - \avgrew^{\pol}(s) + (1-\gamma) \cdot (V^{\polopt}(s) - V^{\pol}(s) + e_{\gamma}^{\polopt}(s) - e_{\gamma}^{\pol}(s)) \geqslant 0\\
  \avgrew^{\polopt}(s) - \avgrew^{\pol}(s) \geqslant 0\\
\end{align*}

That is a policy \(\pol\) which ought to be (\(-1\))-discount-optimal has to maximise the average reward
\(\rho^{\pol}(s)\) over all states \(s\).

\subsection{\(\mathbf{0}\)-Discount-Optimality\cite{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}.}

Note that a policy can only be \(0\)-discount-optimal if it is \(-1\)-discount-optimal, thus
\(\avgrew^{\polopt}(s)=\avgrew^{\pol}(s)\). Furthermore, recall that
\(\lim_{\gamma \to 1} e_{\gamma}^{\pol}(s) = 0\).

\begin{align*}
  \lim_{\gamma \to 1} (1-\gamma)^{0} \cdot (V_{\gamma}^{\polopt}(s) - V_{\gamma}^{\pol}(s)) \geqslant 0\\
  \lim_{\gamma \to 1} (\frac{\avgrew^{\polopt}(s) - \avgrew^{\pol}(s)}{1-\gamma} + V^{\polopt}(s) - V^{\pol}(s) + e_{\gamma}^{\polopt}(s) - e_{\gamma}^{\pol}(s)) \geqslant 0\\
  \lim_{\gamma \to 1} (\frac{0}{1-\gamma} + V^{\polopt}(s) - V^{\pol}(s)) \geqslant 0\\
  V^{\polopt}(s) - V^{\pol}(s) \geqslant 0\\
\end{align*}

Therefore a \(0\)-discount-optimal policy \(\pol\) has to maximise the bias values
\(V^{\pol}(s)\) for all states \(s\).

\vfill
\pagebreak

\subsection{\(\mathbf{n}\)-Discount-Optimality for \(\mathbf{n \geqslant 1}\).}

The following derivation provides an important insight to the analysis. Recall that for unichain
MDPs the average reward \(\avgrew^{\pol}(s)\) for all states \(s\) is equal and stated as
\(\avgrew^{\pol}\). Furthermore, as the policy is \(0\)-discount-optimal
\(V^{\polopt}(s) = V^{\pol}(s)\).

\begin{align*}
  \lim_{\gamma \to 1} (1-\gamma)^{-n} \cdot (V_{\gamma}^{\polopt}(s) - V_{\gamma}^{\pol}(s)) & \geqslant 0\\
  \lim_{\gamma \to 1} (1-\gamma)^{-n} \cdot ( \frac{\avgrew^{\polopt}(s)}{1-\gamma} + V^{\polopt}(s) + e_{\gamma}^{\polopt}(s)  %& \\
  - \frac{\avgrew^{\pol}(s)}{1-\gamma} - V^{\pol}(s) - e_{\gamma}^{\pol}(s) ) & \geqslant 0\\
  \lim_{\gamma \to 1} (1-\gamma)^{-n} \cdot ( e_{\gamma}^{\polopt}(s) - e_{\gamma}^{\pol}(s) ) & \geqslant 0\\
\end{align*}

As stated before \(e_{\gamma}^{\polopt}(s)\) approaches 0 as \(\gamma \to 1\). Therefore, it is
important to note that for \(n>0\) we analyse the case when \(\gamma\) is strictly less than \(1\). This
means that the number of actions to reach a desired goal is taken into account.
        %         The ratio of long-term to short-term is provided by the discount factor \(\gamma\).
Therefore, a \(\infty\)-discount-optimal policy \(\pol\) has to maximise the error term
$e_{\gamma}^{\pol}(s)$ for all states \(s\). That is, an \(\infty\)-discount-optimal algorithm has to
choose the action with the largest error term \(e_{\gamma}^{\pol}(s)\) once \(\gamma \to 1\).
However, as the error term depends on infinitely many sub-terms simply estimating these and summing
up does not work.

\begin{figure*}[t]
  \centering
  \begin{tikzpicture}[scale=8]
    \draw[->] (0,0) -- (1.0,0) node[right] {$\gamma$};
    \draw[->] (0,0) -- (0,0.8) node[above] {$\mathbb{R}$};
    \draw[scale=1,domain=0.001:0.85,smooth,variable=\x,blue] plot ({\x},{0.1/(1-\x)+0.2});
    \draw[scale=1,domain=0.001:0.865,smooth,variable=\x,red] plot ({\x},{0.1/(1-\x)+0.2-0.2*(1-\x)^0.5});

        %         Labels
    \draw[blue] (0.6,0.64) node () {$\frac{\avgrew^{\pol}(s)}{1-\gamma} + V^{\pol}(s)$};
    \draw[red] (0.875,0.58) node () {$V_{\gamma}^{\pol}(s)$};
    \draw (-0.015,0.3) -- (0.015,0.3);
    \draw (-0.2, 0.3) node[align=left] () {$\avgrew^{\pol}(s) + V^{\pol}(s)$};

    \draw (-0.015,0.1) -- (0.015,0.1);
    \draw (-0.111, 0.1) node[] () {$R^{\pol}(s)$};

    \draw (0.9,-0.015) -- (0.9,0.015) node[below, yshift=-0.2cm] () {$1$};
    \draw (0.75,-0.015) -- (0.75,0.015) node[below, yshift=-0.2cm] () {$\gamma_{1}$};
    \draw (0.20,-0.015) -- (0.20,0.015) node[below, yshift=-0.2cm] () {$\gamma_{0}$};

    \draw (0.5,0.4) -- (0.5,0.258578644) node[midway,left, yshift=-0.2cm] () {$e_{\gamma}^{\pol}(s)$};

    \draw (0.2,0.146114562) -- (0.75,0.146114562) node[midway, below] () {$\gamma_{1}-\gamma_{0}$};
    \draw (0.75,0.146114562) -- (0.75,0.5) node[midway,right] () {$V_{\gamma_{1}}^{\pol}(s)-V_{\gamma_{0}}^{\pol}(s)$};

  \end{tikzpicture}
  \caption{Visualisation of the monotonically increasing state values for a state $s$ and the
    corresponding decreasing error term $e_{\gamma}^{\pol}(s)$ as $\gamma$ approaches 1. }
  \label{fig:e}
\end{figure*}

Therefore, the idea is to use two discounted state-values with different $\gamma$'s to infer the
slope of the error term. Figure~\ref{fig:e} illustrates the idea. Here $\gamma_{1}$ and $\gamma_{0}$
are used to estimate the difference of \(V_{\gamma_{1}}^{\pol}(s)\) and
\(V_{\gamma_{0}}^{\pol}(s)\). As the average reward and the bias values are equal (recall
\(0\)-discount-optimality) the difference is a direct estimate of the slope of the error term. Thus
the larger the increase of the difference of the discounted state values when comparing
\(0\)-discount-optimal actions the better the action. However, for this to hold true we have to
prove strict monotonicity of
\(V_{\gamma}^{\pol}(s) = \frac{\avgrew^{\pol}(s)}{1-\gamma} + V^{\pol}(s) + e_{\gamma}^{\pol}(s)\)
under increasing $\gamma$. In the sequel we assume \(\avgrew^{\pol}(s) \geqslant 0\). Note that this
assumption is easily met by adding a constant reward for any action in cases where
\(\avgrew^{\pol}(s) < 0\). This assumption eases the following analysis and makes sense as usually
we are interested in problems which accumulate positive average reward, e.g. profit or points, over
time.

\begin{remark}
  Nonetheless, it is an important insight that if \(\avgrew^{\pol}(s) < 0\) then
  \(\frac{\avgrew^{\pol}(s)}{1-\gamma} + V^{\pol}(s)\) approaches negative infinity (as opposed to
  positive infinity, cf. Figure~\ref{fig:e}). Therefore, in such cases the following optimisation
  objectives would need to be inverted.
\end{remark}

\begin{lemma}
  \label{lem:err}
  The error term
  \(e_{\gamma}^{\pol}(s) = \sum_{m=1}^{\infty}(\frac{1-\gamma}{\gamma})^{m} \cdot y_{m}\) strictly
  monotonically decreases as $\gamma$ increases.
\end{lemma}

\begin{proof}
        %         S[t]{Is the following formula (indices) correct?}
  Note that \(e_{\gamma}^{\pol}(s) = \sum_{m=1}^{\infty}(\frac{1-\gamma}{\gamma})^{m} \cdot y_{m}\),
  where the sign of \(y_{m}\) alternates in each step. This is due to the fact that
  \(y_{m} = \E_{\pol} [ y_{m} ] - y_{m-1}\) for all \(m \geqslant 1\)~\cite[p.346]{Puterman94}, that
  is \(y_{m}\) only depends on \(y_{m-1}\) and the underlying (stationary) policy. Further \(y_{0}\)
  is the bias value \(V^{\pol}(s)\).
        %
  We consider three cases:
  \begin{enumerate}
  \item[1)]
    \begin{minipage}[t]{0.17\textwidth}
      \(\gamma = 0.5\):
    \end{minipage}
    The factor \((\frac{1-\gamma}{\gamma})^{m}\) is $1$.
  \item[2)]
    \begin{minipage}[t]{0.17\textwidth}
      \(0.5 < \gamma < 1\):
    \end{minipage}
    The factor \((\frac{1-\gamma}{\gamma})^{m}\) is $<1$ and strictly monotonically converging towards $0$.
  \item[3)]
    \begin{minipage}[t]{0.17\textwidth}
      \(0 < \gamma < 0.5\):
    \end{minipage}
    The factor \((\frac{1-\gamma}{\gamma})^{m}\) is $>1$ and strictly monotonically converging towards $\infty$.
  \end{enumerate}

  \paragraph{Case 1)}
  { As \(\frac{1-\gamma}{\gamma} = 1 \) the term reduces to
    \(e_{\gamma}^{\pol}(s) = \sum_{m=1}^{\infty} y_{m}^{\pol}\). We use the definition
        %
    \begin{align*}
      y_{m}^{\pol}=(-1)^{m} H_{\pol}^{m+1} h_{\pol} \tcom
    \end{align*}
    %
    where \(H_{\pol}\) is the deviation matrix defined as
    %
    \begin{align*}
      H_{\pol} = \clim_{N \to \infty} \sum_{t=0}^{N-1}{P^{t}-P^{\star}} \tcom
    \end{align*}
    \(\clim\) the Cesaro limit\footnote{Recall that the Cesaro limit (with degree \(1\)) is only
      required in periodic MDPs to ensure stationary matrices. For aperiodic MDPs they can be read
      as normal limits.~\cite[p.592ff]{Puterman94}}, \(P^{\star}\) the stationary matrix defined as
    \(P^{\star} = \clim_{N \to \infty}P^{N}\) and \(h_{\pol}\) a column vector holding the bias values
    of the states \cite[]{Puterman94,MillerVeinott1969}. Thus, we get
        %
    \begin{align*}
      e_{\gamma}^{\pol}(s) & = \sum_{m=1}^{\infty}y_{m}^{\pol}(s) = \sum_{m=1}^{\infty}{y_{2m}^{\pol}(s) + y_{2m+1}^{\pol}(s)}
                             = \sum_{m=1}^{\infty} H^{2m+2} h - H^{2m+1} h \\
                           & = \sum_{m=1}^{\infty} (H^{2m+2} - H^{2m+1}) h \tpkt
    \end{align*}

    % Wlog. assume all entries in \(H\) are positive. Therefore,
    It suffices\footnote{This claim has not been proven yet!} to show that $H^{2m+2} - H^{2m+1} < 0$
    for all $m \geqslant 1$ to prove strict monotonicity of $e_{\gamma}^{\pi}(s)$.
    % \MS[t]{why should that prove monotonicity? Rethink!!!}
    However, this directly follows as for MDPs with at least two states \(H-1 < 0\) in
    %
    \begin{align*}
      e_{\gamma}^{\pol}(s) = \sum_{m=1}^{\infty} (H^{2m+2} - H^{2m+1}) h = \sum_{m=1}^{\infty} H^{2m+1} (H - 1) h \tpkt
    \end{align*}
    %
  }

  \paragraph{Case 2)}{ As the factor \((\frac{1-\gamma}{\gamma})^{m}\) is $<1$ and strictly
    monotonically decreasing for an increasing $m \geqslant 1$, we use the same argument as in the
    previous case (Case 1) with which the claim follows. }

  \paragraph{Case 3)}{ This part of the proof is open!

    % Recall the Laurent Series expansion only holds when \(0 <
    % \frac{1-\gamma}{\gamma} < \norm{H_{\pi}}^{-1}\),
    % where for a finite matrix \(C = (c_{ij})\) the norm \(\norm{C}\) is defined as
    % \(\norm{C} \coloneqq \max_{i} \sum_{j} |c_{ij}|\).

    % \begin{align*}
    %   e_{\gamma}^{\pol}(s) & = \sum_{m=1}^{\infty}y_{m}^{\pol}(s) = \sum_{m=1}^{\infty}{y_{2m}^{\pol}(s) + y_{2m+1}^{\pol}(s)}\\
    %   & = \sum_{m=1}^{\infty} (\frac{1-\gamma}{\gamma})^{2m+1} H^{2m+2} h -
    %   (\frac{1-\gamma}{\gamma})^{2m}H^{2m+1} h \\
    %   & = \sum_{m=1}^{\infty} (\frac{1-\gamma}{\gamma})^{2m} H^{2m+1}((\frac{1-\gamma}{\gamma}) H - 1) h \tpkt
    % \end{align*}

          %           \MS{No! This does not prove that under two different gammas $e_{\gamma_{1}}^{\pol}(s) -
          %           e_{\gamma_{0}}^{\pol}(s)$ is decreasing}

  }

          %           We now investigate \(e^{\pol}(s) = \sum_{m=1}^{\infty}y_{m}^{\pol}(s)\) to infer properties of
          %           \(e_{\gamma}^{\pol}(s)\). Observe that we removed the factor $(\frac{1-\gamma}{\gamma})^{m}$
          %           depending on a concrete \(\gamma\). There are three cases to distinguish for this factor under
          %           increasing $m$:

          %           In all three cases the factor does neither change the monotonicity properties nor the sign of
          %           $y_{m}$ in the term
          %           \(e_{\gamma}^{\pol}(s) = \sum_{m=1}^{\infty}(\frac{1-\gamma}{\gamma})^{m} \cdot y_{m}\).
          %           Therefore, wlog. suppose $\gamma = 0.5$ and thus \(e^{\pol}(s) = e_{\gamma}^{\pol}(s)\).

          %           This factor is positive and strictly and monotonically
          %           converging towards $0$ (resp. $\infty$) as $m$ increases and as long as \(0.5 < \gamma < 1\)
          %           (resp. $0 < \gamma < 0.5)$. Note that we are usually interested in $\gamma$ values close to $1$,
          %           as $\gamma \to 1$. \MS[t]{what happens when $\gamma < 0.5$?}
          %
\end{proof}

\begin{corollary}
  \label{cor:diff}
  For any two \(0\)-discount-optimal stationary policies $\pol_{1}, \pol_{2}$ with stationary
  transition matrices \(P_{1}\) and \(P_{2}\) respectively, the differences of the slopes of the
  error terms \(e_{\gamma}^{\pol_{1}}(s)\) and \(e_{\gamma}^{\pol_{2}}(s)\) for any \(\gamma\) is
  dependent on \(H_{\pol_{1}}\) and \(H_{\pol_{2}}\) and thus \(P_{1}\) and \(P_{2}\) solely.
\end{corollary}

\begin{proof}
  Note that due to \(0\)-discount-optimality the bias values \(h_{1} = h_{2}\) are equal for both
  policies. Therefore as for any given \(H\) the error term is determined by
  \(e_{\gamma}^{\pol}(s) = \sum_{m=1}^{\infty} ((\frac{1-\gamma}{\gamma})^{2m + 1} H^{2m+2} -
  (\frac{1-\gamma}{\gamma})^{2m} H^{2m+1}) h = \sum_{m=1}^{\infty} (\frac{1-\gamma}{\gamma})^{2m}
  H^{2m+1}( \frac{1-\gamma}{\gamma} H - 1) h \) the claim follows directly.
\end{proof}

\begin{corollary}
  \label{cor:cont}
  The error term function \(e_{\gamma}^{\pol}(s)\) for \(\gamma \to 1\) and for any given policy
  \(\pol\) is continuous.
\end{corollary}

\begin{proof}
  The claim follows directly from Corollary~\ref{cor:diff} and due to the shape of the term
  \(e_{\gamma}^{\pol}(s) = \sum_{m=1}^{\infty} (\frac{1-\gamma}{\gamma})^{m} H^{m+1} h\).
\end{proof}

\begin{theorem}
  If \(\avgrew^{\pol}(s) > 0\) then
  \(V_{\gamma}^{\pol}(s) = \frac{\avgrew^{\pol}(s)}{1-\gamma} + V^{\pol}(s) + e_{\gamma}^{\pol}(s)\) is
  strictly monotonically increasing as $\gamma$ approaches \(1\).
\end{theorem}

\begin{proof}
  Due to \(\avgrew^{\pol}(s) > 0\) and \(0 < \gamma < 1\) the term
  \(\frac{\avgrew^{\pol}(s)}{1-\gamma}\) strictly monotonically approaches infinity as
  $\gamma \to 1$. Furthermore, by Lemma~\ref{lem:err} and as
  \(\lim_{\gamma \to 1} e_{\gamma}^{\pol}(s) = 0\) the theorem follows immediately.
\end{proof}

\begin{theorem}
  \label{thm:minmax}
  If \(\avgrew^{\pol}(s) \geqslant 0\) and for \(\gamma\)-values \(\gamma_{0}, \gamma_{1}\) with
  \(0.5 \leqslant \gamma_{0} < \gamma_{1} < 1\) a Blackwell-optimal agent chooses the action \(a\)
  with expected future state \(s'\) that maximises the expected\footnote{To ease readability and as
    we aim for model-free RL we have dropped the expectations in the formulas. } discounted state
  value difference
  \(\Delta V_{\gamma}^{\pol}(s') = V_{\gamma_{0}}^{\pol}(s') - V_{\gamma_{1}}^{\pol}(s')\) of the
  set of \(0\)-discount-optimal actions available and for which
  \begin{itemize}
  \item[a)] \(\frac{\avgrew}{1-\gamma_{0}} + V^{\pol}(s') < V_{\gamma_{0}}^{\pol}(s')\) holds, or if
    no such actions exists, then for which
  \item[b)]
    \(\frac{\avgrew}{1-\gamma_{0}} + V^{\pol}(s') \geqslant V_{\gamma_{0}}^{\pol}(s')\) hold.
  \end{itemize}

  % the discounted state value difference \(V_{\gamma_{1}}^{\pol}(s') - V_{\gamma_{0}}^{\pol}(s')\),
  % where the decision to maximise or minimise the difference is based on the value that minimises
  % \(V_{\gamma_{1}}^{\pol}(s') - V_{\gamma_{0}}^{\pol}(s')\).
  %         %
\end{theorem}

\begin{remark}
  In Theorem~\ref{thm:minmax} we have chosen to base the decision on the values generated with
  \(\gamma_{0}\) as the error term is greater with smaller \(\gamma\)-values.
\end{remark}

\begin{proof}
  First observe that due to strict monotonicity of \(\frac{\avgrew^{\pol}(s')}{1-\gamma}\) and
  \(e_{\gamma}^{\pol}(s')\) and as \(\lim_{\gamma \to 1} e_{\gamma}^{\pol}(s') = 0\) the discounted
  state values \(V_{\gamma}^{\pol}(s')\) and the sum of average reward values
  \(\frac{\avgrew^{\pol}(s')}{1-\gamma} + V^{\pol}(s')\) converge with increasing \(\gamma\).
          %
  Thus for any state \(s'\) either \(e_{\gamma}^{\pol}(s') > 0\) or \(e_{\gamma}^{\pol}(s') < 0\)
  for all \(\gamma\)-values. That is \(V_{\gamma}^{\pol}(s')\) converges to
  \(\frac{\avgrew^{\pol}(s)}{1-\gamma} + V^{\pol}(s')\) for \(\gamma \to 1\) from above or below.
  Clearly just considering the slope of the error term is insufficient as for any states
  \(s_{1}, s_{2}\) it may happen to be that \(e_{\gamma}^{\pol}(s_{1}) > 0\) while
  \(e_{\gamma}^{\pol}(s_{2}) < 0\). Thus we split the decision in cases, where in case \(a)\) we
  consider all states \(s_{1}\) with \(e_{\gamma}^{\pol}(s_{1}) > 0\) and in case \(b)\) we states
  \(s_{2}\) with \(e_{\gamma}^{\pol}(s_{2}) < 0\). Note that
  \(e_{\gamma}^{\pol}(s_{1}) > e_{\gamma}^{\pol}(s_{2})\) which explains the prioritisation of case
  \(a)\) over case \(b)\).

  Within these two sets of states we investigate the slopes of the error term.

  \MS[t]{Todo: Show that two points are sufficient for a proxy of the slope.}
  Thus the slope can be approximated by
  \begin{align*}
    \frac{\Delta V_{\gamma}^{\pol}(s)}{\Delta \gamma} = \frac{e_{\gamma_{1}}(s) - e_{\gamma_0}(s)}{\gamma_{1} - \gamma_{0}} \tpkt
  \end{align*}

  As \(\gamma_{1}(s)-\gamma_{0}(s)\) is constant the difference of the error terms
  \(e_{\gamma_{1}}(s) - e_{\gamma_0}(s)\) is sufficient to compare the slopes of the error term and
  can be computed by
  \begin{align}
    \label{eq:errorAvgRew}
    e_{\gamma_{1}}(s) - e_{\gamma_0}(s) = \frac{\avgrew^{\pol}(s)}{1-\gamma_{1}} -
    \frac{\avgrew^{\pol}(s)}{1-\gamma_{0}} - V_{\gamma_{1}}^{\pol}(s) + V_{\gamma_{0}}^{\pol}(s)\tpkt
  \end{align}

  However, as all possible future states are gain-optimal \(\avgrew^{\pol}(s)\) is equal for all
  states \(s\) under consideration. Therefore, the difference \(V_{\gamma_{0}}^{\pol}(s) -
  V_{\gamma_{1}}^{\pol}(s)\) suffices for comparison purposes between any two states \(s_{1}, s_{2}\):
  \begin{align*}
  e_{\gamma_{1}}(s_{1}) - e_{\gamma_0}(s_{1}) - e_{\gamma_{1}}(s_{2}) + e_{\gamma_0}(s_{2}) =
    V_{\gamma_{0}}^{\pol}(s_1) - V_{\gamma_{1}}^{\pol}(s_1) - V_{\gamma_{0}}^{\pol}(s_2) + V_{\gamma_{1}}^{\pol}(s_2)\tpkt\\
  \end{align*}

  Then due to Corollaries~\ref{cor:diff}~and~\ref{cor:cont} we conclude that
  \(\lim_{\gamma \to 1}(1-\gamma)^{-n}\ (V_{\gamma}^{\polopt}(s) - V_{\gamma}^{\pol}(s)) \geqslant
  0\), where \(\gamma\) is the discount factor and \(V_{\gamma}^{\pol}(s)\) the value function.

\end{proof}

\begin{remark}
  Note that this means a Blackwell-optimal agent collects the rewards as soon as possible, as it
  maximises the error term \(e_{\gamma}^{\pol}(s)\).

  % In the case of model-free reinforcement learning, in which rather state-action pairs as opposed to
  % states are assessed, Theorem~\ref{thm:minmax} does not hold. Observe that in the model-free case
  % the agent has to choose between MDP states, for instance \((1,l)\) or \((1,r)\) in
  % Figure~\ref{fig:three-states2}.
  % Thus, although the gain and
  % bias values are equal,
  % Thus the error terms of different actions \(a, a'\) may cause
  % \(e_{\gamma}^{\pol}(s,a) > 0\) and \(e_{\gamma}^{\pol}(s,a') < 0\).

  % This is due to the fact that by observing one state for each action multiple the error

  % states defined

  % chooses the among different states, cf. Figure~\ref{fig:three-states2}.

  % chooses action based on
  % a state \(s\) state-action pairs are assessed. Thus, multiple values could be available to decide
  % on whether to maximise or minimise the discounted state value difference. The correct approach is
  % to decide on behalf of the action \(a\) which minimises
  % \(\mid V^{\pol}_{\gamma_{0}}(s,a) - \frac{\avgrew^{\pol}}{1-\gamma_{0}} - V^{\pol}(s,a) \mid\) as
  % for all actions the error term \(e_{\gamma}^{\pol}(s)\) is converging towards \(0\) with the same
  % rate (due to the same discount factor \(\gamma_{0}\)).
\end{remark}

Thus under the assumption of correct approximations and by strict monotonicity of $e_{\gamma}^{\pi}$
we conclude Blackwell optimality of the algorithm. Figure~\ref{fig:e} visualises the discounted
state values and the idea of estimating the slope of $e_{\gamma}^{\pi}(s)$ in comparison to
$\frac{\avgrew^{\pol}(s)}{1-\gamma} + V^{\pol}(s)$.

\subsection{Refinements}
\label{subsec:Refinements}

The main limitation of the straightforward approach of Theorem~\ref{thm:minmax} is that it requires
accurate estimations of the bias values when splitting the actions into the two sets with increasing
and deceasing error term value. However, to be able to estimate \(V^{\pol}(s)\) accurately enough
one usually has to go through a tedious process of investigating many different parameterisations,
which is a cumbersome task, especially if the correct values are not known. Furthermore, the process
of finding the correct values is computationally very expensive in comparison to finding the
deviations between any two values.

Therefore, we have developed another approach of selecting the best possible action which does not
depend on correct bias values. In this approach we utilise Equation~\ref{eq:errorAvgRew} which
allows a determination of the slope simply by rectifying the additional slope imposed due to the
term \(\frac{\avgrew^{\pol}(s)}{1-\gamma}\). As \(V_{\gamma}^{\pol}(s)\) and \(\avgrew^{\pol}(s)\)
are computationally cheap they can be usually accurately computed.

Therefore in the implementation we propose to replace Theorem~\ref{thm:minmax} by following the
following one.

\begin{theorem}\label{thm:err}
  If \(\avgrew^{\pol}(s) \geqslant 0\) and for \(\gamma\)-values \(\gamma_{0}, \gamma_{1}\) with
  \(0.5 \leqslant \gamma_{0} < \gamma_{1} < 1\) a Blackwell-optimal agent chooses the action \(a\)
  with expected future state \(s'\) that maximises the expected error term difference
  \(\Delta e_{\gamma_{1},\gamma_{0}}^{\pol}(s') = V_{\gamma_{1}}^{\pol}(s') -
  V_{\gamma_{0}}^{\pol}(s') - \avgrew^{\pol}(s') (\frac{1}{1-\gamma_{1}} - \frac{1}{1-\gamma_{0}})\)
  of the set of \(0\)-discount-optimal actions available and for which
  \begin{itemize}
  \item[a)] \(\Delta e_{\gamma_{1},\gamma_{0}}^{\pol}(s') < 0\) holds, or if
    no such actions exists, then for which
  \item[b)]
    \(\Delta e_{\gamma_{1},\gamma_{0}}^{\pol}(s') \geqslant 0\) hold.
  \end{itemize}
\end{theorem}

\begin{proof}
  We have for any state \(s'\) either \(e_{\gamma}^{\pol}(s') > 0\) or \(e_{\gamma}^{\pol}(s') < 0\)
  for all \(\gamma\)-values (see proof of Theorem~\ref{thm:minmax}). Therefore, by definition states
  \(s'\) with \(e_{\gamma_{1},\gamma_{0}}^{\pol}(s') < 0\) have a decreasing error term value. As
  \(e_{\gamma_{1},\gamma_{0}}^{\pol}(s')\) is a direct measure for the slope of the error term, and
  by the characteristics on the term specified in Corollaries~\ref{cor:diff} and~\ref{cor:cont}, we
  can conclude that the maximum expected error term difference for negative values of
  \(e_{\gamma_{1},\gamma_{0}}^{\pol}(s')\) impose the greatest \(V_{\gamma}^{\pol}(s')\) when
  \(\gamma \to 1\). The proof for \(e_{\gamma_{1},\gamma_{0}}^{\pol}(s') \geqslant 0\) follows the
  same argument. Therefore, the Theorem provides a way to find the policy \(\polopt\) with
  \(\lim_{\gamma \to 1} (1-\gamma)^{-n} \cdot ( e_{\gamma}^{\polopt}(s') - e_{\gamma}^{\pol}(s')) >
  0\) for any other policy \(\pol\).\qed
\end{proof}

\section{Bellman Optimality Formulas and Derivations (Incomplete)}
\label{sec:bellman_optimality}

\paragraph{Bellman optimality equations for the average reward.}

\begin{align*}
  V^{\star}(s) & = \max_{a \in A(s)} R^{\polopt}(s,a) \\
               & = \max_{a \in A(s)} \E_{\polopt} [ \sum_{t=0}^{\infty}(R_{t}(s) - \avgrew^{\polopt}(s)) \mid s=s_{t}, a_{t} = a] \\
               & = \max_{a \in A(s)} \E_{\polopt} [ R_{0}(s) - \avgrew^{\polopt}(s) + \sum_{t=0}^{\infty}(R_{t}(s_{t+1}) - \avgrew^{\polopt}(s_{t+1}))  \mid s=s_{t}, a_{t} = a] \\
               & = \max_{a \in A(s)} \E [ R_{0}(s) - \avgrew^{\polopt}(s) + V^{\star}(s_{t+1})  \mid s=s_{t}, a_{t} = a ] \\
               & = \max_{a \in A(s)} (r(s,a) - \avgrew^{\polopt}(s) + \sum_{s_{t+1}} p(s_{t+1}|s,a) \cdot V^{\star}(s_{t+1}))
\end{align*}

The value of an action in a given state must equal the expected return for following an optimal
policy from that state:

\begin{align*}
  R^{\star}(s,a) & = \E [ r(s,a) - \avgrew^{\polopt}(s) + V^{\star}(s_{t+1})  \mid s=s_{t},a=a_{t} ]\\
                 & = r(s,a) - \avgrew^{\polopt}(s) + \E [ \max_{a_{t+1} \in A(s_{t+1})}R^{\star}(s_{t+1},a_{t+1}) \mid s=s_{t},a=a_{t}]\\
                 & = r(s,a) - \avgrew^{\polopt}(s) + \sum_{s_{t+1}} p(s_{t+1} | s, a) (\max_{a_{t+1} \in A(s_{t+1})} R^{\star}(s_{t+1},a_{t+1}))
\end{align*}

\paragraph*{Bellman optimal equations for the bias value.}

\begin{align*}
  W^{\star}(s) & = \E [ R_{t}(s) + V^{\star}(s_{t+1}) | s = s_{t}, a=a_{t} ]\\
               & = \E [ R_{t}(s) + \max_{a_{t+1} \in A(s_{t+1})} R^{\star}(s_{t+1},a_{t+1}) | s = s_{t} , a=a_{t}]\\
\end{align*}

\begin{align*}
  W^{\star}(s,a) & = \E [ r(s,a) + V^{\star}(s_{t+1}) \mid s=s_{t}, a=a_{t} ] \\
                 & = r(s,a) + \E [  \max_{a_{t+1} \in A(s_{t+1})} R^{\star}(s_{t+1},a_{t+1}) \mid s=s_{t}, a=a_{t} ]\\
                 & = r(s,a) + \sum_{s_{t+1}} p(s_{t+1} | s, a) (\max_{a_{t+1} \in A(s_{t+1})} R^{\star}(s_{t+1},a_{t+1}))
\end{align*}

Difference \(R_{n+1}\) to \(W_{n+1}\)?

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2020-12-02 paper/productionsystem.tex
initial commit from leuven paper

\begin{tikzpicture}[queue/.style={regular polygon,regular polygon sides=4,inner sep=0.075cm},scale=0.89]

  \draw (-3.5,0) node[] {$\ldots$};
  \draw (-4.1,-1.75) node[] {\scriptsize Due Date:};

  \draw (-3.0,-1.5) -- (-3.0,1.5)
        (-2.5,1.5)  -- (-2.5,-1.5)
        (-2.5,-1.5) -- (-3.0,-1.5);
  \draw (-2.75,-1.75) node[] {\tiny $t+3$};

  \draw (-2.5,-1.5) -- (-2.5,1.5)
        (-2,1.5)  -- (-2,-1.5)
        (-2,-1.5) -- (-2.5,-1.5);
  \draw (-2.25,-2) node[] {\tiny $t+2$};
  \draw (-2 ,-1.5) -- (-2,1.5)
        (-1.5,1.5)  -- (-1.5,-1.5)
        (-1.5,-1.5) -- (-2,-1.5);
  \draw (-1.75,-1.75) node[] {\tiny $t+1$};

  \draw (-1.5,0.0) node[] (OP) {};
  \draw [decorate,decoration={brace,amplitude=5pt},xshift=-2pt,yshift=0pt] (-1.0,-2.25) -- (-3.75,-2.25)     node [black,midway,yshift=-0.4cm] {Order Pool};
  \draw [decorate,decoration={brace,amplitude=5pt},xshift=-2pt,yshift=0pt] (5.0,-2.25) -- (-.75,-2.25)  node [black,midway,yshift=-0.4cm] {Production System};
  \draw [decorate,decoration={brace,amplitude=5pt},xshift=-2pt,yshift=0pt] (7,-2.25) -- (5.25,-2.25) node [black,midway,yshift=-0.4cm] {Inventory};

  % \draw (-1.75,-1.3) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 1$};
  % \draw (-1.75,-1.0) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 1$};
  % \draw (-1.75,-0.7) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 2$};
  % \draw (-1.75,-0.4) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 2$};
  % \draw (-1.75,-0.1) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 3$};
  % \draw (-1.75, 0.2) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 3$};
  % \draw (-1.75, 0.5) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 4$};
  % \draw (-1.75, 0.8) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 4$};
  % \draw (-1.75, 1.1) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 6$};
  % \draw (-1.75, 1.4) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 6$};
  % \draw (-1.75, 1.7) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 6$};

  \draw (-2.25,-1.3) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$  \scriptscriptstyle 1$};
  \draw (-2.25,-0.95) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$ \scriptscriptstyle 2$};
  \draw (-2.25,-0.6) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$  \scriptscriptstyle 1$};
  \draw (-2.25,-0.25) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$ \scriptscriptstyle 1$};

  \draw (-2.75,-1.3) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$  \scriptscriptstyle 1$};
  \draw (-2.75,-0.95) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$ \scriptscriptstyle 1$};
  \draw (-2.75,-0.6) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$  \scriptscriptstyle 2$};
  \draw (-2.75,-0.25) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$ \scriptscriptstyle 2$};
  \draw (-2.75, 0.1) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$  \scriptscriptstyle 2$};
  \draw (-2.75, 0.45) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$ \scriptscriptstyle 1$};
  \draw (-2.75, 0.8) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$  \scriptscriptstyle 2$};

  \draw (0.5,0)    node[minimum size=1cm,draw,align=center] (M1) {WC1\\[-1ex]{\tiny $\Unif(70,130)$}};
  \draw (3.25,0.75)   node[minimum size=1cm,draw,align=center] (M2) {WC2\\[-1ex]{\tiny  $\Unif(130,170)$}};
  \draw (3.25,-0.75)  node[minimum size=1cm,draw,align=center] (M3) {WC3\\[-1ex]{\tiny $\Unif(180,200)$}};
  % \draw (6,1.5)    node[minimum size=1cm,draw,align=center] (M4) {M4\\[-1ex]{\tiny $\Exp(210)$ or}\\[-1ex]{\tiny $\Unif(50,370)$}};
  % \draw (6,0)      node[minimum size=1cm,draw,align=center] (M5) {M5\\[-1ex]{\tiny $\Exp(285)$ or}\\[-1ex]{\tiny $\Unif(200,370)$}};
  % \draw (6,-1.5)   node[minimum size=1cm,draw,align=center] (M6) {M6\\[-1ex]{\tiny $\Exp(215)$ or}\\[-1ex]{\tiny $\Unif(110,320)$}};
  \draw (6,0)      node[minimum size=1cm,draw,align=center,fill] (\fgi{}) {\color{white} \fgi{}};

  \draw (M1.west) node[draw,queue,minimum size=0.05cm,fill,xshift=-0.075cm] (Q1) {};
  \draw (M2.west) node[draw,queue,minimum size=0.05cm,fill,xshift=-0.075cm] (Q2) {};
  \draw (M3.west) node[draw,queue,minimum size=0.05cm,fill,xshift=-0.075cm] (Q3) {};
  % \draw (M4.west) node[draw,queue,minimum size=0.05cm,fill,xshift=-0.075cm] (Q4) {};
  % \draw (M5.west) node[draw,queue,minimum size=0.05cm,fill,xshift=-0.075cm] (Q5) {};
  % \draw (M6.west) node[draw,queue,minimum size=0.01cm,fill,xshift=-0.075cm] (Q6) {};

  \draw (8.5, 1) node[] (CustomerTop) {\Huge \Gentsroom};
  \draw (8, 0.5) node[] {\Huge \Ladiesroom};
  \draw (8.5, 0) node[] (Customer) {\Huge \Gentsroom};
  \draw (8,-0.5) node[] {\Huge \Ladiesroom};
  \draw (8.5,-1) node[] {\Huge \Gentsroom};

  % Paths
  \path (OP.center) edge[->, bend left=10, above]
  node[xshift=0.088cm,yshift=0.05cm,align=center] {\scriptsize Order \\[-1ex] \scriptsize Release} (Q1);

  \path (M1) edge[->, bend left=10, above]                 node[xshift=-0.1cm,yshift=-0.05cm] {$\scriptstyle \text{P}1$} (Q2);
  \path (M1) edge[->, bend right=10,above]                 node[xshift=0.1cm,yshift=-0.05cm] {$\scriptstyle \text{P}2$} (Q3);
  % \path (M2) edge[->, bend left=10,above,near start ]      node {$\scriptstyle 1$}     (Q4);
  % \path (M2) edge[->, bend left=10,above,near start ]      node {$\scriptstyle 2$}     (Q5);
  % \path (M2) edge[->, bend left=10,above,near start ]      node[xshift=0.06cm,yshift=-0.1cm] {$\scriptstyle 3$}     (Q6);
  % \path (M3) edge[->, bend right=10,above,very near start] node[xshift=0.20cm,yshift=-0.15cm] {$\scriptstyle 4$}     (Q4);
  % \path (M3) edge[->, bend right=10,above,very near start]      node[yshift=-0.075cm] {$\scriptstyle 5$}     (Q5);
  % \path (M3) edge[->, bend right=10,above,near start]      node {$\scriptstyle 6$}     (Q6);
  % \path (M4) edge[->, bend left=10,above,near start]       node[xshift=0.1cm] {$\scriptstyle 1,4$} (\fgi{});
  \path (M2) edge[->, bend left=10,above,near start]       node[xshift=0.1cm, yshift=-0.07cm] {$\scriptstyle \text{P}1$} (\fgi{});
  \path (M3) edge[->, bend right=10,above,near start]      node[xshift=0.07cm,yshift=0.07cm] {$\scriptstyle \text{P}2$} (\fgi{});

  \path (\fgi{}) edge[->,near start] node[xshift=0.1cm,align=left] {\tiny Not \\[-1ex]
    \tiny before \\[-0.5ex] \tiny due \\[-1ex] \tiny date}    (Customer);

  \path (CustomerTop) edge[densely dashed] (8.5,2.8);
  \path (8.5,2.8) edge[densely dashed,midway,align=center] node[yshift=0.0cm] {\scriptsize Customer places
    an order which initiates the process\\ {\tiny Distribution of incoming orders:
      $\Unif(3,15)$ % or $\Exp(105)$ or $\Unif(65, 145)$
    }
  } (-4.5, 2.8);
  \path (-4.5,2.8) edge[densely dashed] (-4.5,0);
  \path (-4.5,0) edge[->,densely dashed] (-4.0,0);

\end{tikzpicture}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End:

### 2020-12-16 paper/paper.tex
minor updates paper

@@
   research applications as discounted reinforcement learning. Additionally we present an average
-  reward reinforcement learning algorithm which infers
-  (Blackwell-)optimal\footnote{Blackwell-optimality is not yet fully proven! A first version of such
-    a proof is attached in the Appendix.} policies.
-  % \SH{Add RL stuff}
-  % Therefore, we present a flow time estimation procedure to set lead times dynamically
-  % by using an artificial neural network. Additionally, we implement a safety lead time
-  % to incorporate the underlying cost ratio in the order release model. We test our
-  % proposed approach using a simulation model of a three-stage make-to-order flow-
-  % shop and compare the forecast accuracy and the accumulated costs to other forecast-
-  % based order release models from literature. We show that our proposed model using
-  We use a simulation model of a two-stage flow-shop to compare the algorithm
+  reward reinforcement learning algorithm which infers near-Blackwell-optimal policies. We use a
+  simulation model of a \MS[t]{todo}two-stage flow-shop to compare the algorithm
   to % and show that our approach outperforms
@@
 
-\section{Introduction}
-\label{sec:introduction} An important goal in Manufacturing Planning and Control (MPC) systems is to
-achieve short and predictable flow times, especially where high flexibility in meeting customer
-demand is required. Besides achieving short flow times, one should also maintain high output and
-due-date performance while keeping the work-in-process level low. One approach to address this
-problem is to collect all incoming orders in an order-pool and periodically decide which orders to
-release to the shop floor. Once orders are released, costs start to accumulate as planned orders
-materialise as actual jobs in the production system. The main challenge is to find a good compromise
-of balancing the shop floor and timely completion of jobs. Although the performance of such systems
-can be measured manifold the most overarching objective is to maximise profits by adequately
-assigning holding and lateness costs.
-
-One of the key modeling parameters for order release mechanisms is the \textit{lead time} which
+\section{Introduction}\label{sec:introduction}
+An important goal in Manufacturing Planning and Control (MPC) systems is to achieve short and
+predictable flow times, especially where high flexibility in meeting customer demand is required.
+Besides achieving short flow times, one should also maintain high output and due-date performance
+while keeping the work-in-process level low. One approach to address this problem is to collect all
+incoming orders in an order-pool and periodically decide which orders to release to the shop floor.
+Once orders are released, costs start to accumulate as planned orders materialise as actual jobs in
+the production system. The main challenge is to find a good compromise of balancing the shop floor
+and timely completion of jobs. Although the performance of such systems can be measured manifold the
+most overarching objective is to maximise profits by adequately assigning holding and lateness
+costs.
+
+One of the key modeling parameters for order release mechanisms is the \textit{lead time}, which
 refers to the planned time that elapses between the release of an order and its arrival in the
 finished goods inventory. This planning parameter is often based on the observable time an order
-needs to traverse through the production system, which we denote as \textit{flow time}. Flow times
-consist of processing, setup, control, transport, and waiting times, whereas the latter is the
-governing factor. Waiting times are a result from queuing (e.g. jobs queue before and after
-processing) which depend heavily on the amount of jobs in the system (WIP) and thus, waiting times
-are relatively difficult to estimate which makes the setting of favorable lead times so difficult
-(e.g., \citealt{Tatsiopoulos1983, Wiendahl1995}). Most state-of-the-art order release mechanisms use
-static or fixed lead times to address the order release problem, and thus neglect the nonlinear
-relationship between resource utilisation and flow times which is well known from practice and
-queuing theory (\citealt{Pahl2007}).
-
-One way to address this nonlinear interaction effects is to set lead times dynamically. Thus,
-intuitively the order release problem is solved by perfectly matching the lead times to the flow
-times, but one faces ``sampling issues'' meaning that the flow times depend on the lead times. An
-extreme scenario of this problem is the so called ``lead time syndrome'' which describes a vicious
-cycle where increasing flow times perpetually inflate the lead times which leads to worse
-performance (see e.g., \citealt{Mather1978, knollmann2013control, Selcuk2006}. Therefore,
-\cite{SchneckenreitherHaeussler2019} propose to use reinforcement learning as approach to
+needs to traverse the production system, which in contrast is denoted as the \textit{flow time}.
+Flow times consist of processing, setup, control, transport, and waiting times, whereas the latter
+is the governing factor \citep[][p.223]{zapfel1982produktionswirtschaft}. Waiting times are a result
+from queuing (e.g.\@ jobs queue before and after processing), depend heavily on the amount of jobs in
+the system (WIP) and thus are relatively difficult to estimate, which makes the
+setting of favorable lead times so difficult (e.g., \citealt{Tatsiopoulos1983, Wiendahl1995}). Most
+state-of-the-art order release mechanisms use static or fixed lead times to address the order
+release problem, and thus neglect the nonlinear relationship between resource utilisation and flow
+times, which is well known from practice and queuing theory (\citealt{Pahl2007}).
+
+One way to address this nonlinear interaction effects is to set lead times dynamically. Intuitively
+the order release problem is solved by perfectly matching the lead times to the flow times, but the
+corresponding optimisation problem faces ``sampling issues'' meaning that the flow times depend on
+the lead times. An extreme scenario of this problem is the so called ``lead time syndrome'', which
+describes a vicious cycle where increasing flow times perpetually inflate the lead times which leads
+to worse performance (see e.g., \citealt{Mather1978, knollmann2013control, Selcuk2006}. Therefore,
+\citet{SchneckenreitherHaeussler2019} propose to use reinforcement learning as approach to
 incorporate these effects in this optimisation problem. Reinforcement learning is an optimisation
-technique that stems from dynamic programming. The goal in reinforcement learning is to find the
-best stationary policy for a given problem. This policy is usually provided by assessing the states
-(or state-action pairs) of an underlying Markov Decision Process (MDP). The advantage of
-reinforcement learning over dynamic programming is that (i) the problem space is explored by an
-agent and thus only expectantly interesting parts of the problem space need to be assessed and (ii)
-the knowledge (acquisition) of transition probabilities becomes unnecessary as the states are
-evaluated by consecutively observed states solely.
+technique that stems from dynamic programming and its goal is to find the best stationary policy for
+a given problem. This policy is usually provided by assessing the states (or state-action pairs) of
+an underlying Markov Decision Process (MDP). The advantage of reinforcement learning over dynamic
+programming is that (i) the problem space is explored by an agent and thus only expectantly
+interesting parts of the problem space need to be assessed and (ii) the knowledge (acquisition) of
+transition probabilities becomes unnecessary as the states are evaluated by consecutively observed
+states solely.
 
 Over the past decades reinforcement learning has been applied to various problems, for which
-astonishing results have been reported. E.g. only recently \cite{mnih2015human} have presented a novel
-value-iteration reinforcement learning agent which exceeds human-level abilities in playing many
-classic Atari 2600 games~\citep{bellemare2012investigating}. \cite{mnih2016asynchronous} presents
-improved results with asynchronous actor-critic reinforcement learning. Also games like
-Go~\citep{silver2016mastering} and Chess~\citep{silver2017mastering} have been mastered with
-superhuman performance by \textit{tabula rasa} reinforcement learning agents. Furthermore, the
-method has also been applied in the setting of manufacturing system, e.g. to improve the ramp-up
-process~\citep{doltsinis2012reinforcement}, in locally selecting appropriate dispatching rules
-\citep{zhang1995reinforcement,wang2005application} or scheduling
+astonishing results have been reported. E.g.\@ only recently \citet{mnih2015human} have presented a
+novel value-iteration reinforcement learning agent which exceeds human-level abilities in playing
+many classic Atari 2600 games~\citep{bellemare2012investigating}. Further,
+\citet{mnih2016asynchronous} present improved results with asynchronous actor-critic reinforcement
+learning. Also games like Go~\citep{silver2016mastering} and Chess~\citep{silver2017mastering} have
+been mastered with superhuman performance by \textit{tabula rasa} reinforcement learning agents.
+Furthermore, the method has also been applied in the setting of manufacturing system, e.g.\@ to
+improve the ramp-up process~\citep{doltsinis2012reinforcement}, in locally selecting appropriate
+dispatching rules \citep{zhang1995reinforcement,wang2005application} or scheduling
 \citep{zhang1995reinforcement, waschneck2018optimization}.
@@
 However, all these applications use discounted reinforcement learning and are either designed to
-investigate a rather simple MDP, e.g. by selecting heuristics instead of optimising the underlying
+investigate a rather simple MDP, e.g.\@ by selecting heuristics instead of optimising the underlying
 problem itself, or by mapping the actual objective in a reward function that is approximately \(0\)
@@
 %
-This is due to the fact that state value is largely composed of a term defined by the policy's
+This is due to the fact that state value is largely composed of a term defined by the policys'
 average reward value \citep{MillerVeinott1969,Blackwell62} which would otherwise dilute the state
-values and thus decrease the solution quality, as for instance in
-\cite{SchneckenreitherHaeussler2019} and \cite{gijsbrechts2018can}.
+values and thus decrease the solution quality, as can for instance be observed in
+\citet{SchneckenreitherHaeussler2019} and \citet{gijsbrechts2018can}.
 %
@@
 Therefore, most applications that incorporate and directly reflect costs or profit in the reward
-function use average reward reinforcement learning. For instance \cite{aydin2000dynamic} use it in
-the setting of scheduling, while in a series of papers Mahadevan et al. investigated several problem
-domains starting with simple MDPs
+function use average reward reinforcement learning. \citet{aydin2000dynamic} use it in the setting
+of scheduling, while in a series of papers Mahadevan et al. investigated several problem domains
+starting with simple MDPs
 \citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults,Mahadevan96_OptimalityCriteriaInReinforcementLearning}.
@@
 low as possible \citep{Mahadevan98_OptimizingProductionManufacturingUsingReinforcementLearning}.
+However, in practise usually decision have to be made on a daily basis \citep{enns2004work}.
+Therefore, we refrain from this adaption and concentrate on standard MDPs only. Furthermore, often
+continuous-time semi-MDP problems can be converted through uniformisation into equivalent discrete
+time instances \citep[see][]{Puterman94,bertsekas1995dynamic}.
 
 Like discounted reinforcement learning also average reward reinforcement learning is based on an
-oracle function, in our case the accumulated profit of a period, to assess the decisions taken by
-the agent. By repeatedly choosing different actions the agent examines the problem space and rates
+oracle function, in our case the accumulated costs of a period, to assess the decisions taken by the
+agent. By repeatedly choosing different actions the agent examines the problem space and rates
 possible actions for any observed state. The advantage of average reward reinforcement learning over
 the widely applied discounted reinforcement learning framework is that the underlying optimisation
-technique is able to find better policies. This yields from the fact that in the latter method the
-states are assessed by a single value, while in the former the values are split up by the average
-reward accumulated over time and a bias value which optimises the number of steps to reach the best
-possible states of the examined process
+technique is able to find better policies. This yields from the fact that in standard discounted
+reinforcement learning method the states are assessed independently and by a single value. In
+contrast to that average reward reinforcement learning splits up the evaluation of the average
+reward per step, a bias value that specifies the amount of collected rewards to reach the optimal
+path when starting in a suboptimal state and an error term which defines the number of steps to
+reach the optimal path
 \citep{Howard64,Puterman94,Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults}.
 Thus, while in average reward reinforcement learning these terms are learned separately in the
-discounted framework one value consisting of the addition of these values plus an additional error
-term is estimated. This incautious combining of different kinds of state values as done in
-discounted reinforcement learning leads to the problems that (i) the average reward, which is equal
-for all states of usually investigated unichain MDPs, is dominating and thus diluting the bias
-values, and (ii) the state values are deteriorated by the error term that is only imposed due to the
-discounting technique.
-
-Thus, as opposed to the commonly applied discounted reinforcement learning algorithm and as proposed
-as future work by \cite{SchneckenreitherHaeussler2019} we use an average reward reinforcement
-learning algorithm to adaptively release orders based on the assessed state values of the production
-system. In contrast to the aforementioned works on average reward reinforcement learning our
-algorithm incorporates the optimisation of not only the average reward over time, but also the bias
-values. Only the work by
+discounted framework one value consisting of the addition of these values is estimated, where
+however the average reward is scaled by \(1/(1-\gamma)\) with the discount factor \(\gamma\) usually
+being set very close to \(1\), e.g. \(0.99\). In commonly applied unichain MDPs the average reward
+per step is equal for all states. Thus, independently assessing it is computationally unwisely.
+Therefore, we adapt an algorithm that uses a scalar value for the estimation of the average reward.
+Furthermore, as the average reward is scaled in the discounted state values it dominates the other
+two terms. This, combined with the iterative evaluation and the independently assessing of the
+average reward for each state lets standard discounted reinforcement learning struggle to find good
+policies.
+% Thus, this incautious combining of different kinds of state values as done in discounted
+% reinforcement learning leads to the problems that (i) the average reward, which is equal for all
+% states of usually investigated unichain MDPs, is dominating and thus diluting the bias values, and
+% (ii) the state values are deteriorated by the error term that is only imposed due to the
+% discounting technique.
+
+Thus, as opposed to the commonly applied discounted reinforcement learning algorithm we use an
+average reward adjusted reinforcement learning algorithm to adaptively release orders based on the
+assessed state values of the production system. In contrast to the aforementioned works on average
+reward reinforcement learning our algorithm incorporates the optimisation of not only the average
+reward over time, but also the bias values, which is an important adaption in highly stochastic
+systems. Only the work by
 \cite{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}
@@
 the lack of feedback in the iterative process of optimising the different refinement optimisation
-levels.
+levels. Furthermore, they asses the average reward independently.
 
@@
 average reward reinforcement learning and presents the used algorithm (see also
-\citealt{Schneckenreither2019}) including a proof-of-concept.
-Section~\ref{sec:Experimental_Evaluation} describes the simulation model we use to evaluate the
-approach. Then~\ref{sec:Preliminary} provides preliminary result data, wheras
-Section~\ref{sec:conclusion} concludes this working paper.
-
-\section{Average Reward Reinforcement Learning}
-\label{sec:Average_Reward_Reinforcement_Learning}
-
-This section introduces average reward reinforcement learning in comparison to the discounted
+\citealt{schneckenreither2020average}). Section~\ref{sec:Experimental_Evaluation} describes the
+simulation model we use to evaluate the approach. \MS[t]{todo}
+% Then~\ref{sec:Preliminary} presents preliminary
+% result data, whereas Section~\ref{sec:conclusion} concludes this working paper.
+
+\section{Average Reward Adjusted Reinforcement Learning}
+
+This section introduces average reward adjusted reinforcement learning in comparison to the discounted
 framework, elaborates on optimiality criteria and provides insights of the underlying algorithm.
@@
 
-In the discounted framework the value of a state \(V_{\gamma}^{\pol}(s)\) is defined as the expected
-discounted sum of rewards under the stationary policy \(\pol\) when starting in state \(s\). That is
-\begin{align*}
-  V_{\gamma}^{\pol}(s)=\lim_{N \to \infty} E[\sum_{t=0}^{N-1} \gamma^{t} R_{t}^{\pol}(s)]\tcom
+In the widely applied discounted framework the value of a state \(V_{\gamma}^{\pol_{\gamma}}(s)\) is
+defined as the expected discounted sum of rewards under the stationary policy \(\pol_{\gamma}\) when
+starting in state \(s\). Note that the policy \(\pol_{\gamma}\) depends on the selected discount
+factor. That is
+\begin{align*}
+  V_{\gamma}^{\pol_{\gamma}}(s)=\lim_{N \to \infty} E[\sum_{t=0}^{N-1} \gamma^{t} R_{t}^{\pol_{\gamma}}(s)]\tcom
 \end{align*}
@@
 where \(0 \leqslant \gamma < 1\) is the discount factor and
-\(R_{t}^{\pol}(s) = E_{\pol}[ r(s_{t},a_{t}) \mid s_{t} = s, a_{t} = a]\) % \MS[t]{check if correct!}
+\(R_{t}^{\pol}(s) = E_{\pol}[ r(s_{t},a_{t}) \mid s_{t} = s, a_{t} =
+a]\) % \MS[t]{check if correct!}
 the reward received at time \(t\) upon starting in state \(s\) by following policy \(\pol\)
@@
 
-The aim in the discounted framework is to find an optimal policy \(\polopt\), which when followed,
-maximises the state value for all states \(s\) as compared to any other policy \(\pol\):
-\(V_{\gamma}^{\polopt} - V_{\gamma}^{\pol} \geqslant 0\). This criteria is usually referred to as
-\(\gamma\)-optimality as the discount factor \(\gamma\) is fixed. However, this also means that the
-actual value set for \(\gamma\) defines the policy that is optimised for. For instance, as can be
-seen in Figure~\ref{fig:printer} setting \(\gamma\) value to \(<0.8027\) the printer-loop is
-preferred over the mail-loop by the agent, although the mail-loop accumulates more reward over time,
-i.e. selecting the mail-loop is a sub-optimal choice.
+The aim in the discounted framework is to find an optimal policy \(\polopt_{\gamma}\), which when
+followed, maximises the state value for all states \(s\) as compared to any other policy
+\(\pol_{\gamma}\): \(V_{\gamma}^{\polopt_{\gamma}} - V_{\gamma}^{\pol_{\gamma}} \geqslant 0\). This
+criteria is usually referred to as \(\gamma\)-optimality as the discount factor \(\gamma\) is fixed.
+However, this also means that the actual value set for \(\gamma\) defines the policy that is
+optimised for. For instance, as can be seen in Figure~\ref{fig:printer} setting \(\gamma\) value to
+\(<0.8027\) the printer-loop is preferred over the mail-loop by the agent, although the mail-loop
+accumulates more reward over time, i.e. selecting the mail-loop is a sub-optimal choice.
 
@@
   \end{tikzpicture}
-  \caption{\label{fig:printer} A MDP with two different determinstic policies. The only action choice is in
+  \caption{\label{fig:printer} A MDP with two different deterministic policies. The only action choice is in
     state \(1\), in which the agent can choose between doing the printer-loop or the mail-loop.
@@
 
-The main idea behind the \(\gamma\)-parameter of the discounted framework is to be able to balance
+One idea behind the \(\gamma\)-parameter of the discounted framework is to be able to balance
 short-term rewards (low \(\gamma\)-values) and long-term rewards (high \(\gamma\)-values). However,
@@
 \subsection{Average Reward Reinforcement Learning}
-\label{subsec:Average_Reward_Reinforcement_Learning}
 
@@
 \(R^{\pol}(s) = E_{\pol}[ r(s, a) ] = \sum_{a \in \mathcal{A}(s)}p(a | s) r(s, a)\) as shown by
-\cite{MillerVeingsott1969} and \citet[p.346]{Puterman94}:
+\cite{MillerVeinott1969} and \citet[p.346]{Puterman94}:
 
@@
 closed irreducible classes of \(P\), that is for ergodic MDPs \(u\) is determined by a single
-constant. Our algorithm is based on the above formulation.
+constant. Average reward learning is based on the above formulation.
+
+A major problem occurring at average reward RL is that the bias values are not uniquely defined
+without solving the first set of constraints defined by the error term addends
+\citep[see][p.346]{Puterman94,Mahadevan96_SensitiveDiscountOptimalityUnifyingDiscountedAndAverageRewardReinforcementLearning}.
+% We could overcome this issue by simply requiring \(\gamma\) to be strictly less than \(1\).
+%
+Our algorithm, based on the tabular version of \citet{schneckenreither2020average}, does not require
+the exact solution for \(V^{\pol}(s)\), but a solution which is offset suffices. Clearly this
+observation reduces the required iteration steps tremendously as finding the exact solution,
+especially for large discount factors, is tedious. Therefore, we allow to set \(\gamma = 1\), which
+induces \(\X_{\gamma}^{\pol}(s) = V^{\pol}(s) + u\), where \(u\) is for unichain MDPs a scalar value
+independent of \(s\), i.e.\@ equivalent for all states of the MDP~\cite[p.346]{Puterman94}.
+%
+%% Nonetheless, we require \(\gamma < 1\) to stabilise the process.
+%
+If we are interested in correct bias values, i.e. \(\gamma\) is sufficiently close but strictly less
+than \(1\), our approach is a tremendous advantage over average reward RL as it reduces the number
+of iterative learning steps by requiring only a single constraint per state plus one for the scalar
+average reward value. That is, for an MDP with \(N\) states only one more constraint (\(N+1)\) has
+to be solved in \ARA{} as compared to (at least) \(2N+1\) nested constraints for average reward RL.
+Therefore, it is cheap to compute \(\X_{\gamma}^{\pol}(s)\), while it is rather expensive to find
+the correct values of \(V^{\pol}(s)\) directly, especially in an iterative manner as RL is.
 
@@
   \begin{algorithmic}[1]
-    \State{}Initialize values to \(0\) \While{the stopping criterion is not fulfilled}
-    \State{}\begin{minipage}[t]{0.9\textwidth} With probability \(p_{exp}\) choose a random action
-      and probability \(1-p_{exp}\) one that maximizes \MS[t]{max V$_{\gamma_{1}}(s,a)$ if more than
-      one e?}
-      \(\lex(\avgrew^{\pol}(s,a),\V^{\pol}(s,a),\Delta e^\pol_{\gamma_{1},\gamma_{0}}(s,a))\) at the
-      current state \(s\) where
-      \(\Delta e_{\gamma_{1},\gamma_{0}}^{\pol}(s,a) = V_{\gamma_{1}}^{\pol}(s,a) -
-      V_{\gamma_{0}}^{\pol}(s,a) - \avgrew^{\pol}(s,a) (\frac{1}{1-\gamma_{1}} -
-      \frac{1}{1-\gamma_{0}})\)
-      % \(\Delta e^\pol_\gamma_{1},\gamma_{0}(s,a) = V^\pol_{\gamma_0}(s,a) - V^\pol_{\gamma_1}(s,a)\)
-      for all actions for which
-      % \(\frac{\avgrew^{\pol}(s,a)}{1-\gamma_{0}} + \V^{\pol}(s,a) < V^\pol_{\gamma_0}(s,a)\)
-      \(\Delta e_{\gamma_{1},\gamma_{0}}^{\pol}(s,a) < 0\) holds or if no such actions exists, from
-      all actions. Set \(a^{\pol} = 0\) if a random action was chosen,
-      \(a^{\pol}=1\) otherwise.
+    \State{}Initialise state \(s_{0}\) and network parameters \(\netTarget, \netWorker\) randomly, set
+    an exploration rate \(0 \leqslant \plearn \leqslant \pexp \leqslant 1\), exponential smoothing learning rates
+    \(0 < \alpha, \gamma < 1\), and discount factors
+    \(0 < \gamma_{0} < \gamma_{1} \leqslant 1\).
+    \While{the stopping criterion is not fulfilled}
+    \State{}\begin{minipage}[t]{0.9\textwidth} With probability \(\pexp\) choose a random action and
+      probability \(1-\pexp\) one that fulfills
+      \(\max_{a}\lexeq(\X^{\pol}_{\gamma_{1}}(s_{t},a; \netWorker{}),\X^\pol_ {\gamma_{0}}(s_{t},a; \netWorker{}))\). Let
+      \(\isRandAct{t}\) indicate if the action was chosen randomly.
     \end{minipage}
-    \State{}Carry out action \(a\), observe reward \(r\) and resulting state \(s'\)
-    \State{}Update values using exponential smoothing
-      \MS[t]{why only random/non-random for rho and enforcing? very inconsistent!!!}
-    \begin{align*}
-      \qquad a) & & \avgrew^{\pol}(s,a)  & \gets (1- a^{\pol} \alpha) \avgrew^{\pol}(s,a) + a^{\pol} \alpha [r + \max_{a}\V^{\pol}(s',a) - \V^{\pol}(s,a)]     \\
-      b) & & \V^{\pol}(s,a)        & \gets (1-\beta) \V^{\pol}(s,a) + \beta  \cdot \\
-         &        &                      & [r - \avgrew^{\pol}(s,a) + \max_{a} \V^{\pol}(s',a) +
-                                           a^{\pol} (\Psi_{V}^{\pol}(s,a) + \zeta \Psi_{W}^{\pol}(s,a))] \\
-      c) & & W^{\pol}(s,a)        & \gets (1-\delta) W^{\pol}(s,a) + \delta [-\V^{\pol}(s,a) +
-                                    \max_{a} W^{\pol}(s',a)]              \\
-      d) & & \Psi_{V}^{\pol}(s,a) & \gets (1-\xi \beta) \Psi_{V}^{\pol}(s,a) + \xi \beta [r - \avgrew^{\pol}(s,a) - V(s,a) + \max_{a}V^{\pol}(s',a)]\\
-      e) & & \Psi_{W}^{\pol}(s,a) & \gets (1-\xi \delta) \Psi_{W}^{\pol}(s,a) + \xi \delta [- V^{\pol}(s,a) - W^{\pol}(s,a) + \max_{a} W^{\pol}(s',a)]
-    \end{align*}
-    \State{Update discounted state-values}
-    \begin{align*}
-      V^\pol_{\gamma_0}(s,a)    & \gets (1-\gamma) V^\pol_{\gamma_0}(s,a) + \gamma (r + \gamma_{0} \max_{a} V^\pol_{\gamma_0}(s',a)) \\
-      V^\pol_{\gamma_1}(s,a)    & \gets (1-\gamma) V^\pol_{\gamma_1}(s,a) + \gamma (r + \gamma_{1} \max_{a} V^\pol_{\gamma_1}(s',a))
-    \end{align*}
-    \State{}Set \(s \gets s'\), \(t \gets t+1\) and decay parameters accordingly
+    \State{}\begin{minipage}[t]{0.9\textwidth}
+    Carry out action \(a_{t}\), observe reward \(r_{t}\) and resulting state \(s_{t+1}\).
+    Store the experience \((s_{t}, \isRandAct{t}, a_{t}, r_{t}, s_{t+1})\) in the experience replay
+    memory \(M\).
+    \end{minipage}
+    \If{a non-random action was chosen or \(\pexp > \plearn\)}
+    \begin{align*}
+      \avgrew^{\pol}  \gets (1- \alpha) \avgrew^{\pol} + \alpha [r_{t} +
+      \max_{a}\X^\pol_{\gamma_1}(s_{t+1},a; \netWorker{}) - \X^\pol_{\gamma_1}(s_{t},a_{t}; \netWorker{})]
+    \end{align*}
+    \EndIf
+
+    \State{}Sample random mini-batch of experiences \((s_{i}, \isRandAct{i}, a_{i}, r_{i}, s_{i})\)
+    from \(M\) and do
+    \begin{align*}
+      y_{i,\gamma_0} & \gets r_{t} + \gamma_{0} \max_{a}
+                                        \X^\pol_{\gamma_0}(s_{t+1},a; \netTarget{}) - \avgrew^{\pol} \\
+      y_{i,\gamma_{1}} & \gets r_{t} + \gamma_{1} \max_{a}
+                                        \X^\pol_{\gamma_1}(s_{t+1},a; \netTarget{}) - \avgrew^{\pol}
+    \end{align*}
+    \State{} \begin{minipage}[t]{0.9\textwidth} Update the average reward adjusted discounted
+      state-values using the sum of the gradients on \({(y_{i,\cdot} -
+      \X^\pol_{\cdot}(s_{t+1},a; \netWorker{}))}^{2}\) wrt. network parameters
+    \(\netWorker{}\).
+    \end{minipage}
+    \State{} Every \(C\) steps exponential smoothly set target network:
+      \(\netTarget \gets (1 - \gamma) \netTarget{} + \gamma \netWorker{}\)
+    \State{} Set \(s \gets s'\), \(t \gets t+1\) and decay parameters
     \EndWhile{}
   \end{algorithmic}
-  \caption{\label{alg}Model-free tabular version of the Blackwell-optimal RL algorithm.}
+  \caption{\label{alg:near}Near-Blackwell-optimal deep RL for unichain MDPs}
 \end{algorithm}
 
-% \MS{
-%   Optimisations:
-%   \begin{enumerate}
-%   \item Use Lemma 3 of Miller \& Veinott to get rid of W2
-%   \item Use Reference state as in Mahadevan to get rid of W
-%   \item Average reward freed reinforcement learning
-%   \end{enumerate}
-% }
-
-The reinforcement learning algorithm is depicted in Algorithm~\ref{alg}. After initialising all
-values the agent enters the loop in which the first task is to choose an action (step 3). In this
-action selection process we utilise an \(\epsilon\)-sensitive lexicographic order
-\(a=(a_{1},\ldots,a_{n}) \lex (b_{1},\ldots,b_{n})=b\) defined as \(a \lex b\) if and only if
+The model-free average reward adjusted deep RL algorithm is based on the tabular
+version of \citet{schneckenreither2020average} and depicted in Algorithm~\ref{alg:near}.
+%
+%
+%
+\MS[t]{describe network initialisation method (Uniform, HeAtAl, etc.)}
+%
+\MS[t]{Extensions: overestimate rho, rho minimum, Multiple Agents (shared state,  shared rho?),
+  init-phase (0.5 for expSmthRewRate, always adapt rho, same rho for all agents): exp smooth reward
+  as rho, replay memory actions, n-step (describe replay memory), gradient clipping }
+\MS[t]{Side info: Replay memory also stores (disallowed) filtered action indices for optimisation purposes and on
+episodic tasks needs to store a bool indicating an episode end. Uses Int8 for state features values.  }
+
+%
+%
+After initialising all values the agent enters the loop in which the first task is to choose an
+action (step 3). In this action selection process we utilise an \(\epsilon\)-sensitive lexicographic
+order \(a=(a_{1},\ldots,a_{n}) \lex (b_{1},\ldots,b_{n})=b\) defined as \(a \lex b\) if and only if
 \(| a_{j} - b_{j} | \leqslant \epsilon\) for all \(j < i\) and \(|a_{i} - b_{i}| > \epsilon\). Note
@@
 This part proves (or rather ought to prove, as this is only the first version of such a proof)
-Blackwell-optimality of the average reward reinforcement learning algorithm shown in
-Algorithm~\ref{alg}. The idea is to prove \(n\)-discount optimality for \(n=-1\), \(n=0\), and
+Blackwell-optimality of the average reward adjusted reinforcement learning algorithm shown in
+Algorithm~\ref{alg:near}. The idea is to prove \(n\)-discount optimality for \(n=-1\), \(n=0\), and
 finally for \(n \geqslant 1\). But first recall the definition of \(n\)-discount-optimality.

### 2020-12-18 paper/paper.tex
update literature

@@
 describes a vicious cycle where increasing flow times perpetually inflate the lead times which leads
-to worse performance (see e.g., \citealt{Mather1978, knollmann2013control, Selcuk2006}. Therefore,
-\citet{SchneckenreitherHaeussler2019} propose to use reinforcement learning as approach to
-incorporate these effects in this optimisation problem. Reinforcement learning is an optimisation
-technique that stems from dynamic programming and its goal is to find the best stationary policy for
-a given problem. This policy is usually provided by assessing the states (or state-action pairs) of
-an underlying Markov Decision Process (MDP). The advantage of reinforcement learning over dynamic
-programming is that (i) the problem space is explored by an agent and thus only expectantly
-interesting parts of the problem space need to be assessed and (ii) the knowledge (acquisition) of
-transition probabilities becomes unnecessary as the states are evaluated by consecutively observed
-states solely.
+to worse performance (see e.g., \citealt{Mather1978, knollmann2013control, Selcuk2006}.
+%
+Thus, setting lead times dynamically harbors optimisation potential \citep{hoyt1978dynamic}, but may
+also substantially degrade the system performance.
+%
+\citet{schneckenreither2020order} have established
+following categorisation of dynamic lead time management approaches:
+
+\begin{itemize}
+\item \textsf{Reactive lead time management} approaches set lead times by reacting on
+  earlier flow times (e.g., \citealt{enns2004work, Selcuk2006}). Note that the forecast is always
+  based on \textit{past} data as the most recent system changes cannot be reflected by flow times
+  until the corresponding orders arrive in the FGI, which might take several
+  periods.
+\item
+  \textsf{Proactive lead time management} may incorporate \textit{past} data in conjunction with
+  the \textit{current} system state to set lead times (e.g., \citealt{Bertrand1983,
+    ChungHuang2002}). Put differently, these methods aim to find a function that provides lead times
+  based on the current state of the production system and possibly information from the past.
+
+\item \textsf{Predictive lead time management} may not only incorporate \textit{past} data and the
+  \textit{current} system state to set lead times, but also utilises the anticipated \textit{future}
+  system state to detect arising issues of future periods and react accordingly (e.g.
+  \citealt{PaterninaArboleda2001,schneckenreither2020order,Schneckenreither2019}). Thus, it extends
+  proactive lead time management from a flow time forecasting or simple lead time setting technique
+  to a lead time management approach that integrates the future behaviour of the system when setting
+  lead times. This allows reasoning of the system dynamics, as for instance triggering the lead time
+  system, and thus such an algorithm can react accordingly to find a more farsighted optimal lead
+  time update.
+\end{itemize}
+
+We hypothesis that dynamic lead time management approaches need to aim for a predictive design in
+order to be able to compete with state-of-the-art order release methods from literature. However,
+there only exist three papers that propose a predictive lead time management algorithm in
+literature.
+
+The first approach by \citet{PaterninaArboleda2001} introduces a predictive order release model by
+using reinforcement learning in a single product, four-station serial flow line and compare its
+performance (WIP costs) with conventional order release policies (e.g., Kanban and CONWIP).
+%
+% Reinforcement learning is an optimisation technique that stems from dynamic programming and its goal
+% is to find the best stationary policy for a given problem. This policy is usually provided by
+% assessing current and future states (or state-action pairs) of an underlying Markov Decision Process
+% (MDP).
+% The advantage of reinforcement learning over dynamic programming is that (i) the problem
+% space is explored by an agent and thus only expectantly interesting parts of the problem space need
+% to be assessed and (ii) the knowledge (acquisition) of transition probabilities becomes unnecessary
+% as the states are evaluated by consecutively observed states solely.
+%
+The algorithm of \citet{PaterninaArboleda2001} decides on whether or not to release an order after
+each system change, the completion of an operation of any order at any stage or a new order arrival,
+and assumes that any unsatisfied demand is lost. Thus, they use a continuous order release method
+although in practice order release decisions often need to be made on a periodical basis, e.g.\@
+daily (see \citealt{enns2004work,GeldersvanW1982}). They outperform existing control policies with their tabular
+based reinforcement learning agent.
+%
+Then, \citet{Schneckenreither2019} use several different reinforcement learning algorithms to make
+periodic order release decisions for a flow shop production system. The algorithm directly sets lead
+times for each product type, which are then used to release the orders in the order pool. They show
+that their approach outperforms static order release mechanisms by yielding lower costs, lateness
+and standard deviation of lateness, but conclude that research using average reward reinforcement
+learning methods harbor optimisation potential for the order release problem.
+%
+And finally, \citet{schneckenreither2020order} present a flow time estimation procedure to set lead
+times dynamically using an artificial neural network, which is used to forecast flow times. By
+implementing a rolling horizon order release simulation of the proceeding periods they lift their
+approach to a predictive lead time management approach, which is able to detect backorders of future
+periods and reacts by releasing orders earlier. Nonetheless, their method is unable to foresee the
+triggering of the lead time syndrome and therefore they introduce an upper lead time bound to
+prevent the negative effects of the lead time syndrome. Their model outperforms static and reactive
+lead time management approaches, especially for scenarios with high utilisation and high variability
+in processing times.
+
+As reinforcement learning stems from dynamic programming the future system state is by design
+considered as a main driver of decision making in the current period.
+% Reinforcement learning is an optimisation technique that stems from dynamic programming and
+Its goal is to find the best stationary policy for a given problem. % This policy is usually provided
+% by assessing current and future states (or state-action pairs) of an underlying Markov Decision
+% Process (MDP).
+The advantage of reinforcement learning over dynamic programming is that (i) the problem
+space is explored by an agent and thus only expectantly interesting parts of the problem space need
+to be assessed and (ii) the knowledge (acquisition) of transition probabilities becomes unnecessary
+as the states are evaluated by consecutively observed states solely.
 
@@
 low as possible \citep{Mahadevan98_OptimizingProductionManufacturingUsingReinforcementLearning}.
-However, in practise usually decision have to be made on a daily basis \citep{enns2004work}.
-Therefore, we refrain from this adaption and concentrate on standard MDPs only. Furthermore, often
-continuous-time semi-MDP problems can be converted through uniformisation into equivalent discrete
-time instances \citep[see][]{Puterman94,bertsekas1995dynamic}.
+However, in practise usually decision have to be made on a daily basis
+\citep{enns2004work,GeldersvanW1982}. Therefore, we refrain from this adaption and concentrate on
+standard MDPs only. Furthermore, often continuous-time semi-MDP problems can be converted through
+uniformisation into equivalent discrete time instances
+\citep[see][]{Puterman94,bertsekas1995dynamic}.
 

### 2020-12-18 paper/figures/printer.tex
update

\begin{tikzpicture}[thin, scale=0.65]
  % Nodes
  \foreach \x in {1,...,5}{
    \draw ({cos(\x*72-72)*2.5},{sin(\x*72-72)*2.5}) node(\x) [circle,draw,minimum size=25] {\footnotesize \(\x\)};
  };
  \foreach[evaluate={
    \y=int(\x+5);
  }]  \x in {2,...,3}{
    \draw ({-cos(\x*72-72)*2.5+5},{sin(\x*72-72)*2.5}) node(\y) [circle,draw,minimum size=25]
    {\footnotesize \(\x'\)};
  };
  \foreach [evaluate={
    \y=int(\x+5);
    \z=int(\x+5);
  }] \x in {4,...,5}{
    \draw ({-cos(\x*72-72)*2.5+5},{sin(\x*72-72)*2.5}) node(\z) [circle,draw,minimum size=25]
    {\footnotesize \(\y'\)};
  };

  \newcommand\bend{22.5}
  % Edges
  \path[thin, ->, bend right=\bend, >=stealth] (1) edge[below left] node {\footnotesize\(\)} (2);
  \path[thin, ->, bend right=\bend, >=stealth] (2) edge[above]      node {\footnotesize\(\)} (3);
  \path[thin, ->, bend right=\bend, >=stealth] (3) edge[left ]      node {\footnotesize\(\)} (4);
  \path[thin, ->, bend right=\bend, >=stealth] (4) edge[above]      node {\footnotesize\(\)} (5);
  \path[thin, ->, bend right=\bend, >=stealth] (5) edge[above left] node[yshift=-2] {\footnotesize\(5\)} (1);

  \path[thin, ->, bend left=\bend, >=stealth]         (1)  edge[below right] node {\footnotesize\(\)} (7);
  \path[thin, ->, bend left=\bend, >=stealth]         (7)  edge[above right] node {\footnotesize\(\)} (8);
  \path[thin, ->, dashed, bend left=\bend, >=stealth] (8)  edge[right]       node {\(\ldots\)} (9) ;
  \path[thin, ->, bend left=\bend, >=stealth]         (9)  edge[above]       node {\footnotesize\(\)} (10);
  \path[thin, ->, bend left=\bend, >=stealth]         (10) edge[above right] node[yshift=-2] {\footnotesize\(20\)} (1);

  \draw (0,0) node[] { Printer };
  \draw (5,0) node[] { Mail };

\end{tikzpicture}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../paper"
%%% End:

### 2020-12-18 paper/paper.tex
update

@@
 function use average reward reinforcement learning. \citet{aydin2000dynamic} use it in the setting
-of scheduling, while in a series of papers Mahadevan et al. investigated several problem domains
+of scheduling, while in a series of papers Mahadevan et al.\@ investigated several problem domains
 starting with simple MDPs
@@
 technique is able to find better policies. This yields from the fact that in standard discounted
-reinforcement learning method the states are assessed independently and by a single value. In
-contrast to that average reward reinforcement learning splits up the evaluation of the average
-reward per step, a bias value that specifies the amount of collected rewards to reach the optimal
-path when starting in a suboptimal state and an error term which defines the number of steps to
-reach the optimal path
-\citep{Howard64,Puterman94,Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults}.
-Thus, while in average reward reinforcement learning these terms are learned separately in the
-discounted framework one value consisting of the addition of these values is estimated, where
-however the average reward is scaled by \(1/(1-\gamma)\) with the discount factor \(\gamma\) usually
-being set very close to \(1\), e.g. \(0.99\). In commonly applied unichain MDPs the average reward
-per step is equal for all states. Thus, independently assessing it is computationally unwisely.
-Therefore, we adapt an algorithm that uses a scalar value for the estimation of the average reward.
-Furthermore, as the average reward is scaled in the discounted state values it dominates the other
-two terms. This, combined with the iterative evaluation and the independently assessing of the
-average reward for each state lets standard discounted reinforcement learning struggle to find good
-policies.
+reinforcement learning method the states are assessed independently and by a single value. To be
+more precise, average reward reinforcement learning splits up the evaluation of the average reward
+per step, a bias value that specifies the amount of collected rewards to reach the optimal path when
+starting in a suboptimal state and an error term which defines the number of steps to reach the
+optimal path \citep{Howard64,Puterman94,
+  Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults}. % Thus,
+% while in average reward reinforcement learning these terms are learned separately
+In contrast to that in the standard discounted framework one single scalar value consisting of the
+addition of these subterms is estimated, where however the average reward is scaled by
+\(1/(1-\gamma)\). The discount factor \(\gamma\) is usually set very close to \(1\), e.g. \(0.99\),
+which leads to the fact that this subterm dominates the other two.
+%
+Further, in commonly applied unichain (and thus ergodic) MDPs the average reward per step is equal
+for all states. Thus, independently assessing it is not only computationally unwisely, but also
+problematic when iteratively annealed as done in reinforcement learning. Therefore, we adapt an
+algorithm that uses a scalar value for the estimation of the average reward over all states, and
+estimates for every state-action pair that incorporate the bias and error term. The later values are
+adjusted by the average reward.
+% This, combined with the iterative evaluation and the independently assessing of the average reward
+% for each state lets standard discounted reinforcement learning struggle to find good policies.
 % Thus, this incautious combining of different kinds of state values as done in discounted
@@
 systems. Only the work by
-\cite{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}
+\citet{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}
 integrates this second-level refinement optimisation. However, their algorithm requires the
@@
 the lack of feedback in the iterative process of optimising the different refinement optimisation
-levels. Furthermore, they asses the average reward independently.
-
-Therefore, we propose a novel average reward reinforcement learning algorithm, which is the first
-that solves the occurring cyclic constraint problem in the setting of average reward reinforcement
-learning, to assess orders for their release to the shop floor. These cyclic constraint problem
-emerges as the underlying constraint structure is based on the Laurent series expansion of the state
-values as shown by \cite{MillerVeinott1969}, and thus easily imposes an infinite number of
-interconnected constraints when handled unwisely.
-%
-The agent assesses the expected costs for the possible releases imposed by adapting the lead time.
-According to the agent's estimates it sets a planned lead time and with that releases orders into
-the production system.
-% upon releasing an order for the current and future periods
-% and, based on its estimates decides whether to release the order or not.
+levels. Furthermore, they asses the average reward independently for each state.
+
+In summary, we propose a novel average reward adjusted reinforcement learning algorithm to assess
+orders for their release to the shop floor. This is the first average reward adjusted reward
+learning algorithm that operates using a artificial neural network as function approximation. To
+ensure scalability we adapt a multi-agent approach, where each agent optimises the lead time
+management of a single product type. The agents do so by assessesing the expected costs for the
+possible releases imposed by adapting the lead time. According to the learned estimates each agent
+sets a planned lead time for the corresponding product type and with that releases orders into the
+production system. The highly optimised value-iteration algorithm uses improved n-step reinforcement
+learning with small action-based replay memories and worker agents to infer order release policies
+which outerperform \MS[t]{todo}.
+
 %
@@
 
-Thus, the action space is defined as \(F = \times_{s=1}^{\size{\States}} A_{s}\), where \(A_{s}\) is a
-finite set of possible actions. A \emph{policy} is a sequence \(\pol = (f_{1},f_{2},\ldots)\) of
+Thus, the action space is defined as \(F = \times_{s=1}^{\size{\States}} A_{s}\), where \(A_{s}\) is
+a finite set of possible actions. A \emph{policy} is a sequence \(\pol = (f_{1},f_{2},\ldots)\) of
 elements \(f_{t} \in F\). Using the policy \(\pol\) means that if the system is in state \(s\) at
@@
 the \(s\)-th component of \(f_{t}\), is chosen. A stationary policy \(\pol = (f,f,\ldots)\) does not
-depend on time. In the sequel we are concerned with stationary policies only. Thus the goal in
+depend on time. In the sequel we are concerned with stationary policies only. The goal in
 reinforcement learning (RL) is to find the best stationary policy \(\pol\) for a given problem,
@@
 criteria is usually referred to as \(\gamma\)-optimality as the discount factor \(\gamma\) is fixed.
-However, this also means that the actual value set for \(\gamma\) defines the policy that is
-optimised for. For instance, as can be seen in Figure~\ref{fig:printer} setting \(\gamma\) value to
+However, this also means that the actual value set for \(\gamma\) determines the best achievable
+policy. For instance, as can be seen in Figure~\ref{fig:printer} setting \(\gamma\) value to
 \(<0.8027\) the printer-loop is preferred over the mail-loop by the agent, although the mail-loop
-accumulates more reward over time, i.e. selecting the mail-loop is a sub-optimal choice.
+accumulates more reward over time, i.e.\@ selecting the mail-loop is a sub-optimal choice. Observe
+that the average reward received per step equals \(1\) for the printer-loop and \(2\) for the
+mail-loop. Thus, the (Blackwell-)optimal policy is to choose the mail-loop. However, if
+\(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted reinforcement learning
+chooses the printer loop.
 
@@
   \centering
-  \begin{tikzpicture}[thin, scale=0.75]
-    % Nodes
-    \foreach \x in {1,...,5}{
-      \draw ({cos(\x*72-72)*2.5},{sin(\x*72-72)*2.5}) node(\x) [circle,draw,minimum size=25] {\footnotesize \(\x\)};
-    };
-    \foreach[evaluate={
-      \y=int(\x+5);
-    }]  \x in {2,...,3}{
-      \draw ({-cos(\x*72-72)*2.5+5},{sin(\x*72-72)*2.5}) node(\y) [circle,draw,minimum size=25]
-      {\footnotesize \(\x'\)};
-    };
-    \foreach [evaluate={
-      \y=int(\x+5);
-      \z=int(\x+5);
-    }] \x in {4,...,5}{
-      \draw ({-cos(\x*72-72)*2.5+5},{sin(\x*72-72)*2.5}) node(\z) [circle,draw,minimum size=25]
-      {\footnotesize \(\y'\)};
-    };
-
-    % Edges
-    \path[thin, ->, bend right, >=stealth] (1) edge[below left] node {\footnotesize\(\)} (2);
-    \path[thin, ->, bend right, >=stealth] (2) edge[above]      node {\footnotesize\(\)} (3);
-    \path[thin, ->, bend right, >=stealth] (3) edge[left ]      node {\footnotesize\(\)} (4);
-    \path[thin, ->, bend right, >=stealth] (4) edge[above]      node {\footnotesize\(\)} (5);
-    \path[thin, ->, bend right, >=stealth] (5) edge[above left] node {\footnotesize\(5\)} (1);
-
-    \path[thin, ->, bend left, >=stealth]         (1)  edge[below right] node {\footnotesize\(\)} (7);
-    \path[thin, ->, bend left, >=stealth]         (7)  edge[above right] node {\footnotesize\(\)} (8);
-    \path[thin, ->, dashed, bend left, >=stealth] (8)  edge[right]       node {\(\ldots\)} (9) ;
-    \path[thin, ->, bend left, >=stealth]         (9)  edge[above]       node {\footnotesize\(\)} (10);
-    \path[thin, ->, bend left, >=stealth]         (10) edge[above right] node {\footnotesize\(20\)} (1);
-
-    \draw (0,0) node[] { Printer };
-    \draw (5,0) node[] { Mail };
-
-  \end{tikzpicture}
-  \caption{\label{fig:printer} A MDP with two different deterministic policies. The only action choice is in
-    state \(1\), in which the agent can choose between doing the printer-loop or the mail-loop.
-    Observe that the average reward received per step equals \(1\) for the printer-loop and \(2\)
-    for the mail-loop. Thus, the (Blackwell-)optimal policy is to choose the mail-loop. However,
-    if \(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted reinforcement
-    learning chooses the printer loop.
-    %
-    Adapted from \cite{Mahadevan96_OptimalityCriteriaInReinforcementLearning}.}
+  \input{figures/printer}
+  \caption{\label{fig:printer} A MDP with a single action choice in
+    state \(1\), i.e.\@ two different deterministic policies
+    % , in which the agent can choose between doing the printer-loop or the mail-loop  . The
+    % Observe that the average reward received per step equals \(1\) for the printer-loop and \(2\)
+    % for the mail-loop. Thus, the (Blackwell-)optimal policy is to choose the mail-loop. However,
+    % if \(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted reinforcement
+    % learning chooses the printer loop.
+    % %
+    \citep[][adapted]{Mahadevan96_OptimalityCriteriaInReinforcementLearning}.
+  }
 \end{figure}
@@
 almost all cases the aim is to perform well over time which in terms of reward means to first
-maximise for a the policy with highest average reward before more selectively choosing actions. That
-is, if several processes exist, we are searching for the ones with highest average reward before
-considering other criteria.
+maximise for a policy with highest average reward before more selectively choosing actions. That is,
+we are searching for the policies with highest average reward before considering other criteria.
 %
 Therefore, in almost all RL studies the discount factor is set to a value very close to \(1\), for
-instance to \(0.99\)~\cite[e.g.]{mnih2015human,mnih2016asynchronous,Lillicrap15}.
+instance to \(0.99\)~\citep[e.g.][]{mnih2015human,mnih2016asynchronous,Lillicrap15}.
 %
@@
 \citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults,Puterman94}.
-For simplicity we concentrate on unichain MDPs in this work and thus may simply refer to it as
+In the sequel we focus on unichain MDPs in this work and thus may simply refer to it as
 \(\avgrew^{\pol}\). A policy that maximises the average reward in every state is called
-gain-optimal. Gain optimality is the least selective criteria an average reward reinforcement
+gain-optimal. Gain-optimality is the least selective criteria an average reward reinforcement
 learning aims for.
@@
 average reward. Thus the bias value can be seen as the rewards that additionally sum up in case the
-process starts in state \(s\). A policy that is gain-optimal is also bias-optimal if it maximises
+process starts in state \(s\). A policy that is gain-optimal is also bias-optimal, if it maximises
 the bias values in every state and compared to every other policy.
@@
 only choice in state \(1\). Both policies have the same average reward of \(1\). However, only
-selecting the A-loop is bias-optimal, as the actions with non-zero rewards are selected earlier.
-E.g.~consider starting in state 1. Under the policy \(\pol_{A}\) which takes the A-loop the reward
-sequence is \((2,0,2,0,\ldots)\), while for the other policy \(\pol_{B}\) it is
-\((0,2,0,2,\ldots)\).
+taking the A loop is bias-optimal, as its policy \(\pol_{A}\) leads to bias values
+\(V^{\pol_{A}}(1)=0.5\), \(V^{\pol_{A}}(0)=-0.5\) and \(V^{\pol_{A}}(2)=1.5\), while policy
+\(\pol_{B}\) which selects the B loop generates bias values \(V^{\pol_{B}}(1)=-0.5\),
+\(V^{\pol_{B}}(0)=-1.5\) and \(V^{\pol_{B}}(2)=0.5\).
+%
+This yields from the fact that the actions with non-zero rewards are selected earlier in policy
+\(\pol_{A}\). Consider starting in state 1. Under the policy \(\pol_{A}\) the reward sequence is
+\((2,0,2,0,\ldots)\), while for the other policy \(\pol_{B}\) it is \((0,2,0,2,\ldots)\).
 
@@
   \end{tikzpicture}
-  \caption{\label{fig:three-states} A task with two different MDPs, A (going left in 1) and B
-    (going right). As in the previous example the only action choice is in state \(1\). Observe
-    that the average reward received per step equals \(1\) for both policies. However, only taking
-    the A loop is bias-optimal, as its policy \(\pol_{A}\) leads to bias values
-    \(V^{\pol_{A}}(1)=0.5\), \(V^{\pol_{A}}(0)=-0.5\) and \(V^{\pol_{A}}(2)=1.5\), while policy
-    \(\pol_{B}\) which selects the B loop generates bias values \(V^{\pol_{B}}(1)=-0.5\),
-    \(V^{\pol_{B}}(0)=-1.5\) and \(V^{\pol_{B}}(2)=0.5\). Adapted from
-    \cite{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults}.}
+  \caption{\label{fig:three-states} A MDP with two gain-optimal deterministic policies, \(\pol_{A}\)
+    going left in 1 is also bias-optimal, while \(\pol_{B}\) is not \citep[Adapted
+    from][]{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults}.}
 \end{figure}
 
-\subsection{More Refined Optimality Criteria}
+\subsection{Optimality Criteria}
 \label{subsec:More_Refined_Optimality_Criteria}
 
-The aforementioned and informally introduced notions of gain-optimality and bias-optimality can be
-generalised to \(n\)-discount-optimality and will now be defined formally.
+The aforementioned and informally introduced notions of gain- and bias-optimality are formalised and
+generalised to \(n\)-discount-optimality.
 
 \begin{definition}
-  Due to \cite{Veinott69} for MDPs a policy \(\polopt\) is \emph{\(n\)-discount-optimal} for
+  Due to \citet{Veinott69} for MDPs a policy \(\polopt\) is \emph{\(n\)-discount-optimal} for
   \(n=-1,0,1,\ldots\) for all states \(s \in \States\) with discount factor
@@
 \end{definition}
+
+Only if a policy is \(n\)-discount-optimal for all \(n < m\) it can be \(m\)-discount
+optimal \citep{Puterman94,Veinott69}.
+%
+If a policy is \(\infty\)-discount-optimal then it is said to be Blackwell-optimal
+\citep{Blackwell62}.
+% weakening the argument!!!
+That is, for Blackwell-optimal policies \(\polopt\) there exists a discount factor \(\gammaopt < 1\)
+such that \(V_{\gamma}^{\polopt}(s) \geqslant V_{\gamma}^{\pol}(s)\) for all
+\(\gamma \geqslant \gammaopt\) and under all policies \(\pol\)
+\citep{Mahadevan96_SensitiveDiscountOptimalityUnifyingDiscountedAndAverageRewardReinforcementLearning,Blackwell62}.
+Informally that means there exists a discount factor \(<1\) which finds Blackwell-optimal policies.
+However, (i) in more complex, i.e.\@ real world, MDPs this value can be arbitrary close to \(1\) and
+(ii) the difference in state values may be very small, which (iii) due to the need of state value
+function approximation likely causes errors when choosing among different actions in the discounted
+framework.
 
@@
 the case of \(n=0\) it describes bias-optimality with \(V^{\polopt}(s) - V^{\pol}(s) \geqslant 0\)
-forall policies \(\pol\) and states \(s\)
+for all policies \(\pol\) and states \(s\)
 \citep{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}.
-%
-Only if a policy is \(n\)-discount-optimal for all \(n < m\) it can be \(m\)-discount
-optimal \citep{Puterman94,Veinott69}.
-%
-If a policy is \(\infty\)-discount-optimal then it is said to be
-Blackwell-optimal \citep{Blackwell62}.
-% weakening the argument!!!
-I.e.~for Blackwell-optimal policies \(\polopt\) there exists a discount factor \(\gammaopt < 1\)
-such that \(V_{\gamma}^{\polopt}(s) \geqslant V_{\gamma}^{\pol}(s)\) for all
-\(\gamma \geqslant \gammaopt\) and under all policies \(\pol\)
-\citep{Mahadevan96_SensitiveDiscountOptimalityUnifyingDiscountedAndAverageRewardReinforcementLearning,Blackwell62}.
-Informally that means there exists a discount factor \(<1\) which finds Blackwell-optimal policies.
-However, (i) in more complex, i.e. real world, MDPs this value can be arbitrary close to \(1\) and
-(ii) the difference in state values may be very small, which (iii) due to the need of state value
-function approximation likely causes errors when choosing among different actions in the discounted
-framework.
-% Real world applications underlie much more complex MDPs as the ones of Figures~\ref{fig:printer}
-% and~\ref{fig:three-states}. Thus \(\gamma\)-values even very close to \(1\) can lead to sub-optimal
-% policies. To investigate these differences we will now present a more elective optimality criteria.
-%
 %
@@
 
-where \cite{Puterman94} shows that \(\lim_{\gamma \to 1} e_{\gamma}^{\pol}(s) = 0\). Note how the
-first term depending on the average reward \(\avgrew^{\pol}(s)\) converges to infinity as \(\gamma\)
+where the error term \(e_{\gamma}^{\pol}(s)\) exists of infinitely many subterms and
+\citet{Puterman94} shows that \(\lim_{\gamma \to 1} e_{\gamma}^{\pol}(s) = 0\). Note how the first
+term depending on the average reward \(\avgrew^{\pol}(s)\) converges to infinity as \(\gamma\)
 increases.
@@
 %
-Finally \(n\)-discount-optimality for \(n \geqslant 1\) requires to maximise the term
+Finally \(n\)-discount-optimality for \(n \geqslant 1\) requires to maximise the error term
 \(e_{\gamma}^{\pol}(s)\) which only exists as \(\gamma\) is strictly less than \(1\). Thus this term
-incorporates the number of steps its expected rewards on the path which are required for reaching
-the highest rewards. That is, it optimises the expected reward according to the occurrence on the
-paths, where shorter paths and those which collect the rewards sooner are preferred.
+incorporates the number of expected steps and their corresponding reward on the path to the
+Blackwell-optimal policy. That is, it optimises the expected reward collected to reach the Blackwell
+optimal policy according to the occurrence on the paths, where shorter paths and those which collect
+the rewards sooner are preferred.
 

### 2020-12-26 paper/paper.tex
shortened standard discounted RL

@@
 
-This section introduces average reward adjusted reinforcement learning in comparison to the discounted
-framework, elaborates on optimiality criteria and provides insights of the underlying algorithm.
+This section briefly reintroduces the most important concepts of average reward adjusted
+reinforcement learning, elaborates on optimality criteria and provides insights of the underlying
+algorithm. For a more extensive introduction to average reward reinforcement learning we refer to
+\citet{schneckenreither2020average}.
 
@@
 \begin{align*}
-  V_{\gamma}^{\pol_{\gamma}}(s)=\lim_{N \to \infty} E[\sum_{t=0}^{N-1} \gamma^{t} R_{t}^{\pol_{\gamma}}(s)]\tcom
+  V_{\gamma}^{\pol_{\gamma}}(s)=\lim_{N \to \infty} \E[\sum_{t=0}^{N-1} \gamma^{t} R_{t}^{\pol_{\gamma}}(s)]\tcom
 \end{align*}
@@
 where \(0 \leqslant \gamma < 1\) is the discount factor and
-\(R_{t}^{\pol}(s) = E_{\pol}[ r(s_{t},a_{t}) \mid s_{t} = s, a_{t} =
-a]\) % \MS[t]{check if correct!}
-the reward received at time \(t\) upon starting in state \(s\) by following policy \(\pol\)
+\(R_{t}^{\pol}(s) = \E_{\pol}[ r(s_{t},a_{t}) \mid s_{t} = s, a_{t} = a]\) the reward received at
+time \(t\) upon starting in state \(s\) by following policy \(\pol\)
 \citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults}.
-
-The aim in the discounted framework is to find an optimal policy \(\polopt_{\gamma}\), which when
-followed, maximises the state value for all states \(s\) as compared to any other policy
-\(\pol_{\gamma}\): \(V_{\gamma}^{\polopt_{\gamma}} - V_{\gamma}^{\pol_{\gamma}} \geqslant 0\). This
-criteria is usually referred to as \(\gamma\)-optimality as the discount factor \(\gamma\) is fixed.
-However, this also means that the actual value set for \(\gamma\) determines the best achievable
-policy. For instance, as can be seen in Figure~\ref{fig:printer} setting \(\gamma\) value to
-\(<0.8027\) the printer-loop is preferred over the mail-loop by the agent, although the mail-loop
-accumulates more reward over time, i.e.\@ selecting the mail-loop is a sub-optimal choice. Observe
-that the average reward received per step equals \(1\) for the printer-loop and \(2\) for the
-mail-loop. Thus, the (Blackwell-)optimal policy is to choose the mail-loop. However, if
-\(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted reinforcement learning
-chooses the printer loop.
-
-\begin{figure}[t!]
-  \centering
-  \input{figures/printer}
-  \caption{\label{fig:printer} A MDP with a single action choice in
-    state \(1\), i.e.\@ two different deterministic policies
-    % , in which the agent can choose between doing the printer-loop or the mail-loop  . The
-    % Observe that the average reward received per step equals \(1\) for the printer-loop and \(2\)
-    % for the mail-loop. Thus, the (Blackwell-)optimal policy is to choose the mail-loop. However,
-    % if \(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted reinforcement
-    % learning chooses the printer loop.
+\MS[t]{how are the update rules?}
+
+The Bellman Equation defined as
+\(V_{\gamma}^{\pol}(s) = \E_{\pol}[r + \gamma V_{\gamma}^{\pol}(s')]\) \citep[see
+e.g.][p.70]{sutton1998introduction} provides a way to assess the state values using two
+consecutively observed states \(s, s'\) and the observed reward \(r\) returned by the system. When
+this formula is used as an update rule it provides an algorithm that converges the state values by
+iteratively adapting the state value estimates.
+%
+There are four major issues with discounted RL, which results in the fact that it is inapplicable
+for many problems of operations research.
+
+\begin{enumerate}
+\item Standard discounted RL can only infer suboptimal policies as the discount factor is strictly
+  less than one, i.e.\@ \(\gamma < 1\).
+\item It is very difficult to specify a desired balance between the short-term and long-term
+  (average reward) rewards. This results from the fact that the long-term rewards are scaled
+  exponentially, where the factor depends on the chosen \(\gamma\) value.
+\item Episodic MDPs with an average reward per step that is non-zero cannot be solved, as the
+  average reward is ignored in the terminal states.
+\item The average reward is independently assessed for each state, even though it is the dominating
+  part for all state values and thus is shared between more than one or, for unichain MDPs, even all
+  states. This usually leads to an exponentially increased number of learning steps required for
+  policies to converge.
+\end{enumerate}
+
+% The aim in the discounted framework is to find an optimal policy \(\polopt_{\gamma}\), which when
+% followed, maximises the state value for all states \(s\) as compared to any other policy
+% \(\pol_{\gamma}\): \(V_{\gamma}^{\polopt_{\gamma}} - V_{\gamma}^{\pol_{\gamma}} \geqslant 0\). This
+% criteria is usually referred to as \(\gamma\)-optimality as the discount factor \(\gamma\) is fixed.
+% However, this also means that the actual value set for \(\gamma\) determines the best achievable
+% policy. For instance, as can be seen in Figure~\ref{fig:printer} setting \(\gamma\) value to
+% \(<0.8027\) the printer-loop is preferred over the mail-loop by the agent, although the mail-loop
+% accumulates more reward over time, i.e.\@ selecting the mail-loop is a sub-optimal choice. Observe
+% that the average reward received per step equals \(1\) for the printer-loop and \(2\) for the
+% mail-loop. Thus, the (Blackwell-)optimal policy is to choose the mail-loop. However, if
+% \(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted reinforcement learning
+% chooses the printer loop.
+
+% \begin{figure}[t!]
+%   \centering
+%   \input{figures/printer}
+%   \caption{\label{fig:printer} A MDP with a single action choice in
+%     state \(1\), i.e.\@ two different deterministic policies
+%     % , in which the agent can choose between doing the printer-loop or the mail-loop  . The
+%     % Observe that the average reward received per step equals \(1\) for the printer-loop and \(2\)
+%     % for the mail-loop. Thus, the (Blackwell-)optimal policy is to choose the mail-loop. However,
+%     % if \(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted reinforcement
+%     % learning chooses the printer loop.
+%     % %
+%     \citep[][adapted]{Mahadevan96_OptimalityCriteriaInReinforcementLearning}.
+%   }
+% \end{figure}
+
+% One idea behind the \(\gamma\)-parameter of the discounted framework is to be able to balance
+% short-term rewards (low \(\gamma\)-values) and long-term rewards (high \(\gamma\)-values). However,
+% what seems to be an advantage rather becomes a disadvantage for most applications. The issue is that
+% the average reward value is non-linearly increased when \(\gamma\) approaches one. However, in
+% almost all cases the aim is to perform well over time which in terms of reward means to first
+% maximise for a policy with highest average reward before more selectively choosing actions. That is,
+% we are searching for the policies with highest average reward before considering other criteria.
 % %
-    \citep[][adapted]{Mahadevan96_OptimalityCriteriaInReinforcementLearning}.
-  }
-\end{figure}
-
-One idea behind the \(\gamma\)-parameter of the discounted framework is to be able to balance
-short-term rewards (low \(\gamma\)-values) and long-term rewards (high \(\gamma\)-values). However,
-what seems to be an advantage rather becomes a disadvantage for most applications. The issue is that
-the average reward value is non-linearly increased when \(\gamma\) approaches one. However, in
-almost all cases the aim is to perform well over time which in terms of reward means to first
-maximise for a policy with highest average reward before more selectively choosing actions. That is,
-we are searching for the policies with highest average reward before considering other criteria.
-%
-Therefore, in almost all RL studies the discount factor is set to a value very close to \(1\), for
-instance to \(0.99\)~\citep[e.g.][]{mnih2015human,mnih2016asynchronous,Lillicrap15}.
-%
-By doing so the above mentioned non-linear relationship leads to the fact that the state value
-consists almost solely of the up-scaled average reward term, whereas the bias values can be
-neglected. This leads to diluted state values and thus difficulties in distinguishing actions which
-impose policies that possess the same average reward.
-%
-In contrast to that, average reward RL separately assesses these values and according to these
-selectively chooses the best action, comparing one after the other.
-%
-To illustrate the idea reconsider Figure~\ref{fig:printer}. First the agent picks the policy
-according to the highest average reward. Thus, all actions are assessed by an average reward of
-\(2\) which is inferred from the Mail-loop. This however, makes choosing the Printer-loop
-unattractive; i.e. the policy converges to the optimal policy of choosing the Mail loop.
+% Therefore, in almost all RL studies the discount factor is set to a value very close to \(1\), for
+% instance to \(0.99\)~\citep[e.g.][]{mnih2015human,mnih2016asynchronous,Lillicrap15}.
+% %
+% By doing so the above mentioned non-linear relationship leads to the fact that the state value
+% consists almost solely of the up-scaled average reward term, whereas the bias values can be
+% neglected. This leads to diluted state values and thus difficulties in distinguishing actions which
+% impose policies that possess the same average reward.
+% %
+% In contrast to that, average reward RL separately assesses these values and according to these
+% selectively chooses the best action, comparing one after the other.
+% %
+% To illustrate the idea reconsider Figure~\ref{fig:printer}. First the agent picks the policy
+% according to the highest average reward. Thus, all actions are assessed by an average reward of
+% \(2\) which is inferred from the Mail-loop. This however, makes choosing the Printer-loop
+% unattractive; i.e. the policy converges to the optimal policy of choosing the Mail loop.
+
+Especially due to the fourth issue standard discounted RL is not well applicable in the operations
+research domain. This mainly results from the fact that a continuous assessing of the state values
+using costs or profit is often required, which leads to intractable long learning phases with high
+exploration rate even for small sized problems. But as high exploration rates produce errors in the
+state value estimations \citep{MillerVeinott1969} finding a well working parameterisation is
+difficult. Average reward reinforcement learning overcomes these issues by separately assessing
+addends of the state values.
 
@@
 \(\avgrew^{\pol}\). A policy that maximises the average reward in every state is called
-gain-optimal. Gain-optimality is the least selective criteria an average reward reinforcement
-learning aims for.
+gain-optimal. % Gain-optimality is the least selective criteria an average reward reinforcement
+% learning aims for.
 
@@
 can be reformulated to following constraint problem with expected reward
-\(R^{\pol}(s) = E_{\pol}[ r(s, a) ] = \sum_{a \in \mathcal{A}(s)}p(a | s) r(s, a)\) as shown by
+\(R^{\pol}(s) = \E_{\pol}[ r(s, a) ] = \sum_{a \in \mathcal{A}(s)}p(a | s) r(s, a)\) as shown by
 \cite{MillerVeinott1969} and \citet[p.346]{Puterman94}:
@@
 \begin{align}
-  \label{eq:constr1}\avgrew^{\pol}(s) - E[\avgrew^{\pol}(s)] & = 0 &  & \text{for } n = -1 \\
-  \label{eq:constr2}\avgrew^{\pol}(s) + V^{\pol}(s) - E[V^{\pol}(s)] & = R^{\pol}(s) && \text{for } n = 0 \\
-  \label{eq:constr3}W^{\pol}_{n-1}(s) + W^{\pol}_{n}(s) - E[W^{\pol}_{n}(s)] & = 0 && \text{for } n \geqslant 1, \text{where } W_{0}^{\pol}(s) = V^{\pol}(s)
+  \label{eq:constr1}\avgrew^{\pol}(s) - \E[\avgrew^{\pol}(s)] & = 0 &  & \text{for } n = -1 \\
+  \label{eq:constr2}\avgrew^{\pol}(s) + V^{\pol}(s) - \E[V^{\pol}(s)] & = R^{\pol}(s) && \text{for } n = 0 \\
+  \label{eq:constr3}W^{\pol}_{n-1}(s) + W^{\pol}_{n}(s) - \E[W^{\pol}_{n}(s)] & = 0 && \text{for } n \geqslant 1, \text{where } W_{0}^{\pol}(s) = V^{\pol}(s)
 \end{align}

### 2020-12-27 paper/paper.tex
todo: algorithm, title

@@
 
+\newcommand\theauthor{
+  Manuel Schneckenreither% \inst{1}
+  % \orcidID{0000-0002-4812-4665}
+  \and Stefan Haeussler% \inst{2}
+  % \orcidID{0000-0003-2589-1367}
+}
+\newcommand\thetitle{Average Reward Adjusted Deep Reinforcement Learning: Near-Blackwell-Optimal
+  Policies for Order Release}
+
 % \usepackage{academicons}
@@
 plainpages=false,
-pdftitle={Reinforcement Learning Methods for Operations Research Applications -- Reloaded},
-pdfauthor={Manuel Schneckenreither and Stefan Haeussler},
-pdfsubject={Reinforcement Learning Methods for Operations Research Applications -- Reloaded},
-pdfkeywords={production planning, order release, machine learning, reinforcement learning,
+pdftitle={\thetitle{}},
+pdfauthor={\theauthor{}},
+pdfsubject={\thetitle{}},
+pdfkeywords={average reward adjusted reinforcement learning, production planning, order release, machine learning,
   operations research},
@@
 % \newcolumntype{x}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}
-\title{
-  Reinforcement Learning Methods for Operations Research Applications -- Reloaded
-}
-\author{Manuel Schneckenreither% \inst{1}
-  % \orcidID{0000-0002-4812-4665}
-  \and Stefan Haeussler% \inst{2}
-  % \orcidID{0000-0003-2589-1367}
-}
+
+\author{\theauthor}
+\date{\today}
+\title{\thetitle}
 
@@
 the \(s\)-th component of \(f_{t}\), is chosen. A stationary policy \(\pol = (f,f,\ldots)\) does not
-depend on time. In the sequel we are concerned with stationary policies only. The goal in
-reinforcement learning (RL) is to find the best stationary policy \(\pol\) for a given problem,
-where the phrase ``best'' will be elaborated in detail below.
+depend on time. In the sequel we are concerned with stationary policies only.
 
@@
 \citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults}.
-\MS[t]{how are the update rules?}
+%
+The aim in the discounted framework is to find an optimal policy \(\polopt_{\gamma}\), which when
+followed, maximises the state value for all states \(s\) as compared to any other policy
+\(\pol_{\gamma}\): \(V_{\gamma}^{\polopt_{\gamma}} - V_{\gamma}^{\pol_{\gamma}} \geqslant 0\). This
+criteria is usually referred to as \(\gamma\)-optimality as the discount factor \(\gamma\) is fixed.
+However, this also means that the actual value set for \(\gamma\) determines the best achievable
+policy. For instance, as can be seen in Figure~\ref{fig:printer} setting \(\gamma <0.8027\) the
+printer-loop is preferred over the mail-loop, although the mail-loop accumulates more
+reward over time, i.e.\@ selecting the mail-loop is a sub-optimal choice. Observe that the average
+reward received per step equals \(1\) for the printer-loop and \(2\) for the mail-loop. Thus, the
+(Blackwell-)optimal policy is to choose the mail-loop. However, if
+\(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted reinforcement learning
+chooses the printer loop as going the printer loop has a higher evaluation.
+%
+\begin{figure}[t!]
+  \centering
+  \input{figures/printer}
+  \caption{\label{fig:printer} A MDP with a single action choice in
+    state \(1\), i.e.\@ two different deterministic policies
+    % , in which the agent can choose between doing the printer-loop or the mail-loop  . The
+    % Observe that the average reward received per step equals \(1\) for the printer-loop and \(2\)
+    % for the mail-loop. Thus, the (Blackwell-)optimal policy is to choose the mail-loop. However,
+    % if \(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted reinforcement
+    % learning chooses the printer loop.
+    % %
+    \citep[][adapted]{Mahadevan96_OptimalityCriteriaInReinforcementLearning}.
+  }
+\end{figure}
+%
+
+Therefore, in the seek of a more general optimality criteria for determining the best stationary
+policy \(\pol\) for a given problem, we introduce \(n\)-discount-optimality.
+
+% \subsection{Optimality Criteria}
+% \label{subsec:More_Refined_Optimality_Criteria}
+
+% The aforementioned and informally introduced notions of gain- and bias-optimality are formalised and
+% generalised to \(n\)-discount-optimality as follows.
+
+\begin{definition}[\(n\)-Discount-Optimality]
+  Due to \citet{Veinott69} for a MDP a policy \(\polopt\) is \emph{\(n\)-discount-optimal} for
+  \(n=-1,0,1,\ldots\) for all states \(s \in \States\) with discount factor
+  \(\gamma\) % and \(V_{\gamma}^{\pol}(s)\) being the value function as defined above
+  %
+  if and only if
+  \begin{align*}
+    \lim_{\gamma \to 1}(1-\gamma)^{-n}\ (V_{\gamma}^{\polopt}(s) - V_{\gamma}^{\pol}(s)) \geqslant 0 \tpkt
+  \end{align*}
+\end{definition}
+
+Only if a policy is \(n\)-discount-optimal for all \(n < m\) it can be \(m\)-discount
+optimal \citep{Puterman94,Veinott69}.
+%
+If a policy is \(\infty\)-discount-optimal then it is said to be Blackwell-optimal
+\citep{Blackwell62}.
+% weakening the argument!!!
+That is, for Blackwell-optimal policies \(\polopt\) there exists a discount factor \(\gammaopt < 1\)
+such that \(V_{\gamma}^{\polopt}(s) \geqslant V_{\gamma}^{\pol}(s)\) for all
+\(\gamma \geqslant \gammaopt\) and under all policies \(\pol\)
+\citep{Mahadevan96_SensitiveDiscountOptimalityUnifyingDiscountedAndAverageRewardReinforcementLearning,Blackwell62}.
+Informally that means there exists a discount factor \(\gamma <1\) which finds Blackwell-optimal
+policies. However, (i) in more complex, i.e.\@ real world, MDPs this value can be arbitrary close to
+\(1\) and (ii) the difference in state values may be very small, which (iii) due to the need of
+state value function approximation likely causes errors when choosing among different actions in the
+discounted framework.
+
+% In the sequel we present the Laurent series expansion which not only links discounted reinforcement
+% learning with average reward reinforcement learning, but also connects to \(n\)-discount-optimality.
+
+% Furthermore, for the rest of the paper we occasionally write drop the \(V(s)\) and $V_{\gamma}$ instead of
+% \(V^{\pol}(s)\) and \(V^{\pol}_{\gamma}\) when the policy is clear from the context.
 
@@
 \end{enumerate}
-
-% The aim in the discounted framework is to find an optimal policy \(\polopt_{\gamma}\), which when
-% followed, maximises the state value for all states \(s\) as compared to any other policy
-% \(\pol_{\gamma}\): \(V_{\gamma}^{\polopt_{\gamma}} - V_{\gamma}^{\pol_{\gamma}} \geqslant 0\). This
-% criteria is usually referred to as \(\gamma\)-optimality as the discount factor \(\gamma\) is fixed.
-% However, this also means that the actual value set for \(\gamma\) determines the best achievable
-% policy. For instance, as can be seen in Figure~\ref{fig:printer} setting \(\gamma\) value to
-% \(<0.8027\) the printer-loop is preferred over the mail-loop by the agent, although the mail-loop
-% accumulates more reward over time, i.e.\@ selecting the mail-loop is a sub-optimal choice. Observe
-% that the average reward received per step equals \(1\) for the printer-loop and \(2\) for the
-% mail-loop. Thus, the (Blackwell-)optimal policy is to choose the mail-loop. However, if
-% \(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted reinforcement learning
-% chooses the printer loop.
-
-% \begin{figure}[t!]
-%   \centering
-%   \input{figures/printer}
-%   \caption{\label{fig:printer} A MDP with a single action choice in
-%     state \(1\), i.e.\@ two different deterministic policies
-%     % , in which the agent can choose between doing the printer-loop or the mail-loop  . The
-%     % Observe that the average reward received per step equals \(1\) for the printer-loop and \(2\)
-%     % for the mail-loop. Thus, the (Blackwell-)optimal policy is to choose the mail-loop. However,
-%     % if \(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted reinforcement
-%     % learning chooses the printer loop.
-%     % %
-%     \citep[][adapted]{Mahadevan96_OptimalityCriteriaInReinforcementLearning}.
-%   }
-% \end{figure}
 
@@
 In the sequel we focus on unichain MDPs in this work and thus may simply refer to it as
-\(\avgrew^{\pol}\). A policy that maximises the average reward in every state is called
-gain-optimal. % Gain-optimality is the least selective criteria an average reward reinforcement
+\(\avgrew^{\pol}\). A policy \(\polopt\) that maximises the average reward
+\(\avgrew^{\polopt}(s) - \avgrew^{\pol}(s) \geqslant 0\) in every state \(s\) as compared to any
+other policy \(\pol\) is called gain-optimal.
+%
+% A policy that maximises the average reward in every state is called
+% gain-optimal
+% \citep{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}.
+% Gain-optimality is the least selective criteria an average reward reinforcement
 % learning aims for.
+\citet{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}
+shows that \(n\)-discount-optimality for \(n=-1\) describes gain-optimality.
 
@@
 average reward. Thus the bias value can be seen as the rewards that additionally sum up in case the
-process starts in state \(s\). A policy that is gain-optimal is also bias-optimal, if it maximises
-the bias values in every state and compared to every other policy.
+process starts in state \(s\). A policy \(\polopt\) that is gain-optimal is also bias-optimal, if it
+maximises the bias values \(V^{\polopt}(s) - V^{\pol}(s) \geqslant 0\) in every state \(s\) and
+compared to every other policy \(\pol\). Bias-optimality is given by setting \(n=0\) in the
+definition of \(n\)-discount-optimality
+\citep{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}.
 
@@
 
-\subsection{Optimality Criteria}
-\label{subsec:More_Refined_Optimality_Criteria}
-
-The aforementioned and informally introduced notions of gain- and bias-optimality are formalised and
-generalised to \(n\)-discount-optimality.
-
-\begin{definition}
-  Due to \citet{Veinott69} for MDPs a policy \(\polopt\) is \emph{\(n\)-discount-optimal} for
-  \(n=-1,0,1,\ldots\) for all states \(s \in \States\) with discount factor
-  \(\gamma\) % and \(V_{\gamma}^{\pol}(s)\) being the value function as defined above
-  %
-  if and only if
-  \begin{align*}
-    \lim_{\gamma \to 1}(1-\gamma)^{-n}\ (V_{\gamma}^{\polopt}(s) - V_{\gamma}^{\pol}(s)) \geqslant 0 \tpkt
-  \end{align*}
-\end{definition}
-
-Only if a policy is \(n\)-discount-optimal for all \(n < m\) it can be \(m\)-discount
-optimal \citep{Puterman94,Veinott69}.
-%
-If a policy is \(\infty\)-discount-optimal then it is said to be Blackwell-optimal
-\citep{Blackwell62}.
-% weakening the argument!!!
-That is, for Blackwell-optimal policies \(\polopt\) there exists a discount factor \(\gammaopt < 1\)
-such that \(V_{\gamma}^{\polopt}(s) \geqslant V_{\gamma}^{\pol}(s)\) for all
-\(\gamma \geqslant \gammaopt\) and under all policies \(\pol\)
-\citep{Mahadevan96_SensitiveDiscountOptimalityUnifyingDiscountedAndAverageRewardReinforcementLearning,Blackwell62}.
-Informally that means there exists a discount factor \(<1\) which finds Blackwell-optimal policies.
-However, (i) in more complex, i.e.\@ real world, MDPs this value can be arbitrary close to \(1\) and
-(ii) the difference in state values may be very small, which (iii) due to the need of state value
-function approximation likely causes errors when choosing among different actions in the discounted
-framework.
-
-For the case of \(n=-1\) this lead to gain-optimality
-\(\avgrew^{\polopt}(s) - \avgrew^{\pol}(s) \geqslant 0\) for all policies \(\pol\) and states \(s\)
-\citep{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}. In
-the case of \(n=0\) it describes bias-optimality with \(V^{\polopt}(s) - V^{\pol}(s) \geqslant 0\)
-for all policies \(\pol\) and states \(s\)
-\citep{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}.
-%
-
-In the sequel we present the Laurent series expansion which not only links discounted reinforcement
-learning with average reward reinforcement learning, but also connects to \(n\)-discount-optimality.
-
-% Furthermore, for the rest of the paper we occasionally write drop the \(V(s)\) and $V_{\gamma}$ instead of
-% \(V^{\pol}(s)\) and \(V^{\pol}_{\gamma}\) when the policy is clear from the context.
-
 \subsection{The Laurent Series Expansion of Discounted State Values}
@@
 \end{align*}
-
+%
 where the error term \(e_{\gamma}^{\pol}(s)\) exists of infinitely many subterms and
@@
 series expansion of \(V_{\gamma}^{\pol}(s)\). Each addend corresponds to one step in
-\(n\)-discount-optimality. That is for \((-1)\)-discount-optimality the average reward must be
-maximised.
-%
-Then for \(0\)-discount-optimality the agent has to choose the highest bias values.
-%
-Finally \(n\)-discount-optimality for \(n \geqslant 1\) requires to maximise the error term
-\(e_{\gamma}^{\pol}(s)\) which only exists as \(\gamma\) is strictly less than \(1\). Thus this term
-incorporates the number of expected steps and their corresponding reward on the path to the
-Blackwell-optimal policy. That is, it optimises the expected reward collected to reach the Blackwell
-optimal policy according to the occurrence on the paths, where shorter paths and those which collect
-the rewards sooner are preferred.
-
-Note how the Laurent series expansion dissipates these values in single terms and therefore defines
-a divide and conquer methodological approach to reinforcement learning. Thus in average reward
-reinforcement learning these terms are learned separately, while in the discounted framework the
-combined value for a fixed \(\gamma\)-value is learned.
-
-The addends, where \(n=-1,0,\ldots\) denote the coefficients of the Laurent series expansion,
-can be reformulated to following constraint problem with expected reward
-\(R^{\pol}(s) = \E_{\pol}[ r(s, a) ] = \sum_{a \in \mathcal{A}(s)}p(a | s) r(s, a)\) as shown by
-\cite{MillerVeinott1969} and \citet[p.346]{Puterman94}:
-
+\(n\)-discount-optimality. That is for \((-1)\)-discount-optimality the addend describes (a scaled
+version of) the term to maximise for gain-optimality,
+% the average reward must be
+% maximised.
+%
+for \(0\)-discount-optimality the term to maximise for bias-optimality, and
+%
+finally \(n\)-discount-optimality for \(n \geqslant 1\) requires to maximise the error term
+\(e_{\gamma}^{\pol}(s)\). The later only exists as \(\gamma\) is strictly less than \(1\). Thus this
+term incorporates the number of expected steps and their corresponding reward on the path to the
+Blackwell-optimal policy. Put differently, it optimises the expected reward collected to reach the
+Blackwell optimal policy according to the occurrence on the paths, where shorter paths and those
+which collect the rewards sooner are preferred.
+
+% Note how the Laurent series expansion dissipates these values in single terms and therefore defines
+% a divide and conquer methodological approach to reinforcement learning. Thus in average reward
+% reinforcement learning these terms are learned separately, while in the discounted framework a
+% single scalar value for each state with a fixed \(\gamma\)-value is estimated.
+
+These addends, where \(n=-1,0,\ldots\) denote the coefficients of the Laurent series expansion and
+thus correspond to the \(n\) of the definition of \(n\)-discount-optimality, can be reformulated to
+following constraint problem \citep[p.346]{MillerVeinott1969,Puterman94}.
+%
 \begin{align}
@@
 \end{align}
-
+%
+The error term \(e_{\gamma}^{\pol}(s)\) is given by
+\(e_{\gamma}^{\pol}(s) = \sum_{n=1}^{\infty}W^{\pol}_{n}(s)\) and the expected reward
+\(R^{\pol}(s) = \E_{\pol}[ r(s, a) ]\). % = \sum_{a \in \mathcal{A}(s)}p(a | s) r(s, a)\)
+%
 \citet[p.343ff]{Puterman94} shows that due to the given degree of freedom if \(n=-1,0,\ldots,M\)
@@
 
-\subsection{Algorithm}
-\label{subsec:Algorithm}
+\subsection{Average Reward Adjusted Reinforcement Learning}
+\label{subsec:Average_Reward_Adjusted_Reinforcement_Learning}
+
+In the sequel we briefly present the needed formalism of average reward adjusted RL, where for a
+more comprehensive version see \citep{schneckenreither2020average}. \MS[t]{todo}
+
+\begin{definition}
+  Due to \citet{schneckenreither2020average} the \textit{average reward adjusted (discounted) state
+    value} $\X_{\gamma}^{\pol}(s)$ of a state $s$ under policy $\pol$ and with discount factor
+  $0 \leqslant \gamma \leqslant 1$ is given as
+  \(\X_{\gamma}^{\pol}(s) \defsym V^{\pol}(s) + e_{\gamma}^{\pol}(s)\).
+\end{definition}
+%
+With the Laurent Series expensive this reformulates to
+\(\X_{\gamma}^{\pol}(s) = V_{\gamma}^{\pol}(s) - \frac{\avgrew^{\pol}}{1-\gamma}\). Therefore,
+\(X_{\gamma}^{\pol}(s)\) describes the discounted state value adjusted by the average reward term.
+The Bellman Equation is given by
+\(\X_{\gamma}^{\pol}(s) = \E_{\pol}[r_{t} + \gamma \X_{\gamma}^{\pol}(s_{t+1}) - \avgrew^{\pol}
+\mid s_{t} = s]\). By turning this formula into an update rule it builds the foundation of our
+algorithm.
+
+% \subsection{Algorithm}
+% \label{subsec:Algorithm}
 
@@
     \(0 < \alpha, \gamma < 1\), and discount factors
-    \(0 < \gamma_{0} < \gamma_{1} \leqslant 1\).
+    \(0 < \gamma_{0} < \gamma_{1} \leqslant 1\), where \(\gamma_{1} = 1\) is
+    usually a good choice.
     \While{the stopping criterion is not fulfilled}
@@
 
-The model-free average reward adjusted deep RL algorithm is based on the tabular
-version of \citet{schneckenreither2020average} and depicted in Algorithm~\ref{alg:near}.
+\paragraph{Average Reward Adjusted Deep RL Algorithm.} %
+The model-free average reward adjusted RL algorithm is based on the tabular version of
+\citet{schneckenreither2020average} and depicted in Algorithm~\ref{alg:near}.
 %

### 2020-12-27 paper/paper.tex
fixed pdfauthor

@@
 pdftitle={\thetitle{}},
-pdfauthor={\theauthor{}},
+pdfauthor={Manuel Schneckenreither, Stefan Haeussler},
 pdfsubject={\thetitle{}},

### 2020-12-28 paper/paper.tex
algorithm, near-Blackwell-optimality

@@
 %
-If a policy is \(\infty\)-discount-optimal then it is said to be Blackwell-optimal
+If a policy is \(-1\)-discount-optimal it is called gain-optimal, if it is \(0\)-discount-optimal it
+is also called bias-optimal.
+%
+Furthermore, if a policy is \(\infty\)-discount-optimal then it is said to be Blackwell-optimal
 \citep{Blackwell62}.
-% weakening the argument!!!
+%
 That is, for Blackwell-optimal policies \(\polopt\) there exists a discount factor \(\gammaopt < 1\)
@@
 state value function approximation likely causes errors when choosing among different actions in the
-discounted framework.
-
+discounted framework. \citet{schneckenreither2020average} establishes a practically relevant
+definition for near-Blackwell-optimality.
+
+\begin{definition}
+  If an algorithm infers for any MDP bias-optimal policies, and for a given MDP can in theory be
+  configured to infer Blackwell-optimal policies, but in practise this ability is naturally limited
+  due to the finite accuracy of floating-point representation of modern computer systems, it is said
+  to be \emph{near-Blackwell-optimal}. An according to a near-Blackwell-optimal algorithm inferred
+  Blackwell-optimal policy is called near-Blackwell-optimal.
+\end{definition}
+
+Note that standard discounted RL does not meet the requirements of near-Blackwell-optimality, as it
+generally does not infer bias-optimal policies. Only by chance the resulting policy could be
+bias-optimal.
+%
 % In the sequel we present the Laurent series expansion which not only links discounted reinforcement
 % learning with average reward reinforcement learning, but also connects to \(n\)-discount-optimality.
-
+%
 % Furthermore, for the rest of the paper we occasionally write drop the \(V(s)\) and $V_{\gamma}$ instead of
 % \(V^{\pol}(s)\) and \(V^{\pol}_{\gamma}\) when the policy is clear from the context.
-
-The Bellman Equation defined as
+%
+The corresponding Bellman equation defined as
 \(V_{\gamma}^{\pol}(s) = \E_{\pol}[r + \gamma V_{\gamma}^{\pol}(s')]\) \citep[see
@@
 \begin{enumerate}
-\item Standard discounted RL can only infer suboptimal policies as the discount factor is strictly
-  less than one, i.e.\@ \(\gamma < 1\).
+\item In general standard discounted RL can only infer suboptimal policies as the discount factor is
+  strictly less than one, i.e.\@ \(\gamma < 1\).
 \item It is very difficult to specify a desired balance between the short-term and long-term
-  (average reward) rewards. This results from the fact that the long-term rewards are scaled
-  exponentially, where the factor depends on the chosen \(\gamma\) value.
-\item Episodic MDPs with an average reward per step that is non-zero cannot be solved, as the
-  average reward is ignored in the terminal states.
+  (average) rewards. This results from the fact that the long-term rewards are scaled exponentially,
+  where the factor depends on the chosen \(\gamma\) value.
+\item Episodic MDPs with an average reward per step that is non-zero cannot be solved correctly, as
+  the average reward is ignored in the terminal states.
 \item The average reward is independently assessed for each state, even though it is the dominating
@@
 state value estimations \citep{MillerVeinott1969} finding a well working parameterisation is
-difficult. Average reward reinforcement learning overcomes these issues by separately assessing
-addends of the state values.
+difficult. Average reward reinforcement learning overcomes these issues by separately assessing the
+subterms of the state values directly.
 
@@
 
-In the sequel we briefly present the needed formalism of average reward adjusted RL, where for a
-more comprehensive version see \citep{schneckenreither2020average}. \MS[t]{todo}
+Thus, average reward adjusted RL can be seen as a specialised version of average reward RL, where it
+targets unichain MDPs only. In the sequel we briefly present the needed formalism of average reward
+adjusted RL. For more comprehensive version we refer to \citet{schneckenreither2020average}.
 
 \begin{definition}
-  Due to \citet{schneckenreither2020average} the \textit{average reward adjusted (discounted) state
-    value} $\X_{\gamma}^{\pol}(s)$ of a state $s$ under policy $\pol$ and with discount factor
-  $0 \leqslant \gamma \leqslant 1$ is given as
+  The \textit{average reward adjusted (discounted) state value} $\X_{\gamma}^{\pol}(s)$ of a state
+  $s$ under policy $\pol$ and with discount factor $0 \leqslant \gamma \leqslant 1$ is given as
   \(\X_{\gamma}^{\pol}(s) \defsym V^{\pol}(s) + e_{\gamma}^{\pol}(s)\).
@@
 %
-With the Laurent Series expensive this reformulates to
+With the Laurent Series expansion this reformulates to
 \(\X_{\gamma}^{\pol}(s) = V_{\gamma}^{\pol}(s) - \frac{\avgrew^{\pol}}{1-\gamma}\). Therefore,
@@
 The Bellman Equation is given by
-\(\X_{\gamma}^{\pol}(s) = \E_{\pol}[r_{t} + \gamma \X_{\gamma}^{\pol}(s_{t+1}) - \avgrew^{\pol}
-\mid s_{t} = s]\). By turning this formula into an update rule it builds the foundation of our
-algorithm.
-
-% \subsection{Algorithm}
-% \label{subsec:Algorithm}
+\(\X_{\gamma}^{\pol}(s) = \E_{\pol}[r_{t} + \gamma \X_{\gamma}^{\pol}(s_{t+1}) - \avgrew^{\pol} \mid
+s_{t} = s]\). The corresponding Bellman optimality equation then is given by
+\(\X_{\gamma}^{\polopt}(s) = \max_{a} \E_{\polopt}[r_{t} + \gamma \X_{\gamma}^{\polopt}(s_{t+1}) -
+\avgrew^{\pol} \mid s_{t} = s]\). By turning this formula into an update rule it builds the
+foundation of our algorithm.
 
@@
     Carry out action \(a_{t}\), observe reward \(r_{t}\) and resulting state \(s_{t+1}\).
-    Store the experience \((s_{t}, \isRandAct{t}, a_{t}, r_{t}, s_{t+1})\) in the experience replay
+    Store the experience \((s_{t}, a_{t}, \isRandAct{t}, r_{t}, s_{t+1})\) in the experience replay
     memory \(M\).
@@
 The model-free average reward adjusted RL algorithm is based on the tabular version of
-\citet{schneckenreither2020average} and depicted in Algorithm~\ref{alg:near}.
+\citet{schneckenreither2020average} and depicted in Algorithm~\ref{alg:near}. The algorithm operates
+on a target network \(\netTarget{}\) and a worker network \(\netWorker{}\) as in
+\citet{mnih2015human}. % The target network reduces fluctuations in the state value estimates and
+% therefore reduces the
+%
+In the action selection process a \(\epsilon\)-sensitive lexicographic order
+\(a=(a_{1},\ldots,a_{n}) \lex (b_{1},\ldots,b_{n})=b\) is used. It is defined as \(a \lex b\) if and
+only if \(| a_{j} - b_{j} | \leqslant \epsilon\) for all \(j < i\) and
+\(|a_{i} - b_{i}| > \epsilon\). Note that the resulting sets of actions may not be disjoint, but
+taking the maximum as required in our algorithm is straight-forward and thus cheap to compute. The
+first action selection criteria selects the actions which maximise the bias values, the second the
+error term.
+%
+Equation~\ref{eq:constr2} of the average reward RL constraints is used to estimate the average
+reward based on the \MS[t]{reformulate avg RL constraints using s' etc.?} currently inferred state
+values of two consecutive states and the returned reward. % As the average
+% reward is equal among all states estimating the average reward is based on many observations.
+This infers a solid average reward prediction of the current policy in few steps only and thus
+quickly leads to better policies, as ones with higher gain are preferred.
+\citet{schneckenreither2020average} and
+\citet{Tadepalli98_ModelbasedAverageRewardReinforcementLearning} find that updating the average
+reward by Equation~\ref{eq:constr2} is superior as compared to exponentially smoothing the actual
+observed rewards. This makes sense as the later rather evaluates past policies.
+
+The average reward adjusted state value, which is approximated based on the Bellman optimality
+equation, is estimated twice. Once to infer the bias values with a very high discount factor
+\(\gamma_{1}\), e.g.\@ \(\gamma_{1} = 1.0\), and a second time with a discount factor
+\(\gamma_{0} < \gamma_{1}\). The later is used to more selectively choose actions in case more than
+one action leads to bias-optimal policies. The predetermined discount value \(\gamma_{0}\) is a
+measure for farsightedness. Lower values prefer to partially collect reward in fewer steps and
+higher values are used to prefer actions for which the full bias is collected earlier.
+%
+The corresponding update values are saved to calculate the gradients based on the quadratic errors.
+Every \(C\) steps the target network exponentially adapts the parameters of the worker network.
+%
 %
@@
 episodic tasks needs to store a bool indicating an episode end. Uses Int8 for state features values.  }
-
-%
-%
-After initialising all values the agent enters the loop in which the first task is to choose an
-action (step 3). In this action selection process we utilise an \(\epsilon\)-sensitive lexicographic
-order \(a=(a_{1},\ldots,a_{n}) \lex (b_{1},\ldots,b_{n})=b\) defined as \(a \lex b\) if and only if
-\(| a_{j} - b_{j} | \leqslant \epsilon\) for all \(j < i\) and \(|a_{i} - b_{i}| > \epsilon\). Note
-that the resulting sets of actions may not be disjoint. Although this is an unusual order in
-programming, taking the maximum as in our algorithm is straight-forward and thus cheap to compute.
-%
-The first action selection criteria maximises the average reward. In case of unichain and thus
-ergodic MDPs only one scalar average reward value over all states is learned and thus the problem
-simplified by abolishing the comparison in the first component. Then the set of actions which
-maximise the bias values are selected. Finally, \(\Delta V^\pol_\gamma(s,a)\) is used to estimate
-the slope of the error term \(e_{\gamma}^{\pol}(s,a)\), which can be either positive or negative.
-Therefore, there is the need of a distinction of which actions to maximise according to the
-discounted state value difference
-\(\Delta V^{\pol}_{\gamma}(s,a) = V^\pol_{\gamma_1}(s,a) - V^\pol_{\gamma_0}(s,a)\). As the aim is
-to collect rewards as soon as possible, a Blackwell-optimal reinforcement learning agents maximises
-the error term, thus any actions converging to
-\(\frac{\avgrew^{\pol}(s,a)}{1-\gamma_{0}} + \V^{\pol}(s,a)\) from above are preferred over those
-that converge from below. Within these sets the action which maximises
-\(\Delta V^{\pol}_{\gamma}(s,a)\) is selected. If multiple such actions the agent can choose
-randomly among these.
-
-As usual in reinforcement learning the equations are integrated by exponentially smoothing the
-values utilising parameters \(\alpha, \beta, \gamma, \delta\).
-% Further, as in this paper we concentrate on unichain MDPs the average reward
-% \(\avgrew^{\pol}(s,a)\) can be learned as a scalar value \(\avgrew^{\pol}\), whereas the algorithm
-% is supports multichain MDPs, cf. step \(5a\).
-Step \(5a\) uses the Bellman equation for average reward reinforcement learning
-(cf.~\citealt{Howard64, tadepalli1998model}) to find the average reward. Then step \(5b\), \(5c\)
-and \(5d\) depict Equations (\ref{eq:constr2}) and (\ref{eq:constr3}) for \(n=0\), \(n=1\) and
-\(n=2\). The need of learning values for \(n=2\) stems from the fact that the actual values
-corresponding to the highest \(n\) are offset by vector \(u\).
-
-% For \(n=0\) (step \(5b\)) the bias values are learned and further to ensure a unique solution of
-% \(V^{\pol}\) step \(5c\) solves the constraints for \(n=1\). Thus, the first three assignments of
-% step \(5\) correspond to the above constraint system for \(n=-1,0,1\). The last line of step \(5\)
-% evaluates the deviation of \(W^{\pol}(s,a)\), which ought to be \(0\) once the system is correctly
-% solved and if an action of the optimal policy \(\polopt\) was taken.
-Step \(6\) infers the discounted state values using learning rate \(\gamma\) and two discount rates
-\(0.5 \leqslant \gamma_{0} < \gamma_{1} < 1\) to estimate the state value difference
-\(\Delta V^{\pol}_{\gamma}(s,a)\).
-
-Steps \(7\) and \(8\) are crucial, but were not considered in average reward reinforcement learning
-before. They feed back the deviations \(\Psi_{\cdot}(s,a)\) into the constraint problem creating a
-recursive system of equations. Depending on the underlying MDP when this constraint is omitted one
-likely observes steadily increasing deviations and therefore a wrong solution. This feedback loop
-however is only activated in case a non-random action was chosen, as otherwise the deviations
-\(\Psi_{\cdot}(s,a) > 0\) \citep{MillerVeinott1969}.
-
-%
-%
-% Steps \(5d\) and \(5e\) calculate the deviations of actual values and expected values. In case the
-% agent performs a non-random action
-%
-Recall that the algorithm evaluates the slope of \(e_{\gamma}^{\pol}(s)\) by learning the state
-values using two different \(\gamma\) settings and discounted reinforcement learning. The slope of
-\(e_{\gamma}^{\pol}(s)\) combined with the observation that for Blackwell-optimal policies
-\(\polopt\) there exists a discount factor \(\gammaopt < 1\) such that
-\(V_{\gamma}^{\polopt}(s) \geqslant V_{\gamma}^{\pol}(s)\) for all \(\gamma \geqslant \gammaopt\)
-and under all policies \(\pol\)
-\citep{Mahadevan96_SensitiveDiscountOptimalityUnifyingDiscountedAndAverageRewardReinforcementLearning,Blackwell62}
-provides the most selective criteria for our algorithm, which theoretically can find
-Blackwell-optimal policies\footnote{This claim is only not fully proven yet. See the appendix for a
-  first version of the proof. } (see Section~\ref{sec:methods} in Appendix).
 
@@
 
+\MS[t]{Algorithm adaptions}
+
 \paragraph{Markov Decision Process.} The underlying MDP is unichain and looks as follows.

### 2020-12-28 paper/figures/productionsystem.tex
simulation setup. todo: algorithm setup

\begin{tikzpicture}[queue/.style={regular polygon,regular polygon sides=4,inner sep=0.075cm},scale=0.89]

  \draw (-3.5,0) node[] {$\ldots$};
  \draw (-4.1,-1.75) node[] {\scriptsize Due Date:};

  \draw (-3.0,-1.5) -- (-3.0,1.5)
        (-2.5,1.5)  -- (-2.5,-1.5)
        (-2.5,-1.5) -- (-3.0,-1.5);
  \draw (-2.75,-1.75) node[] {\tiny $t+3$};

  \draw (-2.5,-1.5) -- (-2.5,1.5)
        (-2,1.5)  -- (-2,-1.5)
        (-2,-1.5) -- (-2.5,-1.5);
  \draw (-2.25,-2) node[] {\tiny $t+2$};
  \draw (-2 ,-1.5) -- (-2,1.5)
        (-1.5,1.5)  -- (-1.5,-1.5)
        (-1.5,-1.5) -- (-2,-1.5);
  \draw (-1.75,-1.75) node[] {\tiny $t+1$};

  \draw (-1.5,0.0) node[] (OP) {};
  \draw [decorate,decoration={brace,amplitude=5pt},xshift=-2pt,yshift=0pt] (-0.75,-2.25) -- (-3.75,-2.25)     node [black,midway,yshift=-0.4cm] {Order Pool};
  \draw [decorate,decoration={brace,amplitude=5pt},xshift=-2pt,yshift=0pt] (7,-2.25) -- (-.5,-2.25)  node [black,midway,yshift=-0.4cm] {Production System};
  \draw [decorate,decoration={brace,amplitude=5pt},xshift=-2pt,yshift=0pt] (9,-2.25) -- (7.25,-2.25) node [black,midway,yshift=-0.4cm] {Inventory};

  % \draw (-1.75,-1.3) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 1$};
  % \draw (-1.75,-1.0) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 1$};
  % \draw (-1.75,-0.7) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 2$};
  % \draw (-1.75,-0.4) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 2$};
  % \draw (-1.75,-0.1) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 3$};
  % \draw (-1.75, 0.2) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 3$};
  % \draw (-1.75, 0.5) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 4$};
  % \draw (-1.75, 0.8) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 4$};
  % \draw (-1.75, 1.1) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 6$};
  % \draw (-1.75, 1.4) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 6$};
  % \draw (-1.75, 1.7) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 6$};

  \draw (-2.25,-1.3) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 1$};
  \draw (-2.25,-0.95) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 3$};
  \draw (-2.25,-0.6) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 3$};
  \draw (-2.25,-0.25) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 4$};

  \draw (-2.75,-1.3) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 1$};
  \draw (-2.75,-0.95) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 1$};
  \draw (-2.75,-0.6) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 3$};
  \draw (-2.75,-0.25) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 4$};
  \draw (-2.75, 0.1) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 4$};
  \draw (-2.75, 0.45) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 4$};
  \draw (-2.75, 0.8) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 5$};
  \draw (-2.75, 1.15) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 5$};
  \draw (-2.75, 1.5) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 6$};
  \draw (-2.75, 1.85) node[circle,inner sep=1pt,minimum size=0.2cm,draw] {$\scriptscriptstyle 6$};

  \draw (0.5,0)    node[minimum size=1cm,draw,align=center] (M1) {M1\\[-1ex]{\tiny $\Exp(80) $ or} \\[-1ex]{\tiny $\Unif(30,130)$}};
  \draw (3,0.75)   node[minimum size=1cm,draw,align=center] (M2) {M2\\[-1ex]{\tiny $\Exp(160)$ or}\\[-1ex]{\tiny $\Unif(80,240)$}};
  \draw (3,-0.75)  node[minimum size=1cm,draw,align=center] (M3) {M3\\[-1ex]{\tiny $\Exp(155)$ or}\\[-1ex]{\tiny $\Unif(50,260)$}};
  \draw (6,1.5)    node[minimum size=1cm,draw,align=center] (M4) {M4\\[-1ex]{\tiny $\Exp(210)$ or}\\[-1ex]{\tiny $\Unif(50,370)$}};
  \draw (6,0)      node[minimum size=1cm,draw,align=center] (M5) {M5\\[-1ex]{\tiny $\Exp(285)$ or}\\[-1ex]{\tiny $\Unif(200,370)$}};
  \draw (6,-1.5)   node[minimum size=1cm,draw,align=center] (M6) {M6\\[-1ex]{\tiny $\Exp(215)$ or}\\[-1ex]{\tiny $\Unif(110,320)$}};
  \draw (8,0)      node[minimum size=1cm,draw,align=center,fill] (\fgi{}) {\color{white} \fgi{}};

  \draw (M1.west) node[draw,queue,minimum size=0.05cm,fill,xshift=-0.075cm] (Q1) {};
  \draw (M2.west) node[draw,queue,minimum size=0.05cm,fill,xshift=-0.075cm] (Q2) {};
  \draw (M3.west) node[draw,queue,minimum size=0.05cm,fill,xshift=-0.075cm] (Q3) {};
  \draw (M4.west) node[draw,queue,minimum size=0.05cm,fill,xshift=-0.075cm] (Q4) {};
  \draw (M5.west) node[draw,queue,minimum size=0.05cm,fill,xshift=-0.075cm] (Q5) {};
  \draw (M6.west) node[draw,queue,minimum size=0.01cm,fill,xshift=-0.075cm] (Q6) {};

  \draw (10.5, 1) node[] (CustomerTop) {\Huge \Gentsroom};
  \draw (10, 0.5) node[] {\Huge \Ladiesroom};
  \draw (10.5, 0) node[] (Customer) {\Huge \Gentsroom};
  \draw (10,-0.5) node[] {\Huge \Ladiesroom};
  \draw (10.5,-1) node[] {\Huge \Gentsroom};

  % Paths
  \path (OP.center) edge[->, bend left=10, above]
  node[xshift=0.088cm,yshift=0.05cm,align=center] {\scriptsize Order \\[-1ex] \scriptsize Release} (Q1);

  \path (M1) edge[->, bend left=10, above]                 node[xshift=-0.1cm,yshift=-0.05cm] {$\scriptstyle 1,2,3$} (Q2);
  \path (M1) edge[->, bend right=10,below]                 node[xshift=-0.1cm,yshift=-0.05cm] {$\scriptstyle 4,5,6$} (Q3);
  \path (M2) edge[->, bend left=10,above,near start ]      node {$\scriptstyle 1$}     (Q4);
  \path (M2) edge[->, bend left=10,above,near start ]      node {$\scriptstyle 2$}     (Q5);
  \path (M2) edge[->, bend left=10,above,near start ]      node[xshift=0.07cm,yshift=-0.12cm] {$\scriptstyle 3$}     (Q6);
  \path (M3) edge[->, bend right=10,above,very near start] node[xshift=0.22cm,yshift=-0.09cm] {$\scriptstyle 4$}     (Q4);
  \path (M3) edge[->, bend right=10,above,very near start]      node[yshift=-0.075cm] {$\scriptstyle 5$}     (Q5);
  \path (M3) edge[->, bend right=10,above,near start]      node {$\scriptstyle 6$}     (Q6);
  \path (M4) edge[->, bend left=10,above,near start]       node[xshift=0.1cm] {$\scriptstyle 1,4$} (\fgi{});
  \path (M5) edge[->, bend left=10,above,near start]       node[xshift=0.1cm, yshift=-0.07cm] {$\scriptstyle 2,5$} (\fgi{});
  \path (M6) edge[->, bend right=10,below,near start]      node[xshift=0.15cm,yshift=0.07cm] {$\scriptstyle 3,6$} (\fgi{});

  \path (\fgi{}) edge[->,near start] node[xshift=0.1cm,align=left] {\tiny Not \\[-1ex]
    \tiny before \\[-0.5ex] \tiny due \\[-1ex] \tiny date}    (Customer);

  \path (CustomerTop) edge[densely dashed] (10.5,2.8);
  \path (10.5,2.8) edge[densely dashed,midway,align=center] node[yshift=0.0cm] {\scriptsize Customer places
    an order which initiates the process\\ {\tiny Interarrival-time distributions:
      $\Exp(135)$ or $\Unif(95,175)$ or $\Exp(118)$ or $\Unif(78,158)$ % or $\Exp(105)$ or $\Unif(65, 145)$
    }
  } (-4.5, 2.8);
  \path (-4.5,2.8) edge[densely dashed] (-4.5,0);
  \path (-4.5,0) edge[->,densely dashed] (-4.0,0);

\end{tikzpicture}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../paper"
%%% End:

### 2020-12-28 paper/paper.tex
simulation setup. todo: algorithm setup

@@
 
+\usepackage{lineno}
+
 \usepackage{xcolor}
@@
 
-\subsection{Proof-of-Concept}
-\label{subsec:Proof-of-Concept}
-
-Reconsider the task depicted in Figure~\ref{fig:three-states}. The Blackwell-optimal MDP
-\(\pol_{A}\), that is the one that chooses the A-loop, is shown on the left side of
-Figure~\ref{fig:three-states2}. Recall that the bias values for \(\pol_{A}\) are
-\(V^{\pol_{A}}(0)=-0.5\), \(V^{\pol_{A}}(1)=0.5\) and \(V^{\pol_{A}}(2)=1.5\). Our model-based
-implementation infers the values \(\avgrew^{\pol_{A}} = 1\), \(V^{\pol_{A}}(0)=-0.5\), \(V^{\pol_{A}}(1)=0.5\)
-and \(V^{\pol_{A}}(2)=1.4391\) fully automatically in \(200k\) steps. Furthermore, the addition of
-\(V^{\pol_{A}}(2) + \Psi^{\pol_{A}}_{V}(2) = 1.4998\), where \(\Psi^{\pol_{A}}_{V}(2)\) is the exponentially
-smoothed deviation for state \(2\). Thus the model-based algorithm is able to correctly approximate
-the mathematical solution.
-
-The right side of Figure~\ref{fig:three-states2} shows the same MDP for the model-free version. In
-model-free reinforcement learning state-action pairs as opposed to state values are estimated. Thus
-each state is replicated as often, as actions are available in that state. Therefore, state \(1\) is
-replaced by states-action pairs \((1,l)\) and \((1,r)\) for moving to the left (taking the A-loop)
-and to the right respectively. The dashed line indicates that these states are connected and thus
-the agent has to choose among them when reaching either. In this case the inferred bias values are
-\(V^{\pol_{A}}(0,r) = -0.5\), \(V^{\pol_{A}}(1,l)=V^{\pol_{A}}(1,r)=0.5\), and \(V^{\pol_{A}}(2,l)=1.5\). The above given tabular model-free
-average reward reinforcement learning algorithm is able to exactly infer these values in less than
-100k steps. This provides evidence that the algorithm in general infers correct solutions in regard to
-the underlying constraints.
-
-% \MS{Occasionally we write \(V^{\pol_{A}}(s)\) and $V_{\gamma}$ instead of \(V^{\pol}(s)\) and
-% \(V^{\pol}_{\gamma}\) when the policy is clear from the context.}
-
-Furthermore, in the later case \(V^{\pol_{A}}(1,l)=V^{\pol_{A}}(1,r)=0.5\), that is if the agent
-resides in state \(1\) it has to choose among the both available actions. This cannot be done on
-basis of the average reward, nor on the bias values. Therefore, it infers the slopes of the error
-terms using the difference of the discounted state values \(V^{\pol_{A}}_{\gamma_{1}}(1,l) = 5.56\)
-and \(V^{\pol_{A}}_{\gamma_{0}}(1,l) = 2.67\) for state-action pair \((1,l)\) and
-\(V^{\pol_{A}}_{\gamma_{1}}(1,r) = 5.16\) and \(V^{\pol_{A}}_{\gamma_{0}}(1,r) = 1.67\) for
-\((1,r)\), yielding \(\Delta V^{\pol_{A}}_\gamma(1,l) = -2.89\) and
-\(\Delta V^{\pol_{A}}_\gamma(1,r) = -3.49\), where \(\gamma_{0} = 0.5\) and \(\gamma_{1}=0.8\).
-
-The idea is that rewards are collected as soon as possible, thus the error term has to be maximised
-when \(\gamma \to 1\). As we have
-\(2.5 = \frac{\avgrew^{\pol_{A}}}{1-\gamma_{0}} + V^{\pol_{A}}(1,a) < V^{\pol_{A}}_{\gamma_{0}}(1,l)
-= 2.67\), while \(2.5 \geqslant 1.67\), we choose action \(l\) as it is the only action with
-positive error term.
-
-% Current state:                               A
-% Period:                                      5000
-% Alpha:                                       0.0050
-% Beta:                                        0.0700
-% Delta:                                       0.0700
-% Gamma:                                       0.0700
-% Epsilon:                                     1.2913
-% Exploration:                                 0.8609
-% Algorithm:                                   BORL with gammas  (0.5,0.8); state values for rho; Deciding on V
-% Zeta (for forcing V instead of W):           1.0000
-% Xi (ratio of W error forcing to V):          0.2000
-% Scaling (V,W,R0,R1) by V Config:             Tabular representation (no scaling needed)
-% Psi Rho/Psi V/Psi W:                         (0.0000,0.0000,0.0000)
-% Rho:                                         1.0000
-% V                                       W
-% [+0.0000]right: -0.5000                 [+0.0000]right: 0.2061
-% [+1.0000]left : 0.5000                  [+1.0000]left : -0.2939
-% [+1.0000]right: 0.5000                  [+1.0000]right: -2.2939
-% [+2.0000]left : 1.5000                  [+2.0000]left : -1.7939
-% R0                                      R1
-% [+0.0000]right: 1.3333                  [+0.0000]right: 4.4444
-% [+1.0000]left : 2.6667                  [+1.0000]left : 5.5556
-% [+1.0000]right: 1.6667                  [+1.0000]right: 5.1556
-% [+2.0000]left : 3.3333                  [+2.0000]left : 6.4444
-% PsiV
-% [+0.0000]right: -0.0000
-% [+1.0000]left : -0.0000
-% [+1.0000]right: -0.0000
-% [+2.0000]left : -0.0000
-
-\begin{figure}[t!]
-  \centering
-  \begin{tikzpicture}[thin, scale=0.75]
-    % Nodes
-    \draw (-3,0) node(0) [circle,draw,minimum size=25] {\footnotesize \(0\)};
-    \draw (0,0)  node(1) [circle,draw,minimum size=25] {\footnotesize \(1\)};
-    \draw (3,0)  node(2) [circle,draw,minimum size=25] {\footnotesize \(2\)};
-
-    % Edges
-    \path[thin, ->, bend right, >=stealth] (1) edge[above] node {\footnotesize\(2\)} (0);
-    \path[thin, ->, bend right, >=stealth] (2) edge[above] node {\footnotesize\(2\)} (1);
-    \path[thin, ->, bend right, >=stealth] (0) edge[below] node {\footnotesize\(0\)} (1);
-    % \path[thin, ->, bend right, >=stealth] (1) edge[below] node {\footnotesize\(0\)} (2);
-
-    \draw (-1.5,0) node[] { A };
-    \draw (1.5,0) node[] { B };
-
-    \draw (6,0) node(B0) [circle,draw,minimum size=25] {\footnotesize \((0,r)\)};
-    \draw (9,1.0)  node(B1l) [circle,draw,minimum size=25] {\footnotesize \((1,l)\)};
-    \draw (9,-1.0)  node(B1r) [circle,draw,minimum size=25] {\footnotesize \((1,r)\)};
-    \draw (12,0)  node(B2) [circle,draw,minimum size=25] {\footnotesize \((2,l)\)};
-
-    \path[thin, ->, bend right, >=stealth] (B1l) edge[above] node {\footnotesize\(2\)} (B0);
-    \path[thin, ->, bend right, >=stealth] (B2) edge[above] node {\footnotesize\(2\)} (B1l);
-    \path[thin, ->, bend right, >=stealth] (B0) edge[below] node {\footnotesize\(0\)} (B1l);
-    \path[thin, ->, bend right, >=stealth] (B1r) edge[below] node {\footnotesize\(0\)} (B2);
-
-    \path[thin, dashed] (B1r.west) edge[below] node {} (B1l.west);
-    \path[thin, dashed] (B1r.east) edge[below] node {} (B1l.east);
-
-  \end{tikzpicture}
-  \caption{\label{fig:three-states2} The left side depicts the Blackwell-optimal MDP of
-    Figure~\ref{fig:three-states}, that is the MDP that takes the A loop. On the right side is the
-    model-free version of the same MDP with policy \(\pol_{A}\). As in the model-free scenario
-    state-action pairs are evaluated state \(1\) is split into a state that takes the left loop
-    and one that traverses to the right. They are linked with dashed lines to indicate that these
-    states are connected and the agent has to decide for one among them. The bias values for the
-    model-free version are \(V^{\pol_{A}}(0,r) = -0.5\),
-    \(V^{\pol_{A}}(1,l)=V^{\pol_{A}}(1,r)=0.5\), and \(V^{\pol_{A}}(2,l)=1.5\).}
-\end{figure}
-
 \section{Experimental Evaluation}
@@
 
-This section introduces the simulation model and the parameterisation of the algorithm. To be able
-to compare the performance of the algorithm we chose to implement the same setup as has been used by
-\cite{SchneckenreitherHaeussler2019}.
+This section introduces the simulation model and the parameterisation of the algorithm. To provide a
+proof-of-concept for the proposed algorithm we adapt the flow shop presented in
+\citet{schneckenreither2020order}. However, our algorithms operates on discrete lead times on a
+product type basis and thus is not directly comparable to their order based lead time management
+approach. Nonetheless, this section is largely based on the corresponding section of
+\citet{schneckenreither2020order}.
 
@@
 
+We adapt the simulation model of a make-to-order (MTO) flow shop with six different
+products and six work-centers of \citet{schneckenreither2020order} and examine the system under
+moderate and high utilisation in combination with moderate or high variability in demand and
+processing times. This section provides the details of the simulation model and the experimental
+setup.
+
+% \subsection{Make-To-Order Flow Shop}
+% \label{subsec:make-to-order_flow_shop}
+% \subsection{Simulation model}
+% \label{subsec:Simulation_model}
+
+The simulation model is depicted in Figure~\ref{fig:ps}, where the customers initiate the process by
+placing orders as indicated by the dashed line.
+%
 \begin{figure*}[t!]
   \centering
-  \input{productionsystem}                      % Figure production system
-  \caption{Production system of the simulation model with routing, processing time distributions and
-    demand.} \label{fig:ps}
+  \input{figures/productionsystem}                      % Figure production system
+  \caption{Production System of the Simulation Model with routing, processing time distributions and
+    demand interarrival time distributions.} \label{fig:ps}
 \end{figure*}
-
-Therefore, we use the hypothetical flow shop make-to-order manufacturing system similar to the
-system analyzed by \cite{Lee1997} and later by \cite{Schneckenreither2019}. The simulation model
-consists of three work centres; each consists of a single machine and can process only one order at
-a time. Figure~\ref{fig:ps} depicts the make-to-order production system setup.
-%
-The number of orders arriving is uniformly distributed between 3 and 15 order per period and
-thus with a mean of 9 order per period.
-%
-% Orders arrive
-% uniformly between 3 and 15 orders per period and thus a mean of 9 orders per period, or
-In other words, on average every $0.11$ periods one order arrives in the order pool.
-%
-Incoming orders are queued at the order pool until released. The due date slack ($\DDS$), that is
-the periods until incoming orders are due, is set to 7 periods.
-%
-Once orders are released they are queued at each work center and wait until being processed by the
-machine. The queuing priority is first-come-first-serve. No preemptions are allowed. Order
-routings are deterministic and embrace two production stages with diverging shape and no return
-visits. This results in 2 different products. The operation times of the work centres are
-uniformly distributed, cf. Table~\ref{tbl:costs}. These characteristics lead to a utilization rate
-of 90\% for the bottleneck work center (WC3) in steady state. % Planning periods were set to 960
-% minutes (16 hours).
-
-% \begin{figure}[b!]
-%   \centering
-%   \def\arraystretch{1.2}          %  1 is the default, change whatever you need
-%   \begin{table}[t!]
-%     \centering
-%     \begin{tabular}{|l||c|c|c|}
-        %         \hline
-        %     %         BEGIN RECEIVE ORGTBL optime
-        %         Work Center & WC1 & WC2 & WC3\\
-        %         \hline
-        %         Operation time & \(\mathcal{U}(70,130)\) & \(\mathcal{U}(130,170)\) & \(\mathcal{U}(180,200)\)\\
-        %     %         END RECEIVE ORGTBL optime
-        %         \hline
-        %       \end{tabular}
-        %   %       \egroup
-
-        %         \caption{\label{tab:OTs}Operation times.}
-        %   %       \end{figure}
-        %         \end{table}
-\begin{table}[t]
-  \centering
-  \begin{tabular}{|l||c|c|c|}
-    % BEGIN RECEIVE ORGTBL values
-    \hline
-    Work Center & WC1 & WC2 & WC3\\
-    \hline
-    Operation time & \(\mathcal{U}(70,130)\) & \(\mathcal{U}(130,170)\) & \(\mathcal{U}(180,200)\)\\
-    % END RECEIVE ORGTBL optime
-    \hline
-    \multicolumn{4}{c}{}\\
-    \hline
-    Cost & Wc p. Order/Period & Fc p. Order/Period & Bc p. Order/Period \\
-    \hline
-    Value & 3 & 10 & 20 \\
-    \hline
-    % END RECEIVE ORGTBL values
-  \end{tabular}
-
-  \caption{\label{tbl:costs}Operation times (upper half) and costs setup (in monetary unit).}
-\end{table}
-
-        %         \begin{figure}[b!]
-        %         \centering
-        %         \def\arraystretch{1.2}          %  1 is the default, change whatever you need
-        %         \setlength\tabcolsep{1.7pt}
-        %         \begin{tabular}{|l||c|c|c|c|c|c|}
-
-        %       \end{tabular}
-        %         \egroup
-        %         \caption{Operation times.}
-        %         \label{tab:OTs}
-        %         \end{figure}
-
-To evaluate the performance of the different order release models we define
-following performance measures similar to literature
-(e.g.,~\citealt{Baykasoglu2011,Thuerer2012}):
-\begin{itemize}
-\item \textbf{Cost related measures:} average total holding costs for  WIP and finished goods inventory,
-  costs for backorders. There are no earnings,
-  thus the algorithms minimizes the costs. The setup for the actual values of the costs is given in
-  Table~\ref{tbl:costs} and shows that late deliveries are especially expensive. All costs are per
-  order and period, and are measured and reported at the end of each period.
-
-  % \begin{table}[t]
-  %   \centering
-  %   \begin{tabular}{|l||c|c|c|}
-  %     %     BEGIN RECEIVE ORGTBL values
-        %         \hline
-        %         Work Center & WC1 & WC2 & WC3\\
-        %         \hline
-        %         Operation time & \(\mathcal{U}(70,130)\) & \(\mathcal{U}(130,170)\) & \(\mathcal{U}(180,200)\)\\
-        %     %         END RECEIVE ORGTBL optime
-        %         \hline
-        %         \multicolumn{4}{c}{}\\
-        %         \hline
-        %         Cost & Wc p. Order/Period & Fc p. Order/Period & Bc p. Order/Period \\
-        %         \hline
-        %         Value & 3 & 10 & 20 \\
-        %         \hline
-        %     %         END RECEIVE ORGTBL values
-        %       \end{tabular}
-
-        %         \caption{\label{tbl:costs}Operation times and costs setup (in monetary unit).}
-        %         \end{table}
-
-        %         \SH{The costs were set as follows: $Wc:Fc:Bc = 1:1.5:2.5$?}
-\item \textbf{Delivery related measures:} % mean service level -- percentage of
-        %         orders on time (\%SL),
-  mean tardiness of late orders (TA), standard deviation of
-  lateness ($\sigma$TA).
-\item \textbf{Flow time related measures:} mean shop floor throughput time (time
-  duration from release until entry of finished goods inventory; SFTT).
-\end{itemize}
-
-The length of each simulation run to evaluate the performance was 6000 periods including a warm-up
-period of 1000 periods. Welch’s procedure was applied to approximate the length of the warm-up
-period (see~\citealt{law:simulationc}).
-
-        %         Each order release method was evaluated on 25 predefined demand
-        %         streams.
-
-        %         \MS{TODO}
+%
+%
+The manufacturer produces six different product types, each of which is identified by a natural
+number ${1,2,\ldots,6}$ and has a different routing through the production system. The routing setup
+is provided by corresponding labels of the edges. For instance, product types $1-3$ are routed from
+machine $M1$ to machine $M2$, while the other products are transferred to machine $M3$. The
+production system consists of three production stages with diverging material flow and no return
+visits.
+%
+The incoming orders are uniformly distributed among the product types.
+%
+A period lasts $960$ minutes ($=16$ hours) which represent two \(8\)-hour shifts. At the
+beginning of each period orders can be released into the production system assuming material
+availability for all orders at their release date. Upon release, the orders are
+placed in the buffer at machine $M1$. Buffers and inventories are plotted as filled squares. Thus,
+each workstation is equipped with a buffer which queues the orders until processing is started. All
+buffers use a first-come-first-serve dispatching rule.
+%
+Furthermore, each workstation processes only one order at a time and early deliveries are
+prohibited. Orders completed ahead of their due date remain in the finished goods inventory (\fgi{})
+until they are due and sent to the customer.
+
+\paragraph{Processing Times.} The machine processing time distributions are given under the
+corresponding node labels of the machines. To simulate high or moderate variance the processing
+times for all machines are  drawn from an exponential or an uniform distribution,
+respectively. There is one bottleneck machine ($M5$) and therefore we refer to product
+types $2$ and $5$, those routed through machine $M5$, as
+bottleneck products whereas products $1,3,4,6$ are non-bottleneck products.
+
+\paragraph{Demand.} The incoming orders are placed in the order pool with a due date slack of
+\dds{} periods, that is each order arriving within period $t$ will be due at the end of period
+$t+\dds{}$. The due date slack of \(\dds{}\) was chosen to (i) ensure sufficient time between the
+first occurrence of orders in the order pool and the latest possible release of orders and (ii) to
+have a clear cause and effect relationship between the order release decision and the cost
+performance in the analysis.
+%
+To simulate high and moderate variability of the demand process the interarrival time between
+consecutive incoming orders is either drawn from an exponential ($\Exp$) or an uniform ($\Unif$)
+distribution, which are adjusted to yield the desired bottleneck utilisation level of either $70\%$
+with $\Exp(135)$ and $\Unif(95,175)$, or $80\%$ with $\Exp(118)$ and $\Unif(78,158)$. These values
+were chosen as they comprise the non-linear relationship between flow times and high utilisation
+levels.
+
+\paragraph{Order Release.}
+%
+Before the order release decision is made, all orders in the order pool are sorted by due date. This
+portrays the implementation of a sequencing rule. According to this sequence, orders are considered
+for release in the beginning of each period starting with the highest priority order.
+
+In our model, orders are released by specifying lead times
+$\lt_{1}, \lt_{2}, \ldots, \lt_{6} \geqslant 1$, where the index corresponds to the product type and
+$\lt_{i} \in \mathbb{N}$.
+%
+%
+By setting a lead time $\lt_{i}$ a \emph{planned release date} is computed for each
+order $j$ of product type $i$ in the order pool given by
+%
+\begin{linenomath*}
+  \begin{equation}
+    \prd{}_j= \dd{}_j - \lt_{i}\ ,
+    \label{eq:Orel}
+  \end{equation}
+\end{linenomath*}
+
+where $\dd_{j}$ denotes the due date of the order.  For dynamic order releases the lead times
+$\lt_{i}$ may vary from period to period, whereas in static order release methods the lead times are
+predetermined and fixed (\citealt{RagatzMabert1988,KimBobrowski1995}).
+%
+For both approaches a job of product type $i$ with lead time $\lt{}_{i}$ in period $t$ and due date
+$\dd_{j}$ is released at the end of period $t$ if and only if $t \geqslant \prd{}_{j}$.
+%
+Thus, in the dynamic setting  the planned
+release date $\prd{}_{j}$ of order $j$ can be updated several times before it is actually released
+to the production system.
+%
+However, once an order is released its release cannot be revoked.
+%
+% If there are several orders with the same due date the order release
+% sequence is randomized.
+%
+Hence, if the set lead time corresponds with the actual flow time the product is finished in the
+same period as it is shipped. If the lead times are set either too long or short,
+the order has to wait in the \fgi{} until it is due or the order is late and backorder (BO) costs
+occur.
+
+\paragraph{Costs.}
+%
+The cost parameters are set by assuming an increase in value from raw material ($1$~Dollar per order
+and period) to the final product ($4$~Dollar per order and period) and the backorder costs are set
+very high ($16$~Dollar per order and period) due to the MTO environment. All costs are assessed
+based on the production system state at the end of each period.
+
 \subsection{Conventional Order Release Rules}
 
-As external benchmark for comparison we use
-        %         use two different order release rules. First, we use immediate release which releases orders as
-        %         soon as they arrive at the order pool. And second, we
-different parameterized backward infinite loading (\BIL{}) techniques:
-
-\begin{equation*}
-  RD_{j} = DD_{j} - LT_{j} \tcom
-  \label{eq:BIL}
-\end{equation*}
-
-where the release date \(RD_{j}\) of product type \(j\) is calculated by the difference of the due
-date \(DD_{j}\) and the lead time \(LT_{j}\). Note that we set lead times for each product type and
-not for every order as done by \cite{Ackerman1963}.
-
-        %         The
-        %         variants are named according to the used lead time $LT$. For example in case of $\BILOne$ we have
-        %         $LT=1$. \MS{@Equation 1: Use LT$_{j}$ or LT?}
-
-        %         \MS[t]{do we?}
+As external benchmark for comparison we use different parameterized backward infinite loading
+(\BIL{}) techniques \citep{Ackerman1963}. \BIL{} uses Equation~\ref{eq:Orel} with a predetermined
+and fixed lead times \(\lt_{i}\).
+%
+% \begin{equation*}
+%   \rd_{j} = \dd_{j} - \lt_{j} \tcom
+%   \label{eq:BIL}
+% \end{equation*}
+%
+% where the release date \(RD_{j}\) of product type \(j\) is calculated by the difference of the due
+% date \(DD_{j}\) and the lead time \(LT_{j}\).
+Note that we set lead times for each product type and not for every order as done by
+\cite{Ackerman1963}.
 

### 2021-01-10 paper/figures/indep_agents.tex
indep agents + n-step algorithm

\begin{tikzpicture}[state/.style={draw,circle,inner sep=0.075cm,minimum size=1cm},scale=1.0]

  \def\spread{1.5}

  \draw (-1,0)  node[] (s00)  {};
  \draw (0,0)  node[state] (s0)  {$(3,3)$};
  \draw (4,\spread)  node[state] (s11) {$(4,4)$};
  \draw (4,0)  node[state] (s12) {$(4,3)$};
  \draw (4,-\spread) node[state] (s13) {$(4,2)$};

  \draw (5,\spread)  node[] (s21) {};
  \draw (5,0)        node[] (s22) {};
  \draw (5,-\spread) node[] (s23) {};

  \path (s0) edge[->, bend left=10, above]  node[yshift=0.10cm,xshift=0.03cm] {$p(\inc,\inc)$} (s11);
  \path (s0) edge[->, above]                node[] {$p(\inc,\pers)$} (s12);
  \path (s0) edge[->, bend right=10, above] node[yshift=0.12cm,xshift=0.03cm] {$p(\inc,\dec)$} (s13);

  \path (s00) edge[->, dashed]  node[yshift=0.10cm,xshift=0.03cm] {} (s0);
  \path (s11) edge[-, dashed]  node[yshift=0.10cm,xshift=0.03cm] {} (s21);
  \path (s12) edge[-, dashed]                node[] {} (s22);
  \path (s13) edge[-, dashed] node[yshift=0.12cm,xshift=0.03cm] {} (s23);

\end{tikzpicture}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../paper"
%%% End:

### 2021-01-10 paper/figures/productionsystem.tex
indep agents + n-step algorithm

@@
-\begin{tikzpicture}[queue/.style={regular polygon,regular polygon sides=4,inner sep=0.075cm},scale=0.89]
+\begin{tikzpicture}[queue/.style={regular polygon,regular polygon sides=4,inner sep=0.075cm},scale=0.78]
 
@@
   \path (OP.center) edge[->, bend left=10, above]
-  node[xshift=0.088cm,yshift=0.05cm,align=center] {\scriptsize Order \\[-1ex] \scriptsize Release} (Q1);
+  node[xshift=0.09cm,yshift=0.05cm,align=center] {\scriptsize Order \\[-1ex] \scriptsize Release} (Q1);
 
@@
   \path (M4) edge[->, bend left=10,above,near start]       node[xshift=0.1cm] {$\scriptstyle 1,4$} (\fgi{});
-  \path (M5) edge[->, bend left=10,above,near start]       node[xshift=0.1cm, yshift=-0.07cm] {$\scriptstyle 2,5$} (\fgi{});
+  \path (M5) edge[->, bend left=10,above,near start]       node[xshift=0.075cm, yshift=-0.07cm] {$\scriptstyle 2,5$} (\fgi{});
   \path (M6) edge[->, bend right=10,below,near start]      node[xshift=0.15cm,yshift=0.07cm] {$\scriptstyle 3,6$} (\fgi{});

### 2021-01-10 paper/paper.tex
indep agents + n-step algorithm

@@
 \newcommand\theauthor{
-  Manuel Schneckenreither% \inst{1}
+  Manuel Schneckenreither\inst{1}
   % \orcidID{0000-0002-4812-4665}
-  \and Stefan Haeussler% \inst{2}
+  \and Stefan Haeussler\inst{1}
   % \orcidID{0000-0003-2589-1367}
+  \and Juanjo Peiró\inst{2}
 }
@@
 \institute{
-  Department of Information Systems,
-  Production and Logistics Management,
-  University of Innsbruck, Austria\\
-  email: \email{manuel.schneckenreither@uibk.ac.at,stefan.haeussler@uibk.ac.at}
+  Department of Information Systems, Production and Logistics Management, University of Innsbruck, Austria\\
+  \email{manuel.schneckenreither@uibk.ac.at,stefan.haeussler@uibk.ac.at}
+  \and
+  Departament d’Estadística i Investigació Operativa, Universitat de València, Spain\\
+  \email{juanjo.peiro@uv.es}
 }
@@
 
+\MS[t]{ideas: Similar Selçuk (2013) detect that a high update frequency has a negative effect and
+  thus should be reduced. They conclude that the lead times ought to control the WIP level, instead
+  of the due date reliability as changes are directly seen (Knollmann and Windt, 2013). For a review
+  of analytical works in the context of the LTS see Bendul and Knollmann (2016).}
+
 \section{Introduction}\label{sec:introduction}
@@
 \citep[p.66]{sutton1998introduction}.
-% \MS{does our system possesses the Markov property at all? :-o }
 
@@
     \While{the stopping criterion is not fulfilled}
-    \State{}\begin{minipage}[t]{0.9\textwidth} With probability \(\pexp\) choose a random action and
+    \State{}\begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth}
+      With probability \(\pexp\) choose a random action and
       probability \(1-\pexp\) one that fulfills
@@
     \end{minipage}
-    \State{}\begin{minipage}[t]{0.9\textwidth}
+    \State{}\begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth}
     Carry out action \(a_{t}\), observe reward \(r_{t}\) and resulting state \(s_{t+1}\).
@@
     \begin{align*}
-      \avgrew^{\pol}  \gets (1- \alpha) \avgrew^{\pol} + \alpha [r_{t} +
-      \max_{a}\X^\pol_{\gamma_1}(s_{t+1},a; \netWorker{}) - \X^\pol_{\gamma_1}(s_{t},a_{t}; \netWorker{})]
+      \avgrew^{\pol}  & \gets (1- \alpha) \avgrew^{\pol} + \alpha [r_{t} +
+      \max_{a}\X^\pol_{\gamma_1}(s_{t+1},a; \netWorker{}) - \X^\pol_{\gamma_1}(s_{t},a_{t};
+                        \netWorker{})]
+      % \\ \avgrew^{\pol}  & \gets \max(\avgrew^{\pol}, \avgrew_{\min}^{\pol})\\
+      % \avgrew_{\min}^{\pol} & \gets (1 - \alpha_{\min}) \avgrew_{\min}^{\pol} + \alpha_{\min} 0.975
+      % \cdot \avgrew^{\pol}
     \end{align*}
@@
     \end{align*}
-    \State{} \begin{minipage}[t]{0.9\textwidth} Update the average reward adjusted discounted
+    \State{} \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth} Update the average reward adjusted discounted
       state-values using the sum of the gradients on \({(y_{i,\cdot} -
@@
 
-\subsection{Conventional Order Release Rules}
+\subsection{Conventional Order Release Rule}
 
@@
 % date \(DD_{j}\) and the lead time \(LT_{j}\).
-Note that we set lead times for each product type and not for every order as done by
-\cite{Ackerman1963}.
+%
+% Note that we set lead times for each product group and not for every order as done by
+% \cite{Ackerman1963}.
 
@@
 
-\paragraph{Markov Decision Process.} The underlying MDP is unichain and looks as follows.
-Both, state space $\mathcal{S}$ and action space $\mathcal{A}$ are discrete. Any state
-$s \in \mathcal{S}$ of the \textit{state space} is composed of the following information for each
-product $p$:
-        %         The following information is presented to the algorithm for both Product 1 and Product 2 as follows.
+\paragraph{Markov Decision Process.} The underlying MDP is unichain and looks as follows. Both,
+state space $\mathcal{S}$ and action space $\mathcal{A}$ are discrete.
+%
+% Recall that we gproducts routed over the bottleneck machine \(M5\) are bottleneck products.
+%
+Any state $s \in \mathcal{S}$ of the \textit{state space} is composed of the following information
+for both groups \(g\), i.e.\@ for bottleneck and non-bottleneck products:
 \begin{itemize}
 
-\item The currently set lead time $LT_{p} \in \{1,2,\ldots,\DDS\}$ (Recall:
+\item The currently set lead time $LT_{g} \in \{1,2,\ldots,\DDS\}$ (Recall:
   $\text{Due Date Period} - \text{Lead Time} = \text{Release Period}$). Note that we bound the
@@
 
-\item Counters $OP_{p,d}\in \N$ for the number of orders in the order pool divided in time buckets
+\item Counters $OP_{g,d}\in \N$ for the number of orders in the order pool divided in time buckets
   with $d \in \{1,2,\ldots,\DDS \}$, which stands for the number of periods until the due date.
@@
 
-\item Counters $Q_{i}\in \N$ standing for the number of orders for each queue $i$.
-
-\item Counters $FGI_{p,d} \in \N$ of orders in the finished goods inventory divided in time
-  buckets with $d \in \{-5,-4,-3,\ldots, \DDS \}$, which stands for the number of periods until the
-  due date. Orders with a due date with more than 5 periods ago are listed in the counter
-  $FGI_{-5}$.
-
-\item Counters $S_{p,d} \in \N$ of shipped orders from the last period divided in time buckets
-  with $d \in\{-5,-4,-3,\ldots, \DDS \}$, which stands for the number of periods until the due date.
-  Orders with a due date with more than 5 periods ago are listed in the counter $S_{-5}$.
+\item Counters $Q_{g,i,d}\in \N$ for the number of orders in the machine and queue \(i \in
+  \{M_{1}, M_{2}, \ldots , M_{6}\}\) divided in
+  time buckets with $d \in \{1,2,\ldots,\DDS \}$, which stands for the number of periods until the
+  due date.
+
+\item Counters $FGI_{g,d} \in \N$ of orders in the finished goods inventory divided in time
+  buckets with $d \in \{1, 2,\ldots, \DDS \}$, which stands for the number of periods until the
+  due date. Orders with a due date with more than 3 periods ago are listed in the counter
+  $FGI_{-3}$.
+
+\item Counters $S_{g,d} \in \N$ of shipped orders from the last period divided in time buckets
+  with $d \in\{-3,-2,\ldots, \DDS \}$, which stands for the number of periods until the due date.
+  Orders with a due date with more than 3 periods ago are listed in the counter $S_{-3}$.
 
@@
 
-\noindent The algorithm implicitly learns a function which maps the current state of the
-production system to a release decision. In an optimal situation it does so by multiply exploring
-every action for each state and assessing its economic viability.
-
+The algorithm implicitly learns a function which maps the current state of the production
+system to a lead time update and thus a release decision. In an optimal situation it does so by
+multiply exploring every action for each state and assessing its economic viability.
+%
 Orders are released once the due date is within the interval $[t,t+\text{lead time}]$, whereas $t$
-is the current period. Clearly, this yields bulk releases (either all or no orders with the same
-due date and product type are released). % Note that we do not present past state
-        %         information to the neural network.
-
-The \textit{actions space} is composed of two independent decisions. These are the relative changes
-of the lead times to the currently set lead times $LT_{p} \in \{1,2,\ldots,\DDS\}$ for each product
-$p$. Recall that $p=2$.
-        %
-Furthermore, we restrict the action space for each state $s_{t+1}$ according to the last set lead
-time $LT_{p}$ from state $s_{t}$ by restricting the change of the lead time for consecutive
-periods to a maximum of 1. Thus, if $LT_{p}$ is the current lead time for product $p$ the action
-space for this product is given by $\{1,2,\ldots,\DDS\} \cap \{LT_{p}-1,LT_{p},LT_{p}+1\}$. Put
-differently the algorithm can increase or decrease the lead time by 1 or leave it as it is, as
-long as it acts within the discrete action space given by the set $\{1,2,\ldots,\DDS\}$ for each
-type of product. Thus the action space over all products is given by the full enumeration of
-available actions of the individual products.
+is the current period. Clearly, this yields bulk releases, meaning that either all or no orders with
+the same due date and of the same product group are
+released.
+
+The \textit{action space} is composed of two decisions, the relative change of the lead times to the
+currently set lead times $LT_{g} \in \{1,2,\ldots,\DDS\}$ for each product group $g$.
+%
+We restrict the lead time update for state $s_{t+1}$ according to the lead time $LT_{g}$ from state
+$s_{t}$ by a maximum change of 1. Thus, if $LT_{g}$ is the current lead time of product group $g$
+the agent can choose a lead time in $\{1,2,\ldots,\DDS\} \cap \{LT_{g}-1,LT_{g},LT_{g}+1\}$. % Put
+% differently the algorithm can increase or decrease the lead time by 1 or leave it as it is, as long
+% as it acts within the discrete action space given by the set $\{1,2,\ldots,\DDS\}$ for each product
+% group.
+A naive approach is to define one action for each combination of actions over all product groups.
+Although applicable in our setup, as we only define two groups that yield \(9\) actions, it simply
+does not scale.
+%
+\begin{figure*}[t!]
+  \centering
+  \input{figures/indep_agents}                      % Figure production system
+  \caption{Illustration of independent agents as agent one increases the lead time}\label{fig:indep-agents}
+\end{figure*}
+%
+Therefore, we invent and implement \textit{independent agents}, each operating on the same system
+state, but choosing actions independently for the designated product group.
+Figure~\ref{fig:indep-agents} illustrates the idea, where wlog. we use two agents and for simplicity
+omit the reward and truncate the state to the lead times. Agent one chooses to increase the lead
+time from \(3\) to \(4\). Thus, the future state, seen from agent one, depends on the probability of
+agent two choosing to increase (\(\inc\)), decrease (\(\dec\)) or not change (\(\pers\)) the lead
+time. Note that the Markov property the agent, that is
+\(p(s_{t+1}, r_{t} \mid s_{t}, a_{t}) = p((4,\cdot), r_{t} \mid (3,3), \inc) \), is not touched by
+this adaption. While for agent one the reward and future state only depends on the current state and
+action taken, the corresponding probabilities for future state and reward change during the learning
+process as the other agent starts to behave differently.
+%
+\MS[t]{independent agents, true?}This allows us to investigate the algorithms behaviour for action
+solution spaces. In particular, we present the algorithm the above defined solution space not only
+on the basis of product groups \(g\), but also for each product type \(p\).
+That means, the state space, as well as the action space operates on a product type basis. Note that
+without independent agents this is infeasible due to \(3^{6} = 729\) possible actions.
 
@@
 period by simulating the production system. The rewards are the accumulated costs at the end of the
-period. These costs are consist of the number of backorders, the current WIP level and the number of
-orders in the inventory.
-        %
-        %         The costs are divided by the normalizing parameter $\eta$ (see Table~\ref{tbl:variants}) and then
-        %         clipped (cut off) to the interval $[-0.5,0.5]$.
+period consisting of the backorders, the WIP orders and the number of orders in the FGI. The
+objective is to minimise the returned reward. Therefore, we adapt the algorithm by swapping all
+maximisation functions with minimisations.
 
 \paragraph*{Adapted Algorithm.}
-The algorithm depicted above is designed as a tabular version. We have adapted it to use neural
-networks for the approximation of the different kind of state-action pair value functions. This is
-required due to the exponential growth of the state space. The implementation collects the observed
-information for each state-action function in a replay memory and randomly selects \(128\) data
-points out of a collection of \(30k\) memories in each period and trains these on the neural
-network. This technique is called experience replay memory
-\citep{lin1993reinforcement,mnih2015human} and ought to overcome we instabilities. Nonetheless, this
-approximation results in the fact that the algorithm can be unstable or may even diverge
-\citep{tsitsiklis1997analysis}. Additionally we use a target and a worker network, whereas the
-target network is overwritten every \(10k\) steps by the worker network. This has been proposed by
-\cite{mnih2015human} and should further stabilize the algorithm.
+Our implementation includes several optimisations, that help the algorithm in finding better
+policies and stabilise the learning process. These are the above discussed independent agents that
+operate on the same state, \(n\)-step learning \citep{mnih2016asynchronous}, overestimating
+\(\avgrew^{\pol}\), using multiple parallel worker agents that operate on the same environment to
+collect a broader range of experiences, storing experiences according to their action and uniformly
+sampling over the actions to ensure each action is constantly trained to prevent catastrophic
+forgetting.
+%
+N-step RL collects a set of consecutive experiences and
+
+\MS[t]{explanation
+  missing}
+
+% The algorithm depicted above is designed as a tabular version. We have adapted it to use neural
+% networks for the approximation of the different kind of state-action pair value functions. This is
+% required due to the exponential growth of the state space. The implementation collects the observed
+% information for each state-action function in a replay memory and randomly selects \(128\) data
+% points out of a collection of \(30k\) memories in each period and trains these on the neural
+% network. This technique is called experience replay memory
+% \citep{lin1993reinforcement,mnih2015human} and ought to overcome we instabilities. Nonetheless, this
+% approximation results in the fact that the algorithm can be unstable or may even diverge
+% \citep{tsitsiklis1997analysis}. Additionally we use a target and a worker network, whereas the
+% target network is overwritten every \(10k\) steps by the worker network. This has been proposed by
+% \cite{mnih2015human} and should further stabilize the algorithm.
+
+\MS[t]{Extensions: overestimate rho, rho minimum, Multiple Agents (shared state,  shared rho?),
+  init-phase (0.5 for expSmthRewRate, always adapt rho, same rho for all agents): exp smooth reward
+  as rho, replay memory actions, n-step (describe replay memory), gradient clipping }
+\MS[t]{Side info: Replay memory also stores (disallowed) filtered action indices for optimisation purposes and on
+episodic tasks needs to store a bool indicating an episode end. Uses Int8 for state features values.  }
+
+\begin{algorithm}[t!]
+  \begin{algorithmic}[1]
+    \State{}Initialise state \(s_{0}\) and network parameters \(\netTarget, \netWorker\) randomly, set
+    an exploration rate \(0 \leqslant \plearn \leqslant \pexp \leqslant 1\), exponential smoothing learning rates
+    \(0 < \alpha, \gamma < 1\), and discount factors
+    \(0 < \gamma_{0} < \gamma_{1} \leqslant 1\), where \(\gamma_{1} = 1\) is
+    usually a good choice.
+    \While{the stopping criterion is not fulfilled}
+    \State{}%
+    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth}
+      With probability \(\pexp\) choose a random action and
+      probability \(1-\pexp\) one that fulfills
+      \(\min_{a}\lexeq(\X^{\pol}_{\gamma_{1}}(s_{t},a; \netWorker{}),\X^\pol_ {\gamma_{0}}(s_{t},a;
+      \netWorker{}))\). Let \(\isRandAct{t} = \mathsf{True}\) indicate that the action was chosen
+      randomly
+    \end{minipage}
+    \State{}%
+    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth}
+    Carry out action \(a_{t}\), observe reward \(r_{t}\) and resulting state \(s_{t+1}\).
+    Store the experience \((s_{t}, a_{t}, \isRandAct{t}, r_{t}, s_{t+1})\) in the experience replay
+    memory \(M\).
+    \end{minipage}
+    \If{\(t \mod{\nstep} \equiv 0\)}
+    \State{Reset gradient: \(d\netWorker{} \gets 0\)}
+    \State{}%
+    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
+      Sample random mini-batch of \(\nstep\) consecutive experiences from \(M\), each with
+      \((s_{i}, \isRandAct{i}, a_{i}, r_{i}, s_{i})\) where \(i \in {1,2,\ldots,\nstep}\).
+    \end{minipage}
+    \For{each mini-batch}
+    \For{each experience, starting with the latest seen experience \(\nstep\)}
+    \If{experience \(\nstep\) or (\(\isRandAct{i}\) and  \(\pexp \leqslant \plearn\))}
+      \State{\(\target_{\gamma0} \gets \min_{a} \X^\pol_{\gamma_0}(s_{t+1},a; \netTarget{})\)}
+      \State{\(\target_{\gamma1} \gets \min_{a} \X^\pol_{\gamma_1}(s_{t+1},a; \netTarget{})\)}
+    \EndIf
+    \If{a non-random action was chosen or \(\pexp > \plearn\)}\\
+    \hspace{2.4cm}\(\avgrew^{\pol} \gets (1- \alpha) \avgrew^{\pol} + \alpha [r_{t} +
+      \min_{a}\X^\pol_{\gamma_1}(s_{t+1},a; \netWorker{}) - \X^\pol_{\gamma_1}(s_{t},a_{t};
+      \netWorker{})]\)
+    \EndIf
+    \State{\(\target_{\gamma_0}  \gets r_{t} + \gamma_{0} \target_{\gamma_{0}} - \avgrew^{\pol}  \)}
+    \State{\(\target_{\gamma_{1}}  \gets r_{t} + \gamma_{1} \target_{\gamma_{1}} - \avgrew^{\pol}\)}
+    \State{\(d\netWorker{} \gets d\netWorker{} + \partial{(\target_{\gamma_{0}} - \X^\pol_{\gamma_{0}}(s_{t+1},a; \netWorker{}))}^{2} / \partial\netWorker{}\)}
+    \State{\(d\netWorker{} \gets d\netWorker{} +
+      \partial{(\target_{\gamma_{1}} - \X^\pol_{\gamma_{1}}(s_{t+1},a; \netWorker{}))}^{2} / \partial\netWorker{}\)}
+    \EndFor{}
+    \EndFor{}
+    \State{Perform update of \(\netWorker{}\) using \(d\netWorker{}\)}
+    \EndIf{}
+    \State{Every \(C\) steps exponential smoothly set target network:
+    \(\netTarget \gets (1 - \gamma) \netTarget{} + \gamma \netWorker{}\)}
+    \State{} Set \(s \gets s'\), \(t \gets t+1\) and decay parameters
+    \EndWhile{}
+  \end{algorithmic}
+  \caption{\label{alg:fin}Near-Blackwell-optimal deep RL for unichain MDPs}
+\end{algorithm}
 

### 2021-01-10 paper/paper.tex
update

@@
     \end{minipage}
+    \State{Sample random mini-batch of experiences from \(M\)}
+    \State{Reset gradients: \(d\netWorker{} \gets 0\)}
+    \For{each experience \(i\) with \((s_{t}, \isRandAct{t}, a_{t}, r_{t}, s_{t+1})\)}
     \If{a non-random action was chosen or \(\pexp > \plearn\)}
-    \begin{align*}
-      \avgrew^{\pol}  & \gets (1- \alpha) \avgrew^{\pol} + \alpha [r_{t} +
+    \State{\(
+      \avgrew^{\pol}  \gets (1- \alpha) \avgrew^{\pol} + \alpha [r_{t} +
       \max_{a}\X^\pol_{\gamma_1}(s_{t+1},a; \netWorker{}) - \X^\pol_{\gamma_1}(s_{t},a_{t};
-                        \netWorker{})]
-      % \\ \avgrew^{\pol}  & \gets \max(\avgrew^{\pol}, \avgrew_{\min}^{\pol})\\
-      % \avgrew_{\min}^{\pol} & \gets (1 - \alpha_{\min}) \avgrew_{\min}^{\pol} + \alpha_{\min} 0.975
-      % \cdot \avgrew^{\pol}
-    \end{align*}
+                        \netWorker{})]\)}
     \EndIf
-
-    \State{}Sample random mini-batch of experiences \((s_{i}, \isRandAct{i}, a_{i}, r_{i}, s_{i})\)
-    from \(M\) and do
-    \begin{align*}
-      y_{i,\gamma_0} & \gets r_{t} + \gamma_{0} \max_{a}
-                                        \X^\pol_{\gamma_0}(s_{t+1},a; \netTarget{}) - \avgrew^{\pol} \\
-      y_{i,\gamma_{1}} & \gets r_{t} + \gamma_{1} \max_{a}
-                                        \X^\pol_{\gamma_1}(s_{t+1},a; \netTarget{}) - \avgrew^{\pol}
-    \end{align*}
-    \State{} \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth} Update the average reward adjusted discounted
-      state-values using the sum of the gradients on \({(y_{i,\cdot} -
-      \X^\pol_{\cdot}(s_{t+1},a; \netWorker{}))}^{2}\) wrt. network parameters
-    \(\netWorker{}\).
+    \State{\(y_{i,\gamma_0} \gets r_{t} + \gamma_{0} \max_{a}
+                                        \X^\pol_{\gamma_0}(s_{t+1},a; \netTarget{}) - \avgrew^{\pol}\)}
+    \State{\(y_{i,\gamma_{1}} \gets r_{t} + \gamma_{1} \max_{a}
+                                        \X^\pol_{\gamma_1}(s_{t+1},a; \netTarget{}) - \avgrew^{\pol}\)}
+    \State{Sum gradients on \({(y_{i,\cdot} - \X^\pol_{\cdot}(s_{t+1},a;
+        \netWorker{}))}^{2}\) wrt. parameters \(\netWorker{}\) in \(d\netWorker{}\)}
+    \EndFor
+    \State{}%
+    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth}
+      Perform update with sum of the gradients \(d\netWorker{}\) on \(\netWorker{}\)
     \end{minipage}
-    \State{} Every \(C\) steps exponential smoothly set target network:
-      \(\netTarget \gets (1 - \gamma) \netTarget{} + \gamma \netWorker{}\)
-    \State{} Set \(s \gets s'\), \(t \gets t+1\) and decay parameters
+    \State{Every \(C\) steps exponential smoothly set target network:
+      \(\netTarget \gets (1 - \gamma) \netTarget{} + \gamma \netWorker{}\)}
+    \State{Set \(s \gets s'\), \(t \gets t+1\) and decay parameters}
     \EndWhile{}
@@
 policies and stabilise the learning process. These are the above discussed independent agents that
-operate on the same state, \(n\)-step learning \citep{mnih2016asynchronous}, overestimating
-\(\avgrew^{\pol}\), using multiple parallel worker agents that operate on the same environment to
-collect a broader range of experiences, storing experiences according to their action and uniformly
-sampling over the actions to ensure each action is constantly trained to prevent catastrophic
-forgetting.
-%
-N-step RL collects a set of consecutive experiences and
-
-\MS[t]{explanation
-  missing}
+operate on the same state, \(n\)-step learning \citep{mnih2016asynchronous}, overestimating the
+average reward \(\avgrew^{\pol}\). Algorithm~\ref{alg:full} depicts the algorithm including the
+extensions.
+
+%
+N-step RL collects a set of consecutive experiences, in our case retrieved from the replay memory,
+and connects the state values starting from the latest observed experience, such that the newly
+computed state value of the corresponding experience propagates to the experience of the previous
+period. This allows a faster convergence, as values bootstrap over multiple periods in contrast to
+just one step.
+
+\MS[t]{rhoMinimum, (shared rho)?, init-phase, (gradient clipping)?}
+
+Furthermore, we use multiple parallel worker agents that operate on the same
+environment to collect a broader range of experiences, storing experiences according to their action
+and uniformly sampling over the actions to ensure each action is constantly trained to prevent
+catastrophic forgetting.
+
+\MS[t]{explanation missing}
 
@@
     \While{the stopping criterion is not fulfilled}
+    \State{\(\isRandAct{t} \gets \mathsf{False}\)}
+    \ForEach{independent agent \(\agent_{1},\agent_{2},\ldots,\agent_{n_{\mathsf{ag}}}\)}
     \State{}%
-    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth}
-      With probability \(\pexp\) choose a random action and
-      probability \(1-\pexp\) one that fulfills
+    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
+      With probability \(\frac{\pexp}{n_{\mathsf{ag}}}\) choose a random action and
+      probability \(1-\frac{\pexp}{n_{\mathsf{ag}}}\) one that fulfills
       \(\min_{a}\lexeq(\X^{\pol}_{\gamma_{1}}(s_{t},a; \netWorker{}),\X^\pol_ {\gamma_{0}}(s_{t},a;
-      \netWorker{}))\). Let \(\isRandAct{t} = \mathsf{True}\) indicate that the action was chosen
-      randomly
+      \netWorker{}))\). Set \(\isRandAct{t} \gets \mathsf{True}\) if the action was chosen randomly
     \end{minipage}
+    \EndFor
     \State{}%
@@
     \end{minipage}
-    \For{each mini-batch}
-    \For{each experience, starting with the latest seen experience \(\nstep\)}
+    \ForEach{mini-batch}
+    \ForEach{experience, starting with the latest seen experience \(\nstep\)}
     \If{experience \(\nstep\) or (\(\isRandAct{i}\) and  \(\pexp \leqslant \plearn\))}
@@
     \EndIf
-    \State{\(\target_{\gamma_0}  \gets r_{t} + \gamma_{0} \target_{\gamma_{0}} - \avgrew^{\pol}  \)}
-    \State{\(\target_{\gamma_{1}}  \gets r_{t} + \gamma_{1} \target_{\gamma_{1}} - \avgrew^{\pol}\)}
+    \State{Compute overestimate of \(\avgrew^{\pol}\) and store in \(\avgrew_{\Overestim}^{\pol}\)}
+    \State{\(\target_{\gamma_0}  \gets r_{t} + \gamma_{0} \target_{\gamma_{0}} - \avgrew_{\Overestim}^{\pol}  \)}
+    \State{\(\target_{\gamma_{1}}  \gets r_{t} + \gamma_{1} \target_{\gamma_{1}} - \avgrew_{\Overestim}^{\pol}\)}
     \State{\(d\netWorker{} \gets d\netWorker{} + \partial{(\target_{\gamma_{0}} - \X^\pol_{\gamma_{0}}(s_{t+1},a; \netWorker{}))}^{2} / \partial\netWorker{}\)}
@@
   \end{algorithmic}
-  \caption{\label{alg:fin}Near-Blackwell-optimal deep RL for unichain MDPs}
+  \caption{\label{alg:full}Near-Blackwell-optimal deep RL for unichain MDPs}
 \end{algorithm}

### 2021-01-11 paper/paper.tex
update algorithm

@@
 period by simulating the production system. The rewards are the accumulated costs at the end of the
-period consisting of the backorders, the WIP orders and the number of orders in the FGI. The
+period consisting of the backorders, the WIP orders and the number of orders in the FGI\@. The
 objective is to minimise the returned reward. Therefore, we adapt the algorithm by swapping all
@@
 operate on the same state, \(n\)-step learning \citep{mnih2016asynchronous}, overestimating the
-average reward \(\avgrew^{\pol}\). Algorithm~\ref{alg:full} depicts the algorithm including the
-extensions.
-
-%
+average reward \(\avgrew^{\pol}\) and ensuring to constantly reduce the average reward as searching
+for policies with higher average costs is unnecessary. Algorithm~\ref{alg:full} depicts a detailed
+view on the algorithm including the extensions and with the minimisation objective.
+
+% N-Step
 N-step RL collects a set of consecutive experiences, in our case retrieved from the replay memory,
@@
 just one step.
+% Overestimating
+The idea behind overestimating the average reward is to constantly seek for better policies than the
+currently known. We overestimate the reward by subtracting \(10\%\) of the computed average reward
+\(\avgrew^{\pol}\) and an exponentially smoothed average reward of the actually returned reward
+values of greedy actions. Thus, if the computed and actually observed average reward coincide, the
+overestimation equals \(0\).
+% Rho Minimum
+Preventing an increase of the average reward with \(\avgrew_{\max}^{\pol}\) reduces the search space
+and provides a methodology for constantly improving the policy
+
+\MS[t]{describe algorithm}
 
@@
   \begin{algorithmic}[1]
-    \State{}Initialise state \(s_{0}\) and network parameters \(\netTarget, \netWorker\) randomly, set
-    an exploration rate \(0 \leqslant \plearn \leqslant \pexp \leqslant 1\), exponential smoothing learning rates
-    \(0 < \alpha, \gamma < 1\), and discount factors
-    \(0 < \gamma_{0} < \gamma_{1} \leqslant 1\), where \(\gamma_{1} = 1\) is
-    usually a good choice.
+    \State{}Initialise state \(s_{0}\) and network parameters \(\netTarget, \netWorker\) randomly,
+    set an exploration rate \(0 \leqslant \plearn \leqslant \pexp \leqslant 1\), exponential
+    smoothing learning rates \(0 < \alpha_{\max}, \alpha, \gamma < 1\), a sufficiently large default
+    value for \(\avgrew_{\max}^{\pol}\) and discount factors \(0 < \gamma_{0} < \gamma_{1} \leqslant
+    1\)
     \While{the stopping criterion is not fulfilled}
@@
     \EndIf
-    \If{a non-random action was chosen or \(\pexp > \plearn\)}\\
-    \hspace{2.4cm}\(\avgrew^{\pol} \gets (1- \alpha) \avgrew^{\pol} + \alpha [r_{t} +
+    \If{a non-random action was chosen or \(\pexp > \plearn\)}
+    \State{\(\avgrew^{\pol} \gets (1- \alpha) \avgrew^{\pol} + \alpha [r_{t} +
       \min_{a}\X^\pol_{\gamma_1}(s_{t+1},a; \netWorker{}) - \X^\pol_{\gamma_1}(s_{t},a_{t};
-      \netWorker{})]\)
+      \netWorker{})]\)}
+    \State{\(\avgrew^{\pol} \gets \min(\avgrew^{\pol}, \avgrew_{\max}^{\pol})\)}
+    \State{\(\avgrew_{\max}^{\pol} \gets (1- \alpha_{\max}) \avgrew_{\max}^{\pol} +
+    \alpha_{\max} 1.025 \avgrew^{\pol}\)}
     \EndIf
-    \State{Compute overestimate of \(\avgrew^{\pol}\) and store in \(\avgrew_{\Overestim}^{\pol}\)}
+    \State{Compute overestimate of \(\avgrew^{\pol}\) and store result in \(\avgrew_{\Overestim}^{\pol}\)}
     \State{\(\target_{\gamma_0}  \gets r_{t} + \gamma_{0} \target_{\gamma_{0}} - \avgrew_{\Overestim}^{\pol}  \)}
@@
     \(\netTarget \gets (1 - \gamma) \netTarget{} + \gamma \netWorker{}\)}
-    \State{} Set \(s \gets s'\), \(t \gets t+1\) and decay parameters
+    \State{Set \(s \gets s'\), \(t \gets t+1\) and decay parameters}
     \EndWhile{}
@@
 %
+\MS[t]{TODO}
 We use the same setup for the neural network as done by \cite{Schneckenreither2019}. However as they

### 2021-01-12 paper/paper.tex
update

@@
 %
-\MS[t]{independent agents, true?}This allows us to investigate the algorithms behaviour for action
-solution spaces. In particular, we present the algorithm the above defined solution space not only
-on the basis of product groups \(g\), but also for each product type \(p\).
-That means, the state space, as well as the action space operates on a product type basis. Note that
-without independent agents this is infeasible due to \(3^{6} = 729\) possible actions.
+\MS[t]{independent agents, true?}This allows us to investigate the algorithms behaviour for even
+larger solution spaces. In particular, this adaption enables us to present the algorithm the above
+defined solution space not only on the basis of product groups \(g\), but also for each product type
+\(p\). That means, the state space, as well as the action space operates on a product type basis.
+Note that without independent agents this is infeasible due to \(3^{6} = 729\) possible actions.
+Removing the dependency between the lead times of different product types reduces the possibilities,
+and thus the number of ANN output layer nodes, to \(18\). Therefore, this is a vital adaption that
+ensures scalability.
 
@@
 \paragraph*{Reward.}
-In each period the agent chooses an action which generates a reward while traversing to the next
+In each period the agents choose actions which generates a reward while traversing to the next
 period by simulating the production system. The rewards are the accumulated costs at the end of the
@@
 \paragraph*{Adapted Algorithm.}
-Our implementation includes several optimisations, that help the algorithm in finding better
-policies and stabilise the learning process. These are the above discussed independent agents that
-operate on the same state, \(n\)-step learning \citep{mnih2016asynchronous}, overestimating the
-average reward \(\avgrew^{\pol}\) and ensuring to constantly reduce the average reward as searching
-for policies with higher average costs is unnecessary. Algorithm~\ref{alg:full} depicts a detailed
-view on the algorithm including the extensions and with the minimisation objective.
+Our implementation includes several optimisations, that help the algorithm in better exploring the
+solution space and stabilise the learning process. These include the above discussed independent
+agents that operate on the same state, \(n\)-step learning \citep{mnih2016asynchronous},
+overestimating the average reward \(\avgrew^{\pol}\) and ensuring to constantly reduce the average
+reward as searching for policies with higher average costs is unnecessary. Algorithm~\ref{alg:full}
+depicts a detailed view on the algorithm including these extensions and with the minimisation
+objective. The idea is the same as in Algorithm~\ref{alg:near}, except that the action selection
+procedure is a loop over all independent agents and the implementation of the n-step logic.
 
@@
 Preventing an increase of the average reward with \(\avgrew_{\max}^{\pol}\) reduces the search space
-and provides a methodology for constantly improving the policy
-
-\MS[t]{describe algorithm}
-
-\MS[t]{rhoMinimum, (shared rho)?, init-phase, (gradient clipping)?}
-
-Furthermore, we use multiple parallel worker agents that operate on the same
-environment to collect a broader range of experiences, storing experiences according to their action
-and uniformly sampling over the actions to ensure each action is constantly trained to prevent
-catastrophic forgetting.
-
-\MS[t]{explanation missing}
-
-% The algorithm depicted above is designed as a tabular version. We have adapted it to use neural
-% networks for the approximation of the different kind of state-action pair value functions. This is
-% required due to the exponential growth of the state space. The implementation collects the observed
-% information for each state-action function in a replay memory and randomly selects \(128\) data
-% points out of a collection of \(30k\) memories in each period and trains these on the neural
-% network. This technique is called experience replay memory
-% \citep{lin1993reinforcement,mnih2015human} and ought to overcome we instabilities. Nonetheless, this
-% approximation results in the fact that the algorithm can be unstable or may even diverge
-% \citep{tsitsiklis1997analysis}. Additionally we use a target and a worker network, whereas the
-% target network is overwritten every \(10k\) steps by the worker network. This has been proposed by
-% \cite{mnih2015human} and should further stabilize the algorithm.
-
-\MS[t]{Extensions: overestimate rho, rho minimum, Multiple Agents (shared state,  shared rho?),
-  init-phase (0.5 for expSmthRewRate, always adapt rho, same rho for all agents): exp smooth reward
-  as rho, replay memory actions, n-step (describe replay memory), gradient clipping }
-\MS[t]{Side info: Replay memory also stores (disallowed) filtered action indices for optimisation purposes and on
-episodic tasks needs to store a bool indicating an episode end. Uses Int8 for state features values.  }
+and provides a methodology for constantly improving policy.
+
+\MS[t]{(shared rho)?, (gradient clipping)?}
+
+Furthermore, we sometimes use multiple parallel workers (parallel RL) that operate on the same
+environment to collect a broader range of experiences. The experiences are stored according to the
+first taken action in the n-step series. We uniformly sample over the actions to ensure each action
+is constantly trained. This reduces catastrophic forgetting.
+
+The algorithm initiation tactics is to first fill the replay memory buffer by skipping any learning
+(steps \(10-25\)) until the period, that corresponds to the replay memory size, is reached. Then
+for \(500\) steps the average reward is set to the last computed exponentially smoothed value to
+prevent divergence due to the untrained network weight initialisation values.
+
+% \MS[t]{Extensions: overestimate rho, rho minimum, Multiple Agents (shared state,  shared rho?),
+%   init-phase (0.5 for expSmthRewRate, always adapt rho, same rho for all agents): exp smooth reward
+%   as rho, replay memory actions, n-step (describe replay memory), gradient clipping }
+% \MS[t]{Side info: Replay memory also stores (disallowed) filtered action indices for optimisation purposes and on
+% episodic tasks needs to store a bool indicating an episode end. Uses Int8 for state features values.  }
 
@@
     \State{\(\isRandAct{t} \gets \mathsf{False}\)}
-    \ForEach{independent agent \(\agent_{1},\agent_{2},\ldots,\agent_{n_{\mathsf{ag}}}\)}
+    \ForEach{independent agent \(\agent \in \agent_{1},\agent_{2},\ldots,\agent_{n_{\mathsf{ag}}}\)}
     \State{}%
@@
       probability \(1-\frac{\pexp}{n_{\mathsf{ag}}}\) one that fulfills
-      \(\min_{a}\lexeq(\X^{\pol}_{\gamma_{1}}(s_{t},a; \netWorker{}),\X^\pol_ {\gamma_{0}}(s_{t},a;
+      \(\min_{a \in \mathcal{A}(\agent)}\lexeq(\X^{\pol}_{\gamma_{1}}(s_{t},a; \netWorker{}),\X^\pol_ {\gamma_{0}}(s_{t},a;
       \netWorker{}))\). Set \(\isRandAct{t} \gets \mathsf{True}\) if the action was chosen randomly
@@
     \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth}
-    Carry out action \(a_{t}\), observe reward \(r_{t}\) and resulting state \(s_{t+1}\).
+    Carry out combined action \(a_{t}\), observe reward \(r_{t}\) and resulting state \(s_{t+1}\).
     Store the experience \((s_{t}, a_{t}, \isRandAct{t}, r_{t}, s_{t+1})\) in the experience replay

### 2021-01-12 paper/paper.tex
results 1

@@
 
-\section{Preliminary Results}
+\section{Computational Experiments}
 \label{sec:Preliminary}
@@
 
-The parameter setup is given in Table~\ref{tbl:params}. The parameters are exponentially decayed
-with rate \(0.05\) and \(350k\) steps\footnote{The decay is defined as \(0.05 ^{(t/350k)}\), where
-  \(t\) is the current period}. We implemented the deep Q-network algorithm proposed by
-\cite{mnih2015human}, which is parameterised by learning rate \(\delta\) and a discount factor of
-\(0.99\).
-
 \begin{table}[t]
   \centering
-        %         \def\arraystretch{1.2}          %  1 is the default, change whatever you need
-
   \begin{tabular}{|l||c|c|c|c|c|c|c|c|c|c|}
@@
   \end{tabular}
-
-        %         \egroup
-
   \caption{\label{tbl:params}Parameter Setup.}
-        %         \end{figure}
 \end{table}
 
-Table~\ref{tbl:res} shows the results, where we evaluate the average reward and discounted
-reinforcement learning algorithm by providing either direct reward feedback, that is the reward
-accumulated in the next period, without keeping track of any orders, as \(\ActShipped\) and
-\(\ActShippedDQN\) respectively. Further both algorithms are evaluated with keeping track of the
-order pool orders, such that each action is responsible for all orders in the order pool. These are
-labelled \(\ActOrdPool\) and \(\ActOrdPoolDQN\). For comparisons reasons we provide
-\(\BILOne-\BILSix\) and immediate release (interval release).
-
-The results show that at the current time we are facing issues with scaling the algorithm to the
-order release problem. Likely this is due to parameterisation issues, which we are investigating
-currently. However, it can be seen that i) the technique of keeping track of the order pool orders
-leads to better policies and ii) that despite the parameterisation issues average reward
-reinforcement learning finds better solutions than its discounted counterpart. However, the least
-costs are generated by \(\BILThree\), which shows that there is room for improvement for the
-reinforcement learning policies.
-
-Interesting is also to see, that all reinforcement learning variants are releasing rather late,
-leading to low finished goods inventory costs (FGIC), but high backorder costs (BOC). Furthermore,
-the agents based on the order-pool orders, namely \(\ActOrdPool\) and \(\ActOrdPoolDQN\), are able
-achieve lower work-in-process costs (WIPC) and also lower tardiness, than the other the agents
-rewarded directly on the shipped orders.
-
-\begin{table}[t!]
-  \centering
-  \def\arraystretch{1.2}          %  1 is the default, change whatever you need
-  \setlength\tabcolsep{4pt}
-  \begin{tabular}{|l||c|c|c|c|c|c|c|c|c|}
-    % BEGIN RECEIVE ORGTBL res2
+The parameter setup is given in Table~\ref{tbl:params}. The parameters are exponentially decayed
+with rate \(0.05\) and \(350k\) steps\footnote{The decay is defined as \(0.05 ^{(t/350k)}\), where
+  \(t\) is the current period}. We implemented the deep Q-network algorithm proposed by
+\cite{mnih2015human}, which is parameterised by learning rate \(\delta\) and a discount factor of
+\(0.99\).
+
+% \definecolor{lightgray}{rgb}{0.83, 0.83, 0.83}
+% \begin{table}[tb]
+%   \begin{center}
+%     \begin{tabular}{c|ccc|c|c||c}
+%       \hline
+%       70-\(\Exp\)-\(\Unif\) & WIPC & FGIC & BOC & FGIC+BOC & SUM & SFTT & SL(\%) \\
+%       \hline                &
+%        ESSLTA{}             &      &      &     &          &     &      &        \\
+%        ANN{}                &      &      &     &          &     &      &        \\
+%        ESSLTB{}             &      &      &     &          &     &      &        \\
+%        BILD{}               &      &      &     &          &     &      &        \\
+%       \hline                &
+% %
+%       \hline                &
+%       70-\(\Unif\)-\(\Exp\) & WIPC & FGIC & BOC & FGIC+BOC & SUM & SFTT & SL(\%) \\
+%       \hline                &
+%        ESSLTA{}             &      &      &     &          &     &      &        \\
+%        ANN{}                &      &      &     &          &     &      &        \\
+%        ESSLTB{}             &      &      &     &          &     &      &        \\
+%        BILD{}               &      &      &     &          &     &      &        \\
+%       \hline                &
+% %
+%       \hline                &
+%       70-\(\Exp\)-\(\Exp\)  & WIPC & FGIC & BOC & FGIC+BOC & SUM & SFTT & SL(\%) \\
+%       \hline                &
+%        ANN{}                &      &      &     &          &     &      &        \\
+%        ESSLTA{}             &      &      &     &          &     &      &        \\
+%        ESSLTB{}             &      &      &     &          &     &      &        \\
+%        BILD{}               &      &      &     &          &     &      &        \\
+%     \end{tabular}
+%   \end{center}
+%   \caption{\label{tbl:70} Results for a bottleneck utilisation of 70\%. }
+% \end{table}
+
+\begin{table}[tb]
+  \begin{center}
+    \begin{tabular}{c|rrr|c|c||c|c|c|c|c}
       \hline
-    Algorithm/KPI & SUM & BOC & FGIC & WIPC & % \%SL &
-                                                       TARD & \(\sigma\)TARD & SFTT\\
+      80-\(\Exp\)-\(\Unif\)   & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTT+FGIT & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
       \hline
-    % Evaluation 1 & \multicolumn{11}{c|}{} \\
-    \ActShipped    & 1777.74 & 1322.04 & 94.49 & 361.20       & 2.24 & 1.52 & 3.44\\
-    \ActOrdPool    & 1593.03 & 1318.65 &14.01 &260.36         & 2.11 & 1.42 & 2.68\\
-    \ActShippedDQN & 1940.62 & 1570.40 &6.30 &363.92          & 2.35 & 1.59 & 3.95 \\
-    \ActOrdPoolDQN & 1618.88 & 1306.20 & 40.68& 272.00        & 2.08 & 1.38 & 2.58\\
-    \BILOne        & 1736.98 & 1510.42  &  0.00  & 226.56     & 2.04 & 1.26 & 2.16 \\
-    \BILTwo        & 1081.79 & 769.86 & 86.89   & 225.03      & 1.87 & 1.18 & 2.15\\
-    \BILThree      &  \textbf{922.36} & 362.59 &  334.47  & 225.30     & 1.74 & 1.07 & 2.15\\
-    \BILFour       & 1062.68 & 155.72  & 681.03   & 225.93    & 1.63 & 0.99 & 2.15\\
-    \BILFive       & 1380.04 & 65.45  & 1088.94    & 225.64   & 1.67 & 1.02 & 2.15\\
-    \BILSix        & 1777.77 & 25.04 &  1529.77  & 222.96     & 1.61 & 0.85 & 2.13\\
-    \imre          & 2209.97 & 12.08 &1974.72 & 223.16        & 1.58 & 0.63 & 2.14 \\
+       \rl{0.995}             & 55.218 & 137.006 & 52.090   &          & 244.315 & 1.506 & 2.597    & 0.915  & 34.902   & 32.791             \\
+      \rl{1.000}              & 67.265 & 102.180 & 175.686  &          & 345.132 & 1.517 & 2.627    & 0.908  & 49.305   & 30.777             \\
+      \BILOne{}               & 45.902 & 0.000   & 734.431  &          & 780.333 & 1.319 & 1.806    & 0.352  &          &                    \\
+      \BILTwo{}               & 45.734 & 80.313  & 1425.221 &          & 268.569 & 1.318 & 2.158    & 0.876  &          &                    \\
+      \BILThree{}             & 45.861 & 279.723 & 30.818   &          & 356.402 & 1.318 & 3.034    & 0.975  &          &                    \\
       \hline
-    % END RECEIVE ORGTBL res2
+%
+%       \hline
+%       80-\(\Unif\)-\(\Exp\) & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTP+FGI & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
+%       \hline                &
+%        ESSLTA{}             &        &         &          &          &         &       &          &        &          &                    \\
+%        ANN{}                &        &         &          &          &         &       &          &        &          &                    \\
+%        ESSLTB{}             &        &         &          &          &         &       &          &        &          &                    \\
+%        BILD{}               &        &         &          &          &         &       &          &        &          &                    \\
+%       \hline
+% %
+%       \hline
+%       80-\(\Exp\)-\(\Exp\)  & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTP+FGI & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
+%       \hline                &
+%        ANN{}                &        &         &          &          &         &       &          &        &          &                    \\
+%        ESSLTA{}             &        &         &          &          &         &       &          &        &          &                    \\
+%        ESSLTB{}             &        &         &          &          &         &       &          &        &          &                    \\
+%        BILD{}               &        &         &          &          &         &       &          &        &          &                    \\
     \end{tabular}
-        %         \egroup
-
-  \caption{\label{tbl:res}Evaluation results where the monetary units are in \(k\)-values.}
-
-        %         \end{figure}
+  \end{center}
+  \caption{\label{tbl:80} Results for a bottleneck utilisation of 80\%. }
 \end{table}
+
+% \definecolor{lightgray}{rgb}{0.83, 0.83, 0.83}
+% \begin{table}[tb]
+%   \begin{center}
+%     \begin{tabular}{c|ccc|c|c||c}
+%       \hline
+%       80-\(\Exp\)-\(\Unif\)   & WIPC      & FGIC      & BOC       & FGIC+BOC  & SUM   & SL(\%) \\
+%       \hline
+%        ESSLTA{}     &           &           &           &           &       &                  \\
+%        ANN{}        &           &           &           &       &       &                      \\
+%        ESSLTB{}     &           &           &           &       &       &                      \\
+%        BILD{}       &           &           &           &       &       &                      \\
+%       \hline
+% %
+%       \hline
+%       80-\(\Unif\)-\(\Exp\)   & WIPC      & FGIC      & BOC       & FGIC+BOC  & SUM    & SL(\%) \\
+%       \hline
+%        ESSLTA{}     &           &           &           &           &       &      \\
+%        ANN{}        &           &           &           &           &       &      \\
+%        ESSLTB{}     &           &           &           &           &       &      \\
+%        BILD{}       &           &           &           &           &       &      \\
+%       \hline
+% %
+%       \hline
+%       80-\(\Exp\)-\(\Exp\) & WIPC     & FGIC      & BOC       & FGIC+BOC  & SUM   & SL(\%) \\
+%       \hline
+%        ANN{}     &          &           &           &           &       &      \\
+%        ESSLTA{}  &          &           &           &           &       &      \\
+%        ESSLTB{}  &          &           &           &           &       &      \\
+%        BILD{}    &          &           &           &           &       &      \\
+%     \end{tabular}
+%   \end{center}
+%   \caption{\label{tbl:80} Results for a bottleneck utilisation of 70\%. }
+% \end{table}
+
+% Table~\ref{tbl:res} shows the results, where we evaluate the average reward and discounted
+% reinforcement learning algorithm by providing either direct reward feedback, that is the reward
+% accumulated in the next period, without keeping track of any orders, as \(\ActShipped\) and
+% \(\ActShippedDQN\) respectively. Further both algorithms are evaluated with keeping track of the
+% order pool orders, such that each action is responsible for all orders in the order pool. These are
+% labelled \(\ActOrdPool\) and \(\ActOrdPoolDQN\). For comparisons reasons we provide
+% \(\BILOne-\BILSix\) and immediate release (interval release).
+
+% The results show that at the current time we are facing issues with scaling the algorithm to the
+% order release problem. Likely this is due to parameterisation issues, which we are investigating
+% currently. However, it can be seen that i) the technique of keeping track of the order pool orders
+% leads to better policies and ii) that despite the parameterisation issues average reward
+% reinforcement learning finds better solutions than its discounted counterpart. However, the least
+% costs are generated by \(\BILThree\), which shows that there is room for improvement for the
+% reinforcement learning policies.
+
+% Interesting is also to see, that all reinforcement learning variants are releasing rather late,
+% leading to low finished goods inventory costs (FGIC), but high backorder costs (BOC). Furthermore,
+% the agents based on the order-pool orders, namely \(\ActOrdPool\) and \(\ActOrdPoolDQN\), are able
+% achieve lower work-in-process costs (WIPC) and also lower tardiness, than the other the agents
+% rewarded directly on the shipped orders.
+
+% \begin{table}[t!]
+%   \centering
+%   \def\arraystretch{1.2}          %  1 is the default, change whatever you need
+%   \setlength\tabcolsep{4pt}
+%   \begin{tabular}{|l||c|c|c|c|c|c|c|c|c|}
+%     % BEGIN RECEIVE ORGTBL res2
+%     \hline
+%     Algorithm/KPI & SUM & BOC & FGIC & WIPC & % \%SL &
+%                                                        TARD & \(\sigma\)TARD & SFTT\\
+%     \hline
+%     % Evaluation 1 & \multicolumn{11}{c|}{} \\
+%     \ActShipped    & 1777.74 & 1322.04 & 94.49 & 361.20       & 2.24 & 1.52 & 3.44\\
+%     \ActOrdPool    & 1593.03 & 1318.65 &14.01 &260.36         & 2.11 & 1.42 & 2.68\\
+%     \ActShippedDQN & 1940.62 & 1570.40 &6.30 &363.92          & 2.35 & 1.59 & 3.95 \\
+%     \ActOrdPoolDQN & 1618.88 & 1306.20 & 40.68& 272.00        & 2.08 & 1.38 & 2.58\\
+%     \BILOne        & 1736.98 & 1510.42  &  0.00  & 226.56     & 2.04 & 1.26 & 2.16 \\
+%     \BILTwo        & 1081.79 & 769.86 & 86.89   & 225.03      & 1.87 & 1.18 & 2.15\\
+%     \BILThree      &  \textbf{922.36} & 362.59 &  334.47  & 225.30     & 1.74 & 1.07 & 2.15\\
+%     \BILFour       & 1062.68 & 155.72  & 681.03   & 225.93    & 1.63 & 0.99 & 2.15\\
+%     \BILFive       & 1380.04 & 65.45  & 1088.94    & 225.64   & 1.67 & 1.02 & 2.15\\
+%     \BILSix        & 1777.77 & 25.04 &  1529.77  & 222.96     & 1.61 & 0.85 & 2.13\\
+%     \imre          & 2209.97 & 12.08 &1974.72 & 223.16        & 1.58 & 0.63 & 2.14 \\
+%     \hline
+%     % END RECEIVE ORGTBL res2
+%   \end{tabular}
+%         %         \egroup
+
+%   \caption{\label{tbl:res}Evaluation results where the monetary units are in \(k\)-values.}
+
+%         %         \end{figure}
+% \end{table}
 
@@
 
-This paper introduces deep theoretical insights in reinfocement learning and provides first evidence
-that average reward reinfocement learning is better applicable to operations research problems than
-the discounted framework.
-        %         adds to the growing body of evidence that machine learning algorithms can contribute
-        %         positively to a company's performance.
-The paper describes an % successful
-application of an order release model based on reinforcement learning. The performance is tested on
-a multi-product, two-stage hypothetical flow shop and is measured by cost, delivery and lead time
-related measures. We show that in the current version the machine learning approach is not able to
-outperform all other tested order release
-approaches. % By yielding lower \textit{UPDATE: total costs,
-        %         less mean and standard deviation of tardiness and a shorter shop floor throughput time (SFTT)}.
-        %         Despite the good performance of our proposed adaptive order release model based on \textit{UPDATE:
-        %         reinforcement learning} we are aware of its limitations. Firstly, the results are limited to the
-        %         simulated case and the validity of the results for other production systems must be assessed in
-        %         future studies. Secondly, adding further experimental factors would be beneficial like including
-        %         scenarios with machine failures and different scheduling rules. In conclusion we found that our
-        %         proposed adaptive order release approach harbours enormous potential.
-In the future we expect to be able to yield better results by investigating the parameterisation and
-resolving current issues when scaling average reinforcement learning to complex scenarios.
+% This paper introduces deep theoretical insights in reinfocement learning and provides first evidence
+% that average reward reinfocement learning is better applicable to operations research problems than
+% the discounted framework. The paper describes an application of an order release model based on
+% reinforcement learning. The performance is tested on a multi-product, two-stage hypothetical flow
+% shop and is measured by cost, delivery and lead time related measures. We show that in the current
+% version the machine learning approach is not able to outperform all other tested order release
+% approaches. In the future we expect to be able to yield better results by investigating the
+% parameterisation and resolving current issues when scaling average reinforcement learning to complex
+% scenarios.
+
+\MS[t]{todo}
 
@@
 \bibliography{references}
-
-\vfill
-\pagebreak
-\appendix
-
-\section{N-Discount-Optimality}
-\label{sec:methods}
-
-This part proves (or rather ought to prove, as this is only the first version of such a proof)
-Blackwell-optimality of the average reward adjusted reinforcement learning algorithm shown in
-Algorithm~\ref{alg:near}. The idea is to prove \(n\)-discount optimality for \(n=-1\), \(n=0\), and
-finally for \(n \geqslant 1\). But first recall the definition of \(n\)-discount-optimality.
-
-\begin{definition*}
-  Due to \cite{Veinott69} for MDPs a policy \(\polopt\) is \emph{\(n\)-discount-optimal} for
-  \(n=-1,0,1,\ldots\) for all states \(s \in \States\) with discount factor
-  \(\gamma\) % and \(V_{\gamma}^{\pol}(s)\) being the value function as defined above
-        %
-  if and only if
-  \begin{align*}
-    \lim_{\gamma \to 1}(1-\gamma)^{-n}\ (V_{\gamma}^{\polopt}(s) - V_{\gamma}^{\pol}(s)) \geqslant 0 \tpkt
-  \end{align*}
-\end{definition*}
-
-\subsection{\(\mathbf{(-1)}\)-Discount-Optimality\cite{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}.}
-
-Here we provide insights in the algorithm and
-\(n\)-discount-optimality, where we concentrate on \(n=-1\). Recall that the discounted state value
-can be expanded by the Laurent series expansion:
-\(V_{\gamma}^{\pol} = \frac{\avgrew^{\pol}(s)}{1-\gamma} + V^{\pol} + e_{\gamma}^{\pol}(s)\), where
-\(\lim_{\gamma \to 1} e_{\gamma}^{\pol}(s) = 0\)
-
-\begin{align*}
-  \lim_{\gamma \to 1} (1-\gamma)^{1} \cdot (V_{\gamma}^{\polopt}(s) - V_{\gamma}^{\pol}(s)) \geqslant 0\\
-  \lim_{\gamma \to 1} (1-\gamma) \cdot (\frac{\avgrew^{\polopt}(s) - \avgrew^{\pol}(s)}{1-\gamma} + V^{\polopt}(s) - V^{\pol}(s) + e_{\gamma}^{\polopt}(s) - e_{\gamma}^{\pol}(s)) \geqslant 0\\
-  \lim_{\gamma \to 1} (\avgrew^{\polopt}(s) - \avgrew^{\pol}(s) + (1-\gamma) \cdot (V^{\polopt}(s) - V^{\pol}(s) + e_{\gamma}^{\polopt}(s) - e_{\gamma}^{\pol}(s)) \geqslant 0\\
-  \avgrew^{\polopt}(s) - \avgrew^{\pol}(s) \geqslant 0\\
-\end{align*}
-
-That is a policy \(\pol\) which ought to be (\(-1\))-discount-optimal has to maximise the average reward
-\(\rho^{\pol}(s)\) over all states \(s\).
-
-\subsection{\(\mathbf{0}\)-Discount-Optimality\cite{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}.}
-
-Note that a policy can only be \(0\)-discount-optimal if it is \(-1\)-discount-optimal, thus
-\(\avgrew^{\polopt}(s)=\avgrew^{\pol}(s)\). Furthermore, recall that
-\(\lim_{\gamma \to 1} e_{\gamma}^{\pol}(s) = 0\).
-
-\begin{align*}
-  \lim_{\gamma \to 1} (1-\gamma)^{0} \cdot (V_{\gamma}^{\polopt}(s) - V_{\gamma}^{\pol}(s)) \geqslant 0\\
-  \lim_{\gamma \to 1} (\frac{\avgrew^{\polopt}(s) - \avgrew^{\pol}(s)}{1-\gamma} + V^{\polopt}(s) - V^{\pol}(s) + e_{\gamma}^{\polopt}(s) - e_{\gamma}^{\pol}(s)) \geqslant 0\\
-  \lim_{\gamma \to 1} (\frac{0}{1-\gamma} + V^{\polopt}(s) - V^{\pol}(s)) \geqslant 0\\
-  V^{\polopt}(s) - V^{\pol}(s) \geqslant 0\\
-\end{align*}
-
-Therefore a \(0\)-discount-optimal policy \(\pol\) has to maximise the bias values
-\(V^{\pol}(s)\) for all states \(s\).
-
-\vfill
-\pagebreak
-
-\subsection{\(\mathbf{n}\)-Discount-Optimality for \(\mathbf{n \geqslant 1}\).}
-
-The following derivation provides an important insight to the analysis. Recall that for unichain
-MDPs the average reward \(\avgrew^{\pol}(s)\) for all states \(s\) is equal and stated as
-\(\avgrew^{\pol}\). Furthermore, as the policy is \(0\)-discount-optimal
-\(V^{\polopt}(s) = V^{\pol}(s)\).
-
-\begin{align*}
-  \lim_{\gamma \to 1} (1-\gamma)^{-n} \cdot (V_{\gamma}^{\polopt}(s) - V_{\gamma}^{\pol}(s)) & \geqslant 0\\
-  \lim_{\gamma \to 1} (1-\gamma)^{-n} \cdot ( \frac{\avgrew^{\polopt}(s)}{1-\gamma} + V^{\polopt}(s) + e_{\gamma}^{\polopt}(s)  %& \\
-  - \frac{\avgrew^{\pol}(s)}{1-\gamma} - V^{\pol}(s) - e_{\gamma}^{\pol}(s) ) & \geqslant 0\\
-  \lim_{\gamma \to 1} (1-\gamma)^{-n} \cdot ( e_{\gamma}^{\polopt}(s) - e_{\gamma}^{\pol}(s) ) & \geqslant 0\\
-\end{align*}
-
-As stated before \(e_{\gamma}^{\polopt}(s)\) approaches 0 as \(\gamma \to 1\). Therefore, it is
-important to note that for \(n>0\) we analyse the case when \(\gamma\) is strictly less than \(1\). This
-means that the number of actions to reach a desired goal is taken into account.
-        %         The ratio of long-term to short-term is provided by the discount factor \(\gamma\).
-Therefore, a \(\infty\)-discount-optimal policy \(\pol\) has to maximise the error term
-$e_{\gamma}^{\pol}(s)$ for all states \(s\). That is, an \(\infty\)-discount-optimal algorithm has to
-choose the action with the largest error term \(e_{\gamma}^{\pol}(s)\) once \(\gamma \to 1\).
-However, as the error term depends on infinitely many sub-terms simply estimating these and summing
-up does not work.
-
-\begin{figure*}[t]
-  \centering
-  \begin{tikzpicture}[scale=8]
-    \draw[->] (0,0) -- (1.0,0) node[right] {$\gamma$};
-    \draw[->] (0,0) -- (0,0.8) node[above] {$\mathbb{R}$};
-    \draw[scale=1,domain=0.001:0.85,smooth,variable=\x,blue] plot ({\x},{0.1/(1-\x)+0.2});
-    \draw[scale=1,domain=0.001:0.865,smooth,variable=\x,red] plot ({\x},{0.1/(1-\x)+0.2-0.2*(1-\x)^0.5});
-
-        %         Labels
-    \draw[blue] (0.6,0.64) node () {$\frac{\avgrew^{\pol}(s)}{1-\gamma} + V^{\pol}(s)$};
-    \draw[red] (0.875,0.58) node () {$V_{\gamma}^{\pol}(s)$};
-    \draw (-0.015,0.3) -- (0.015,0.3);
-    \draw (-0.2, 0.3) node[align=left] () {$\avgrew^{\pol}(s) + V^{\pol}(s)$};
-
-    \draw (-0.015,0.1) -- (0.015,0.1);
-    \draw (-0.111, 0.1) node[] () {$R^{\pol}(s)$};
-
-    \draw (0.9,-0.015) -- (0.9,0.015) node[below, yshift=-0.2cm] () {$1$};
-    \draw (0.75,-0.015) -- (0.75,0.015) node[below, yshift=-0.2cm] () {$\gamma_{1}$};
-    \draw (0.20,-0.015) -- (0.20,0.015) node[below, yshift=-0.2cm] () {$\gamma_{0}$};
-
-    \draw (0.5,0.4) -- (0.5,0.258578644) node[midway,left, yshift=-0.2cm] () {$e_{\gamma}^{\pol}(s)$};
-
-    \draw (0.2,0.146114562) -- (0.75,0.146114562) node[midway, below] () {$\gamma_{1}-\gamma_{0}$};
-    \draw (0.75,0.146114562) -- (0.75,0.5) node[midway,right] () {$V_{\gamma_{1}}^{\pol}(s)-V_{\gamma_{0}}^{\pol}(s)$};
-
-  \end{tikzpicture}
-  \caption{Visualisation of the monotonically increasing state values for a state $s$ and the
-    corresponding decreasing error term $e_{\gamma}^{\pol}(s)$ as $\gamma$ approaches 1. }
-  \label{fig:e}
-\end{figure*}
-
-Therefore, the idea is to use two discounted state-values with different $\gamma$'s to infer the
-slope of the error term. Figure~\ref{fig:e} illustrates the idea. Here $\gamma_{1}$ and $\gamma_{0}$
-are used to estimate the difference of \(V_{\gamma_{1}}^{\pol}(s)\) and
-\(V_{\gamma_{0}}^{\pol}(s)\). As the average reward and the bias values are equal (recall
-\(0\)-discount-optimality) the difference is a direct estimate of the slope of the error term. Thus
-the larger the increase of the difference of the discounted state values when comparing
-\(0\)-discount-optimal actions the better the action. However, for this to hold true we have to
-prove strict monotonicity of
-\(V_{\gamma}^{\pol}(s) = \frac{\avgrew^{\pol}(s)}{1-\gamma} + V^{\pol}(s) + e_{\gamma}^{\pol}(s)\)
-under increasing $\gamma$. In the sequel we assume \(\avgrew^{\pol}(s) \geqslant 0\). Note that this
-assumption is easily met by adding a constant reward for any action in cases where
-\(\avgrew^{\pol}(s) < 0\). This assumption eases the following analysis and makes sense as usually
-we are interested in problems which accumulate positive average reward, e.g. profit or points, over
-time.
-
-\begin{remark}
-  Nonetheless, it is an important insight that if \(\avgrew^{\pol}(s) < 0\) then
-  \(\frac{\avgrew^{\pol}(s)}{1-\gamma} + V^{\pol}(s)\) approaches negative infinity (as opposed to
-  positive infinity, cf. Figure~\ref{fig:e}). Therefore, in such cases the following optimisation
-  objectives would need to be inverted.
-\end{remark}
-
-\begin{lemma}
-  \label{lem:err}
-  The error term
-  \(e_{\gamma}^{\pol}(s) = \sum_{m=1}^{\infty}(\frac{1-\gamma}{\gamma})^{m} \cdot y_{m}\) strictly
-  monotonically decreases as $\gamma$ increases.
-\end{lemma}
-
-\begin{proof}
-        %         S[t]{Is the following formula (indices) correct?}
-  Note that \(e_{\gamma}^{\pol}(s) = \sum_{m=1}^{\infty}(\frac{1-\gamma}{\gamma})^{m} \cdot y_{m}\),
-  where the sign of \(y_{m}\) alternates in each step. This is due to the fact that
-  \(y_{m} = \E_{\pol} [ y_{m} ] - y_{m-1}\) for all \(m \geqslant 1\)~\cite[p.346]{Puterman94}, that
-  is \(y_{m}\) only depends on \(y_{m-1}\) and the underlying (stationary) policy. Further \(y_{0}\)
-  is the bias value \(V^{\pol}(s)\).
-        %
-  We consider three cases:
-  \begin{enumerate}
-  \item[1)]
-    \begin{minipage}[t]{0.17\textwidth}
-      \(\gamma = 0.5\):
-    \end{minipage}
-    The factor \((\frac{1-\gamma}{\gamma})^{m}\) is $1$.
-  \item[2)]
-    \begin{minipage}[t]{0.17\textwidth}
-      \(0.5 < \gamma < 1\):
-    \end{minipage}
-    The factor \((\frac{1-\gamma}{\gamma})^{m}\) is $<1$ and strictly monotonically converging towards $0$.
-  \item[3)]
-    \begin{minipage}[t]{0.17\textwidth}
-      \(0 < \gamma < 0.5\):
-    \end{minipage}
-    The factor \((\frac{1-\gamma}{\gamma})^{m}\) is $>1$ and strictly monotonically converging towards $\infty$.
-  \end{enumerate}
-
-  \paragraph{Case 1)}
-  { As \(\frac{1-\gamma}{\gamma} = 1 \) the term reduces to
-    \(e_{\gamma}^{\pol}(s) = \sum_{m=1}^{\infty} y_{m}^{\pol}\). We use the definition
-        %
-    \begin{align*}
-      y_{m}^{\pol}=(-1)^{m} H_{\pol}^{m+1} h_{\pol} \tcom
-    \end{align*}
-    %
-    where \(H_{\pol}\) is the deviation matrix defined as
-    %
-    \begin{align*}
-      H_{\pol} = \clim_{N \to \infty} \sum_{t=0}^{N-1}{P^{t}-P^{\star}} \tcom
-    \end{align*}
-    \(\clim\) the Cesaro limit\footnote{Recall that the Cesaro limit (with degree \(1\)) is only
-      required in periodic MDPs to ensure stationary matrices. For aperiodic MDPs they can be read
-      as normal limits.~\cite[p.592ff]{Puterman94}}, \(P^{\star}\) the stationary matrix defined as
-    \(P^{\star} = \clim_{N \to \infty}P^{N}\) and \(h_{\pol}\) a column vector holding the bias values
-    of the states \cite[]{Puterman94,MillerVeinott1969}. Thus, we get
-        %
-    \begin{align*}
-      e_{\gamma}^{\pol}(s) & = \sum_{m=1}^{\infty}y_{m}^{\pol}(s) = \sum_{m=1}^{\infty}{y_{2m}^{\pol}(s) + y_{2m+1}^{\pol}(s)}
-                             = \sum_{m=1}^{\infty} H^{2m+2} h - H^{2m+1} h \\
-                           & = \sum_{m=1}^{\infty} (H^{2m+2} - H^{2m+1}) h \tpkt
-    \end{align*}
-
-    % Wlog. assume all entries in \(H\) are positive. Therefore,
-    It suffices\footnote{This claim has not been proven yet!} to show that $H^{2m+2} - H^{2m+1} < 0$
-    for all $m \geqslant 1$ to prove strict monotonicity of $e_{\gamma}^{\pi}(s)$.
-    % \MS[t]{why should that prove monotonicity? Rethink!!!}
-    However, this directly follows as for MDPs with at least two states \(H-1 < 0\) in
-    %
-    \begin{align*}
-      e_{\gamma}^{\pol}(s) = \sum_{m=1}^{\infty} (H^{2m+2} - H^{2m+1}) h = \sum_{m=1}^{\infty} H^{2m+1} (H - 1) h \tpkt
-    \end{align*}
-    %
-  }
-
-  \paragraph{Case 2)}{ As the factor \((\frac{1-\gamma}{\gamma})^{m}\) is $<1$ and strictly
-    monotonically decreasing for an increasing $m \geqslant 1$, we use the same argument as in the
-    previous case (Case 1) with which the claim follows. }
-
-  \paragraph{Case 3)}{ This part of the proof is open!
-
-    % Recall the Laurent Series expansion only holds when \(0 <
-    % \frac{1-\gamma}{\gamma} < \norm{H_{\pi}}^{-1}\),
-    % where for a finite matrix \(C = (c_{ij})\) the norm \(\norm{C}\) is defined as
-    % \(\norm{C} \coloneqq \max_{i} \sum_{j} |c_{ij}|\).
-
-    % \begin{align*}
-    %   e_{\gamma}^{\pol}(s) & = \sum_{m=1}^{\infty}y_{m}^{\pol}(s) = \sum_{m=1}^{\infty}{y_{2m}^{\pol}(s) + y_{2m+1}^{\pol}(s)}\\
-    %   & = \sum_{m=1}^{\infty} (\frac{1-\gamma}{\gamma})^{2m+1} H^{2m+2} h -
-    %   (\frac{1-\gamma}{\gamma})^{2m}H^{2m+1} h \\
-    %   & = \sum_{m=1}^{\infty} (\frac{1-\gamma}{\gamma})^{2m} H^{2m+1}((\frac{1-\gamma}{\gamma}) H - 1) h \tpkt
-    % \end{align*}
-
-          %           \MS{No! This does not prove that under two different gammas $e_{\gamma_{1}}^{\pol}(s) -
-          %           e_{\gamma_{0}}^{\pol}(s)$ is decreasing}
-
-  }
-
-          %           We now investigate \(e^{\pol}(s) = \sum_{m=1}^{\infty}y_{m}^{\pol}(s)\) to infer properties of
-          %           \(e_{\gamma}^{\pol}(s)\). Observe that we removed the factor $(\frac{1-\gamma}{\gamma})^{m}$
-          %           depending on a concrete \(\gamma\). There are three cases to distinguish for this factor under
-          %           increasing $m$:
-
-          %           In all three cases the factor does neither change the monotonicity properties nor the sign of
-          %           $y_{m}$ in the term
-          %           \(e_{\gamma}^{\pol}(s) = \sum_{m=1}^{\infty}(\frac{1-\gamma}{\gamma})^{m} \cdot y_{m}\).
-          %           Therefore, wlog. suppose $\gamma = 0.5$ and thus \(e^{\pol}(s) = e_{\gamma}^{\pol}(s)\).
-
-          %           This factor is positive and strictly and monotonically
-          %           converging towards $0$ (resp. $\infty$) as $m$ increases and as long as \(0.5 < \gamma < 1\)
-          %           (resp. $0 < \gamma < 0.5)$. Note that we are usually interested in $\gamma$ values close to $1$,
-          %           as $\gamma \to 1$. \MS[t]{what happens when $\gamma < 0.5$?}
-          %
-\end{proof}
-
-\begin{corollary}
-  \label{cor:diff}
-  For any two \(0\)-discount-optimal stationary policies $\pol_{1}, \pol_{2}$ with stationary
-  transition matrices \(P_{1}\) and \(P_{2}\) respectively, the differences of the slopes of the
-  error terms \(e_{\gamma}^{\pol_{1}}(s)\) and \(e_{\gamma}^{\pol_{2}}(s)\) for any \(\gamma\) is
-  dependent on \(H_{\pol_{1}}\) and \(H_{\pol_{2}}\) and thus \(P_{1}\) and \(P_{2}\) solely.
-\end{corollary}
-
-\begin{proof}
-  Note that due to \(0\)-discount-optimality the bias values \(h_{1} = h_{2}\) are equal for both
-  policies. Therefore as for any given \(H\) the error term is determined by
-  \(e_{\gamma}^{\pol}(s) = \sum_{m=1}^{\infty} ((\frac{1-\gamma}{\gamma})^{2m + 1} H^{2m+2} -
-  (\frac{1-\gamma}{\gamma})^{2m} H^{2m+1}) h = \sum_{m=1}^{\infty} (\frac{1-\gamma}{\gamma})^{2m}
-  H^{2m+1}( \frac{1-\gamma}{\gamma} H - 1) h \) the claim follows directly.
-\end{proof}
-
-\begin{corollary}
-  \label{cor:cont}
-  The error term function \(e_{\gamma}^{\pol}(s)\) for \(\gamma \to 1\) and for any given policy
-  \(\pol\) is continuous.
-\end{corollary}
-
-\begin{proof}
-  The claim follows directly from Corollary~\ref{cor:diff} and due to the shape of the term
-  \(e_{\gamma}^{\pol}(s) = \sum_{m=1}^{\infty} (\frac{1-\gamma}{\gamma})^{m} H^{m+1} h\).
-\end{proof}
-
-\begin{theorem}
-  If \(\avgrew^{\pol}(s) > 0\) then
-  \(V_{\gamma}^{\pol}(s) = \frac{\avgrew^{\pol}(s)}{1-\gamma} + V^{\pol}(s) + e_{\gamma}^{\pol}(s)\) is
-  strictly monotonically increasing as $\gamma$ approaches \(1\).
-\end{theorem}
-
-\begin{proof}
-  Due to \(\avgrew^{\pol}(s) > 0\) and \(0 < \gamma < 1\) the term
-  \(\frac{\avgrew^{\pol}(s)}{1-\gamma}\) strictly monotonically approaches infinity as
-  $\gamma \to 1$. Furthermore, by Lemma~\ref{lem:err} and as
-  \(\lim_{\gamma \to 1} e_{\gamma}^{\pol}(s) = 0\) the theorem follows immediately.
-\end{proof}
-
-\begin{theorem}
-  \label{thm:minmax}
-  If \(\avgrew^{\pol}(s) \geqslant 0\) and for \(\gamma\)-values \(\gamma_{0}, \gamma_{1}\) with
-  \(0.5 \leqslant \gamma_{0} < \gamma_{1} < 1\) a Blackwell-optimal agent chooses the action \(a\)
-  with expected future state \(s'\) that maximises the expected\footnote{To ease readability and as
-    we aim for model-free RL we have dropped the expectations in the formulas. } discounted state
-  value difference
-  \(\Delta V_{\gamma}^{\pol}(s') = V_{\gamma_{0}}^{\pol}(s') - V_{\gamma_{1}}^{\pol}(s')\) of the
-  set of \(0\)-discount-optimal actions available and for which
-  \begin{itemize}
-  \item[a)] \(\frac{\avgrew}{1-\gamma_{0}} + V^{\pol}(s') < V_{\gamma_{0}}^{\pol}(s')\) holds, or if
-    no such actions exists, then for which
-  \item[b)]
-    \(\frac{\avgrew}{1-\gamma_{0}} + V^{\pol}(s') \geqslant V_{\gamma_{0}}^{\pol}(s')\) hold.
-  \end{itemize}
-
-  % the discounted state value difference \(V_{\gamma_{1}}^{\pol}(s') - V_{\gamma_{0}}^{\pol}(s')\),
-  % where the decision to maximise or minimise the difference is based on the value that minimises
-  % \(V_{\gamma_{1}}^{\pol}(s') - V_{\gamma_{0}}^{\pol}(s')\).
-  %         %
-\end{theorem}
-
-\begin{remark}
-  In Theorem~\ref{thm:minmax} we have chosen to base the decision on the values generated with
-  \(\gamma_{0}\) as the error term is greater with smaller \(\gamma\)-values.
-\end{remark}
-
-\begin{proof}
-  First observe that due to strict monotonicity of \(\frac{\avgrew^{\pol}(s')}{1-\gamma}\) and
-  \(e_{\gamma}^{\pol}(s')\) and as \(\lim_{\gamma \to 1} e_{\gamma}^{\pol}(s') = 0\) the discounted
-  state values \(V_{\gamma}^{\pol}(s')\) and the sum of average reward values
-  \(\frac{\avgrew^{\pol}(s')}{1-\gamma} + V^{\pol}(s')\) converge with increasing \(\gamma\).
-          %
-  Thus for any state \(s'\) either \(e_{\gamma}^{\pol}(s') > 0\) or \(e_{\gamma}^{\pol}(s') < 0\)
-  for all \(\gamma\)-values. That is \(V_{\gamma}^{\pol}(s')\) converges to
-  \(\frac{\avgrew^{\pol}(s)}{1-\gamma} + V^{\pol}(s')\) for \(\gamma \to 1\) from above or below.
-  Clearly just considering the slope of the error term is insufficient as for any states
-  \(s_{1}, s_{2}\) it may happen to be that \(e_{\gamma}^{\pol}(s_{1}) > 0\) while
-  \(e_{\gamma}^{\pol}(s_{2}) < 0\). Thus we split the decision in cases, where in case \(a)\) we
-  consider all states \(s_{1}\) with \(e_{\gamma}^{\pol}(s_{1}) > 0\) and in case \(b)\) we states
-  \(s_{2}\) with \(e_{\gamma}^{\pol}(s_{2}) < 0\). Note that
-  \(e_{\gamma}^{\pol}(s_{1}) > e_{\gamma}^{\pol}(s_{2})\) which explains the prioritisation of case
-  \(a)\) over case \(b)\).
-
-  Within these two sets of states we investigate the slopes of the error term.
-
-  \MS[t]{Todo: Show that two points are sufficient for a proxy of the slope.}
-  Thus the slope can be approximated by
-  \begin{align*}
-    \frac{\Delta V_{\gamma}^{\pol}(s)}{\Delta \gamma} = \frac{e_{\gamma_{1}}(s) - e_{\gamma_0}(s)}{\gamma_{1} - \gamma_{0}} \tpkt
-  \end{align*}
-
-  As \(\gamma_{1}(s)-\gamma_{0}(s)\) is constant the difference of the error terms
-  \(e_{\gamma_{1}}(s) - e_{\gamma_0}(s)\) is sufficient to compare the slopes of the error term and
-  can be computed by
-  \begin{align}
-    \label{eq:errorAvgRew}
-    e_{\gamma_{1}}(s) - e_{\gamma_0}(s) = \frac{\avgrew^{\pol}(s)}{1-\gamma_{1}} -
-    \frac{\avgrew^{\pol}(s)}{1-\gamma_{0}} - V_{\gamma_{1}}^{\pol}(s) + V_{\gamma_{0}}^{\pol}(s)\tpkt
-  \end{align}
-
-  However, as all possible future states are gain-optimal \(\avgrew^{\pol}(s)\) is equal for all
-  states \(s\) under consideration. Therefore, the difference \(V_{\gamma_{0}}^{\pol}(s) -
-  V_{\gamma_{1}}^{\pol}(s)\) suffices for comparison purposes between any two states \(s_{1}, s_{2}\):
-  \begin{align*}
-  e_{\gamma_{1}}(s_{1}) - e_{\gamma_0}(s_{1}) - e_{\gamma_{1}}(s_{2}) + e_{\gamma_0}(s_{2}) =
-    V_{\gamma_{0}}^{\pol}(s_1) - V_{\gamma_{1}}^{\pol}(s_1) - V_{\gamma_{0}}^{\pol}(s_2) + V_{\gamma_{1}}^{\pol}(s_2)\tpkt\\
-  \end{align*}
-
-  Then due to Corollaries~\ref{cor:diff}~and~\ref{cor:cont} we conclude that
-  \(\lim_{\gamma \to 1}(1-\gamma)^{-n}\ (V_{\gamma}^{\polopt}(s) - V_{\gamma}^{\pol}(s)) \geqslant
-  0\), where \(\gamma\) is the discount factor and \(V_{\gamma}^{\pol}(s)\) the value function.
-
-\end{proof}
-
-\begin{remark}
-  Note that this means a Blackwell-optimal agent collects the rewards as soon as possible, as it
-  maximises the error term \(e_{\gamma}^{\pol}(s)\).
-
-  % In the case of model-free reinforcement learning, in which rather state-action pairs as opposed to
-  % states are assessed, Theorem~\ref{thm:minmax} does not hold. Observe that in the model-free case
-  % the agent has to choose between MDP states, for instance \((1,l)\) or \((1,r)\) in
-  % Figure~\ref{fig:three-states2}.
-  % Thus, although the gain and
-  % bias values are equal,
-  % Thus the error terms of different actions \(a, a'\) may cause
-  % \(e_{\gamma}^{\pol}(s,a) > 0\) and \(e_{\gamma}^{\pol}(s,a') < 0\).
-
-  % This is due to the fact that by observing one state for each action multiple the error
-
-  % states defined
-
-  % chooses the among different states, cf. Figure~\ref{fig:three-states2}.
-
-  % chooses action based on
-  % a state \(s\) state-action pairs are assessed. Thus, multiple values could be available to decide
-  % on whether to maximise or minimise the discounted state value difference. The correct approach is
-  % to decide on behalf of the action \(a\) which minimises
-  % \(\mid V^{\pol}_{\gamma_{0}}(s,a) - \frac{\avgrew^{\pol}}{1-\gamma_{0}} - V^{\pol}(s,a) \mid\) as
-  % for all actions the error term \(e_{\gamma}^{\pol}(s)\) is converging towards \(0\) with the same
-  % rate (due to the same discount factor \(\gamma_{0}\)).
-\end{remark}
-
-Thus under the assumption of correct approximations and by strict monotonicity of $e_{\gamma}^{\pi}$
-we conclude Blackwell optimality of the algorithm. Figure~\ref{fig:e} visualises the discounted
-state values and the idea of estimating the slope of $e_{\gamma}^{\pi}(s)$ in comparison to
-$\frac{\avgrew^{\pol}(s)}{1-\gamma} + V^{\pol}(s)$.
-
-\subsection{Refinements}
-\label{subsec:Refinements}
-
-The main limitation of the straightforward approach of Theorem~\ref{thm:minmax} is that it requires
-accurate estimations of the bias values when splitting the actions into the two sets with increasing
-and deceasing error term value. However, to be able to estimate \(V^{\pol}(s)\) accurately enough
-one usually has to go through a tedious process of investigating many different parameterisations,
-which is a cumbersome task, especially if the correct values are not known. Furthermore, the process
-of finding the correct values is computationally very expensive in comparison to finding the
-deviations between any two values.
-
-Therefore, we have developed another approach of selecting the best possible action which does not
-depend on correct bias values. In this approach we utilise Equation~\ref{eq:errorAvgRew} which
-allows a determination of the slope simply by rectifying the additional slope imposed due to the
-term \(\frac{\avgrew^{\pol}(s)}{1-\gamma}\). As \(V_{\gamma}^{\pol}(s)\) and \(\avgrew^{\pol}(s)\)
-are computationally cheap they can be usually accurately computed.
-
-Therefore in the implementation we propose to replace Theorem~\ref{thm:minmax} by following the
-following one.
-
-\begin{theorem}\label{thm:err}
-  If \(\avgrew^{\pol}(s) \geqslant 0\) and for \(\gamma\)-values \(\gamma_{0}, \gamma_{1}\) with
-  \(0.5 \leqslant \gamma_{0} < \gamma_{1} < 1\) a Blackwell-optimal agent chooses the action \(a\)
-  with expected future state \(s'\) that maximises the expected error term difference
-  \(\Delta e_{\gamma_{1},\gamma_{0}}^{\pol}(s') = V_{\gamma_{1}}^{\pol}(s') -
-  V_{\gamma_{0}}^{\pol}(s') - \avgrew^{\pol}(s') (\frac{1}{1-\gamma_{1}} - \frac{1}{1-\gamma_{0}})\)
-  of the set of \(0\)-discount-optimal actions available and for which
-  \begin{itemize}
-  \item[a)] \(\Delta e_{\gamma_{1},\gamma_{0}}^{\pol}(s') < 0\) holds, or if
-    no such actions exists, then for which
-  \item[b)]
-    \(\Delta e_{\gamma_{1},\gamma_{0}}^{\pol}(s') \geqslant 0\) hold.
-  \end{itemize}
-\end{theorem}
-
-\begin{proof}
-  We have for any state \(s'\) either \(e_{\gamma}^{\pol}(s') > 0\) or \(e_{\gamma}^{\pol}(s') < 0\)
-  for all \(\gamma\)-values (see proof of Theorem~\ref{thm:minmax}). Therefore, by definition states
-  \(s'\) with \(e_{\gamma_{1},\gamma_{0}}^{\pol}(s') < 0\) have a decreasing error term value. As
-  \(e_{\gamma_{1},\gamma_{0}}^{\pol}(s')\) is a direct measure for the slope of the error term, and
-  by the characteristics on the term specified in Corollaries~\ref{cor:diff} and~\ref{cor:cont}, we
-  can conclude that the maximum expected error term difference for negative values of
-  \(e_{\gamma_{1},\gamma_{0}}^{\pol}(s')\) impose the greatest \(V_{\gamma}^{\pol}(s')\) when
-  \(\gamma \to 1\). The proof for \(e_{\gamma_{1},\gamma_{0}}^{\pol}(s') \geqslant 0\) follows the
-  same argument. Therefore, the Theorem provides a way to find the policy \(\polopt\) with
-  \(\lim_{\gamma \to 1} (1-\gamma)^{-n} \cdot ( e_{\gamma}^{\polopt}(s') - e_{\gamma}^{\pol}(s')) >
-  0\) for any other policy \(\pol\).\qed
-\end{proof}
-
-\section{Bellman Optimality Formulas and Derivations (Incomplete)}
-\label{sec:bellman_optimality}
-
-\paragraph{Bellman optimality equations for the average reward.}
-
-\begin{align*}
-  V^{\star}(s) & = \max_{a \in A(s)} R^{\polopt}(s,a) \\
-               & = \max_{a \in A(s)} \E_{\polopt} [ \sum_{t=0}^{\infty}(R_{t}(s) - \avgrew^{\polopt}(s)) \mid s=s_{t}, a_{t} = a] \\
-               & = \max_{a \in A(s)} \E_{\polopt} [ R_{0}(s) - \avgrew^{\polopt}(s) + \sum_{t=0}^{\infty}(R_{t}(s_{t+1}) - \avgrew^{\polopt}(s_{t+1}))  \mid s=s_{t}, a_{t} = a] \\
-               & = \max_{a \in A(s)} \E [ R_{0}(s) - \avgrew^{\polopt}(s) + V^{\star}(s_{t+1})  \mid s=s_{t}, a_{t} = a ] \\
-               & = \max_{a \in A(s)} (r(s,a) - \avgrew^{\polopt}(s) + \sum_{s_{t+1}} p(s_{t+1}|s,a) \cdot V^{\star}(s_{t+1}))
-\end{align*}
-
-The value of an action in a given state must equal the expected return for following an optimal
-policy from that state:
-
-\begin{align*}
-  R^{\star}(s,a) & = \E [ r(s,a) - \avgrew^{\polopt}(s) + V^{\star}(s_{t+1})  \mid s=s_{t},a=a_{t} ]\\
-                 & = r(s,a) - \avgrew^{\polopt}(s) + \E [ \max_{a_{t+1} \in A(s_{t+1})}R^{\star}(s_{t+1},a_{t+1}) \mid s=s_{t},a=a_{t}]\\
-                 & = r(s,a) - \avgrew^{\polopt}(s) + \sum_{s_{t+1}} p(s_{t+1} | s, a) (\max_{a_{t+1} \in A(s_{t+1})} R^{\star}(s_{t+1},a_{t+1}))
-\end{align*}
-
-\paragraph*{Bellman optimal equations for the bias value.}
-
-\begin{align*}
-  W^{\star}(s) & = \E [ R_{t}(s) + V^{\star}(s_{t+1}) | s = s_{t}, a=a_{t} ]\\
-               & = \E [ R_{t}(s) + \max_{a_{t+1} \in A(s_{t+1})} R^{\star}(s_{t+1},a_{t+1}) | s = s_{t} , a=a_{t}]\\
-\end{align*}
-
-\begin{align*}
-  W^{\star}(s,a) & = \E [ r(s,a) + V^{\star}(s_{t+1}) \mid s=s_{t}, a=a_{t} ] \\
-                 & = r(s,a) + \E [  \max_{a_{t+1} \in A(s_{t+1})} R^{\star}(s_{t+1},a_{t+1}) \mid s=s_{t}, a=a_{t} ]\\
-                 & = r(s,a) + \sum_{s_{t+1}} p(s_{t+1} | s, a) (\max_{a_{t+1} \in A(s_{t+1})} R^{\star}(s_{t+1},a_{t+1}))
-\end{align*}
-
-Difference \(R_{n+1}\) to \(W_{n+1}\)?
 

### 2021-01-13 paper/paper.tex
ANN setup

@@
 %
-\MS[t]{TODO}
-We use the same setup for the neural network as done by \cite{Schneckenreither2019}. However as they
-use a actor-critic agent, whereas we approximate state-action-values the output of the network is a
-one dimensional vector as opposed to a matrix. Therefore we use a single three layer fully-connected
-network where the number of nodes and activation functions are 41-ReLU\footnote{ReLU stands for
-  rectified linear unit.}-89-ReLU-20-ReLU-9, with the output activations being Tanh (hyperbolic
-tangent function).
-        %
-The output for both networks consist of $3^{2}=9$ nodes. This is due to the fact that all
-combinations of increase, decrease and no change of the lead time for each product type have to be
-represented.
-        %
-For back-propagation we use the Adam optimiser with learning rate \(0.001\), and beta-values
-\(\beta_{1}=0.9\), and \(\beta_{2}=0.999\).
+Due to the vast possibilities the neural network architecture was manually optimised, starting with
+less stochastic systems and aiming for a coffin shape that can be used in all setups. During this
+process we constantly adapted and reevaluated the best network on various production system
+configurations. The resulting architecture depends on the number of features \(\anninp\) and the
+number of output nodes \(\annout\), which itself depends on the number of actions
+\(\acts = \size{\States}\), the independent agent configuration, and the number of functions that
+are approximated. For the later, in case of the proposed algorithm this is \(2\),
+\(\X^\pol_{\gamma_{0}}\) and \(\X^\pol_{\gamma_{1}}\), while deep Q-Learning only approximates the
+function \(Q_{\gamma}^\pol\).
+%
+The resulting feedforward deep ANN archtiecture is given as %
+\(\anninp %
+\rightarrow 3 \anninp % \rightarrow \relu
+\rightarrow \anninp % \rightarrow \relu
+\rightarrow 2 \annout % \rightarrow \relu
+\rightarrow \annout %
+% \rightarrow   \leakyTanh%
+\), where each arrow symbolises a fully-connected layer with a rectified linear unit (\(\relu\)) as
+activation function. As output activation function we use the hyperbolic tangent function
+(\(\Tanh\)), but similar to leaky \(\relu\), leak values greater than \(0.98\) (less than \(-0.98\))
+with \(5\%\). The idea is that we do not want to restrict the approximations to values inside this
+range, but values outside of it are less important and thus can be less precise.
+
+The values \(\anninp\) and \(\annout\) depend on the feature extraction method. In case of an
+aggregation to product groups \(g\) we have \(\anninp=180\), and due to \(6\) possible actions for
+the two independent agents, \(\annout=12\).
+% SpecFullyConnected 180 540 :=> SpecRelu (540,1,1) :=> SpecFullyConnected 540 180 :=> SpecRelu
+% (180,1,1) :=> SpecFullyConnected 180 24 :=> SpecRelu (24,1,1) :=> SpecFullyConnected 24 12 :=>
+% SpecReshape (12,1,1) (6,2,1) :=> SpecLeakyTanh 0.98 (6,2,1) :=> SpecNNil2D 6x2
+In the case of the full representation using 6 independent agents the setup is \(\anninp=540\) and
+\(\annout=36\).
+% SpecFullyConnected 540 1620 :=> SpecRelu (1620,1,1) :=> SpecFullyConnected 1620 540 :=> SpecRelu
+% (540,1,1) :=> SpecFullyConnected 540 72 :=> SpecRelu (72,1,1) :=> SpecFullyConnected 72 36 :=>
+% SpecReshape (36,1,1) (18,2,1) :=> SpecLeakyTanh 0.98 (18,2,1) :=> SpecNNil2D 18x2
+For the Q-Learning algorithm, which only approximates one function, the output nodes
+\(\annout\) are halved for both setups.
+
+The ANN is trained using the Adam optimiser \citep{kingma2014adam} with learning rate \(0.0001\),
+and the default beta-values \(\beta_{1}=0.9\), and \(\beta_{2}=0.999\). As neural network
+initialisation method we use the commonly known Xavier initialisation
+\citep{glorot2010understanding}, i.e.\@ the initial value for each weight \(i\) is sampled from a
+uniform distribution \(W_{l,i} \sim \Unif(-1/\sqrt{n_l},1/\sqrt{n_l})\), where \(n_{l}\) is the
+number of nodes in layer \(l\).
 
@@
       \hline
-       \rl{0.995}             & 55.218 & 137.006 & 52.090   &          & 244.315 & 1.506 & 2.597    & 0.915  & 34.902   & 32.791             \\
-      \rl{1.000}              & 67.265 & 102.180 & 175.686  &          & 345.132 & 1.517 & 2.627    & 0.908  & 49.305   & 30.777             \\
-      \BILOne{}               & 45.902 & 0.000   & 734.431  &          & 780.333 & 1.319 & 1.806    & 0.352  &          &                    \\
-      \BILTwo{}               & 45.734 & 80.313  & 1425.221 &          & 268.569 & 1.318 & 2.158    & 0.876  &          &                    \\
-      \BILThree{}             & 45.861 & 279.723 & 30.818   &          & 356.402 & 1.318 & 3.034    & 0.975  &          &                    \\
+      \rl{0.995}              & 55.22 & 137.00 & 52.09   &          & 244.32 & 1.51 & 2.60    & 0.915  & 34.902   & 32.791             \\
+      \rl{1.000}              & 67.27 & 102.18 & 175.69  &          & 345.13 & 1.52 & 2.63    & 0.908  & 49.305   & 30.777             \\
+      \BILOne{}               & 45.90 & 0.00   & 734.43  &          & 780.33 & 1.32 & 1.81    & 0.352  &          &                    \\
+      \BILTwo{}               & 45.73 & 80.31  & 1425.22 &          & 268.57 & 1.32 & 2.158    & 0.876  &          &                    \\
+      \BILThree{}             & 45.86 & 279.72 & 30.82   &          & 356.40 & 1.32 & 3.034    & 0.975  &          &                    \\
       \hline

### 2021-01-13 paper/paper.tex
paper

@@
   \end{algorithmic}
-  \caption{\label{alg:near}Near-Blackwell-optimal deep RL for unichain MDPs}
+  \caption{\label{alg:near}Basic near-Blackwell-optimal deep RL for unichain MDPs}
 \end{algorithm}
@@
 The corresponding update values are saved to calculate the gradients based on the quadratic errors.
-Every \(C\) steps the target network exponentially adapts the parameters of the worker network.
-%
-%
-%
-%
-\MS[t]{describe network initialisation method (Uniform, HeAtAl, etc.)}
-%
-\MS[t]{Extensions: overestimate rho, rho minimum, Multiple Agents (shared state,  shared rho?),
-  init-phase (0.5 for expSmthRewRate, always adapt rho, same rho for all agents): exp smooth reward
-  as rho, replay memory actions, n-step (describe replay memory), gradient clipping }
-\MS[t]{Side info: Replay memory also stores (disallowed) filtered action indices for optimisation purposes and on
-episodic tasks needs to store a bool indicating an episode end. Uses Int8 for state features values.  }
+Every \(C\) steps the target network smoothly adapts the parameters of the worker network.
 
@@
 %
-Any state $s \in \mathcal{S}$ of the \textit{state space} is composed of the following information
-for both groups \(g\), i.e.\@ for bottleneck and non-bottleneck products:
+For the \textit{state space} we adapt the result of \citep{knollmann2013control}, which state that
+the lead times ought to control the WIP level, instead of the due date reliability as changes are
+directly seen. Thus, any state $s \in \mathcal{S}$ of the state space is composed of the following
+information for both groups \(g\), i.e.\@ for bottleneck and non-bottleneck products:
 \begin{itemize}
@@
 A naive approach is to define one action for each combination of actions over all product groups.
-Although applicable in our setup, as we only define two groups that yield \(9\) actions, it simply
-does not scale.
-%
+Although applicable in our setup, as we aggregate the data into the groups of bottleneck and
+non-bottleneck products and thus yields \(9\) actions, it simply does not scale.
+
 \begin{figure*}[t!]
@@
 \paragraph*{Reward.}
-In each period the agents choose actions which generates a reward while traversing to the next
-period by simulating the production system. The rewards are the accumulated costs at the end of the
-period consisting of the backorders, the WIP orders and the number of orders in the FGI\@. The
-objective is to minimise the returned reward. Therefore, we adapt the algorithm by swapping all
-maximisation functions with minimisations.
-
-\paragraph*{Adapted Algorithm.}
+In each period the agents choose actions, which when executed generate a reward while traversing to
+the next period by simulating the production system. The rewards are the accumulated costs at the
+end of the period consisting of the backorders, the WIP orders and the number of orders in the
+FGI\@. The objective is to minimise the returned reward. Therefore, we adapt the algorithm by
+swapping all maximisation functions with minimisations.
+
+\paragraph*{Optimised Algorithm.}
 Our implementation includes several optimisations, that help the algorithm in better exploring the
@@
   \end{algorithmic}
-  \caption{\label{alg:full}Near-Blackwell-optimal deep RL for unichain MDPs}
+  \caption{\label{alg:full}Optimised near-Blackwell-optimal deep RL for unichain MDPs}
 \end{algorithm}

### 2021-01-28 paper/paper.tex
update alg

@@
   due date.
+  % \MS[t]{depend}If a product type is not routed over the corresponding machine, then no entry for this
+  % product type is specified.
 
@@
     \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
-      Sample random mini-batch of \(\nstep\) consecutive experiences from \(M\), each with
-      \((s_{i}, \isRandAct{i}, a_{i}, r_{i}, s_{i})\) where \(i \in {1,2,\ldots,\nstep}\).
+      Sample random mini-batch of \(\nstep\) consecutive experiences
+      from \(M\) numbered by \(1 \leqslant i \leqslant \nstep\), each
+      with \((s_{i}, \isRandAct{i}, a_{i}, r_{i}, s_{i+1})\).
     \end{minipage}
@@
     \If{experience \(\nstep\) or (\(\isRandAct{i}\) and  \(\pexp \leqslant \plearn\))}
-      \State{\(\target_{\gamma0} \gets \min_{a} \X^\pol_{\gamma_0}(s_{t+1},a; \netTarget{})\)}
-      \State{\(\target_{\gamma1} \gets \min_{a} \X^\pol_{\gamma_1}(s_{t+1},a; \netTarget{})\)}
+      \State{\(\target_{\gamma0} \gets \min_{a} \X^\pol_{\gamma_0}(s_{i+1},a; \netTarget{})\)}
+      \State{\(\target_{\gamma1} \gets \min_{a} \X^\pol_{\gamma_1}(s_{i+1},a; \netTarget{})\)}
     \EndIf
-    \If{a non-random action was chosen or \(\pexp > \plearn\)}
-    \State{\(\avgrew^{\pol} \gets (1- \alpha) \avgrew^{\pol} + \alpha [r_{t} +
-      \min_{a}\X^\pol_{\gamma_1}(s_{t+1},a; \netWorker{}) - \X^\pol_{\gamma_1}(s_{t},a_{t};
+    \If{\(\isRandAct{i}\) or \(\pexp > \plearn\)}
+    \State{\(\avgrew^{\pol} \gets (1- \alpha) \avgrew^{\pol} + \alpha [r_{i} +
+      \min_{a}\X^\pol_{\gamma_1}(s_{i+1},a; \netWorker{}) - \X^\pol_{\gamma_1}(s_{i},a_{i};
       \netWorker{})]\)}
@@
     \State{Compute overestimate of \(\avgrew^{\pol}\) and store result in \(\avgrew_{\Overestim}^{\pol}\)}
-    \State{\(\target_{\gamma_0}  \gets r_{t} + \gamma_{0} \target_{\gamma_{0}} - \avgrew_{\Overestim}^{\pol}  \)}
-    \State{\(\target_{\gamma_{1}}  \gets r_{t} + \gamma_{1} \target_{\gamma_{1}} - \avgrew_{\Overestim}^{\pol}\)}
-    \State{\(d\netWorker{} \gets d\netWorker{} + \partial{(\target_{\gamma_{0}} - \X^\pol_{\gamma_{0}}(s_{t+1},a; \netWorker{}))}^{2} / \partial\netWorker{}\)}
+    \State{\(\target_{\gamma_0}  \gets r_{i} + \gamma_{0} \target_{\gamma_{0}} - \avgrew_{\Overestim}^{\pol}  \)}
+    \State{\(\target_{\gamma_{1}}  \gets r_{i} + \gamma_{1} \target_{\gamma_{1}} - \avgrew_{\Overestim}^{\pol}\)}
+    \State{\(d\netWorker{} \gets d\netWorker{} + \partial{(\target_{\gamma_{0}} - \X^\pol_{\gamma_{0}}(s_{i+1},a; \netWorker{}))}^{2} / \partial\netWorker{}\)}
     \State{\(d\netWorker{} \gets d\netWorker{} +
-      \partial{(\target_{\gamma_{1}} - \X^\pol_{\gamma_{1}}(s_{t+1},a; \netWorker{}))}^{2} / \partial\netWorker{}\)}
+      \partial{(\target_{\gamma_{1}} - \X^\pol_{\gamma_{1}}(s_{i+1},a; \netWorker{}))}^{2} / \partial\netWorker{}\)}
     \EndFor{}
@@
 %
-The resulting feedforward deep ANN archtiecture is given as %
+The resulting feedforward deep ANN architecture is given as %
 \(\anninp %

### 2021-04-23 paper/paper.tex
minor

@@
     \end{minipage}
-    \State{Sample random mini-batch of experiences from \(M\)}
+    \State{Sample random mini-batch of \(m\) experiences from \(M\)}
     \State{Reset gradients: \(d\netWorker{} \gets 0\)}
@@
     \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth}
-      Perform update with sum of the gradients \(d\netWorker{}\) on \(\netWorker{}\)
+      Perform update with average of the gradients \(d\netWorker{}/m\)  on \(\netWorker{}\)
     \end{minipage}
@@
     \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
-      Sample random mini-batch of \(\nstep\) consecutive experiences
-      from \(M\) numbered by \(1 \leqslant i \leqslant \nstep\), each
-      with \((s_{i}, \isRandAct{i}, a_{i}, r_{i}, s_{i+1})\).
+      Sample \(m\) random mini-batches of \(\nstep\) consecutive experiences from \(M\) numbered by
+      \(1 \leqslant i \leqslant \nstep\), each with
+      \((s_{i}, \isRandAct{i}, a_{i}, r_{i}, s_{i+1})\).
     \end{minipage}
@@
     \EndFor{}
-    \State{Perform update of \(\netWorker{}\) using \(d\netWorker{}\)}
+    \State{Perform update of \(\netWorker{}\) using \(d\netWorker{} / (m \cdot \nstep)\)}
     \EndIf{}

### 2021-04-28 paper/paper.tex
update

@@
 \usepackage{listings}
+\usepackage{booktabs}
 \usepackage{algorithm}
@@
 %
-For the \textit{state space} we adapt the result of \citep{knollmann2013control}, which state that
+For the \textit{state space} we adapt the result of \citet{knollmann2013control}, which state that
 the lead times ought to control the WIP level, instead of the due date reliability as changes are
 directly seen. Thus, any state $s \in \mathcal{S}$ of the state space is composed of the following
-information for both groups \(g\), i.e.\@ for bottleneck and non-bottleneck products:
+information for each product type \(\pt\):
 \begin{itemize}
@@
 is the current period. Clearly, this yields bulk releases, meaning that either all or no orders with
-the same due date and of the same product group are
-released.
+the same due date and of the same product type are released.
 
 The \textit{action space} is composed of two decisions, the relative change of the lead times to the
-currently set lead times $LT_{g} \in \{1,2,\ldots,\DDS\}$ for each product group $g$.
+currently set lead times $LT_{g} \in \{1,2,\ldots,\DDS\}$ for each product type $\pt$.
 %
 We restrict the lead time update for state $s_{t+1}$ according to the lead time $LT_{g}$ from state
-$s_{t}$ by a maximum change of 1. Thus, if $LT_{g}$ is the current lead time of product group $g$
+$s_{t}$ by a maximum change of 1. Thus, if $LT_{g}$ is the current lead time of product type $\pt$
 the agent can choose a lead time in $\{1,2,\ldots,\DDS\} \cap \{LT_{g}-1,LT_{g},LT_{g}+1\}$. % Put
@@
 % as it acts within the discrete action space given by the set $\{1,2,\ldots,\DDS\}$ for each product
-% group.
-A naive approach is to define one action for each combination of actions over all product groups.
-Although applicable in our setup, as we aggregate the data into the groups of bottleneck and
-non-bottleneck products and thus yields \(9\) actions, it simply does not scale.
+% type.
+A naive approaches is to define one action for each combination of actions over all product type.
+This however, due to the exponential growth, results in only being able to solve very small
+production systems with only few product types. A common solution in production planning is to
+aggregate the products into product groups, cf.\@ hierarchical production
+planning~\citep{schneeweibeta1995hierarchical}.
+%
+Although applicable in our setup, if we group the products into bottleneck and non-bottleneck
+products as done by \citet{schneckenreither2020order}, also this approach simply does not scale.
 
@@
 %
-Therefore, we invent and implement \textit{independent agents}, each operating on the same system
-state, but choosing actions independently for the designated product group.
-Figure~\ref{fig:indep-agents} illustrates the idea, where wlog. we use two agents and for simplicity
-omit the reward and truncate the state to the lead times. Agent one chooses to increase the lead
-time from \(3\) to \(4\). Thus, the future state, seen from agent one, depends on the probability of
-agent two choosing to increase (\(\inc\)), decrease (\(\dec\)) or not change (\(\pers\)) the lead
-time. Note that the Markov property the agent, that is
-\(p(s_{t+1}, r_{t} \mid s_{t}, a_{t}) = p((4,\cdot), r_{t} \mid (3,3), \inc) \), is not touched by
-this adaption. While for agent one the reward and future state only depends on the current state and
-action taken, the corresponding probabilities for future state and reward change during the learning
-process as the other agent starts to behave differently.
-%
-\MS[t]{independent agents, true?}This allows us to investigate the algorithms behaviour for even
-larger solution spaces. In particular, this adaption enables us to present the algorithm the above
-defined solution space not only on the basis of product groups \(g\), but also for each product type
-\(p\). That means, the state space, as well as the action space operates on a product type basis.
-Note that without independent agents this is infeasible due to \(3^{6} = 729\) possible actions.
-Removing the dependency between the lead times of different product types reduces the possibilities,
-and thus the number of ANN output layer nodes, to \(18\). Therefore, this is a vital adaption that
-ensures scalability.
-
-        %         We allow the agent a higher increase
-        %         of the lead time, compared to the decrease, to give it the chance to react quickly to high
-        %         number of
-        %         incoming demand.
+Therefore, in the seek of a broader solution we establish the idea of \textit{independent agents}.
+Each independent agent operates on the same system state, but chooses only a part of the actual
+executed action. In our case each independent agent is responsible for the lead time update of a
+specific product type. To the best of the author's knowledge this is the first work to reduce the
+action space in this way. Figure~\ref{fig:indep-agents} illustrates the idea, where wlog.\@ we use
+two agents and for simplicity omit the reward and truncate the state to the lead times. Assume that
+the first independent agent chooses to increase the lead time from \(3\) to \(4\). Thus, the future
+state, seen from this agent, depends on the probability of the other independent agent choosing to
+increase (\(\inc\)), decrease (\(\dec\)) or not change (\(\pers\)) the lead time. The Markov
+property, that is
+\(p(s_{t+1}, r_{t} \mid s_{t}, a_{t}) = p((4,\cdot), r_{t} \mid (3,3), (\inc, \cdot)) \), is not
+touched by this adaption. While for the first agent the reward and future state only depends on the
+current state and action taken, the corresponding probabilities for future state and reward change
+during the learning process as the the agent starts to behave differently.
+%
+This allows us to investigate the algorithms behaviour for even larger solution spaces. In
+particular, this adaption enables us to present the algorithm the above defined state space not only
+on the basis of product groups, but directly for each product type \(\pt\). Note that without this
+extension the problem is computationally infeasible due to \(3^{6} = 729\) possible actions.
+Removing the dependency between the lead times of different product types reduces the set of
+possible actions, and thus the number of ANN output layer nodes, to \(3 \cdot 6 = 18\). In
+preliminary experiments on small examples we only observed a very slight increase of required steps
+during training. Therefore, this is a vital adaption to ensure scalability.
 
@@
 end of the period consisting of the backorders, the WIP orders and the number of orders in the
-FGI\@. The objective is to minimise the returned reward. Therefore, we adapt the algorithm by
-swapping all maximisation functions with minimisations.
+FGI\@. The objective is to minimise the returned reward. % Therefore, we adapt the algorithm by
+% swapping all maximisation functions with minimisations.
 
@@
 solution space and stabilise the learning process. These include the above discussed independent
-agents that operate on the same state, \(n\)-step learning \citep{mnih2016asynchronous},
-overestimating the average reward \(\avgrew^{\pol}\) and ensuring to constantly reduce the average
-reward as searching for policies with higher average costs is unnecessary. Algorithm~\ref{alg:full}
-depicts a detailed view on the algorithm including these extensions and with the minimisation
-objective. The idea is the same as in Algorithm~\ref{alg:near}, except that the action selection
-procedure is a loop over all independent agents and the implementation of the n-step logic.
-
-% N-Step
-N-step RL collects a set of consecutive experiences, in our case retrieved from the replay memory,
-and connects the state values starting from the latest observed experience, such that the newly
-computed state value of the corresponding experience propagates to the experience of the previous
-period. This allows a faster convergence, as values bootstrap over multiple periods in contrast to
-just one step.
+agents that operate on the same state, \(n\)-step learning \citep{mnih2016asynchronous}
+% overestimating the average reward \(\avgrew^{\pol}\)
+and ensuring to constantly reduce the average reward as searching for policies with higher average
+costs is unnecessary. Algorithm~\ref{alg:full} depicts a detailed view on the algorithm including
+these extensions and with the objective to minimise the returned rewards. The idea is the same as in
+Algorithm~\ref{alg:near}, except that the action selection procedure is a loop over all independent
+agents and the implementation of the n-step logic.
+
+% N-Step: 5
+\textit{N-step learning} collects a set of consecutive experiences, in our case retrieved from the
+replay memory, and connects the state values starting from the latest observed experience, such that
+the newly computed state value of the corresponding experience propagates to the experience of the
+previous period. This allows a faster convergence, as values bootstrap over multiple periods in
+contrast to just one step. The experiences are stored according to the first taken action in the
+n-step series. We uniformly sample over the actions to ensure each action is constantly trained.
+This adaption reduces catastrophic forgetting of already accumulated knowledge.
+%
 % Overestimating
-The idea behind overestimating the average reward is to constantly seek for better policies than the
-currently known. We overestimate the reward by subtracting \(10\%\) of the computed average reward
-\(\avgrew^{\pol}\) and an exponentially smoothed average reward of the actually returned reward
-values of greedy actions. Thus, if the computed and actually observed average reward coincide, the
-overestimation equals \(0\).
+% The idea behind overestimating the average reward is to constantly seek for better policies than the
+% currently known. We overestimate the reward by subtracting \(10\%\) of the computed average reward
+% \(\avgrew^{\pol}\) and an exponentially smoothed average reward of the actually returned reward
+% values of greedy actions. Thus, if the computed and actually observed average reward coincide, the
+% overestimation equals \(0\).
 % Rho Minimum
-Preventing an increase of the average reward with \(\avgrew_{\max}^{\pol}\) reduces the search space
-and provides a methodology for constantly improving policy.
-
-\MS[t]{(shared rho)?, (gradient clipping)?}
-
-Furthermore, we sometimes use multiple parallel workers (parallel RL) that operate on the same
-environment to collect a broader range of experiences. The experiences are stored according to the
-first taken action in the n-step series. We uniformly sample over the actions to ensure each action
-is constantly trained. This reduces catastrophic forgetting.
-
-The algorithm initiation tactics is to first fill the replay memory buffer by skipping any learning
-(steps \(10-25\)) until the period, that corresponds to the replay memory size, is reached. Then
-for \(500\) steps the average reward is set to the last computed exponentially smoothed value to
-prevent divergence due to the untrained network weight initialisation values.
+% Preventing an increase of the average reward with \(\avgrew_{\max}^{\pol}\) reduces the search space
+% and provides a methodology for constantly improving policy.
+%
+% Workers Min Exploration [0.01, 0.02, 0.03, 0.04, 0.05, 0.10, 0.15, 0.20]
+Furthermore, we use \(8\) additional \textsf{worker threads} (parallel RL) that operate on the same
+environment to collect a broader range of experiences~\citep[as e.g. in][]{mnih2016asynchronous}.
+
+% Shared Rho (Yes)
+The average reward value is shared and computed as average over the main agent and all workers.
+% \MS[t]{shared rho}
+%
+%
+% \MS[t]{worker min exploration, epsilon greedy exploration}
+%
+%
+% Learning rate decay: ExponentialDecay (Just 5e-6) 0.25 100000
+% Learning rate: OptAdam 0.005 0.9 0.999 1e-8 1e-3,
+
+% Algorithms tested:
+ %            AlgDQNAvgRewAdjusted 0.8 1.0 ByStateValues
+ %          , AlgDQNAvgRewAdjusted 0.8 0.99 ByStateValues
+ %          , AlgDQN 0.99 Exact
+
+% Alpha: 0.01, Decay: ExponentialDecay (Just 5e-5) 0.25 50000
+% AlphaRhoMin: 2e-5, Decay: ExponentialDecay (Just 2e-5) 0.25 50000
+% Gamma: 0.01, Decay: ExponentialDecay (Just 1e-3) 0.25 100000
+% Epsilon: 0.25, 0.25, Decay: No decay
+% Exploration: 1.0, Decay: ExponentialDecay (Just 0.005) 0.25 100000
+% LearnRandomAbove: 0.5
+%
+%
+% UNUSED: Delta: 0.005, Decay: ExponentialDecay (Just 5e-4) 0.25 100000
+
+The algorithm initiation strategy is to first fill the replay memory buffer by skipping any learning
+(steps \(10-25\)) until the period, that corresponds to the replay memory size, is reached. Then for
+\(500\) steps the average reward is set to the last computed exponentially smoothed value to prevent
+divergence due to the untrained network weight initialisation values.
 
@@
 \(\acts = \size{\States}\), the independent agent configuration, and the number of functions that
-are approximated. For the later, in case of the proposed algorithm this is \(2\),
-\(\X^\pol_{\gamma_{0}}\) and \(\X^\pol_{\gamma_{1}}\), while deep Q-Learning only approximates the
-function \(Q_{\gamma}^\pol\).
+are approximated. For the later, in case of \ARA{} this is \(2\), as we represent both functions
+\(\X^\pol_{\gamma_{0}}\) and \(\X^\pol_{\gamma_{1}}\) with the same neural network, while deep
+Q-Learning only approximates the function \(Q_{\gamma}^\pol\).
 %
@@
 \(\anninp %
-\rightarrow 3 \anninp % \rightarrow \relu
-\rightarrow \anninp % \rightarrow \relu
-\rightarrow 2 \annout % \rightarrow \relu
+\rightarrow 1.75 \anninp % \rightarrow \relu
+\rightarrow 1.5 \anninp % \rightarrow \relu
+% \rightarrow 2 \annout % \rightarrow \relu
 \rightarrow \annout %
 % \rightarrow   \leakyTanh%
-\), where each arrow symbolises a fully-connected layer with a rectified linear unit (\(\relu\)) as
-activation function. As output activation function we use the hyperbolic tangent function
-(\(\Tanh\)), but similar to leaky \(\relu\), leak values greater than \(0.98\) (less than \(-0.98\))
-with \(5\%\). The idea is that we do not want to restrict the approximations to values inside this
-range, but values outside of it are less important and thus can be less precise.
-
-The values \(\anninp\) and \(\annout\) depend on the feature extraction method. In case of an
-aggregation to product groups \(g\) we have \(\anninp=180\), and due to \(6\) possible actions for
-the two independent agents, \(\annout=12\).
-% SpecFullyConnected 180 540 :=> SpecRelu (540,1,1) :=> SpecFullyConnected 540 180 :=> SpecRelu
-% (180,1,1) :=> SpecFullyConnected 180 24 :=> SpecRelu (24,1,1) :=> SpecFullyConnected 24 12 :=>
-% SpecReshape (12,1,1) (6,2,1) :=> SpecLeakyTanh 0.98 (6,2,1) :=> SpecNNil2D 6x2
-In the case of the full representation using 6 independent agents the setup is \(\anninp=540\) and
-\(\annout=36\).
-% SpecFullyConnected 540 1620 :=> SpecRelu (1620,1,1) :=> SpecFullyConnected 1620 540 :=> SpecRelu
-% (540,1,1) :=> SpecFullyConnected 540 72 :=> SpecRelu (72,1,1) :=> SpecFullyConnected 72 36 :=>
-% SpecReshape (36,1,1) (18,2,1) :=> SpecLeakyTanh 0.98 (18,2,1) :=> SpecNNil2D 18x2
-For the Q-Learning algorithm, which only approximates one function, the output nodes
-\(\annout\) are halved for both setups.
-
-The ANN is trained using the Adam optimiser \citep{kingma2014adam} with learning rate \(0.0001\),
-and the default beta-values \(\beta_{1}=0.9\), and \(\beta_{2}=0.999\). As neural network
+\), where each arrow symbolises a fully-connected layer with a leaky rectified linear unit
+(\(\relu\)) as activation function, which leaks values smaller than \(0\) with rate \(0.02\). Note
+that we explicitly do not use an output activation function. The idea is that we do not want to
+restrict the approximations to values inside this range. This is especially important for
+Q-Learning, as the range of the state values are difficult to forecast.
+
+The values \(\anninp\) and \(\annout\) depend on the feature extraction method. In our case case
+using 6 independent agents we gain \MS[t]{todo: check} \(\anninp=540\) and \(\annout=36\)
+\MS[t]{check setup}. For the Q-Learning algorithm, which only approximates one function, the output
+nodes \(\annout\) are halved. \MS[t]{what is the correct size?}
+
+% Min-Max Scaling of input/output. Output: ScalingNetOutParameters (-800) 800 (-300) 300 (-400) 800 (-400) 800
+\MS[t]{scaling of input layer} The features representing order counts are scaled using min-max
+scaling with minimum \(0\) and maximum \(5\), the lead times with minimum \(0\) and maximum \(7\)
+respectively.
+% Output layer scaling
+To further ensure smaller ANN parameter values the output is transformed using min-max scaling to
+the interval \((-400, 800)\), s.t.\@ an ANN output of \(1\) represents a value of \(800\).
+%
+% Training Mini-Batch Size:  4
+On each training step, i.e.\@ every n-step steps, randomly \(m\) sets of experiences the training
+mini-batch size each consisting of n-step experiences are selected, there gradients computed and the
+sum of gradients back-propagated though the worker network using Adam \citep{kingma2014adam}, where
+the weight decay with parameter \(\lambda\) is implemented as presented by
+\citet{loshchilov2017decoupled}.
+%
+% Smooth target network period: 100 with update 0.01
+Every \(C\) steps the target network is updated by exponentially smoothing the parameters with rate
+\(\gamma\).
+%
+% ANN Init
+As neural network
 initialisation method we use the commonly known Xavier initialisation
@@
 number of nodes in layer \(l\).
+%
+%
 
@@
 
-This section provides preliminary results for the described setup and gives an overview of the
-current performance of the algorithm compared to the conventional static lead time setting
-algorithms.
-
-Currently we evaluate two different kinds of reward reporting as they have been proposed by
+This section provides results for the described setup and gives an overview of the current
+performance of the algorithm compared to the conventional static lead time setting algorithms.
+
+\MS[t]{todo} We evaluate two different kinds of reward reporting as they have been proposed by
 \cite{Schneckenreither2019}. They find that it is beneficial to overcome the time offset imposed by
@@
   \centering
-  \begin{tabular}{|l||c|c|c|c|c|c|c|c|c|c|}
-    \hline
-    Parameter & $\alpha$ & $\beta$ & $\delta$ & $\gamma$ & $\epsilon$ & $p_{exp}$ & $\xi$ \\
-    \hline
-    Value & 0.5 & 0.15 & 0.15 & 0.15 & 0.5 & 1.0 & 0.15 \\
-    \hline
+  \footnotesize
+  \begin{tabular}{lP{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}}
+    \toprule
+    Parameter                      & $\alpha$   & $\alpha_{\max}$ & $\gamma_{0}$ & $\gamma_{1}$ & $\gamma$ & $\epsilon$ & $p_{exp}$ & $\plearn$ \\
+    \midrule
+    Start value                    & 0.01       & \(2^{-5}\)      & 0.8          & 0.99/1.0     & 0.01     & 0.25       & 1.0       & 0.5       \\
+    Exp. Smth. Rate                & 0.25       &                 &              &              & 0.25     &            & 0.25      &           \\
+    Exp. Smth. Steps in \(10^{3}\) & 50         &                 &              &              & 100      &            & 100       &           \\
+    Minimum value                  & \(5^{-5}\) &                 &              &              & 0.001         &            & 0.005     &           \\
+    \bottomrule
+    \toprule
+    Parameter                      & \multicolumn{8}{c}{Value} \\
+    \midrule
+    Workers Min. Exploration & \multicolumn{8}{c}{0.01, 0.02, 0.03, 0.04, 0.05, 0.10, 0.15, 0.20}\\
+    Training Batch Size & \multicolumn{8}{c}{4}\\
+    Target Network Update & \multicolumn{8}{c}{Exp. smooth. with rate \(\gamma\) every \(C=100\) steps}\\
+    N-Step & \multicolumn{8}{c}{5}\\
+    Training Mini-Batch Size & \multicolumn{8}{c}{\(m = 4\)}\\
+    Adam Back-Propagation & \multicolumn{8}{c}{\(\alpha = 0.005, \beta_{1}=0.9,
+                       \beta_{2}=0.999, \epsilon=10^{-8}, \lambda=0.001}\\ % OptAdam 0.005 0.9 0.999 1e-8 1e-3,
+    \bottomrule
   \end{tabular}
@@
 with rate \(0.05\) and \(350k\) steps\footnote{The decay is defined as \(0.05 ^{(t/350k)}\), where
-  \(t\) is the current period}. We implemented the deep Q-network algorithm proposed by
-\cite{mnih2015human}, which is parameterised by learning rate \(\delta\) and a discount factor of
+  \(t\) is the current period}. \MS[t]{TODO} We implemented the deep Q-network algorithm proposed by
+\cite{mnih2015human},  a discount factor of
 \(0.99\).
@@
 
+% Therefore, in a current working paper I apply an enhanced state/action-value iteration version with
+% ANN as function approximation on bigger sized problems, where preliminary promising results could be
+% achieved. In the future I want to adapt and investigate the use of an actor-critic reinforcement
+% learning version to problems from research and industry, where I plan to focus on production
+% planning. We hypothesize that due to more correct state value estimations from the critic, the
+% resulting policies from the actor are will outperform the currently known actor-critic algorithms.
+% This would be in line with the observation of overestimated state values of standard discounted
+% reinforcement learning algorithms, which lead to the development of
+% extensions~\cite[]{hessel2018rainbow}. We believe that instead of developing extensions to resolve
+% structurally fundamental problems a better approach is to aim for correct state value estimations,
+% which will consequently lead to better policies.\\
+
+\MS[t]{It remains to be shown that independent agents find the same (or better) policies.}
+
 % This paper introduces deep theoretical insights in reinfocement learning and provides first evidence

### 2021-04-29 paper/paper.tex
update

@@
 
-\MS[t]{Algorithm adaptions}
-
 \paragraph{Markov Decision Process.} The underlying MDP is unichain and looks as follows. Both,
@@
 %
-% Recall that we gproducts routed over the bottleneck machine \(M5\) are bottleneck products.
+% Recall that we products routed over the bottleneck machine \(M5\) are bottleneck products.
 %
@@
 % N-Step: 5
+As in \citep{mnih2016asynchronous, mnih2015human} the algorithm uses a \textit{Replay Memory} to
+save experiences and then randomly selects experiences from there for learning. This ensures that
+the independent and identically distributed (i.i.d.) assumption of the samples for the ANN
+backpropagation algorithms are satisfied. We extend this idea by splitting the available size of the
+replay memory into subsets, where each subset represents one action and sample uniformly over these
+subsets first. This prevents overfitting when one action is overrepresented in the policy, e.g.\@ no
+lead time change occurs much more often than the other actions.
+
 \textit{N-step learning} collects a set of consecutive experiences, in our case retrieved from the
@@
 \), where each arrow symbolises a fully-connected layer with a leaky rectified linear unit
-(\(\relu\)) as activation function, which leaks values smaller than \(0\) with rate \(0.02\). Note
+(\(\relu\)) as activation function, which leaks values smaller than \(0\) with rate \(0.02\).
+%
+Note
 that we explicitly do not use an output activation function. The idea is that we do not want to
@@
 Q-Learning, as the range of the state values are difficult to forecast.
-
-The values \(\anninp\) and \(\annout\) depend on the feature extraction method. In our case case
-using 6 independent agents we gain \MS[t]{todo: check} \(\anninp=540\) and \(\annout=36\)
-\MS[t]{check setup}. For the Q-Learning algorithm, which only approximates one function, the output
-nodes \(\annout\) are halved. \MS[t]{what is the correct size?}
+%
+The actual values \(\anninp\) and \(\annout\) depend on the used algorithm. In our case case using 6
+independent agents we gain \(\anninp=324\) and \(\annout=36\). For the Q-Learning algorithm, which
+only approximates one function, the output nodes \(\annout\) are halved.
+%
+% SpecFullyConnected 324 567 :=> SpecLeakyRelu (567,1,1) :=> SpecFullyConnected 567 486 :=>
+% SpecLeakyRelu (486,1,1) :=> SpecFullyConnected 486 36 :=> SpecReshape (36,1,1) (18,2,1) :=>
+% SpecNNil2D18x2
+%
 
 % Min-Max Scaling of input/output. Output: ScalingNetOutParameters (-800) 800 (-300) 300 (-400) 800 (-400) 800
-\MS[t]{scaling of input layer} The features representing order counts are scaled using min-max
-scaling with minimum \(0\) and maximum \(5\), the lead times with minimum \(0\) and maximum \(7\)
-respectively.
+The features representing order counts are scaled using min-max scaling with minimum \(0\) and
+maximum \(5\), the lead times with minimum \(0\) and maximum \(7\) respectively.
 % Output layer scaling
@@
 mini-batch size each consisting of n-step experiences are selected, there gradients computed and the
-sum of gradients back-propagated though the worker network using Adam \citep{kingma2014adam}, where
+sum of gradients backpropagated though the worker network using Adam \citep{kingma2014adam}, where
 the weight decay with parameter \(\lambda\) is implemented as presented by
@@
 
-\MS[t]{todo} We evaluate two different kinds of reward reporting as they have been proposed by
-\cite{Schneckenreither2019}. They find that it is beneficial to overcome the time offset imposed by
-the production system. That is, although orders are immediately released to the production system
-and thus generate a response in terms of a reward to the agent, the full impact of the chosen action
-takes several periods to materialise. For instance, by deciding not to release an order, which later
-becomes a backorder, it is beneficial to reward the actions actually responsible for the backorder
-instead of the last action (only). Thus they propose to keep track of the orders currently in the
-order pool until all fully traversed through the production system and reward the actions according
-to all orders in the order pool.
+% \MS[t]{todo} We evaluate two different kinds of reward reporting as they have been proposed by
+% \cite{Schneckenreither2019}. They find that it is beneficial to overcome the time offset imposed by
+% the production system. That is, although orders are immediately released to the production system
+% and thus generate a response in terms of a reward to the agent, the full impact of the chosen action
+% takes several periods to materialise. For instance, by deciding not to release an order, which later
+% becomes a backorder, it is beneficial to reward the actions actually responsible for the backorder
+% instead of the last action (only). Thus they propose to keep track of the orders currently in the
+% order pool until all fully traversed through the production system and reward the actions according
+% to all orders in the order pool.
 
@@
     Training Mini-Batch Size & \multicolumn{8}{c}{\(m = 4\)}\\
-    Adam Back-Propagation & \multicolumn{8}{c}{\(\alpha = 0.005, \beta_{1}=0.9,
-                       \beta_{2}=0.999, \epsilon=10^{-8}, \lambda=0.001}\\ % OptAdam 0.005 0.9 0.999 1e-8 1e-3,
+    Adam ANN Backpropagation & \multicolumn{8}{c}{\(\alpha = 0.005, \beta_{1}=0.9,
+                       \beta_{2}=0.999, \epsilon=10^{-8}, \lambda=0.001}\\ % OptAdam 0.005 0.9 0.999
+% 1e-8 1e-3,
+        Replay Memory Size & \multicolumn{8}{c}{\(10800\)}\\
     \bottomrule
   \end{tabular}
-  \caption{\label{tbl:params}Parameter Setup.}
+  \caption{\label{tbl:params}Overview of Parameter Setup.}
 \end{table}
 
-The parameter setup is given in Table~\ref{tbl:params}. The parameters are exponentially decayed
-with rate \(0.05\) and \(350k\) steps\footnote{The decay is defined as \(0.05 ^{(t/350k)}\), where
-  \(t\) is the current period}. \MS[t]{TODO} We implemented the deep Q-network algorithm proposed by
-\cite{mnih2015human},  a discount factor of
-\(0.99\).
+Table~\ref{tbl:params} presents an overview of the parameter setup. The parameters \(\alpha\),
+\(\gamma\) and \(p_{exp}\) are exponentially decayed as indicated.% \footnote{The decay is defined as
+  % \(0.05 ^{(t/350k)}\), where \(t\) is the current period}.
+\MS[t]{TODO} As further comparison we implement the deep Q-Learning network algorithm proposed
+by~\cite{mnih2015human}, where for comparison reasons we also integrate n-Step learning, the
+experience replay memory adaptions and independent agents. \MS[t]{Discount factor in table?}
+
+% a discount factor of \(0.99\).
 
@@
 
+\MS[t]{Multiple agents on different tasks (hierarchical PP): e.g. Order acceptance + release,
+  scheduling + release, etc.}
+
 % This paper introduces deep theoretical insights in reinfocement learning and provides first evidence

### 2021-04-30 paper/figures/productionsystem.tex
update

@@
 
-  \draw (0.5,0)    node[minimum size=1cm,draw,align=center] (M1) {M1\\[-1ex]{\tiny $\Exp(80) $ or} \\[-1ex]{\tiny $\Unif(30,130)$}};
-  \draw (3,0.75)   node[minimum size=1cm,draw,align=center] (M2) {M2\\[-1ex]{\tiny $\Exp(160)$ or}\\[-1ex]{\tiny $\Unif(80,240)$}};
-  \draw (3,-0.75)  node[minimum size=1cm,draw,align=center] (M3) {M3\\[-1ex]{\tiny $\Exp(155)$ or}\\[-1ex]{\tiny $\Unif(50,260)$}};
-  \draw (6,1.5)    node[minimum size=1cm,draw,align=center] (M4) {M4\\[-1ex]{\tiny $\Exp(210)$ or}\\[-1ex]{\tiny $\Unif(50,370)$}};
-  \draw (6,0)      node[minimum size=1cm,draw,align=center] (M5) {M5\\[-1ex]{\tiny $\Exp(285)$ or}\\[-1ex]{\tiny $\Unif(200,370)$}};
-  \draw (6,-1.5)   node[minimum size=1cm,draw,align=center] (M6) {M6\\[-1ex]{\tiny $\Exp(215)$ or}\\[-1ex]{\tiny $\Unif(110,320)$}};
+  \draw (0.5,0)    node[minimum size=1cm,draw,align=center] (M1) {M1\\[-1ex]{\tiny $\Exp(80) $ }}; % or}\\[-1ex]{\tiny $\Unif(30,130)$}};
+  \draw (3,0.75)   node[minimum size=1cm,draw,align=center] (M2) {M2\\[-1ex]{\tiny $\Exp(160)$ }}; % or}\\[-1ex]{\tiny $\Unif(80,240)$}};
+  \draw (3,-0.75)  node[minimum size=1cm,draw,align=center] (M3) {M3\\[-1ex]{\tiny $\Exp(155)$ }}; % or}\\[-1ex]{\tiny $\Unif(50,260)$}};
+  \draw (6,1.5)    node[minimum size=1cm,draw,align=center] (M4) {M4\\[-1ex]{\tiny $\Exp(210)$ }}; % or}\\[-1ex]{\tiny $\Unif(50,370)$}};
+  \draw (6,0)      node[minimum size=1cm,draw,align=center] (M5) {M5\\[-1ex]{\tiny $\Exp(285)$ }}; % or}\\[-1ex]{\tiny $\Unif(200,370)$}};
+  \draw (6,-1.5)   node[minimum size=1cm,draw,align=center] (M6) {M6\\[-1ex]{\tiny $\Exp(215)$ }}; % or}\\[-1ex]{\tiny $\Unif(110,320)$}};
   \draw (8,0)      node[minimum size=1cm,draw,align=center,fill] (\fgi{}) {\color{white} \fgi{}};

### 2021-04-30 paper/paper.tex
update

@@
   % teilw. von IJPR paper uebernommen
+  \MS[t]{proof-of-concept work}
   An essential task in manufacturing planning and control is to determine when to release orders to
@@
     \State{}Initialise state \(s_{0}\) and network parameters \(\netTarget, \netWorker\) randomly, set
-    an exploration rate \(0 \leqslant \plearn \leqslant \pexp \leqslant 1\), exponential smoothing learning rates
+    an exploration rate \(0 \leqslant \plearn < \pexp \leqslant 1\), exponential smoothing learning rates
     \(0 < \alpha, \gamma < 1\), and discount factors
@@
 
-\paragraph{Processing Times.} The machine processing time distributions are given under the
-corresponding node labels of the machines. To simulate high or moderate variance the processing
-times for all machines are  drawn from an exponential or an uniform distribution,
-respectively. There is one bottleneck machine ($M5$) and therefore we refer to product
-types $2$ and $5$, those routed through machine $M5$, as
-bottleneck products whereas products $1,3,4,6$ are non-bottleneck products.
+\paragraph{Processing Times.} The machine processing time distributions are specified under the
+corresponding node labels of the machines. To simulate high % or moderate
+variance the processing times for all machines are drawn from an exponential % or an uniform
+distribution% , respectively
+. There is one bottleneck machine ($M5$) and therefore we refer to product types $2$ and $5$, those
+routed through machine $M5$, as bottleneck products whereas products $1,3,4,6$ are non-bottleneck
+products.
 
@@
 %
-To simulate high and moderate variability of the demand process the interarrival time between
-consecutive incoming orders is either drawn from an exponential ($\Exp$) or an uniform ($\Unif$)
-distribution, which are adjusted to yield the desired bottleneck utilisation level of either $70\%$
-with $\Exp(135)$ and $\Unif(95,175)$, or $80\%$ with $\Exp(118)$ and $\Unif(78,158)$. These values
-were chosen as they comprise the non-linear relationship between flow times and high utilisation
-levels.
+To simulate moderate, medium and high variability of the demand process the interarrival time
+between consecutive incoming orders is either drawn from an exponential ($\Exp$) or an uniform
+($\Unif$) distribution, which are adjusted to yield the desired bottleneck utilisation level of
+either $70\%$ with $\Exp(135)$ and $\Unif(95,175)$, $80\%$ with $\Exp(118)$ and $\Unif(78,158)$, or
+$90\%$ with $\Exp(105)$ and $\Unif(65,145)$. These values were chosen as they comprise the
+non-linear relationship between flow times and high utilisation levels.
 
@@
 solution space and stabilise the learning process. These include the above discussed independent
-agents that operate on the same state, \(n\)-step learning \citep{mnih2016asynchronous}
+agents that operate on the same state, \(n\)-step learning \citep{mnih2016asynchronous},
 % overestimating the average reward \(\avgrew^{\pol}\)
 and ensuring to constantly reduce the average reward as searching for policies with higher average
-costs is unnecessary. Algorithm~\ref{alg:full} depicts a detailed view on the algorithm including
-these extensions and with the objective to minimise the returned rewards. The idea is the same as in
-Algorithm~\ref{alg:near}, except that the action selection procedure is a loop over all independent
-agents and the implementation of the n-step logic.
+costs is unnecessary. Algorithm~\ref{alg:full}
+depicts a detailed view on the algorithm including these extensions and with the objective to
+minimise the returned rewards. The idea is the same as in Algorithm~\ref{alg:near}, except that the
+action selection procedure is a loop over all independent agents and the implementation of the
+n-step logic.
+%
+The algorithm is based around The agent first choose an action (steps \(4-5\)), executes the
+selected action and stores the experience in the replay memory (step \(6\)) and then samples a
+mini-batch of experiences each of length \(nstep\) to improve the policy (steps \(7-21\)). Step
+\(17\) uses an exponentially smoothed overestimation (by \(2.5\%\)) of the average reward and rate
+\(2^{-5}\). All average reward updates above this overestimation are rejected.
 
@@
 environment to collect a broader range of experiences~\citep[as e.g. in][]{mnih2016asynchronous}.
-
+%
 % Shared Rho (Yes)
@@
 The algorithm initiation strategy is to first fill the replay memory buffer by skipping any learning
-(steps \(10-25\)) until the period, that corresponds to the replay memory size, is reached. Then for
-\(500\) steps the average reward is set to the last computed exponentially smoothed value to prevent
-divergence due to the untrained network weight initialisation values.
+(steps \(7-21\) in Algorithm~\ref{alg:full}) until the period, that corresponds to the replay
+memory size, is reached. Then for \(500\) steps the average reward is set to the last computed
+exponentially smoothed value to prevent divergence due to the untrained network weight
+initialisation values.
 
@@
     \State{\(\isRandAct{t} \gets \mathsf{False}\)}
-    \ForEach{independent agent \(\agent \in \agent_{1},\agent_{2},\ldots,\agent_{n_{\mathsf{ag}}}\)}
+    \ForEach{independent agent \(i \in \agent_{1},\agent_{2},\ldots,\agent_{n}\)}
     \State{}%
     \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
-      With probability \(\frac{\pexp}{n_{\mathsf{ag}}}\) choose a random action and
-      probability \(1-\frac{\pexp}{n_{\mathsf{ag}}}\) one that fulfills
-      \(\min_{a \in \mathcal{A}(\agent)}\lexeq(\X^{\pol}_{\gamma_{1}}(s_{t},a; \netWorker{}),\X^\pol_ {\gamma_{0}}(s_{t},a;
-      \netWorker{}))\). Set \(\isRandAct{t} \gets \mathsf{True}\) if the action was chosen randomly
+      With probability \(\frac{\pexp}{n}\) choose action \(a_{t,i}\) randomly or otherwise one of
+      the set defined by
+      \(\min_{a \in \mathcal{A}(i)}\lexeq(\X^{\pol}_{\gamma_{1}}(s_{t},a; \netWorker{}),\X^\pol_
+      {\gamma_{0}}(s_{t},a; \netWorker{}))\). Set \(\isRandAct{t} \gets \mathsf{True}\) if the
+      action was chosen randomly.
     \end{minipage}
-    \EndFor
+    \EndFor{}
     \State{}%
     \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth}
-    Carry out combined action \(a_{t}\), observe reward \(r_{t}\) and resulting state \(s_{t+1}\).
-    Store the experience \((s_{t}, a_{t}, \isRandAct{t}, r_{t}, s_{t+1})\) in the experience replay
-    memory \(M\).
+      Carry out action \(a_{t}=(a_{t,\agent_{1}},\ldots,a_{t,\agent_{n}})\), observe reward
+      \(r_{t}\) and resulting state \(s_{t+1}\). Store experience
+      \((s_{t}, a_{t}, \isRandAct{t}, r_{t}, s_{t+1})\) in the experience replay memory \(M\).
     \end{minipage}
@@
     \ForEach{experience, starting with the latest seen experience \(\nstep\)}
-    \If{experience \(\nstep\) or (\(\isRandAct{i}\) and  \(\pexp \leqslant \plearn\))}
+    \If{experience \(\nstep\) or \(\neg \isRandAct{i}\) or  \(\pexp > \plearn\)}
       \State{\(\target_{\gamma0} \gets \min_{a} \X^\pol_{\gamma_0}(s_{i+1},a; \netTarget{})\)}
@@
     \EndIf
-    \If{\(\isRandAct{i}\) or \(\pexp > \plearn\)}
+    \If{\(\neg \isRandAct{i}\) or \(\pexp > \plearn\)}
       \State{\(\avgrew^{\pol} \gets (1- \alpha) \avgrew^{\pol} + \alpha [r_{i} +
@@
         \netWorker{})]\)}
-    \State{\(\avgrew^{\pol} \gets \min(\avgrew^{\pol}, \avgrew_{\max}^{\pol})\)}
-    \State{\(\avgrew_{\max}^{\pol} \gets (1- \alpha_{\max}) \avgrew_{\max}^{\pol} +
-    \alpha_{\max} 1.025 \avgrew^{\pol}\)}
+      \State{Ensure improvement of \(\avgrew^{\pol}\), otherwise reset to old value}
+
+    % \State{\(\avgrew^{\pol} \gets \min(\avgrew^{\pol}, \avgrew_{\max}^{\pol})\)}
+    % \State{\(\avgrew_{\max}^{\pol} \gets (1- \alpha_{\max}) \avgrew_{\max}^{\pol} +
+    % \alpha_{\max} 1.025 \avgrew^{\pol}\)}
     \EndIf
-    \State{Compute overestimate of \(\avgrew^{\pol}\) and store result in \(\avgrew_{\Overestim}^{\pol}\)}
-    \State{\(\target_{\gamma_0}  \gets r_{i} + \gamma_{0} \target_{\gamma_{0}} - \avgrew_{\Overestim}^{\pol}  \)}
-    \State{\(\target_{\gamma_{1}}  \gets r_{i} + \gamma_{1} \target_{\gamma_{1}} - \avgrew_{\Overestim}^{\pol}\)}
+    % \State{Compute overestimate of \(\avgrew^{\pol}\) and store result in \(\avgrew_{\Overestim}^{\pol}\)}
+    \State{\(\target_{\gamma_0}  \gets r_{i} + \gamma_{0} \target_{\gamma_{0}} - \avgrew^{\pol}  \)}
+    \State{\(\target_{\gamma_{1}}  \gets r_{i} + \gamma_{1} \target_{\gamma_{1}} - \avgrew^{\pol}\)}
     \State{\(d\netWorker{} \gets d\netWorker{} + \partial{(\target_{\gamma_{0}} - \X^\pol_{\gamma_{0}}(s_{i+1},a; \netWorker{}))}^{2} / \partial\netWorker{}\)}
@@
 %
-%
+
+\subsection{Deep Q-Learning}
+\MS[t]{explain here, alg in appendix?}
 
@@
 
-This section provides results for the described setup and gives an overview of the current
-performance of the algorithm compared to the conventional static lead time setting algorithms.
+This section presents the experimental results and gives an overview of the performance of the
+established algorithm compared to the conventional static lead time setting algorithms.
 
@@
   \footnotesize
-  \begin{tabular}{lP{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}}
+  \begin{tabular}{lP{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}}
     \toprule
-    Parameter                      & $\alpha$   & $\alpha_{\max}$ & $\gamma_{0}$ & $\gamma_{1}$ & $\gamma$ & $\epsilon$ & $p_{exp}$ & $\plearn$ \\
+    Parameter                      & $\alpha$   & $\gamma_{0}$ & $\gamma_{1}$ & $\gamma$ & $\epsilon$ & $p_{exp}$ & $\plearn$ \\
     \midrule
-    Start value                    & 0.01       & \(2^{-5}\)      & 0.8          & 0.99/1.0     & 0.01     & 0.25       & 1.0       & 0.5       \\
-    Exp. Smth. Rate                & 0.25       &                 &              &              & 0.25     &            & 0.25      &           \\
-    Exp. Smth. Steps in \(10^{3}\) & 50         &                 &              &              & 100      &            & 100       &           \\
-    Minimum value                  & \(5^{-5}\) &                 &              &              & 0.001         &            & 0.005     &           \\
+    Start value                    & 0.01       & 0.8          & 0.99/1.0     & 0.01     & 0.25       & 1.0       & 0.5       \\
+    Exp. Smth. Rate                & 0.25       &              &              & 0.25     &            & 0.25      &           \\
+    Exp. Smth. Steps in \(10^{3}\) & 50         &              &              & 100      &            & 100       &           \\
+    Minimum value                  & \(5^{-5}\) &              &              & 0.001    &            & 0.005     &           \\
     \bottomrule
     \toprule
-    Parameter                      & \multicolumn{8}{c}{Value} \\
+    Parameter                      & \multicolumn{7}{c}{Value}                                                                                  \\
     \midrule
-    Workers Min. Exploration & \multicolumn{8}{c}{0.01, 0.02, 0.03, 0.04, 0.05, 0.10, 0.15, 0.20}\\
-    Training Batch Size & \multicolumn{8}{c}{4}\\
-    Target Network Update & \multicolumn{8}{c}{Exp. smooth. with rate \(\gamma\) every \(C=100\) steps}\\
-    N-Step & \multicolumn{8}{c}{5}\\
-    Training Mini-Batch Size & \multicolumn{8}{c}{\(m = 4\)}\\
-    Adam ANN Backpropagation & \multicolumn{8}{c}{\(\alpha = 0.005, \beta_{1}=0.9,
-                       \beta_{2}=0.999, \epsilon=10^{-8}, \lambda=0.001}\\ % OptAdam 0.005 0.9 0.999
+    Workers Min. Exploration       & \multicolumn{7}{c}{0.01, 0.02, 0.03, 0.04, 0.05, 0.10, 0.15, 0.20}                                         \\
+    Training Batch Size            & \multicolumn{7}{c}{4}                                                                                      \\
+    Target Network Update          & \multicolumn{7}{c}{Exp. smooth. with rate \(\gamma\) every \(C=100\) steps}                                \\
+    N-Step                         & \multicolumn{7}{c}{5}                                                                                      \\
+    Training Mini-Batch Size       & \multicolumn{7}{c}{\(m = 4\)}                                                                              \\
+    Adam ANN Backpropagation       & \multicolumn{7}{c}{\(\alpha = 0.005, \beta_{1}=0.9, \beta_{2}=0.999, \epsilon=10^{-8}, \lambda=0.001\)}      \\ % OptAdam 0.005 0.9 0.999
 % 1e-8 1e-3,
-        Replay Memory Size & \multicolumn{8}{c}{\(10800\)}\\
+        Replay Memory Size         & \multicolumn{7}{c}{\(10800\)}                                                                              \\
     \bottomrule
@@
 \end{table}
-
+%
 Table~\ref{tbl:params} presents an overview of the parameter setup. The parameters \(\alpha\),
@@
   % \(0.05 ^{(t/350k)}\), where \(t\) is the current period}.
-\MS[t]{TODO} As further comparison we implement the deep Q-Learning network algorithm proposed
+\MS[t]{TODO} As further comparison we implement the deep Q-Learning algorithm proposed
 by~\cite{mnih2015human}, where for comparison reasons we also integrate n-Step learning, the
@@
 
+\MS[t]{Exp - Exp in simulation model?}
 \begin{table}[tb]
@@
     \begin{tabular}{c|rrr|c|c||c|c|c|c|c}
-      \hline
+      \toprule
       80-\(\Exp\)-\(\Unif\)   & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTT+FGIT & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
-      \hline
-      \rl{0.995}              & 55.22 & 137.00 & 52.09   &          & 244.32 & 1.51 & 2.60    & 0.915  & 34.902   & 32.791             \\
-      \rl{1.000}              & 67.27 & 102.18 & 175.69  &          & 345.13 & 1.52 & 2.63    & 0.908  & 49.305   & 30.777             \\
-      \BILOne{}               & 45.90 & 0.00   & 734.43  &          & 780.33 & 1.32 & 1.81    & 0.352  &          &                    \\
-      \BILTwo{}               & 45.73 & 80.31  & 1425.22 &          & 268.57 & 1.32 & 2.158    & 0.876  &          &                    \\
-      \BILThree{}             & 45.86 & 279.72 & 30.82   &          & 356.40 & 1.32 & 3.034    & 0.975  &          &                    \\
-      \hline
+      \midrule
+      \rl{0.995}              &  &  &  &          &  &  &    &   &    &              \\
+      \rl{1.000}              &  &  &  &          &  &  &    &   &    &              \\
+      \ql{1.000}              &  &  &  &          &  &  &    &   &    &              \\
+      \BILOne{}               &  &  &  &          &  &  &    &   &    &              \\
+      \BILTwo{}               &  &  &  &          &  &  &    &   &    &              \\
+      \BILThree{}             &  &  &  &          &  &  &    &   &    &              \\
+      \bottomrule
 %
@@
 \bibliography{references}
+\vfill
+\pagebreak[4!]
+\appendix
+
+\section{Deep Q-Learning Algorithm}
+
+Algorithm~\ref{alg:ql} presents the detailed deep Q-Learning algorithm. It is very similar to the
+established \ARA{} algorithm that is given in Algorithm~\ref{alg:full}. However, it does not compute
+the average reward. Furthermore, it also operates on a single estimation of the overall future
+discounted reward.
+
+\begin{algorithm}[h!]
+  \begin{algorithmic}[1]
+    \State{}Initialise state \(s_{0}\) and network parameters \(\netTarget, \netWorker\) randomly,
+    set an exploration rate \(0 < \pexp \leqslant 1\), exponential smoothing learning rates
+    \(0 < \alpha_{\max}, \alpha, \gamma < 1\) and discount factor \(0 < < \gamma_{1} < 1\)
+    \While{the stopping criterion is not fulfilled} \ForEach{independent agent
+      \(i \in \agent_{1},\agent_{2},\ldots,\agent_{n}\)} \State{}%
+    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
+      With probability \(\frac{\pexp}{n}\) choose action \(a_{t,i}\) randomly or otherwise one of
+      the set defined by
+      \(\min_{a \in \mathcal{A}(i)}\X^{\pol}_{\gamma_{1}}(s_{t},a; \netWorker{})\).
+    \end{minipage}
+    \EndFor{}
+    \State{}%
+    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth}
+      Carry out action \(a_{t}=(a_{t,\agent_{1}},\ldots,a_{t,\agent_{n}})\), observe reward
+      \(r_{t}\) and resulting state \(s_{t+1}\). Store experience
+      \((s_{t}, a_{t}, \isRandAct{t}, r_{t}, s_{t+1})\) in the experience replay memory \(M\).
+    \end{minipage}
+    \If{\(t \mod{\nstep} \equiv 0\)}
+    \State{Reset gradient: \(d\netWorker{} \gets 0\)}
+    \State{}%
+    \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
+      Sample \(m\) random mini-batches of \(\nstep\) consecutive experiences from \(M\) numbered by
+      \(1 \leqslant i \leqslant \nstep\), each with
+      \((s_{i}, \isRandAct{i}, a_{i}, r_{i}, s_{i+1})\).
+    \end{minipage}
+    \ForEach{mini-batch}
+    \ForEach{experience, starting with the latest seen experience \(\nstep\)}
+    \If{experience \(\nstep\) or \(\neg \isRandAct{i}\) or  \(\pexp > \plearn\)}
+      \State{\(\target_{\gamma1} \gets \min_{a} \X^\pol_{\gamma_1}(s_{i+1},a; \netTarget{})\)}
+    \EndIf
+    \State{\(\target_{\gamma_{1}}  \gets r_{i} + \gamma_{1} \target_{\gamma_{1}} \)}
+    \State{\(d\netWorker{} \gets d\netWorker{} +
+      \partial{(\target_{\gamma_{1}} - \X^\pol_{\gamma_{1}}(s_{i+1},a; \netWorker{}))}^{2} / \partial\netWorker{}\)}
+    \EndFor{}
+    \EndFor{}
+    \State{Perform update of \(\netWorker{}\) using \(d\netWorker{} / (m \cdot \nstep)\)}
+    \EndIf{}
+    \State{Every \(C\) steps exponential smoothly set target network:
+    \(\netTarget \gets (1 - \gamma) \netTarget{} + \gamma \netWorker{}\)}
+    \State{Set \(s \gets s'\), \(t \gets t+1\) and decay parameters}
+    \EndWhile{}
+  \end{algorithmic}
+  \caption{\label{alg:ql}Deep Q-Learning}
+\end{algorithm}
 

### 2021-04-30 paper/paper.tex
update

@@
   \begin{center}
-    \begin{tabular}{c|rrr|c|c||c|c|c|c|c}
+    \begin{tabular}{crrrccccccc}
       \toprule
-      80-\(\Exp\)-\(\Unif\)   & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTT+FGIT & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
+      70-\(\Unif\)   & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTT+FGIT & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
       \midrule
-      \rl{0.995}              &  &  &  &          &  &  &    &   &    &              \\
-      \rl{1.000}              &  &  &  &          &  &  &    &   &    &              \\
-      \ql{1.000}              &  &  &  &          &  &  &    &   &    &              \\
+      \ql{0.99}              & 38776.300 &133872.800  &82673.600  &          &255322.700  &1.340  & 2.558   & \tard{0.095}  &    &              \\
+      \rl{0.99}              & 37111.000 &113425.600  &98654.400  &          &249191.000  &1.305  & 2.412   & \tard{0.110}  &    &              \\
+      \rl{1.00}              & 38755.000 &156444.000  &59718.400  &          &254917.400  &1.339  & 2.679   & \tard{0.068}  &    &              \\
+      \BILOne{}               &32159.900 &     0.000  &514558.400  &         &546718.300  &1.202  & 1.693   & \tard{0.552}  &    &              \\
+      \BILTwo{}               &32169.900 &83391.200   &106219.200  &         &221780.300  &1.202  & 2.143   & \tard{0.114}  &    &              \\
+      \BILThree{}             &32239.800 &247416.800  &21382.400  &          &301039.000  &1.204  & 3.029   & \tard{0.022}  &    & \\
+      \BILFour{}              &32262.600 &428865.200  &5483.200   &          &466611.000  &1.204  & 4.007   & \tard{0.006}  &    &              \\
+ \bottomrule
+      \toprule
+      80-\(\Unif\)   & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTT+FGIT & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
+      \midrule
+      \ql{0.99}              & 62787.000 & 181322.800 & 119542.400 &          & 363652.200 & 1.674 & 3.017   & \tard{0.102}  &    &              \\
+      \ql{0.99}              &  &  &  &          &  &  &    &   &    &              \\
+      \rl{1.00}              &  &  &  &          &  &  &    &   &    &              \\
       \BILOne{}               &  &  &  &          &  &  &    &   &    &              \\
@@
       \BILThree{}             &  &  &  &          &  &  &    &   &    &              \\
+      \BILFour{}             &  &  &  &          &  &  &    &   &    &              \\
       \bottomrule
-%
-%       \hline
-%       80-\(\Unif\)-\(\Exp\) & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTP+FGI & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
-%       \hline                &
-%        ESSLTA{}             &        &         &          &          &         &       &          &        &          &                    \\
-%        ANN{}                &        &         &          &          &         &       &          &        &          &                    \\
-%        ESSLTB{}             &        &         &          &          &         &       &          &        &          &                    \\
-%        BILD{}               &        &         &          &          &         &       &          &        &          &                    \\
-%       \hline
-% %
-%       \hline
-%       80-\(\Exp\)-\(\Exp\)  & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTP+FGI & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
-%       \hline                &
-%        ANN{}                &        &         &          &          &         &       &          &        &          &                    \\
-%        ESSLTA{}             &        &         &          &          &         &       &          &        &          &                    \\
-%        ESSLTB{}             &        &         &          &          &         &       &          &        &          &                    \\
-%        BILD{}               &        &         &          &          &         &       &          &        &          &                    \\
+      \toprule
+      90-\(\Unif\)   & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTT+FGIT & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
+      \midrule
+      \ql{0.99}              &  &  &  &          &  &  &    &   &    &              \\
+      \rl{0.99}              &  &  &  &          &  &  &    &   &    &              \\
+      \rl{1.00}              &  &  &  &          &  &  &    &   &    &              \\
+      \BILOne{}              &  &  &  &          &  &  &    &   &    &              \\
+      \BILTwo{}              &  &  &  &          &  &  &    &   &    &              \\
+      \BILThree{}            &  &  &  &          &  &  &    &   &    &              \\
+      \BILFour{}             &  &  &  &          &  &  &    &   &    &              \\
+      \bottomrule
     \end{tabular}
   \end{center}
-  \caption{\label{tbl:80} Results for a bottleneck utilisation of 80\%. }
+  \caption{\label{tbl:unif} Results with moderate variability in the system due to uniformly
+    distributed demand interarrival times. }
+\end{table}
+
+\begin{table}[tb]
+  \begin{center}
+    \begin{tabular}{crrrccccccc}
+      \toprule
+      70-\(\Exp\)   & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTT+FGIT & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
+      \midrule
+      \ql{0.99}              & 38776.300 & 133872.800 & 82673.600 &          & 255322.700 &1.340  & 2.558   & \tard{0.095}  &    &              \\
+      \rl{0.99}              & 37111.000 & 113425.600 & 98654.400 &          & 249191.000 &1.305  & 2.412   &  \tard{0.110} &    &              \\
+      \rl{1.00}              & 38755.000 & 156444.000 & 59718.400 &          & 254917.400 &1.339  & 2.679   &  \tard{0.068} &    &              \\
+      \BILOne{}              & 32159.900 & 0.000 & 514558.400 &          & 546718.300 & 1.202 & 1.693   &  \tard{0.552} &    &              \\
+      \BILTwo{}              & 32169.900 & 83391.200 & 106219.200 &          & 221780.300 & 1.202 &  2.143   &  \tard{0.114} &    &              \\
+      \BILThree{}            & 32239.800 & 247416.800 & 21382.400 &          & 301039.000 & 1.204 & 3.029   &  \tard{0.022} &    &              \\
+      \BILFour{}             & 32262.600 & 428865.200 & 5483.200 &          & 466611.000 & 1.204 & 4.007   & \tard{0.006}  &    &              \\
+      \bottomrule
+      \toprule
+      80-\(\Exp\)   & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTT+FGIT & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
+      \midrule
+      \ql{0.99}              &  &  &  &          &  &  &    &   &    &              \\
+      \rl{0.99}              &  &  &  &          &  &  &    &   &    &              \\
+      \rl{1.00}              &  &  &  &          &  &  &    &   &    &              \\
+      \BILOne{}               &  &  &  &          &  &  &    &   &    &              \\
+      \BILTwo{}               &  &  &  &          &  &  &    &   &    &              \\
+      \BILThree{}             &  &  &  &          &  &  &    &   &    &              \\
+      \BILFour{}             &  &  &  &          &  &  &    &   &    &              \\
+      \bottomrule
+      \toprule
+      90-\(\Exp\)   & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTT+FGIT & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
+      \midrule
+      \ql{0.99}              &  &  &  &          &  &  &    &   &    &              \\
+      \rl{0.99}              &  &  &  &          &  &  &    &   &    &              \\
+      \rl{1.00}              &  &  &  &          &  &  &    &   &    &              \\
+      \BILOne{}               &  &  &  &          &  &  &    &   &    &              \\
+      \BILTwo{}               &  &  &  &          &  &  &    &   &    &              \\
+      \BILThree{}             &  &  &  &          &  &  &    &   &    &              \\
+      \BILFour{}             &  &  &  &          &  &  &    &   &    &              \\
+      \bottomrule
+    \end{tabular}
+  \end{center}
+  \caption{\label{tbl:exp} Results with high variability in the system due to exponentially
+    distributed demand interarrival times. }
 \end{table}

### 2021-05-03 paper/paper.tex
update

--- 2021-04-30 paper/paper.tex
@@
   % a discount factor of \(0.99\).
-
-% \definecolor{lightgray}{rgb}{0.83, 0.83, 0.83}
-% \begin{table}[tb]
-%   \begin{center}
-%     \begin{tabular}{c|ccc|c|c||c}
-%       \hline
-%       70-\(\Exp\)-\(\Unif\) & WIPC & FGIC & BOC & FGIC+BOC & SUM & SFTT & SL(\%) \\
-%       \hline                &
-%        ESSLTA{}             &      &      &     &          &     &      &        \\
-%        ANN{}                &      &      &     &          &     &      &        \\
-%        ESSLTB{}             &      &      &     &          &     &      &        \\
-%        BILD{}               &      &      &     &          &     &      &        \\
-%       \hline                &
-% %
-%       \hline                &
-%       70-\(\Unif\)-\(\Exp\) & WIPC & FGIC & BOC & FGIC+BOC & SUM & SFTT & SL(\%) \\
-%       \hline                &
-%        ESSLTA{}             &      &      &     &          &     &      &        \\
-%        ANN{}                &      &      &     &          &     &      &        \\
-%        ESSLTB{}             &      &      &     &          &     &      &        \\
-%        BILD{}               &      &      &     &          &     &      &        \\
-%       \hline                &
-% %
-%       \hline                &
-%       70-\(\Exp\)-\(\Exp\)  & WIPC & FGIC & BOC & FGIC+BOC & SUM & SFTT & SL(\%) \\
-%       \hline                &
-%        ANN{}                &      &      &     &          &     &      &        \\
-%        ESSLTA{}             &      &      &     &          &     &      &        \\
-%        ESSLTB{}             &      &      &     &          &     &      &        \\
-%        BILD{}               &      &      &     &          &     &      &        \\
-%     \end{tabular}
-%   \end{center}
-%   \caption{\label{tbl:70} Results for a bottleneck utilisation of 70\%. }
-% \end{table}
+  %    demandUnif65_145            -- 90%   -- 6.
+  % -- demandUnif78_158 -- 80%   -- 3.
+  % -- demandUnif95_175  -- 70%   -- 1.
+  % -- demandExp105  -- 90%           -- 5.
+  % -- demandExp118  -- 80%          -- 4.
+  % -- demandExp135  -- 70%          -- 2.
 
@@
   \begin{center}
-    \begin{tabular}{c|rrr|c|c||c|c|c|c|c}
+    \begin{tabular}{crrrccccccc}
       \toprule
-      80-\(\Exp\)-\(\Unif\)   & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTT+FGIT & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
+      70-\(\Unif\)   & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & FGIT & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
       \midrule
-      \rl{0.995}              &  &  &  &          &  &  &    &   &    &              \\
-      \rl{1.000}              &  &  &  &          &  &  &    &   &    &              \\
-      \ql{1.000}              &  &  &  &          &  &  &    &   &    &              \\
-      \BILOne{}               &  &  &  &          &  &  &    &   &    &              \\
-      \BILTwo{}               &  &  &  &          &  &  &    &   &    &              \\
-      \BILThree{}             &  &  &  &          &  &  &    &   &    &              \\
+      \ql{0.99}              & \cost{38776.300} & \cost{133872.800}  & \cost{82673.600 } &          & \cost{255322.700}  & \timem{1.340}  & \timem{2.558}   & \tard{0.095}  &    &              \\
+      \rl{0.99}              & \cost{37111.000} & \cost{113425.600}  & \cost{98654.400 } &          & \cost{249191.000}  & \timem{1.305}  & \timem{2.412}   & \tard{0.110}  &    &              \\
+      \rl{1.00}              & \cost{38755.000} & \cost{156444.000}  & \cost{59718.400 } &          & \cost{254917.400}  & \timem{1.339}  & \timem{2.679}   & \tard{0.068}  &    &              \\
+      \BILOne{}              & \cost{32159.900} & \cost{     0.000}  & \cost{514558.400} &          & \cost{546718.300}  & \timem{1.202}  & \timem{1.693}   & \tard{0.552}  &    &              \\
+      \BILTwo{}              & \cost{32169.900} & \cost{ 83391.200}  & \cost{106219.200} &          & \cost{221780.300}  & \timem{1.202}  & \timem{2.143}   & \tard{0.114}  &    &              \\
+      \BILThree{}            & \cost{32239.800} & \cost{247416.800}  & \cost{21382.400 } &          & \cost{301039.000}  & \timem{1.204}  & \timem{3.029}   & \tard{0.022}  &    & \\
+      \BILFour{}             & \cost{32262.600} & \cost{428865.200}  & \cost{5483.200  } &          & \cost{466611.000}  & \timem{1.204}  & \timem{4.007}   & \tard{0.006}  &    &              \\
       \bottomrule
-%
-%       \hline
-%       80-\(\Unif\)-\(\Exp\) & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTP+FGI & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
-%       \hline                &
-%        ESSLTA{}             &        &         &          &          &         &       &          &        &          &                    \\
-%        ANN{}                &        &         &          &          &         &       &          &        &          &                    \\
-%        ESSLTB{}             &        &         &          &          &         &       &          &        &          &                    \\
-%        BILD{}               &        &         &          &          &         &       &          &        &          &                    \\
-%       \hline
-% %
-%       \hline
-%       80-\(\Exp\)-\(\Exp\)  & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTP+FGI & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
-%       \hline                &
-%        ANN{}                &        &         &          &          &         &       &          &        &          &                    \\
-%        ESSLTA{}             &        &         &          &          &         &       &          &        &          &                    \\
-%        ESSLTB{}             &        &         &          &          &         &       &          &        &          &                    \\
-%        BILD{}               &        &         &          &          &         &       &          &        &          &                    \\
+      % 70-\(\Unif\) & WIPC    & FGIC    & BOC     & FGIC+BOC & SUM     & SFTT     & FGIT & SL(\%)  & Cost.p.P & \(\avgrew^{\pol}\) \\
+      % \midrule
+      % \ql{0.99}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \rl{0.99}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \rl{1.00}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \BILOne{}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \BILTwo{}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \BILThree{} & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \BILFour{}  & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \bottomrule
+      % \toprule
+      % 80-\(\Unif\) & WIPC    & FGIC    & BOC     & FGIC+BOC & SUM     & SFTT     & FGIT & SL(\%)  & Cost.p.P & \(\avgrew^{\pol}\) \\
+      % \midrule
+      % \ql{0.99}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \rl{0.99}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \rl{1.00}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \BILOne{}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \BILTwo{}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \BILThree{} & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \BILFour{}  & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \bottomrule
+      % \toprule
+      % 90-\(\Unif\) & WIPC    & FGIC    & BOC     & FGIC+BOC & SUM     & SFTT     & FGIT & SL(\%)  & Cost.p.P & \(\avgrew^{\pol}\) \\
+      % \midrule
+      % \ql{0.99}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \rl{0.99}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \rl{1.00}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \BILOne{}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \BILTwo{}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \BILThree{} & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \BILFour{}  & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      \bottomrule
     \end{tabular}
   \end{center}
-  \caption{\label{tbl:80} Results for a bottleneck utilisation of 80\%. }
+  \caption{\label{tbl:unif} Results with moderate variability in the system due to uniformly
+    distributed demand interarrival times. }
+\end{table}
+
+\begin{table}[tb]
+  \begin{center}
+    \begin{tabular}{crrrccccccc}
+      \toprule
+      70-\(\Exp\) & WIPC    & FGIC    & BOC     & FGIC+BOC & SUM     & SFTT     & FGIT & SL(\%)  & Cost.p.P & \(\avgrew^{\pol}\) \\
+      \midrule
+      \ql{0.99}   & \cost{53873.500} & \cost{142398.800} & \cost{164427.200} & \fgiboc{142398.800}{164427.200}  & \cost{360699.500} & \timem{1.585} & \fgit{1.585}{2.796}  & \tard{0.162} & \costpp{360699.500} & \\
+      \rl{0.99}   & \cost{61419.850} & \cost{181702.200} & \cost{115287.200} & \fgiboc{181702.200}{115287.200}  & \cost{358409.250} & \timem{1.734} & \fgit{1.734}{3.145}  & \tard{0.107} & \costpp{358409.250} & \rnd{50.787} \\
+      \rl{1.00}   & \cost{56578.800} & \cost{199310.000} & \cost{115714.400} & \fgiboc{199310.000}{115714.400}  & \cost{371603.200} & \timem{1.639} & \fgit{1.639}{3.136}  & \tard{0.113} & \costpp{371603.200} & \rnd{50.575} \\
+      \BILOne{}   & \cost{48547.100} & \cost{     0.000} & \cost{776753.600} & \fgiboc{     0.000}{776753.600}  & \cost{825300.700} & \timem{1.479} & \fgit{1.479}{1.975}  & \tard{0.663} & \costpp{825300.700} & \\
+      \BILTwo{}   & \cost{48618.850} & \cost{ 66927.200} & \cost{248638.400} & \fgiboc{ 66927.200}{248638.400}  & \cost{364184.450} & \timem{1.480} & \fgit{1.480}{2.312}  & \tard{0.220} & \costpp{364184.450} & \\
+      \BILThree{} & \cost{48632.050} & \cost{222603.200} & \cost{ 74568.800} & \fgiboc{222603.200}{ 74568.800}  & \cost{345804.050} & \timem{1.481} & \fgit{1.481}{3.094}  & \tard{0.063} & \costpp{345804.050} & \\
+      \BILFour{}  & \cost{48514.200} & \cost{409531.200} & \cost{ 23406.400} & \fgiboc{409531.200}{ 23406.400}  & \cost{481451.800} & \timem{1.478} & \fgit{1.478}{4.029}  & \tard{0.019} & \costpp{481451.800} & \\
+      \bottomrule
+      \toprule
+      80-\(\Exp\) & WIPC    & FGIC    & BOC     & FGIC+BOC & SUM     & SFTT     & FGIT & SL(\%)  & Cost.p.P & \(\avgrew^{\pol}\) \\
+      \midrule
+      \ql{0.99}   & \cost{87706.700} & \cost{227485.400} & \cost{225150.400} & \cost{}  & \cost{540342.500} & \timem{2.043} & \fgit{2.043}{3.541}  & \tard{0.166} & \costpp{540342.500} & \\
+      \rl{0.99}   & \cost{89838.950} & \cost{253476.000} & \cost{185687.200} & \cost{}  & \cost{529002.150} & \timem{2.078} & \fgit{2.078}{3.689}  & \tard{0.124} & \costpp{529002.150} & \rnd{70.527} \\
+      \rl{1.00}   & \cost{87171.500} & \cost{242915.600} & \cost{198456.000} & \cost{}  & \cost{528543.100} & \timem{2.034} & \fgit{2.034}{3.599} & \tard{0.134} & \costpp{528543.100} & \rnd{72.944} \\
+      \BILOne{}   & \cost{78731.850} & \cost{    0.000} & \cost{1259709.600} & \cost{}  & \cost{1338441.450} & \timem{1.884} & \fgit{1.884}{2.382}  & \tard{0.771} & \costpp{1338441.450} &                   \\
+      \BILTwo{}   & \cost{78203.150} & \cost{52303.200} & \cost{548736.000} & \cost{}  & \cost{679242.350} & \timem{1.875} & \fgit{1.875}{2.602}  & \tard{0.350} & \costpp{679242.350} &                    \\
+      \BILThree{} & \cost{79029.850} & \cost{200273.200} & \cost{242100.800} & \cost{}  & \cost{521403.850} & \timem{1.890} & \fgit{1.890}{3.265}  & \tard{0.142} & \costpp{521403.850} &                    \\
+      \BILFour{}  & \cost{79384.750} & \cost{395493.000} & \cost{116943.200} & \cost{}  & \cost{591820.950} & \timem{1.896} & \fgit{1.896}{4.128}  & \tard{0.062} & \costpp{591820.950} &                    \\
+      \bottomrule      % \toprule
+      % 90-\(\Exp\) & WIPC    & FGIC    & BOC     & FGIC+BOC & SUM     & SFTT     & FGIT & SL(\%)  & Cost.p.P & \(\avgrew^{\pol}\) \\
+      % \midrule
+      % \ql{0.99}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \rl{0.99}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \rl{1.00}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \BILOne{}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \BILTwo{}   & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \BILThree{} & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      % \BILFour{}  & \cost{} & \cost{} & \cost{} & \cost{}  & \cost{} & \timem{} & \timem{}  & \tard{} &          &                    \\
+      \bottomrule
+    \end{tabular}
+  \end{center}
+  \caption{\label{tbl:exp} Results with high variability in the system due to exponentially
+    distributed demand interarrival times. }
 \end{table}

### 2021-05-04 paper/paper.tex
paper

--- 2021-04-30 paper/paper.tex
@@
   \begin{center}
-    \begin{tabular}{c|rrr|c|c||c|c|c|c|c}
+    \begin{tabular}{crrrccccccc}
+      % \toprule
+      % 70-\(\Unif\) & WIPC              & FGIC              & BOC                & FGIC+BOC                        & SUM                & SFTT          & FGIT                & SL(\%)       & Cost.p.P             & \(\avgrew^{\pol}\) \\
+      % \midrule
+      % \ql{0.99}    & \cost{38776.300}  & \cost{133872.800} & \cost{82673.600 }  &                                 & \cost{255322.700}  & \timem{1.340} & \timem{2.558}       & \tard{0.095} &                      &                    \\
+      % \rl{0.99}    & \cost{37111.000}  & \cost{113425.600} & \cost{98654.400 }  &                                 & \cost{249191.000}  & \timem{1.305} & \timem{2.412}       & \tard{0.110} &                      &                    \\
+      % \rl{1.00}    & \cost{38755.000}  & \cost{156444.000} & \cost{59718.400 }  &                                 & \cost{254917.400}  & \timem{1.339} & \timem{2.679}       & \tard{0.068} &                      &                    \\
+      % \BILOne{}    & \cost{32159.900}  & \cost{     0.000} & \cost{514558.400}  &                                 & \cost{546718.300}  & \timem{1.202} & \timem{1.693}       & \tard{0.552} &                      &                    \\
+      % \BILTwo{}    & \cost{32169.900}  & \cost{ 83391.200} & \cost{106219.200}  &                                 & \cost{221780.300}  & \timem{1.202} & \timem{2.143}       & \tard{0.114} &                      &                    \\
+      % \BILThree{}  & \cost{32239.800}  & \cost{247416.800} & \cost{21382.400 }  &                                 & \cost{301039.000}  & \timem{1.204} & \timem{3.029}       & \tard{0.022} &                      &                    \\
+      % \BILFour{}   & \cost{32262.600}  & \cost{428865.200} & \cost{5483.200  }  &                                 & \cost{466611.000}  & \timem{1.204} & \timem{4.007}       & \tard{0.006} &                      &                    \\
+      % \bottomrule
       \toprule
-      80-\(\Exp\)-\(\Unif\)   & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTT+FGIT & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
+      70-\(\Unif\)   & WIPC              & FGIC              & BOC                & FGIC+BOC                        & SUM                & SFTT          & FGIT                & SL(\%)       & Cost.p.P             & \(\avgrew^{\pol}\) \\
       \midrule
-      \rl{0.995}              &  &  &  &          &  &  &    &   &    &              \\
-      \rl{1.000}              &  &  &  &          &  &  &    &   &    &              \\
-      \ql{1.000}              &  &  &  &          &  &  &    &   &    &              \\
-      \BILOne{}               &  &  &  &          &  &  &    &   &    &              \\
-      \BILTwo{}               &  &  &  &          &  &  &    &   &    &              \\
-      \BILThree{}             &  &  &  &          &  &  &    &   &    &              \\
+      \ql{0.99}      & \cost{38563.000}  & \cost{134703.600} & \cost{81704.800}   & \fgiboc{134703.600}{81704.800}  & \cost{254971.400}  & \timem{1.335} & \fgit{1.335}{2.558} & \tard{0.094} & \costpp{254971.400}  &                    \\
+      \rl{0.99}      & \cost{37197.750}  & \cost{113100.800} & \cost{100398.400}  & \fgiboc{113100.800}{100398.400} & \cost{250696.950}  & \timem{1.307} & \fgit{1.307}{2.412} & \tard{0.111} & \costpp{250696.950}  & \rnd{35.040}       \\
+      \rl{1.00}      & \cost{38659.100}  & \cost{156659.200} & \cost{59743.200}   & \fgiboc{156659.200}{59743.200}  & \cost{255061.500}  & \timem{1.338} & \fgit{1.338}{2.678} & \tard{0.068} & \costpp{255061.500}  & \rnd{36.501}       \\
+      \BILOne{}      & \cost{32056.500}  & \cost{0.000}      & \cost{512904.000}  & \fgiboc{0.000}{512904.000}      & \cost{544960.500}  & \timem{1.200} & \fgit{1.200}{1.691} & \tard{0.551} & \costpp{544960.500}  &                    \\
+      \BILTwo{}      & \cost{32063.850}  & \cost{83461.000}  & \cost{105009.600}  & \fgiboc{83461.000}{105009.600}  & \textbf{\cost{220534.450}}  & \timem{1.200} & \fgit{1.200}{2.142} & \tard{0.114} & \textbf{\costpp{220534.450}}  &                    \\
+      \BILThree{}    & \cost{32155.150}  & \cost{247727.000} & \cost{21680.800 }  & \fgiboc{247727.000}{21680.800 } & \cost{301562.950}  & \timem{1.202} & \fgit{1.202}{3.029} & \tard{0.022} & \costpp{301562.950}  &                    \\
+      \BILFour{}     & \cost{32140.550}  & \cost{429097.800} & \cost{5076.800 }   & \fgiboc{429097.800}{5076.800 }  & \cost{466315.150}  & \timem{1.202} & \fgit{1.202}{4.007} & \tard{0.005} & \costpp{466315.150}  &                    \\
       \bottomrule
-%
-%       \hline
-%       80-\(\Unif\)-\(\Exp\) & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTP+FGI & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
-%       \hline                &
-%        ESSLTA{}             &        &         &          &          &         &       &          &        &          &                    \\
-%        ANN{}                &        &         &          &          &         &       &          &        &          &                    \\
-%        ESSLTB{}             &        &         &          &          &         &       &          &        &          &                    \\
-%        BILD{}               &        &         &          &          &         &       &          &        &          &                    \\
-%       \hline
-% %
-%       \hline
-%       80-\(\Exp\)-\(\Exp\)  & WIPC   & FGIC    & BOC      & FGIC+BOC & SUM     & SFTT  & SFTP+FGI & SL(\%) & Cost.p.P & \(\avgrew^{\pol}\) \\
-%       \hline                &
-%        ANN{}                &        &         &          &          &         &       &          &        &          &                    \\
-%        ESSLTA{}             &        &         &          &          &         &       &          &        &          &                    \\
-%        ESSLTB{}             &        &         &          &          &         &       &          &        &          &                    \\
-%        BILD{}               &        &         &          &          &         &       &          &        &          &                    \\
+      \toprule
+      80-\(\Unif\)   & WIPC              & FGIC              & BOC                & FGIC+BOC                        & SUM                & SFTT          & FGIT                & SL(\%)       & CostpP               & \(\avgrew^{\pol}\) \\
+      \midrule
+      \ql{0.99}      & \cost{63095.100}  & \cost{180403.800} & \cost{121528.000}  & \fgiboc{180403.800}{121528.000} & \textbf{\cost{365026.900}}  & \timem{1.680} & \fgit{1.680}{3.019} & \tard{0.103} & \textbf{\costpp{365026.900}} &                    \\
+      \rl{0.99}      & \cost{65310.650}  & \cost{168489.200} & \cost{148593.600}  & \fgiboc{168489.200}{148593.600} & \cost{382393.450}  & \timem{1.721} & \fgit{1.721}{3.005} & \tard{0.136} & \costpp{382393.450}  & \rnd{52.918}       \\
+      \rl{1.00}      & \cost{60778.500}  & \cost{234045.000} & \cost{76168.000}   & \fgiboc{234045.000}{76168.000}  & \cost{370991.500}  & \timem{1.637} & \fgit{1.637}{3.226} & \tard{0.066} & \costpp{370991.500}  & \rnd{52.291}       \\
+      \BILOne{}      & \cost{51888.650}  & \cost{0.000}      & \cost{830218.400}  & \fgiboc{0.000}{830218.400}      & \cost{882107.050}  & \timem{1.474} & \fgit{1.474}{1.968} & \tard{0.672} & \costpp{882107.050}  &                    \\
+      \BILTwo{}      & \cost{51998.950}  & \cost{70108.800}  & \cost{255166.400}  & \fgiboc{70108.800}{255166.400}  & \cost{377274.150}  & \timem{1.476} & \fgit{1.476}{2.298} & \tard{0.206} & \costpp{377274.150}  &                    \\
+      \BILThree{}    & \cost{51976.550}  & \cost{240175.000} & \cost{77822.400}   & \fgiboc{240175.000}{77822.400}  & \cost{369973.950}  & \timem{1.475} & \fgit{1.475}{3.091} & \tard{0.060} & \costpp{369973.950}  &                    \\
+      \BILFour{}     & \cost{51827.050}  & \cost{442307.600} & \cost{26711.200}   & \fgiboc{442307.600}{26711.200}  & \cost{520845.850}  & \timem{1.472} & \fgit{1.472}{4.031} & \tard{0.019} & \costpp{520845.850}  &                    \\
+      \bottomrule
+      \toprule
+      90-\(\Unif\)   & WIPC              & FGIC              & BOC                & FGIC+BOC                        & SUM                & SFTT          & FGIT                & SL(\%)       & CostpP               & \(\avgrew^{\pol}\) \\
+      \midrule
+      \ql{0.99}      & \cost{98909.150}  & \cost{190681.000} & \cost{297333.600}  & \fgiboc{190681.000}{297333.600} & \cost{586923.750}  & \timem{2.133} & \fgit{2.133}{3.417} & \tard{0.173} & \costpp{586923.750}  &                    \\
+      \rl{0.99}      & \cost{105348.200} & \cost{263950.600} & \cost{207183.200}  & \fgiboc{263950.600}{207183.200} & \textbf{\cost{576482.000}}  & \timem{2.238} & \fgit{2.238}{3.825} & \tard{0.121} & \textbf{\costpp{576482.000}}  & \rnd{79.826}       \\
+      \rl{1.00}      & \cost{107873.800} & \cost{240798.800} & \cost{253865.600}  & \fgiboc{240798.800}{253865.600} & \cost{602538.200}  & \timem{2.280} & \fgit{2.280}{3.771} & \tard{0.153} & \costpp{602538.200}  & \rnd{80.425}       \\
+      \BILOne{}      & \cost{88658.100}  & \cost{0.000}      & \cost{1418529.600} & \fgiboc{0.000}{1418529.600}     & \cost{1507187.700} & \timem{1.964} & \fgit{1.964}{2.462} & \tard{0.790} & \costpp{1507187.700} &                    \\
+      \BILTwo{}      & \cost{88317.700}  & \cost{50854.400}  & \cost{646000.800}  & \fgiboc{50854.400}{646000.800}  & \cost{785172.900}  & \timem{1.958} & \fgit{1.958}{2.666} & \tard{0.355} & \costpp{785172.900}  &                    \\
+      \BILThree{}    & \cost{88252.450}  & \cost{207210.600} & \cost{299885.600}  & \fgiboc{207210.600}{299885.600} & \cost{595348.650}  & \timem{1.957} & \fgit{1.957}{3.309} & \tard{0.148} & \costpp{595348.650}  &                    \\
+      \BILFour{}     & \cost{87774.600}  & \cost{415178.200} & \cost{153612.800}  & \fgiboc{415178.200}{153612.800} & \cost{656565.600}  & \timem{1.950} & \fgit{1.950}{4.158} & \tard{0.071} & \costpp{656565.600}  &                    \\
+      \bottomrule
     \end{tabular}
   \end{center}
-  \caption{\label{tbl:80} Results for a bottleneck utilisation of 80\%. }
+  \caption{\label{tbl:unif} Results with moderate variability in the system due to uniformly
+    distributed demand interarrival times. }
 \end{table}
 
-% \definecolor{lightgray}{rgb}{0.83, 0.83, 0.83}
-% \begin{table}[tb]
-%   \begin{center}
-%     \begin{tabular}{c|ccc|c|c||c}
-%       \hline
-%       80-\(\Exp\)-\(\Unif\)   & WIPC      & FGIC      & BOC       & FGIC+BOC  & SUM   & SL(\%) \\
-%       \hline
-%        ESSLTA{}     &           &           &           &           &       &                  \\
-%        ANN{}        &           &           &           &       &       &                      \\
-%        ESSLTB{}     &           &           &           &       &       &                      \\
-%        BILD{}       &           &           &           &       &       &                      \\
-%       \hline
-% %
-%       \hline
-%       80-\(\Unif\)-\(\Exp\)   & WIPC      & FGIC      & BOC       & FGIC+BOC  & SUM    & SL(\%) \\
-%       \hline
-%        ESSLTA{}     &           &           &           &           &       &      \\
-%        ANN{}        &           &           &           &           &       &      \\
-%        ESSLTB{}     &           &           &           &           &       &      \\
-%        BILD{}       &           &           &           &           &       &      \\
-%       \hline
-% %
-%       \hline
-%       80-\(\Exp\)-\(\Exp\) & WIPC     & FGIC      & BOC       & FGIC+BOC  & SUM   & SL(\%) \\
-%       \hline
-%        ANN{}     &          &           &           &           &       &      \\
-%        ESSLTA{}  &          &           &           &           &       &      \\
-%        ESSLTB{}  &          &           &           &           &       &      \\
-%        BILD{}    &          &           &           &           &       &      \\
-%     \end{tabular}
-%   \end{center}
-%   \caption{\label{tbl:80} Results for a bottleneck utilisation of 70\%. }
-% \end{table}
+\begin{comment}
+* Unif-70%
+
+           QL	AvgQL0.99	AvgQL1.0	BIL1	    BIL2	  BIL3	  BIL4
+P_1    251303	   243743	  259512	540549	213541	295577	467766
+P_2    264665	   247607	  253436	533630	217064	296323	466820
+P_3    260043	   257121	  256038	564332	232896	303723	465187
+P_4    250869	   256036	  254776	541620	221602	301724	465467
+P_5    254079	   250814	  252465	546975	225815	300616	468210
+P_6    247780	   248298	  254246	549508	220674	300177	467039
+P_7    265459	   248659	  257828	546295	220635	305236	464263
+P_8    251896	   245786	  254077	544850	219209	301524	466451
+P_9    254502	   245544	  254622	545734	221235	302580	466944
+P_10   252631	   248302	  252174	553690	225132	302910	467963
+P_11   253068	   258304	  253873	539393	216099	302757	465670
+P_12   260997	   254848	  250599	547757	227971	303835	463119
+P_13   253329	   248691	  256341	549066	219881	306322	469024
+P_14   254715	   267491	  259271	559487	216614	304241	465459
+P_15   251007	   248119	  259367	528088	213454	301483	464478
+P_16   254686	   250503	  257132	552262	219856	304220	466426
+P_17   250830	   249925	  255187	529669	214652	298888	468042
+P_18   259411	   249452	  252322	545394	217661	298885	464701
+P_19   261476	   240635	  256826	535296	224470	300230	465799
+P_20   246682	   254061	  251138	545615	222228	300008	467475
+
+- Friedman chi-squared statistic: 112.778571
+- p-value: 5.335902e-22
+- This p-value is for rejection of the omnibus null hypothesis, that all samples (groups) are from
+  the same distribution, i.e. there is no effect of groups. The alternate hypothesis that one or
+  more of the correlated samples (groups) is different, i.e. there is a group effect.
+- The omnibus p-value is at or below the respectable critical threshold of 0.05, so post-hoc
+  pairwise multiple comparison tests are conducted to discern which of the pairs have significantly
+  differences. Post-hoc tests conducted: Conover, where the p-value is adjusted by the false
+  discovery rate (FDR) procedure of Benjaminyi-Hochberg.
+
+  Post-hoc p-values of all possible pairs (of samples/ groups) are compactly represented as a lower
+  triangular matrix. Each numerical entry is the p-value of row/column pair, i.e. the null
+  hypotheses that the group represented by a particular column name is different from the group
+  represented by a particular row name. Conover p-values, further adjusted by the
+  Benjaminyi-Hochberg FDR method
+
+  * Unif-80%
+
+      QL	AvgQL0.99	AvgQL1.0	BIL1	    BIL2	  BIL3	  BIL4
+      \ql{0.99} \rl{0.99} \rl{1.00} \BILOne{} \BILTwo{} \BILThree{} \BILFour{}
+P_1  387809	388915	383835	885904	393992	398202	514142
+P_2  359157	379690	363164	886975	384114	376660	515336
+P_3  383065	376724	360581	894013	379834	360186	520716
+P_4  376629	379476	381806	882742	360301	358114	528800
+P_5  347754	383895	370965	864484	375807	366606	523167
+P_6  357321	384470	369683	871624	366181	364609	530412
+P_7  358944	383526	358428	911472	366868	363209	514686
+P_8  347414	376654	385126	895509	392616	383019	532197
+P_9  360972	383002	369881	899334	378537	355148	513209
+P_10 357457	374808	368923	866065	355299	359736	513984
+P_11 366141	373215	369628	860455	396973	377169	520757
+P_12 368787	388541	361227	894268	378112	376437	513099
+P_13 370668	397366	371919	900575	368307	357662	514924
+P_14 388434	384845	375453	861305	399392	369540	522818
+P_15 367029	381957	369656	884187	390026	362042	520379
+P_16 367760	383548	371404	860149	363596	370414	530716
+P_17 349081	372732	369545	855763	344912	353143	515141
+P_18 361824	382886	372029	904230	376210	400139	531858
+P_19 366121	398561	379479	889253	403883	378528	515531
+P_20 358171	373058	367098	873834	370523	368916	525045
+
+- Friedman chi-squared statistic: 90.578571
+- p-value: 2.297453e-17
+- rejection of the omnibus null hypothesis
+
+* Unif-90%
+
+      \ql{0.99} \rl{0.99} \rl{1.00} \BILOne{} \BILTwo{} \BILThree{} \BILFour{}
+P_1   653676	521714	583292	1411442	701676	548748	673075
+P_2   678801	606952	610287	1599513	862870	608274	655640
+P_3   600219	602535	594998	1496442	864121	605376	630665
+P_4   662569	518558	612290	1417290	835203	580285	657643
+P_5   557821	607801	610968	1522469	859054	570829	680786
+P_6   561714	553561	569823	1493178	776052	612004	621577
+P_7   577479	538003	539180	1473951	790533	564949	704889
+P_8   538715	603415	575945	1598306	798817	602557	610314
+P_9   569077	528191	571533	1479493	713889	602314	631613
+P_10   615113	527510	650835	1522554	744289	551199	598620
+P_11   558858	561689	580324	1526107	755665	596688	643571
+P_12   564356	564868	594560	1517267	730110	587742	616851
+P_13   567569	619918	619763	1433219	755946	647755	730470
+P_14   576338	619782	636887	1512337	800148	543174	681310
+P_15   585967	581851	658242	1548853	907391	615699	652752
+P_16   594172	609886	596591	1561824	836903	616500	718513
+P_17   535899	570047	600196	1531139	707526	537206	621174
+P_18   560293	612670	645596	1488537	782649	584652	673597
+P_19   633629	571462	627070	1523982	745810	659741	683916
+P_20   546210	609227	572384	1485851	734806	671281	644336
+
+- Friedman chi-squared statistic: 96.150000
+- p-value: 1.592809e-18
+- rejection of the omnibus null hypothesis
+
+\end{comment}
+
+\begin{table}[tb]
+  \begin{center}
+    \begin{tabular}{crrrccccccc}
+      \toprule
+      70-\(\Exp\) & WIPC    & FGIC    & BOC     & FGIC+BOC & SUM     & SFTT     & FGIT & SL(\%)  & Cost.p.P & \(\avgrew^{\pol}\) \\
+      \midrule
+      \ql{0.99}   & \cost{53873.500} & \cost{142398.800} & \cost{164427.200} & \fgiboc{142398.800}{164427.200}  & \cost{360699.500} & \timem{1.585} & \fgit{1.585}{2.796}  & \tard{0.162} & \costpp{360699.500} & \\
+      \rl{0.99}   & \cost{61419.850} & \cost{181702.200} & \cost{115287.200} & \fgiboc{181702.200}{115287.200}  & \cost{358409.250} & \timem{1.734} & \fgit{1.734}{3.145}  & \tard{0.107} & \costpp{358409.250} & \rnd{50.787} \\
+      \rl{1.00}   & \cost{56578.800} & \cost{199310.000} & \cost{115714.400} & \fgiboc{199310.000}{115714.400}  & \cost{371603.200} & \timem{1.639} & \fgit{1.639}{3.136}  & \tard{0.113} & \costpp{371603.200} & \rnd{50.575} \\
+      \BILOne{}   & \cost{48547.100} & \cost{     0.000} & \cost{776753.600} & \fgiboc{     0.000}{776753.600}  & \cost{825300.700} & \timem{1.479} & \fgit{1.479}{1.975}  & \tard{0.663} & \costpp{825300.700} & \\
+      \BILTwo{}   & \cost{48618.850} & \cost{ 66927.200} & \cost{248638.400} & \fgiboc{ 66927.200}{248638.400}  & \cost{364184.450} & \timem{1.480} & \fgit{1.480}{2.312}  & \tard{0.220} & \costpp{364184.450} & \\
+      \BILThree{} & \cost{48632.050} & \cost{222603.200} & \cost{ 74568.800} & \fgiboc{222603.200}{ 74568.800}  & \textbf{\cost{345804.050}} & \timem{1.481} & \fgit{1.481}{3.094}  & \tard{0.063} & \textbf{\costpp{345804.050}} & \\
+      \BILFour{}  & \cost{48514.200} & \cost{409531.200} & \cost{ 23406.400} & \fgiboc{409531.200}{ 23406.400}  & \cost{481451.800} & \timem{1.478} & \fgit{1.478}{4.029}  & \tard{0.019} & \costpp{481451.800} & \\
+      \bottomrule
+      \toprule
+      80-\(\Exp\) & WIPC    & FGIC    & BOC     & FGIC+BOC & SUM     & SFTT     & FGIT & SL(\%)  & Cost.p.P & \(\avgrew^{\pol}\) \\
+      \midrule
+      \ql{0.99}   & \cost{87706.700} & \cost{227485.400} & \cost{225150.400} & \fgiboc{227485.400}{225150.400}  & \cost{540342.500} & \timem{2.043} & \fgit{2.043}{3.541}  & \tard{0.166} & \costpp{540342.500} & \\
+      \rl{0.99}   & \cost{89838.950} & \cost{253476.000} & \cost{185687.200} & \fgiboc{253476.000}{185687.200}  & \cost{529002.150} & \timem{2.078} & \fgit{2.078}{3.689}  & \tard{0.124} & \costpp{529002.150} & \rnd{70.527} \\
+      \rl{1.00}   & \cost{87171.500} & \cost{242915.600} & \cost{198456.000} & \fgiboc{242915.600}{198456.000}  & \cost{528543.100} & \timem{2.034} & \fgit{2.034}{3.599} & \tard{0.134} & \costpp{528543.100} & \rnd{72.944} \\
+      \BILOne{}   & \cost{78731.850} & \cost{     0.000} & \cost{1259709.60} & \fgiboc{     0.000}{1259709.60}  & \cost{1338441.45} & \timem{1.884} & \fgit{1.884}{2.382}  & \tard{0.771} & \costpp{1338441.450} & \\
+      \BILTwo{}   & \cost{78203.150} & \cost{ 52303.200} & \cost{548736.000} & \fgiboc{ 52303.200}{548736.000}  & \cost{679242.350} & \timem{1.875} & \fgit{1.875}{2.602}  & \tard{0.350} & \costpp{679242.350} & \\
+      \BILThree{} & \cost{79029.850} & \cost{200273.200} & \cost{242100.800} & \fgiboc{200273.200}{242100.800}  & \textbf{\cost{521403.850}} & \timem{1.890} & \fgit{1.890}{3.265}  & \tard{0.142} & \textbf{\costpp{521403.850}} & \\
+      \BILFour{}  & \cost{79384.750} & \cost{395493.000} & \cost{116943.200} & \fgiboc{395493.000}{116943.200}  & \cost{591820.950} & \timem{1.896} & \fgit{1.896}{4.128}  & \tard{0.062} & \costpp{591820.950} & \\
+      \bottomrule
+      \toprule
+      90-\(\Exp\) & WIPC    & FGIC    & BOC     & FGIC+BOC & SUM     & SFTT     & FGIT & SL(\%)  & Cost.p.P & \(\avgrew^{\pol}\) \\
+      \midrule
+      \ql{0.99}   & \cost{168655.950} & \cost{222145.800} & \cost{788662.400} & \fgiboc{222145.800}{788662.400}  & \cost{1179464.150} & \timem{3.134} & \fgit{3.134}{4.501}  & \tard{0.285} & \costpp{1179464.150} & \\
+      \rl{0.99}   & \cost{158034.800} & \cost{212726.200} & \cost{701300.800} & \fgiboc{212726.200}{701300.800}  & \cost{1072061.800} & \timem{2.970} & \fgit{2.970}{4.300}  & \tard{0.281} & \costpp{1072061.800} & \rnd{115.801} \\
+      \rl{1.00}   & \cost{160453.950} & \cost{286065.800} & \cost{539803.200} & \fgiboc{286065.800}{539803.200}  & \textbf{\cost{986322.950}} & \timem{3.007} & \fgit{3.007}{4.623}  & \tard{0.204} & \textbf{\costpp{986322.950}} & \rnd{109.907} \\
+      \BILOne{}   & \cost{146835.900} & \cost{     0.000} & \cost{2349374.40} & \fgiboc{     0.000}{2349374.40}  & \cost{2496210.300} & \timem{2.796} & \fgit{2.796}{3.295}  & \tard{0.867} & \costpp{2496210.300} & \\
+      \BILTwo{}   & \cost{143756.000} & \cost{ 34112.800} & \cost{1412752.00} & \fgiboc{ 34112.800}{1412752.00}  & \cost{1590620.800} & \timem{2.748} & \fgit{2.748}{3.380}  & \tard{0.529} & \costpp{1590620.800} & \\
+      \BILThree{} & \cost{144278.250} & \cost{154146.800} & \cost{877450.400} & \fgiboc{154146.800}{877450.400}  & \cost{1175875.450} & \timem{2.756} & \fgit{2.756}{3.857}  & \tard{0.294} & \costpp{1175875.450} & \\
+      \BILFour{}  & \cost{147077.800} & \cost{335070.400} & \cost{622141.600} & \fgiboc{335070.400}{622141.600}  & \cost{1104289.800} & \timem{2.800} & \fgit{2.800}{4.608}  & \tard{0.175} & \costpp{1104289.800} & \\
+      \bottomrule
+    \end{tabular}
+  \end{center}
+  \caption{\label{tbl:exp} Results with high variability in the system due to exponentially
+    distributed interarrival times of the demand. }
+\end{table}
+
+\begin{comment}
+
+* 70-Exp
+
+      \ql{0.99} \rl{0.99} \rl{1.00} \BILOne{} \BILTwo{} \BILThree{} \BILFour{}
+P_1   351947	374561	379567	834003	386432	362168	492681
+P_2   347668	352692	378234	822817	364552	333854	479469
+P_3   367401	390552	366704	864960	381170	361897	482144
+P_4   368674	359179	387919	835193	381497	349521	479063
+P_5   348052	359409	367586	818363	368443	346388	481219
+P_6   372165	357376	385714	820607	381450	351540	477010
+P_7   358072	353692	375693	799663	354775	342411	478984
+P_8   359080	367498	371778	863328	387183	361328	484675
+P_9   362323	351285	364367	839001	371559	335595	489409
+P_10   363594	363521	361902	819961	367956	355976	482128
+P_11   351111	351607	359002	806276	345417	336380	478077
+P_12   360053	358579	372742	823191	343266	337228	481849
+P_13   362083	355953	375080	800258	347386	334618	484658
+P_14   362155	380252	375057	843183	354535	358173	479977
+P_15   362034	352998	365647	817717	351729	338243	477084
+P_16   368261	356893	379373	839426	367362	359818	471416
+P_17   368115	346666	367777	806327	357377	342397	483288
+P_18   356904	343756	367760	816272	361651	341250	489722
+P_19   370278	342598	365683	818686	357727	328406	477375
+P_20   354020	349118	364479	816782	352222	338890	478808
+
+- Friedman chi-squared statistic: 100.242857
+- p-value: 2.232967e-19
+- rejection of the omnibus null hypothesis
+
+* 80-Exp
+
+      \ql{0.99} \rl{0.99} \rl{1.00} \BILOne{} \BILTwo{} \BILThree{} \BILFour{}
+P_1  524150	491399	492299	1253359	627883	492535	576044
+P_2  551981	537349	534693	1378972	702719	546737	629820
+P_3  568969	541461	548726	1323977	685412	522401	579507
+P_4  547969	596543	592121	1364182	705144	536206	590773
+P_5  595489	518330	531775	1334075	658382	532177	639777
+P_6  535908	566964	536670	1429615	731598	520384	572011
+P_7  557047	518934	569058	1336778	648986	497168	563390
+P_8  538454	540953	512740	1395377	724787	530955	586574
+P_9  522057	547783	539467	1387591	709643	522574	604104
+P_10  545500	524853	523222	1307334	710386	511576	591671
+P_11  550026	519641	518909	1312791	661352	529574	602428
+P_12  520436	513871	520666	1302676	627306	495800	579365
+P_13  534303	530579	517976	1377731	693986	553154	570858
+P_14  527004	530556	538978	1407617	740972	528730	663494
+P_15  546023	548754	539483	1324827	759339	560228	594110
+P_16  551856	525486	542790	1326459	673565	553607	582936
+P_17  509943	494150	491463	1302608	648931	482086	581970
+P_18  538290	482416	487062	1275391	608823	502715	558249
+P_19  515500	518251	507451	1331100	627861	495598	567573
+P_20  525945	531770	525313	1296369	637772	513872	601765
+
+- Friedman chi-squared statistic: 97.628571
+- p-value: 7.835701e-19
+- rejection of the omnibus null hypothesis
+
+      \ql{0.99} \rl{0.99} \rl{1.00} \BILOne{} \BILTwo{} \BILThree{} \BILFour{}
+P_1   1252433	1173395	1024728	2399890	1655026	1159432	1012319
+P_2   1074569	870660	941377	2341852	1466107	1214910	971734
+P_3   1188157	965337	871242	2510900	1409012	1050054	1260441
+P_4   911949	1044494	758677	2345507	1566310	1118957	1130965
+P_5   1033474	897575	836501	2197182	1219749	998558	946565
+P_6   1216651	1228266	933766	2692086	1685058	1391199	1272896
+P_7   1015338	983438	1040244	2399295	1429570	958909	951637
+P_8   1512067	878268	1430194	2565419	1444940	1067515	906021
+P_9   1108092	1232311	1309534	2641681	1850877	1153341	1079572
+P_10   1344449	1264977	968913	2375393	1761920	1118482	1042407
+P_11   1188536	1274797	1168751	3040365	1669772	1673853	1853585
+P_12   1271137	985585	915254	2327385	1599041	1175113	1205382
+P_13   1337521	1080134	1114726	2410549	1531861	1078113	1003334
+P_14   1347283	1098738	861594	2728058	1566173	1387567	916554
+P_15   1060109	1000957	947592	2422075	1704239	1016923	1070665
+P_16   1095527	922726	957775	2363731	1637310	1232300	1017197
+P_17   1157196	1095892	1011316	2669102	1959411	1245607	1212991
+P_18   1171769	1047445	804936	2506905	1765175	1140291	1122519
+P_19   1162809	1312587	949697	2481932	1361007	1166369	1052711
+P_20   1140217	1083654	879642	2504899	1529858	1170016	1056301
+
+- Friedman chi-squared statistic: 85.800000
+- p-value: 2.253674e-16
+- rejection of the omnibus null hypothesis
+
+\end{comment}
 
@@
 
+\section{}
+
+\begin{table}
+  \begin{center}
+    \begin{tabular}{lP{1.75cm}P{1.75cm}P{1.75cm}P{1.75cm}P{1.75cm}P{1.75cm}}
+      \toprule
+      70-\(\Unif\) & \ql{0.99} & \rl{0.99} & \rl{1.00} & \BILOne{} & \BILTwo{} & \BILThree{} \\
+      \midrule
+      \rl{0.99}    & 3.37e-04  &           &           &           &           &             \\
+      \rl{1.0}     & 3.55e-01  & 1.04e-05  &           &           &           &             \\
+      \BILOne{}    & 1.94e-45  & 2.48e-51  & 6.90e-44  &           &           &             \\
+      \BILTwo{}    & 1.09e-24  & 3.35e-16  & 9.57e-27  & 3.66e-64  &           &             \\
+      \BILThree{}  & 1.78e-20  & 9.27e-29  & 2.44e-18  & 1.35e-22  & 6.26e-47  &             \\
+      \BILFour{}   & 6.30e-34  & 3.85e-41  & 4.83e-32  & 1.16e-08  & 3.00e-56  & 1.16e-08    \\
+      \bottomrule
+      \toprule
+      80-\(\Unif\) & \ql{0.99} & \rl{0.99} & \rl{1.00} & \BILOne{} & \BILTwo{} & \BILThree{} \\
+      \midrule
+      \rl{0.99}    & 4.78e-16  &           &           &           &           &             \\
+      \rl{1.00}    & 5.26e-04  & 3.96e-08  &           &           &           &             \\
+      \BILOne{}    & 1.02e-40  & 1.98e-21  & 3.92e-34  &           &           &             \\
+      \BILTwo{}    & 6.98e-10  & 7.12e-03  & 2.06e-03  & 1.31e-27  &           &             \\
+      \BILThree{}  & 2.16e-02  & 8.93e-11  & 2.03e-01  & 1.57e-36  & 2.46e-05  &             \\
+      \BILFour{}   & 6.16e-33  & 1.10e-11  & 2.74e-25  & 5.33e-05  & 5.38e-18  & 5.30e-28    \\
+      \bottomrule
+      \toprule
+      90-\(\Unif\) & \ql{0.99} & \rl{0.99} & \rl{1.00} & \BILOne{} & \BILTwo{} & \BILThree{} \\
+      \midrule
+      \rl{0.99}    & 5.18e-01  &           &           &           &           &             \\
+      \rl{1.00}    & 1.94e-04  & 1.56e-05  &           &           &           &             \\
+      \BILOne{}    & 8.40e-42  & 9.61e-43  & 1.10e-34  &           &           &             \\
+      \BILTwo{}    & 2.36e-33  & 1.10e-34  & 5.21e-25  & 1.56e-05  &           &             \\
+      \BILThree{}  & 4.16e-04  & 3.64e-05  & 8.19e-01  & 6.27e-35  & 1.81e-25  &             \\
+      \BILFour{}   & 2.33e-20  & 6.57e-22  & 1.79e-11  & 2.93e-18  & 4.39e-08  & 5.97e-12    \\
+      \bottomrule
+    \end{tabular}
+  \end{center}
+  \caption{\label{tbl:stats-unif}Results from the statistical analysis for the results of of
+    moderate variability in the system. }
+\end{table}
+
+\begin{table}
+  \begin{center}
+    \begin{tabular}{lP{1.75cm}P{1.75cm}P{1.75cm}P{1.75cm}P{1.75cm}P{1.75cm}}
+      \toprule
+      70-\(\Exp\) & \ql{0.99} & \rl{0.99} & \rl{1.00} & \BILOne{} & \BILTwo{} & \BILThree{} \\
+      \midrule
+      \rl{0.99}   & 1.75e-02  &           &           &           &           &             \\
+      \rl{1.00}   & 6.15e-07  & 5.50e-12  &           &           &           &             \\
+      \BILOne{}   & 7.62e-35  & 1.74e-39  & 1.32e-23  &           &           &             \\
+      \BILTwo{}   & 3.33e-01  & 1.02e-03  & 3.17e-05  & 6.06e-33  &           &             \\
+      \BILThree{} & 5.96e-17  & 1.79e-11  & 6.09e-29  & 7.00e-52  & 3.47e-19  &             \\
+      \BILFour{}  & 1.16e-24  & 6.25e-30  & 1.68e-12  & 4.68e-06  & 1.56e-22  & 3.43e-44    \\
+      \bottomrule
+      \toprule
+      80-\(\Exp\) & \ql{0.99} & \rl{0.99} & \rl{1.00} & \BILOne{} & \BILTwo{} & \BILThree{} \\
+      \midrule
+      \bottomrule
+      \rl{0.99}   & 7.07e-02  &           &           &           &           &             \\
+      \rl{1.00}   & 1.74e-03  & 1.72e-01  &           &           &           &             \\
+      \BILOne{}   & 2.39e-35  & 6.77e-39  & 2.38e-41  &           &           &             \\
+      \BILTwo{}   & 1.22e-25  & 1.00e-29  & 1.21e-32  & 1.13e-05  &           &             \\
+      \BILThree{} & 2.72e-05  & 1.33e-02  & 2.44e-01  & 3.65e-43  & 5.68e-35  &             \\
+      \BILFour{}  & 2.07e-13  & 1.12e-17  & 6.55e-21  & 3.62e-17  & 6.09e-07  & 1.45e-23    \\
+      \toprule
+      90-\(\Exp\) & \ql{0.99} & \rl{0.99} & \rl{1.00} & \BILOne{} & \BILTwo{} & \BILThree{} \\
+      \midrule
+      \rl{0.99}   & 2.61e-06  &           &           &           &           &             \\
+      \rl{1.00}   & 5.03e-13  & 1.74e-03  &           &           &           &             \\
+      \BILOne{}   & 9.64e-23  & 2.73e-33  & 2.22e-39  &           &           &             \\
+      \BILTwo{}   & 3.46e-12  & 1.32e-23  & 1.30e-30  & 1.23e-05  &           &             \\
+      \BILThree{} & 6.86e-01  & 1.23e-05  & 3.46e-12  & 1.32e-23  & 5.03e-13  &             \\
+      \BILFour{}  & 2.45e-04  & 2.38e-01  & 2.60e-05  & 6.39e-31  & 6.30e-21  & 9.46e-4     \\
+      \bottomrule
+    \end{tabular}
+  \end{center}
+  \caption{\label{tbl:stats-unif}Results from the statistical analysis for the results of of
+    moderate variability in the system. }
+\end{table}
+
 \end{document}

### 2021-05-04 paper/paper.tex
update

@@
 % \MS[t]{todo} We evaluate two different kinds of reward reporting as they have been proposed by
-% \cite{Schneckenreither2019}. They find that it is beneficial to overcome the time offset imposed by
-% the production system. That is, although orders are immediately released to the production system
-% and thus generate a response in terms of a reward to the agent, the full impact of the chosen action
-% takes several periods to materialise. For instance, by deciding not to release an order, which later
-% becomes a backorder, it is beneficial to reward the actions actually responsible for the backorder
-% instead of the last action (only). Thus they propose to keep track of the orders currently in the
-% order pool until all fully traversed through the production system and reward the actions according
-% to all orders in the order pool.
+% \cite{Schneckenreither2019}. They find that it is beneficial to overcome the time offset imposed
+% by the production system. That is, although orders are immediately released to the production
+% system and thus generate a response in terms of a reward to the agent, the full impact of the
+% chosen action takes several periods to materialise. For instance, by deciding not to release an
+% order, which later becomes a backorder, it is beneficial to reward the actions actually
+% responsible for the backorder instead of the last action (only). Thus they propose to keep track
+% of the orders currently in the order pool until all fully traversed through the production system
+% and reward the actions according to all orders in the order pool.
 
@@
       \midrule
-      \ql{0.99}      & \cost{38563.000}  & \cost{134703.600} & \cost{81704.800}   & \fgiboc{134703.600}{81704.800}  & \cost{254971.400}  & \timem{1.335} & \fgit{1.335}{2.558} & \tard{0.094} & \costpp{254971.400}  &                    \\
+      \ql{0.99}      & \cost{38563.000}  & \cost{134703.600} & \cost{81704.800}   & \fgiboc{134703.600}{81704.800}  & \cost{254971.400}$^{1}$  & \timem{1.335} & \fgit{1.335}{2.558} & \tard{0.094} & \costpp{254971.400}  &                    \\
       \rl{0.99}      & \cost{37197.750}  & \cost{113100.800} & \cost{100398.400}  & \fgiboc{113100.800}{100398.400} & \cost{250696.950}  & \timem{1.307} & \fgit{1.307}{2.412} & \tard{0.111} & \costpp{250696.950}  & \rnd{35.040}       \\
-      \rl{1.00}      & \cost{38659.100}  & \cost{156659.200} & \cost{59743.200}   & \fgiboc{156659.200}{59743.200}  & \cost{255061.500}  & \timem{1.338} & \fgit{1.338}{2.678} & \tard{0.068} & \costpp{255061.500}  & \rnd{36.501}       \\
+      \rl{1.00}      & \cost{38659.100}  & \cost{156659.200} & \cost{59743.200}   & \fgiboc{156659.200}{59743.200}  & \cost{255061.500}$^{1}$  & \timem{1.338} & \fgit{1.338}{2.678} & \tard{0.068} & \costpp{255061.500}  & \rnd{36.501}       \\
       \BILOne{}      & \cost{32056.500}  & \cost{0.000}      & \cost{512904.000}  & \fgiboc{0.000}{512904.000}      & \cost{544960.500}  & \timem{1.200} & \fgit{1.200}{1.691} & \tard{0.551} & \costpp{544960.500}  &                    \\
@@
       \rl{0.99}      & \cost{65310.650}  & \cost{168489.200} & \cost{148593.600}  & \fgiboc{168489.200}{148593.600} & \cost{382393.450}  & \timem{1.721} & \fgit{1.721}{3.005} & \tard{0.136} & \costpp{382393.450}  & \rnd{52.918}       \\
-      \rl{1.00}      & \cost{60778.500}  & \cost{234045.000} & \cost{76168.000}   & \fgiboc{234045.000}{76168.000}  & \cost{370991.500}  & \timem{1.637} & \fgit{1.637}{3.226} & \tard{0.066} & \costpp{370991.500}  & \rnd{52.291}       \\
+      \rl{1.00}      & \cost{60778.500}  & \cost{234045.000} & \cost{76168.000}   & \fgiboc{234045.000}{76168.000}  & \cost{370991.500}$^{1}$  & \timem{1.637} & \fgit{1.637}{3.226} & \tard{0.066} & \costpp{370991.500}  & \rnd{52.291}       \\
       \BILOne{}      & \cost{51888.650}  & \cost{0.000}      & \cost{830218.400}  & \fgiboc{0.000}{830218.400}      & \cost{882107.050}  & \timem{1.474} & \fgit{1.474}{1.968} & \tard{0.672} & \costpp{882107.050}  &                    \\
       \BILTwo{}      & \cost{51998.950}  & \cost{70108.800}  & \cost{255166.400}  & \fgiboc{70108.800}{255166.400}  & \cost{377274.150}  & \timem{1.476} & \fgit{1.476}{2.298} & \tard{0.206} & \costpp{377274.150}  &                    \\
-      \BILThree{}    & \cost{51976.550}  & \cost{240175.000} & \cost{77822.400}   & \fgiboc{240175.000}{77822.400}  & \cost{369973.950}  & \timem{1.475} & \fgit{1.475}{3.091} & \tard{0.060} & \costpp{369973.950}  &                    \\
+      \BILThree{}    & \cost{51976.550}  & \cost{240175.000} & \cost{77822.400}   & \fgiboc{240175.000}{77822.400}  & \cost{369973.950}$^{1}$  & \timem{1.475} & \fgit{1.475}{3.091} & \tard{0.060} & \costpp{369973.950}  &                    \\
       \BILFour{}     & \cost{51827.050}  & \cost{442307.600} & \cost{26711.200}   & \fgiboc{442307.600}{26711.200}  & \cost{520845.850}  & \timem{1.472} & \fgit{1.472}{4.031} & \tard{0.019} & \costpp{520845.850}  &                    \\
@@
       \midrule
-      \ql{0.99}      & \cost{98909.150}  & \cost{190681.000} & \cost{297333.600}  & \fgiboc{190681.000}{297333.600} & \cost{586923.750}  & \timem{2.133} & \fgit{2.133}{3.417} & \tard{0.173} & \costpp{586923.750}  &                    \\
-      \rl{0.99}      & \cost{105348.200} & \cost{263950.600} & \cost{207183.200}  & \fgiboc{263950.600}{207183.200} & \textbf{\cost{576482.000}}  & \timem{2.238} & \fgit{2.238}{3.825} & \tard{0.121} & \textbf{\costpp{576482.000}}  & \rnd{79.826}       \\
-      \rl{1.00}      & \cost{107873.800} & \cost{240798.800} & \cost{253865.600}  & \fgiboc{240798.800}{253865.600} & \cost{602538.200}  & \timem{2.280} & \fgit{2.280}{3.771} & \tard{0.153} & \costpp{602538.200}  & \rnd{80.425}       \\
+      \ql{0.99}      & \cost{98909.150}  & \cost{190681.000} & \cost{297333.600}  & \fgiboc{190681.000}{297333.600} & \cost{586923.750}$^{1}$  & \timem{2.133} & \fgit{2.133}{3.417} & \tard{0.173} & \costpp{586923.750}  &                    \\
+      \rl{0.99}      & \cost{105348.200} & \cost{263950.600} & \cost{207183.200}  & \fgiboc{263950.600}{207183.200} & \textbf{\cost{576482.000}}$^{1}$  & \timem{2.238} & \fgit{2.238}{3.825} & \tard{0.121} & \textbf{\costpp{576482.000}}  & \rnd{79.826}       \\
+      \rl{1.00}      & \cost{107873.800} & \cost{240798.800} & \cost{253865.600}  & \fgiboc{240798.800}{253865.600} & \cost{602538.200}$^{2}$  & \timem{2.280} & \fgit{2.280}{3.771} & \tard{0.153} & \costpp{602538.200}  & \rnd{80.425}       \\
       \BILOne{}      & \cost{88658.100}  & \cost{0.000}      & \cost{1418529.600} & \fgiboc{0.000}{1418529.600}     & \cost{1507187.700} & \timem{1.964} & \fgit{1.964}{2.462} & \tard{0.790} & \costpp{1507187.700} &                    \\
       \BILTwo{}      & \cost{88317.700}  & \cost{50854.400}  & \cost{646000.800}  & \fgiboc{50854.400}{646000.800}  & \cost{785172.900}  & \timem{1.958} & \fgit{1.958}{2.666} & \tard{0.355} & \costpp{785172.900}  &                    \\
-      \BILThree{}    & \cost{88252.450}  & \cost{207210.600} & \cost{299885.600}  & \fgiboc{207210.600}{299885.600} & \cost{595348.650}  & \timem{1.957} & \fgit{1.957}{3.309} & \tard{0.148} & \costpp{595348.650}  &                    \\
+      \BILThree{}    & \cost{88252.450}  & \cost{207210.600} & \cost{299885.600}  & \fgiboc{207210.600}{299885.600} & \cost{595348.650}$^{2}$  & \timem{1.957} & \fgit{1.957}{3.309} & \tard{0.148} & \costpp{595348.650}  &                    \\
       \BILFour{}     & \cost{87774.600}  & \cost{415178.200} & \cost{153612.800}  & \fgiboc{415178.200}{153612.800} & \cost{656565.600}  & \timem{1.950} & \fgit{1.950}{4.158} & \tard{0.071} & \costpp{656565.600}  &                    \\
@@
   \caption{\label{tbl:unif} Results with moderate variability in the system due to uniformly
-    distributed demand interarrival times. }
+    distributed demand interarrival times.\newline \(^{1,2}\) Cells marked with the same number are
+    not significant different to each other.}
 \end{table}
@@
       \midrule
-      \ql{0.99}   & \cost{53873.500} & \cost{142398.800} & \cost{164427.200} & \fgiboc{142398.800}{164427.200}  & \cost{360699.500} & \timem{1.585} & \fgit{1.585}{2.796}  & \tard{0.162} & \costpp{360699.500} & \\
+      \ql{0.99}   & \cost{53873.500}  & \cost{142398.800} & \cost{164427.200} & \fgiboc{142398.800}{164427.200} & \cost{360699.500}$^{1}$          & \timem{1.585} & \fgit{1.585}{2.796} & \tard{0.162} & \costpp{360699.500}          &                    \\
       \rl{0.99}   & \cost{61419.850}  & \cost{181702.200} & \cost{115287.200} & \fgiboc{181702.200}{115287.200} & \cost{358409.250}                & \timem{1.734} & \fgit{1.734}{3.145} & \tard{0.107} & \costpp{358409.250}          & \rnd{50.787}       \\
@@
       \BILOne{}   & \cost{48547.100}  & \cost{     0.000} & \cost{776753.600} & \fgiboc{     0.000}{776753.600} & \cost{825300.700}                & \timem{1.479} & \fgit{1.479}{1.975} & \tard{0.663} & \costpp{825300.700}          &                    \\
-      \BILTwo{}   & \cost{48618.850} & \cost{ 66927.200} & \cost{248638.400} & \fgiboc{ 66927.200}{248638.400}  & \cost{364184.450} & \timem{1.480} & \fgit{1.480}{2.312}  & \tard{0.220} & \costpp{364184.450} & \\
+      \BILTwo{}   & \cost{48618.850}  & \cost{ 66927.200} & \cost{248638.400} & \fgiboc{ 66927.200}{248638.400} & \cost{364184.450}$^{1}$          & \timem{1.480} & \fgit{1.480}{2.312} & \tard{0.220} & \costpp{364184.450}          &                    \\
       \BILThree{} & \cost{48632.050}  & \cost{222603.200} & \cost{ 74568.800} & \fgiboc{222603.200}{ 74568.800} & \textbf{\cost{345804.050}}       & \timem{1.481} & \fgit{1.481}{3.094} & \tard{0.063} & \textbf{\costpp{345804.050}} &                    \\
@@
       \midrule
-      \ql{0.99}   & \cost{87706.700} & \cost{227485.400} & \cost{225150.400} & \fgiboc{227485.400}{225150.400}  & \cost{540342.500} & \timem{2.043} & \fgit{2.043}{3.541}  & \tard{0.166} & \costpp{540342.500} & \\
-      \rl{0.99}   & \cost{89838.950} & \cost{253476.000} & \cost{185687.200} & \fgiboc{253476.000}{185687.200}  & \cost{529002.150} & \timem{2.078} & \fgit{2.078}{3.689}  & \tard{0.124} & \costpp{529002.150} & \rnd{70.527} \\
-      \rl{1.00}   & \cost{87171.500} & \cost{242915.600} & \cost{198456.000} & \fgiboc{242915.600}{198456.000}  & \cost{528543.100} & \timem{2.034} & \fgit{2.034}{3.599} & \tard{0.134} & \costpp{528543.100} & \rnd{72.944} \\
+      \ql{0.99}   & \cost{87706.700}  & \cost{227485.400} & \cost{225150.400} & \fgiboc{227485.400}{225150.400} & \cost{540342.500}$^{1}$          & \timem{2.043} & \fgit{2.043}{3.541} & \tard{0.166} & \costpp{540342.500}          &                    \\
+      \rl{0.99}   & \cost{89838.950}  & \cost{253476.000} & \cost{185687.200} & \fgiboc{253476.000}{185687.200} & \cost{529002.150}$^{1,2}$        & \timem{2.078} & \fgit{2.078}{3.689} & \tard{0.124} & \costpp{529002.150}          & \rnd{70.527}       \\
+      \rl{1.00}   & \cost{87171.500}  & \cost{242915.600} & \cost{198456.000} & \fgiboc{242915.600}{198456.000} & \cost{528543.100}$^{2,3}$        & \timem{2.034} & \fgit{2.034}{3.599} & \tard{0.134} & \costpp{528543.100}          & \rnd{72.944}       \\
       \BILOne{}   & \cost{78731.850}  & \cost{     0.000} & \cost{1259709.60} & \fgiboc{     0.000}{1259709.60} & \cost{1338441.45}                & \timem{1.884} & \fgit{1.884}{2.382} & \tard{0.771} & \costpp{1338441.450}         &                    \\
       \BILTwo{}   & \cost{78203.150}  & \cost{ 52303.200} & \cost{548736.000} & \fgiboc{ 52303.200}{548736.000} & \cost{679242.350}                & \timem{1.875} & \fgit{1.875}{2.602} & \tard{0.350} & \costpp{679242.350}          &                    \\
-      \BILThree{} & \cost{79029.850} & \cost{200273.200} & \cost{242100.800} & \fgiboc{200273.200}{242100.800}  & \textbf{\cost{521403.850}} & \timem{1.890} & \fgit{1.890}{3.265}  & \tard{0.142} & \textbf{\costpp{521403.850}} & \\
+      \BILThree{} & \cost{79029.850}  & \cost{200273.200} & \cost{242100.800} & \fgiboc{200273.200}{242100.800} & \textbf{\cost{521403.850}}$^{3}$ & \timem{1.890} & \fgit{1.890}{3.265} & \tard{0.142} & \textbf{\costpp{521403.850}} &                    \\
       \BILFour{}  & \cost{79384.750}  & \cost{395493.000} & \cost{116943.200} & \fgiboc{395493.000}{116943.200} & \cost{591820.950}                & \timem{1.896} & \fgit{1.896}{4.128} & \tard{0.062} & \costpp{591820.950}          &                    \\
@@
       \midrule
-      \ql{0.99}   & \cost{168655.950} & \cost{222145.800} & \cost{788662.400} & \fgiboc{222145.800}{788662.400}  & \cost{1179464.150} & \timem{3.134} & \fgit{3.134}{4.501}  & \tard{0.285} & \costpp{1179464.150} & \\
-      \rl{0.99}   & \cost{158034.800} & \cost{212726.200} & \cost{701300.800} & \fgiboc{212726.200}{701300.800}  & \cost{1072061.800} & \timem{2.970} & \fgit{2.970}{4.300}  & \tard{0.281} & \costpp{1072061.800} & \rnd{115.801} \\
+      \ql{0.99}   & \cost{168655.950} & \cost{222145.800} & \cost{788662.400} & \fgiboc{222145.800}{788662.400} & \cost{1179464.150}$^{1}$               & \timem{3.134} & \fgit{3.134}{4.501} & \tard{0.285} & \costpp{1179464.150}         &                    \\
+      \rl{0.99}   & \cost{158034.800} & \cost{212726.200} & \cost{701300.800} & \fgiboc{212726.200}{701300.800} & \cost{1072061.800}$^{2}$               & \timem{2.970} & \fgit{2.970}{4.300} & \tard{0.281} & \costpp{1072061.800}         & \rnd{115.801}      \\
       \rl{1.00}   & \cost{160453.950} & \cost{286065.800} & \cost{539803.200} & \fgiboc{286065.800}{539803.200} & \textbf{\cost{986322.950}}       & \timem{3.007} & \fgit{3.007}{4.623} & \tard{0.204} & \textbf{\costpp{986322.950}} & \rnd{109.907}      \\
@@
       \BILTwo{}   & \cost{143756.000} & \cost{ 34112.800} & \cost{1412752.00} & \fgiboc{ 34112.800}{1412752.00} & \cost{1590620.800}               & \timem{2.748} & \fgit{2.748}{3.380} & \tard{0.529} & \costpp{1590620.800}         &                    \\
-      \BILThree{} & \cost{144278.250} & \cost{154146.800} & \cost{877450.400} & \fgiboc{154146.800}{877450.400}  & \cost{1175875.450} & \timem{2.756} & \fgit{2.756}{3.857}  & \tard{0.294} & \costpp{1175875.450} & \\
-      \BILFour{}  & \cost{147077.800} & \cost{335070.400} & \cost{622141.600} & \fgiboc{335070.400}{622141.600}  & \cost{1104289.800} & \timem{2.800} & \fgit{2.800}{4.608}  & \tard{0.175} & \costpp{1104289.800} & \\
+      \BILThree{} & \cost{144278.250} & \cost{154146.800} & \cost{877450.400} & \fgiboc{154146.800}{877450.400} & \cost{1175875.450}$^{1}$               & \timem{2.756} & \fgit{2.756}{3.857} & \tard{0.294} & \costpp{1175875.450}         &                    \\
+      \BILFour{}  & \cost{147077.800} & \cost{335070.400} & \cost{622141.600} & \fgiboc{335070.400}{622141.600} & \cost{1104289.800}$^{2}$               & \timem{2.800} & \fgit{2.800}{4.608} & \tard{0.175} & \costpp{1104289.800}         &                    \\
       \bottomrule
@@
   \caption{\label{tbl:exp} Results with high variability in the system due to exponentially
-    distributed interarrival times of the demand. }
+    distributed interarrival times of the demand.\newline \(^{1,2,3}\) Cells marked with the same number are
+    not significant different to each other. }
 \end{table}
@@
 - rejection of the omnibus null hypothesis
+
+* 90-Exp
 
@@
 
-\section{}
+\section{Statistical Analysis of the Results}
+
+Table~\ref{tbl:stats-unif} and Table~\ref{tbl:stats-exp} provide the statistical results of the
+computation results. We used the Friedman test with a significance level of \(p=0.05\) for a
+statistical analysis of the mean sum of costs. As expected the omnibus null hypotheses (all samples
+are from the same distribution) are rejected for all cases. Therefore, we conducted pairwise Conover
+post-hoc tests adjusted by the Benjaminyi-Hochberg FDR method~\citep{benjamini1995controlling} to
+reduce liberality. These results are shown in Table~\ref{tbl:stats-unif} and
+Table~\ref{tbl:stats-exp} for the uniformly and exponentially distributed interarrival times of the
+demand respectively. The values given in bold are below the significance level of \(5\%\) and
+therefore indicate which row-column tuples are not statistically significant.
 
@@
       \rl{0.99}    & 3.37e-04          &           &                   &           &           &             \\
-      \rl{1.0}     & 3.55e-01  & 1.04e-05  &           &           &           &             \\
+      \rl{1.0}     & \textbf{3.55e-01} & 1.04e-05  &                   &           &           &             \\
       \BILOne{}    & 1.94e-45          & 2.48e-51  & 6.90e-44          &           &           &             \\
@@
       \BILTwo{}    & 6.98e-10          & 7.12e-03  & 2.06e-03          & 1.31e-27  &           &             \\
-      \BILThree{}  & 2.16e-02  & 8.93e-11  & 2.03e-01  & 1.57e-36  & 2.46e-05  &             \\
+      \BILThree{}  & 2.16e-02          & 8.93e-11  & \textbf{2.03e-01} & 1.57e-36  & 2.46e-05  &             \\
       \BILFour{}   & 6.16e-33          & 1.10e-11  & 2.74e-25          & 5.33e-05  & 5.38e-18  & 5.30e-28    \\
@@
       \midrule
-      \rl{0.99}    & 5.18e-01  &           &           &           &           &             \\
+      \rl{0.99}    & \textbf{5.18e-01} &           &                   &           &           &             \\
       \rl{1.00}    & 1.94e-04          & 1.56e-05  &                   &           &           &             \\
@@
       \BILTwo{}    & 2.36e-33          & 1.10e-34  & 5.21e-25          & 1.56e-05  &           &             \\
-      \BILThree{}  & 4.16e-04  & 3.64e-05  & 8.19e-01  & 6.27e-35  & 1.81e-25  &             \\
+      \BILThree{}  & 4.16e-04          & 3.64e-05  & \textbf{8.19e-01} & 6.27e-35  & 1.81e-25  &             \\
       \BILFour{}   & 2.33e-20          & 6.57e-22  & 1.79e-11          & 2.93e-18  & 4.39e-08  & 5.97e-12    \\
@@
   \end{center}
-  \caption{\label{tbl:stats-unif}Results from the statistical analysis for the results of of
-    moderate variability in the system. }
+  \caption{\label{tbl:stats-unif}P-Values from the statistical analysis for the results of
+    moderate variability in the system. Tuples marked bold are not significantly different.}
 \end{table}
@@
       \BILOne{}   & 7.62e-35          & 1.74e-39          & 1.32e-23          &           &           &             \\
-      \BILTwo{}   & 3.33e-01  & 1.02e-03  & 3.17e-05  & 6.06e-33  &           &             \\
+      \BILTwo{}   & \textbf{3.33e-01} & 1.02e-03          & 3.17e-05          & 6.06e-33  &           &             \\
       \BILThree{} & 5.96e-17          & 1.79e-11          & 6.09e-29          & 7.00e-52  & 3.47e-19  &             \\
@@
       \midrule
-      \bottomrule
-      \rl{0.99}   & 7.07e-02  &           &           &           &           &             \\
-      \rl{1.00}   & 1.74e-03  & 1.72e-01  &           &           &           &             \\
+      \rl{0.99}   & \textbf{7.07e-02} &                   &                   &           &           &             \\
+      \rl{1.00}   & 1.74e-03          & \textbf{1.72e-01} &                   &           &           &             \\
       \BILOne{}   & 2.39e-35          & 6.77e-39          & 2.38e-41          &           &           &             \\
       \BILTwo{}   & 1.22e-25          & 1.00e-29          & 1.21e-32          & 1.13e-05  &           &             \\
-      \BILThree{} & 2.72e-05  & 1.33e-02  & 2.44e-01  & 3.65e-43  & 5.68e-35  &             \\
+      \BILThree{} & 2.72e-05          & 1.33e-02          & \textbf{2.44e-01} & 3.65e-43  & 5.68e-35  &             \\
       \BILFour{}  & 2.07e-13          & 1.12e-17          & 6.55e-21          & 3.62e-17  & 6.09e-07  & 1.45e-23    \\
+      \bottomrule
       \toprule
@@
       \BILTwo{}   & 3.46e-12          & 1.32e-23          & 1.30e-30          & 1.23e-05  &           &             \\
-      \BILThree{} & 6.86e-01  & 1.23e-05  & 3.46e-12  & 1.32e-23  & 5.03e-13  &             \\
-      \BILFour{}  & 2.45e-04  & 2.38e-01  & 2.60e-05  & 6.39e-31  & 6.30e-21  & 9.46e-4     \\
+      \BILThree{} & \textbf{6.86e-01} & 1.23e-05          & 3.46e-12          & 1.32e-23  & 5.03e-13  &             \\
+      \BILFour{}  & 2.45e-04          & \textbf{2.38e-01} & 2.60e-05          & 6.39e-31  & 6.30e-21  & 9.46e-4     \\
       \bottomrule
@@
   \end{center}
-  \caption{\label{tbl:stats-unif}Results from the statistical analysis for the results of of
-    moderate variability in the system. }
+  \caption{\label{tbl:stats-exp}P-Values from the statistical analysis for the results of
+    high variability in the system. Tuples marked bold are not significantly different.}
 \end{table}

### 2021-05-04 paper/paper.tex
minor

@@
   \caption{\label{tbl:unif} Results with moderate variability in the system due to uniformly
-    distributed demand interarrival times.\newline \(^{1,2}\) Cells marked with the same number are
-    not significant different to each other.}
+    distributed demand interarrival times.\newline \(^{1,2}\) Costs marked with the same number are
+    not sign.\@ different to each other.}
 \end{table}
@@
   \caption{\label{tbl:exp} Results with high variability in the system due to exponentially
-    distributed interarrival times of the demand.\newline \(^{1,2,3}\) Cells marked with the same number are
+    distributed interarrival times of the demand.\newline \(^{1,2,3}\) Costs with the same number are
     not significant different to each other. }

### 2021-05-07 paper/figures/productionsystem.tex
update

@@
     an order which initiates the process\\ {\tiny Interarrival-time distributions:
-      $\Exp(135)$ or $\Unif(95,175)$ or $\Exp(118)$ or $\Unif(78,158)$ % or $\Exp(105)$ or $\Unif(65, 145)$
+      $\Exp(135)$ or $\Unif(95,175)$ or $\Exp(118)$ or $\Unif(78,158)$ or $\Exp(105)$ or $\Unif(65, 145)$
     }

### 2021-05-07 paper/paper.tex
update

@@
 %
-To simulate moderate, medium and high variability of the demand process the interarrival time
-between consecutive incoming orders is either drawn from an exponential ($\Exp$) or an uniform
-($\Unif$) distribution, which are adjusted to yield the desired bottleneck utilisation level of
-either $70\%$ with $\Exp(135)$ and $\Unif(95,175)$, $80\%$ with $\Exp(118)$ and $\Unif(78,158)$, or
-$90\%$ with $\Exp(105)$ and $\Unif(65,145)$. These values were chosen as they comprise the
-non-linear relationship between flow times and high utilisation levels.
+To simulate medium and high variability of the demand process the interarrival time between
+consecutive incoming orders is either drawn from an exponential ($\Exp$) or an uniform ($\Unif$)
+distribution, which are adjusted to yield the desired bottleneck utilisation level of either $70\%$
+with $\Exp(135)$ and $\Unif(95,175)$, $80\%$ with $\Exp(118)$ and $\Unif(78,158)$, or $90\%$ with
+$\Exp(105)$ and $\Unif(65,145)$. These values were chosen as they comprise the non-linear
+relationship between flow times and high utilisation levels.
 
@@
 \(\X^\pol_{\gamma_{0}}\) and \(\X^\pol_{\gamma_{1}}\) with the same neural network, while deep
-Q-Learning only approximates the function \(Q_{\gamma}^\pol\).
+Q-Learning \citep[cf.][]{mnih2015human} only approximates the function \(Q_{\gamma}^\pol\).
 %
@@
 %
-The actual values \(\anninp\) and \(\annout\) depend on the used algorithm. In our case case using 6
-independent agents we gain \(\anninp=324\) and \(\annout=36\). For the Q-Learning algorithm, which
-only approximates one function, the output nodes \(\annout\) are halved.
+The actual values \(\anninp\) and \(\annout\) depend on the used algorithm. In our case using 6
+independent agents we gain \(\anninp=324\) and \(\annout=36\).
 %
@@
 
-\subsection{Deep Q-Learning}
-\MS[t]{explain here, alg in appendix?}
+\subsection{Deep Q-Learning Algorithm}
+
+Furthermore, we adapt the deep Q-Learning algorithm of \citet{mnih2015human} with integrated
+extensions as described above, i.e.\@ the adaptions to n-step learning, multiple workers,
+independent agents, scaling of the output, smoothly adapting the target network. The resulting
+Algorithm~\ref{alg:ql} is provided in the Appendix.
+
+To ensure comparability we use the same neural network setup as for \ARA{}, except that we halve the
+number of output nodes, as the Q-Learning algorithm only approximates one state-action value
+function. Therefore the output nodes \(\annout\) are \(18\), while the input nodes \(\anninp=324\)
+stay the same.
 
@@
     \toprule
-    Parameter                      & \multicolumn{7}{c}{Value}                                                                                  \\
+    Configuration                  & \multicolumn{7}{c}{Value}                                                                                  \\
     \midrule
-    Workers Min. Exploration       & \multicolumn{7}{c}{0.01, 0.02, 0.03, 0.04, 0.05, 0.10, 0.15, 0.20}                                         \\
-    Training Batch Size            & \multicolumn{7}{c}{4}                                                                                      \\
+    Workers Min. Exploration       & \multicolumn{7}{c}{\(0.01, 0.02, 0.03, 0.04, 0.05, 0.10, 0.15, 0.20\)}                                         \\
     Target Network Update          & \multicolumn{7}{c}{Exp. smooth. with rate \(\gamma\) every \(C=100\) steps}                                \\
-    N-Step                         & \multicolumn{7}{c}{5}                                                                                      \\
-    Training Mini-Batch Size       & \multicolumn{7}{c}{\(m = 4\)}                                                                              \\
+    N-Step                         & \multicolumn{7}{c}{\(5\)}                                                                                      \\
+    Training Mini-Batch Size       & \multicolumn{7}{c}{\(4\)}                                                                              \\
     Adam ANN Backpropagation       & \multicolumn{7}{c}{\(\alpha = 0.005, \beta_{1}=0.9, \beta_{2}=0.999, \epsilon=10^{-8}, \lambda=0.001\)}      \\ % OptAdam 0.005 0.9 0.999
@@
 Table~\ref{tbl:params} presents an overview of the parameter setup. The parameters \(\alpha\),
-\(\gamma\) and \(p_{exp}\) are exponentially decayed as indicated.% \footnote{The decay is defined as
-  % \(0.05 ^{(t/350k)}\), where \(t\) is the current period}.
-\MS[t]{TODO} As further comparison we implement the deep Q-Learning algorithm proposed
-by~\cite{mnih2015human}, where for comparison reasons we also integrate n-Step learning, the
-experience replay memory adaptions and independent agents. \MS[t]{Discount factor in table?}
+\(\gamma\) and \(p_{exp}\) are exponentially decayed as indicated. Due to the vast possibilities we
+manually optimised the values. The methodology was to investigate the values on very small
+production system with just one machine and product due to the computational complexity of bigger
+sized problems. The resulting values were copied to the problem at hand and further manually
+optimised in pre-experimental runs. The actual used discount factors \(\gamma_{1}\) differ for each
+method and thus are further specified in the result tables.
+% \footnote{The decay is defined as \(0.05 ^{(t/350k)}\), where \(t\) is the current period}.
+In a similar approach we reached the configuration values, where a further increase of the values
+was hindered by the available computational resources.
+
+For the experimental runs we used personal computers with Intel CPUs i7-7700 (16GB RAM) and i9-7900X
+(32GB RAM). For each method we performed \(500.000\) learning steps, which took about \(2\)-\(3\)
+days. For the same experimental setup the inputs/random variables are the same for each period.
+After learning each order release method was evaluated on \(20\) demand streams, each with \(1000\)
+periods of warm-up and \(7000\) periods of evaluation phase. Welch’s procedure was applied to
+approximate the length of the warm-up period (see~\cite{law:simulationc}).
+
+\paragraph{Medium Variability.} The results for medium variability in the system are presented in
+Table~\ref{tbl:unif}, where a uniform demand interarrival time is used.
 
@@
       % \toprule
+      % For 10 replications
       % 70-\(\Unif\) & WIPC              & FGIC              & BOC                & FGIC+BOC                        & SUM                & SFTT          & FGIT                & SL(\%)       & Cost.p.P             & \(\avgrew^{\pol}\) \\
@@
     \State{}Initialise state \(s_{0}\) and network parameters \(\netTarget, \netWorker\) randomly,
-    set an exploration rate \(0 < \pexp \leqslant 1\), exponential smoothing learning rates
-    \(0 < \alpha_{\max}, \alpha, \gamma < 1\) and discount factor \(0 < < \gamma_{1} < 1\)
+    set exploration rate \(0 < \pexp \leqslant 1\), target network adaption rate
+    \(0 < \gamma \ll 1\) and discount factor \(0 < \gamma_{1} < 1\)
     \While{the stopping criterion is not fulfilled} \ForEach{independent agent

### 2021-05-07 paper/paper.tex
explanation setups

@@
 
-\MS[t]{ideas: Similar Selçuk (2013) detect that a high update frequency has a negative effect and
-  thus should be reduced. They conclude that the lead times ought to control the WIP level, instead
-  of the due date reliability as changes are directly seen (Knollmann and Windt, 2013). For a review
-  of analytical works in the context of the LTS see Bendul and Knollmann (2016).}
+% \MS[t]{ideas: Similar Selçuk (2013) detect that a high update frequency has a negative effect and
+%   thus should be reduced. They conclude that the lead times ought to control the WIP level, instead
+%   of the due date reliability as changes are directly seen (Knollmann and Windt, 2013). For a review
+%   of analytical works in the context of the LTS see Bendul and Knollmann (2016).}
 
@@
 \paragraph{Medium Variability.} The results for medium variability in the system are presented in
-Table~\ref{tbl:unif}, where a uniform demand interarrival time is used.
+Table~\ref{tbl:unif}, where a uniform demand interarrival time is used. We use the tuples
+\(70-\Unif\), \(80-\Unif\), and \(90-\Unif\) to indicate the experimental setup of using uniform
+demand interarrival times with a aimed bottleneck utilisation of \(70\%\), \(80\%\), or \(90\%\)
+respectively. The results state the accumulated costs, averaged over all replications, for the work
+in process (WIPC), the finished goods inventory (FGIC), the backorder costs (BOC), as well as the
+sum of the finished good inventory and backorder costs (FGIC+BOC) and the sum over all three cost
+centers (SUM). For a better insight in the timing performance we also provide the average shopfloor
+throughput time (SFTT), i.e.\@ the average time an order spends in the system after being released
+and until reaching the finished goods inventory, the finished goods inventory time (FGIT), as well
+as the service level (SL). The later is given in percentage and states the percentage of orders that
+were produced on time. Finally, the actual measured costs per period (Cost.p.P) and the by the
+algorithm learned average reward \(\avgrew\) are stated. The tested methods are the deep Q-Learning
+algorithm with discount factor \(0.99\) (\(\ql{0.99}\)), the \ARA{} algorithm with discount factors
+\(0.99\) (\(\rl{0.99}\)) and \(1.00\) (\(\rl{1.00}\)), as well as the backward infinite loading
+technique \BIL{} with fixed lead times between \(1\) and \(4\) as indicated by the number
+($\BILOne$-$\BILFour$).
+
+For the case of medium variability with a utilisation of \(70\%\) we can see that
 
@@
   % -- demandExp135  -- 70%          -- 2.
-
-\MS[t]{Exp - Exp in simulation model?}
  \begin{table}[tb]
@@
   \caption{\label{tbl:unif} Results with moderate variability in the system due to uniformly
-    distributed demand interarrival times.\newline \(^{1,2}\) Costs marked with the same number are
+    distributed demand interarrival times. \(^{1,2}\) Costs marked with the same number are
     not sign.\@ different to each other.}
@@
   \caption{\label{tbl:exp} Results with high variability in the system due to exponentially
-    distributed interarrival times of the demand.\newline \(^{1,2,3}\) Costs with the same number are
+    distributed interarrival times of the demand. \(^{1,2,3}\) Costs with the same number are
     not significant different to each other. }

### 2021-05-07 paper/paper.tex
update

@@
 
-\paragraph{Medium Variability.} The results for medium variability in the system are presented in
-Table~\ref{tbl:unif}, where a uniform demand interarrival time is used. We use the tuples
-\(70-\Unif\), \(80-\Unif\), and \(90-\Unif\) to indicate the experimental setup of using uniform
-demand interarrival times with a aimed bottleneck utilisation of \(70\%\), \(80\%\), or \(90\%\)
-respectively. The results state the accumulated costs, averaged over all replications, for the work
-in process (WIPC), the finished goods inventory (FGIC), the backorder costs (BOC), as well as the
-sum of the finished good inventory and backorder costs (FGIC+BOC) and the sum over all three cost
-centers (SUM). For a better insight in the timing performance we also provide the average shopfloor
-throughput time (SFTT), i.e.\@ the average time an order spends in the system after being released
-and until reaching the finished goods inventory, the finished goods inventory time (FGIT), as well
-as the service level (SL). The later is given in percentage and states the percentage of orders that
-were produced on time. Finally, the actual measured costs per period (Cost.p.P) and the by the
-algorithm learned average reward \(\avgrew\) are stated. The tested methods are the deep Q-Learning
-algorithm with discount factor \(0.99\) (\(\ql{0.99}\)), the \ARA{} algorithm with discount factors
-\(0.99\) (\(\rl{0.99}\)) and \(1.00\) (\(\rl{1.00}\)), as well as the backward infinite loading
-technique \BIL{} with fixed lead times between \(1\) and \(4\) as indicated by the number
-($\BILOne$-$\BILFour$).
-
-For the case of medium variability with a utilisation of \(70\%\) we can see that
-
   % a discount factor of \(0.99\).
@@
   \begin{center}
-    \begin{tabular}{crrrccccccc}
+    \begin{tabular}{crrrrrlrrrrr}
       % \toprule
       % For 10 replications
-      % 70-\(\Unif\) & WIPC              & FGIC              & BOC                & FGIC+BOC                        & SUM                & SFTT          & FGIT                & SL(\%)       & Cost.p.P             & \(\avgrew^{\pol}\) \\
-      % \midrule
-      % \ql{0.99}    & \cost{38776.300}  & \cost{133872.800} & \cost{82673.600 }  &                                 & \cost{255322.700}  & \timem{1.340} & \timem{2.558}       & \tard{0.095} &                      &                    \\
-      % \rl{0.99}    & \cost{37111.000}  & \cost{113425.600} & \cost{98654.400 }  &                                 & \cost{249191.000}  & \timem{1.305} & \timem{2.412}       & \tard{0.110} &                      &                    \\
-      % \rl{1.00}    & \cost{38755.000}  & \cost{156444.000} & \cost{59718.400 }  &                                 & \cost{254917.400}  & \timem{1.339} & \timem{2.679}       & \tard{0.068} &                      &                    \\
-      % \BILOne{}    & \cost{32159.900}  & \cost{     0.000} & \cost{514558.400}  &                                 & \cost{546718.300}  & \timem{1.202} & \timem{1.693}       & \tard{0.552} &                      &                    \\
-      % \BILTwo{}    & \cost{32169.900}  & \cost{ 83391.200} & \cost{106219.200}  &                                 & \cost{221780.300}  & \timem{1.202} & \timem{2.143}       & \tard{0.114} &                      &                    \\
-      % \BILThree{}  & \cost{32239.800}  & \cost{247416.800} & \cost{21382.400 }  &                                 & \cost{301039.000}  & \timem{1.204} & \timem{3.029}       & \tard{0.022} &                      &                    \\
-      % \BILFour{}   & \cost{32262.600}  & \cost{428865.200} & \cost{5483.200  }  &                                 & \cost{466611.000}  & \timem{1.204} & \timem{4.007}       & \tard{0.006} &                      &                    \\
+      % 70-\(\Unif\) & WIPC              & FGIC              & BOC                & \fgibocName                     & SUM                        &                          & SFTT          & FGIT                & SL(\%)       & Cost.p.P                     & \(\avgrew^{\pol}\) \\
+      % \midrule     &
+      % \ql{0.99}    & \cost{38776.300}  & \cost{133872.800} & \cost{82673.600 }  &                                 & \cost{255322.700}          &                          & \timem{1.340} & \timem{2.558}       & \tard{0.095} &                              &                    \\
+      % \rl{0.99}    & \cost{37111.000}  & \cost{113425.600} & \cost{98654.400 }  &                                 & \cost{249191.000}          &                          & \timem{1.305} & \timem{2.412}       & \tard{0.110} &                              &                    \\
+      % \rl{1.00}    & \cost{38755.000}  & \cost{156444.000} & \cost{59718.400 }  &                                 & \cost{254917.400}          &                          & \timem{1.339} & \timem{2.679}       & \tard{0.068} &                              &                    \\
+      % \BILOne{}    & \cost{32159.900}  & \cost{     0.000} & \cost{514558.400}  &                                 & \cost{546718.300}          &                          & \timem{1.202} & \timem{1.693}       & \tard{0.552} &                              &                    \\
+      % \BILTwo{}    & \cost{32169.900}  & \cost{ 83391.200} & \cost{106219.200}  &                                 & \cost{221780.300}          &                          & \timem{1.202} & \timem{2.143}       & \tard{0.114} &                              &                    \\
+      % \BILThree{}  & \cost{32239.800}  & \cost{247416.800} & \cost{21382.400 }  &                                 & \cost{301039.000}          &                          & \timem{1.204} & \timem{3.029}       & \tard{0.022} &                              &                    \\
+      % \BILFour{}   & \cost{32262.600}  & \cost{428865.200} & \cost{5483.200  }  &                                 & \cost{466611.000}          &                          & \timem{1.204} & \timem{4.007}       & \tard{0.006} &                              &                    \\
       % \bottomrule
       \toprule
-      70-\(\Unif\)   & WIPC              & FGIC              & BOC                & FGIC+BOC                        & SUM                & SFTT          & FGIT                & SL(\%)       & Cost.p.P             & \(\avgrew^{\pol}\) \\
+      70-\(\Unif\)   & WIPC              & FGIC              & BOC                & \fgibocName                     & SUM                        & \multicolumn{2}{r}{SFTT} & FGIT          & SL(\%)              & Cost.p.P     & \(\avgrew^{\pol}\)                                \\
       \midrule
-      \ql{0.99}      & \cost{38563.000}  & \cost{134703.600} & \cost{81704.800}   & \fgiboc{134703.600}{81704.800}  & \cost{254971.400}$^{1}$  & \timem{1.335} & \fgit{1.335}{2.558} & \tard{0.094} & \costpp{254971.400}  &                    \\
-      \rl{0.99}      & \cost{37197.750}  & \cost{113100.800} & \cost{100398.400}  & \fgiboc{113100.800}{100398.400} & \cost{250696.950}  & \timem{1.307} & \fgit{1.307}{2.412} & \tard{0.111} & \costpp{250696.950}  & \rnd{35.040}       \\
-      \rl{1.00}      & \cost{38659.100}  & \cost{156659.200} & \cost{59743.200}   & \fgiboc{156659.200}{59743.200}  & \cost{255061.500}$^{1}$  & \timem{1.338} & \fgit{1.338}{2.678} & \tard{0.068} & \costpp{255061.500}  & \rnd{36.501}       \\
-      \BILOne{}      & \cost{32056.500}  & \cost{0.000}      & \cost{512904.000}  & \fgiboc{0.000}{512904.000}      & \cost{544960.500}  & \timem{1.200} & \fgit{1.200}{1.691} & \tard{0.551} & \costpp{544960.500}  &                    \\
-      \BILTwo{}      & \cost{32063.850}  & \cost{83461.000}  & \cost{105009.600}  & \fgiboc{83461.000}{105009.600}  & \textbf{\cost{220534.450}}  & \timem{1.200} & \fgit{1.200}{2.142} & \tard{0.114} & \textbf{\costpp{220534.450}}  &                    \\
-      \BILThree{}    & \cost{32155.150}  & \cost{247727.000} & \cost{21680.800 }  & \fgiboc{247727.000}{21680.800 } & \cost{301562.950}  & \timem{1.202} & \fgit{1.202}{3.029} & \tard{0.022} & \costpp{301562.950}  &                    \\
-      \BILFour{}     & \cost{32140.550}  & \cost{429097.800} & \cost{5076.800 }   & \fgiboc{429097.800}{5076.800 }  & \cost{466315.150}  & \timem{1.202} & \fgit{1.202}{4.007} & \tard{0.005} & \costpp{466315.150}  &                    \\
+      \ql{0.99}      & \cost{38563.000}  & \cost{134703.600} & \cost{81704.800}   & \fgiboc{134703.600}{81704.800}  & \cost{254971.400}          & $^{1}$                   & \timem{1.335} & \fgit{1.335}{2.558} & \tard{0.094} & \costpp{254971.400}          &                    \\
+      \rl{0.99}      & \cost{37197.750}  & \cost{113100.800} & \cost{100398.400}  & \fgiboc{113100.800}{100398.400} & \cost{250696.950}          &                          & \timem{1.307} & \fgit{1.307}{2.412} & \tard{0.111} & \costpp{250696.950}          & \rnd{35.040}       \\
+      \rl{1.00}      & \cost{38659.100}  & \cost{156659.200} & \cost{59743.200}   & \fgiboc{156659.200}{59743.200}  & \cost{255061.500}          & $^{1}$                   & \timem{1.338} & \fgit{1.338}{2.678} & \tard{0.068} & \costpp{255061.500}          & \rnd{36.501}       \\
+      \BILOne{}      & \cost{32056.500}  & \cost{0.000}      & \cost{512904.000}  & \fgiboc{0.000}{512904.000}      & \cost{544960.500}          &                          & \timem{1.200} & \fgit{1.200}{1.691} & \tard{0.551} & \costpp{544960.500}          &                    \\
+      \BILTwo{}      & \cost{32063.850}  & \cost{83461.000}  & \cost{105009.600}  & \fgiboc{83461.000}{105009.600}  & \textbf{\cost{220534.450}} &                          & \timem{1.200} & \fgit{1.200}{2.142} & \tard{0.114} & \textbf{\costpp{220534.450}} &                    \\
+      \BILThree{}    & \cost{32155.150}  & \cost{247727.000} & \cost{21680.800 }  & \fgiboc{247727.000}{21680.800 } & \cost{301562.950}          &                          & \timem{1.202} & \fgit{1.202}{3.029} & \tard{0.022} & \costpp{301562.950}          &                    \\
+      \BILFour{}     & \cost{32140.550}  & \cost{429097.800} & \cost{5076.800 }   & \fgiboc{429097.800}{5076.800 }  & \cost{466315.150}          &                          & \timem{1.202} & \fgit{1.202}{4.007} & \tard{0.005} & \costpp{466315.150}          &                    \\
       \bottomrule
       \toprule
-      80-\(\Unif\)   & WIPC              & FGIC              & BOC                & FGIC+BOC                        & SUM                & SFTT          & FGIT                & SL(\%)       & CostpP               & \(\avgrew^{\pol}\) \\
+      80-\(\Unif\)   & WIPC              & FGIC              & BOC                & \fgibocName                     & SUM                        & \multicolumn{2}{r}{SFTT} & FGIT          & SL(\%)              & Cost.p.P     & \(\avgrew^{\pol}\)                                \\
       \midrule
-      \ql{0.99}      & \cost{63095.100}  & \cost{180403.800} & \cost{121528.000}  & \fgiboc{180403.800}{121528.000} & \textbf{\cost{365026.900}}  & \timem{1.680} & \fgit{1.680}{3.019} & \tard{0.103} & \textbf{\costpp{365026.900}} &                    \\
-      \rl{0.99}      & \cost{65310.650}  & \cost{168489.200} & \cost{148593.600}  & \fgiboc{168489.200}{148593.600} & \cost{382393.450}  & \timem{1.721} & \fgit{1.721}{3.005} & \tard{0.136} & \costpp{382393.450}  & \rnd{52.918}       \\
-      \rl{1.00}      & \cost{60778.500}  & \cost{234045.000} & \cost{76168.000}   & \fgiboc{234045.000}{76168.000}  & \cost{370991.500}$^{1}$  & \timem{1.637} & \fgit{1.637}{3.226} & \tard{0.066} & \costpp{370991.500}  & \rnd{52.291}       \\
-      \BILOne{}      & \cost{51888.650}  & \cost{0.000}      & \cost{830218.400}  & \fgiboc{0.000}{830218.400}      & \cost{882107.050}  & \timem{1.474} & \fgit{1.474}{1.968} & \tard{0.672} & \costpp{882107.050}  &                    \\
-      \BILTwo{}      & \cost{51998.950}  & \cost{70108.800}  & \cost{255166.400}  & \fgiboc{70108.800}{255166.400}  & \cost{377274.150}  & \timem{1.476} & \fgit{1.476}{2.298} & \tard{0.206} & \costpp{377274.150}  &                    \\
-      \BILThree{}    & \cost{51976.550}  & \cost{240175.000} & \cost{77822.400}   & \fgiboc{240175.000}{77822.400}  & \cost{369973.950}$^{1}$  & \timem{1.475} & \fgit{1.475}{3.091} & \tard{0.060} & \costpp{369973.950}  &                    \\
-      \BILFour{}     & \cost{51827.050}  & \cost{442307.600} & \cost{26711.200}   & \fgiboc{442307.600}{26711.200}  & \cost{520845.850}  & \timem{1.472} & \fgit{1.472}{4.031} & \tard{0.019} & \costpp{520845.850}  &                    \\
+      \ql{0.99}      & \cost{63095.100}  & \cost{180403.800} & \cost{121528.000}  & \fgiboc{180403.800}{121528.000} & \textbf{\cost{365026.900}} &                          & \timem{1.680} & \fgit{1.680}{3.019} & \tard{0.103} & \textbf{\costpp{365026.900}} &                    \\
+      \rl{0.99}      & \cost{65310.650}  & \cost{168489.200} & \cost{148593.600}  & \fgiboc{168489.200}{148593.600} & \cost{382393.450}          &                          & \timem{1.721} & \fgit{1.721}{3.005} & \tard{0.136} & \costpp{382393.450}          & \rnd{52.918}       \\
+      \rl{1.00}      & \cost{60778.500}  & \cost{234045.000} & \cost{76168.000}   & \fgiboc{234045.000}{76168.000}  & \cost{370991.500}          & $^{1}$                   & \timem{1.637} & \fgit{1.637}{3.226} & \tard{0.066} & \costpp{370991.500}          & \rnd{52.291}       \\
+      \BILOne{}      & \cost{51888.650}  & \cost{0.000}      & \cost{830218.400}  & \fgiboc{0.000}{830218.400}      & \cost{882107.050}          &                          & \timem{1.474} & \fgit{1.474}{1.968} & \tard{0.672} & \costpp{882107.050}          &                    \\
+      \BILTwo{}      & \cost{51998.950}  & \cost{70108.800}  & \cost{255166.400}  & \fgiboc{70108.800}{255166.400}  & \cost{377274.150}          &                          & \timem{1.476} & \fgit{1.476}{2.298} & \tard{0.206} & \costpp{377274.150}          &                    \\
+      \BILThree{}    & \cost{51976.550}  & \cost{240175.000} & \cost{77822.400}   & \fgiboc{240175.000}{77822.400}  & \cost{369973.950}          & $^{1}$                   & \timem{1.475} & \fgit{1.475}{3.091} & \tard{0.060} & \costpp{369973.950}          &                    \\
+      \BILFour{}     & \cost{51827.050}  & \cost{442307.600} & \cost{26711.200}   & \fgiboc{442307.600}{26711.200}  & \cost{520845.850}          &                          & \timem{1.472} & \fgit{1.472}{4.031} & \tard{0.019} & \costpp{520845.850}          &                    \\
       \bottomrule
       \toprule
-      90-\(\Unif\)   & WIPC              & FGIC              & BOC                & FGIC+BOC                        & SUM                & SFTT          & FGIT                & SL(\%)       & CostpP               & \(\avgrew^{\pol}\) \\
+      90-\(\Unif\)   & WIPC              & FGIC              & BOC                & \fgibocName                     & SUM                        & \multicolumn{2}{r}{SFTT} & FGIT          & SL(\%)              & Cost.p.P     & \(\avgrew^{\pol}\)                                \\
       \midrule
-      \ql{0.99}      & \cost{98909.150}  & \cost{190681.000} & \cost{297333.600}  & \fgiboc{190681.000}{297333.600} & \cost{586923.750}$^{1}$  & \timem{2.133} & \fgit{2.133}{3.417} & \tard{0.173} & \costpp{586923.750}  &                    \\
-      \rl{0.99}      & \cost{105348.200} & \cost{263950.600} & \cost{207183.200}  & \fgiboc{263950.600}{207183.200} & \textbf{\cost{576482.000}}$^{1}$  & \timem{2.238} & \fgit{2.238}{3.825} & \tard{0.121} & \textbf{\costpp{576482.000}}  & \rnd{79.826}       \\
-      \rl{1.00}      & \cost{107873.800} & \cost{240798.800} & \cost{253865.600}  & \fgiboc{240798.800}{253865.600} & \cost{602538.200}$^{2}$  & \timem{2.280} & \fgit{2.280}{3.771} & \tard{0.153} & \costpp{602538.200}  & \rnd{80.425}       \\
-      \BILOne{}      & \cost{88658.100}  & \cost{0.000}      & \cost{1418529.600} & \fgiboc{0.000}{1418529.600}     & \cost{1507187.700} & \timem{1.964} & \fgit{1.964}{2.462} & \tard{0.790} & \costpp{1507187.700} &                    \\
-      \BILTwo{}      & \cost{88317.700}  & \cost{50854.400}  & \cost{646000.800}  & \fgiboc{50854.400}{646000.800}  & \cost{785172.900}  & \timem{1.958} & \fgit{1.958}{2.666} & \tard{0.355} & \costpp{785172.900}  &                    \\
-      \BILThree{}    & \cost{88252.450}  & \cost{207210.600} & \cost{299885.600}  & \fgiboc{207210.600}{299885.600} & \cost{595348.650}$^{2}$  & \timem{1.957} & \fgit{1.957}{3.309} & \tard{0.148} & \costpp{595348.650}  &                    \\
-      \BILFour{}     & \cost{87774.600}  & \cost{415178.200} & \cost{153612.800}  & \fgiboc{415178.200}{153612.800} & \cost{656565.600}  & \timem{1.950} & \fgit{1.950}{4.158} & \tard{0.071} & \costpp{656565.600}  &                    \\
+      \ql{0.99}      & \cost{98909.150}  & \cost{190681.000} & \cost{297333.600}  & \fgiboc{190681.000}{297333.600} & \cost{586923.750}          & $^{1}$                   & \timem{2.133} & \fgit{2.133}{3.417} & \tard{0.173} & \costpp{586923.750}          &                    \\
+      \rl{0.99}      & \cost{105348.200} & \cost{263950.600} & \cost{207183.200}  & \fgiboc{263950.600}{207183.200} & \textbf{\cost{576482.000}} & $^{1}$                   & \timem{2.238} & \fgit{2.238}{3.825} & \tard{0.121} & \textbf{\costpp{576482.000}} & \rnd{79.826}       \\
+      \rl{1.00}      & \cost{107873.800} & \cost{240798.800} & \cost{253865.600}  & \fgiboc{240798.800}{253865.600} & \cost{602538.200}          & $^{2}$                   & \timem{2.280} & \fgit{2.280}{3.771} & \tard{0.153} & \costpp{602538.200}          & \rnd{80.425}       \\
+      \BILOne{}      & \cost{88658.100}  & \cost{0.000}      & \cost{1418529.600} & \fgiboc{0.000}{1418529.600}     & \cost{1507187.700}         &                          & \timem{1.964} & \fgit{1.964}{2.462} & \tard{0.790} & \costpp{1507187.700}         &                    \\
+      \BILTwo{}      & \cost{88317.700}  & \cost{50854.400}  & \cost{646000.800}  & \fgiboc{50854.400}{646000.800}  & \cost{785172.900}          &                          & \timem{1.958} & \fgit{1.958}{2.666} & \tard{0.355} & \costpp{785172.900}          &                    \\
+      \BILThree{}    & \cost{88252.450}  & \cost{207210.600} & \cost{299885.600}  & \fgiboc{207210.600}{299885.600} & \cost{595348.650}          & $^{2}$                   & \timem{1.957} & \fgit{1.957}{3.309} & \tard{0.148} & \costpp{595348.650}          &                    \\
+      \BILFour{}     & \cost{87774.600}  & \cost{415178.200} & \cost{153612.800}  & \fgiboc{415178.200}{153612.800} & \cost{656565.600}          &                          & \timem{1.950} & \fgit{1.950}{4.158} & \tard{0.071} & \costpp{656565.600}          &                    \\
       \bottomrule
@@
     distributed demand interarrival times. \(^{1,2}\) Costs marked with the same number are
-    not sign.\@ different to each other.}
+    not significantly different from each other.}
 \end{table}
@@
 
+\paragraph{Medium Variability.} The results for medium variability in the system are presented in
+Table~\ref{tbl:unif}, where a uniform demand interarrival time is used. We use the tuples
+\(70-\Unif\), \(80-\Unif\), and \(90-\Unif\) to indicate the experimental setup of using uniform
+demand interarrival times with a aimed bottleneck utilisation of \(70\%\), \(80\%\), or \(90\%\)
+respectively. The results state the accumulated costs, averaged over all replications, for the work
+in process (WIPC), the finished goods inventory (FGIC), the backorder costs (BOC), as well as the
+sum of the finished good inventory and backorder costs (\fgibocName{}) and the sum over all three cost
+centers (SUM). For a better insight in the timing performance we also provide the average shopfloor
+throughput time (SFTT), i.e.\@ the average time an order spends in the system after being released
+and until reaching the finished goods inventory, the finished goods inventory time (FGIT), as well
+as the service level (SL). The later is given in percentage and states the percentage of orders that
+were produced on time. Finally, the actual measured costs per period (Cost.p.P) and the by the
+algorithm learned average reward \(\avgrew\) are stated. The tested methods are the deep Q-Learning
+algorithm with discount factor \(0.99\) (\qlt{0.99}), the \ARA{} algorithm with discount factors
+\(0.99\) (\rlt{0.99}) and \(1.00\) (\rlt{1.00}), as well as the backward infinite loading
+technique \BIL{} with fixed lead times between \(1\) and \(4\) as indicated by the number
+($\BILOne$-$\BILFour$).
+
+For the case of medium variability with a utilisation of \(70\%\) (\(70-\Unif\)) we can see that
+\(\BILTwo\) performs best in the sense of accumulated costs. The reinforcement learning algorithm
+follow, where \rlt{0.99} is significantly better then the other two variants. The other \BIL{}
+variants accumulate much more costs. For the timing one can see that \rlt{1.00} performs very good
+with a service level of \(93.2\%\). Furthermore, the learned average reward \(\avgrew\) and actual
+average reward are very similar for both \ARA{} variants. However, with \(70\%\) of utilisation the
+overhead of the reinforcement learning variants are too big to be able to compete with the simple
+rule based release mechanism of \(\BILTwo{}\) as for changing workload the utilisation changes
+rather linearly due to the low variability in the system.
+
+In the case of \(80-\Unif\)
+
 \begin{table}[tb]
+  \footnotesize
   \begin{center}
-    \begin{tabular}{crrrccccccc}
+    \begin{tabular}{crrrrrlrrrrr}
       \toprule
-      70-\(\Exp\) & WIPC              & FGIC              & BOC               & FGIC+BOC                        & SUM                              & SFTT          & FGIT                & SL(\%)       & Cost.p.P                     & \(\avgrew^{\pol}\) \\
+      70-\(\Exp\) & WIPC              & FGIC              & BOC               & \hide{\fgibocName{}}                        & SUM                            &  \multicolumn{2}{r}{SFTT}          & FGIT                & SL(\%)       & Cost.p.P                     & \(\avgrew^{\pol}\) \\
       \midrule
-      \ql{0.99}   & \cost{53873.500}  & \cost{142398.800} & \cost{164427.200} & \fgiboc{142398.800}{164427.200} & \cost{360699.500}$^{1}$          & \timem{1.585} & \fgit{1.585}{2.796} & \tard{0.162} & \costpp{360699.500}          &                    \\
-      \rl{0.99}   & \cost{61419.850}  & \cost{181702.200} & \cost{115287.200} & \fgiboc{181702.200}{115287.200} & \cost{358409.250}                & \timem{1.734} & \fgit{1.734}{3.145} & \tard{0.107} & \costpp{358409.250}          & \rnd{50.787}       \\
-      \rl{1.00}   & \cost{56578.800}  & \cost{199310.000} & \cost{115714.400} & \fgiboc{199310.000}{115714.400} & \cost{371603.200}                & \timem{1.639} & \fgit{1.639}{3.136} & \tard{0.113} & \costpp{371603.200}          & \rnd{50.575}       \\
-      \BILOne{}   & \cost{48547.100}  & \cost{     0.000} & \cost{776753.600} & \fgiboc{     0.000}{776753.600} & \cost{825300.700}                & \timem{1.479} & \fgit{1.479}{1.975} & \tard{0.663} & \costpp{825300.700}          &                    \\
-      \BILTwo{}   & \cost{48618.850}  & \cost{ 66927.200} & \cost{248638.400} & \fgiboc{ 66927.200}{248638.400} & \cost{364184.450}$^{1}$          & \timem{1.480} & \fgit{1.480}{2.312} & \tard{0.220} & \costpp{364184.450}          &                    \\
-      \BILThree{} & \cost{48632.050}  & \cost{222603.200} & \cost{ 74568.800} & \fgiboc{222603.200}{ 74568.800} & \textbf{\cost{345804.050}}       & \timem{1.481} & \fgit{1.481}{3.094} & \tard{0.063} & \textbf{\costpp{345804.050}} &                    \\
-      \BILFour{}  & \cost{48514.200}  & \cost{409531.200} & \cost{ 23406.400} & \fgiboc{409531.200}{ 23406.400} & \cost{481451.800}                & \timem{1.478} & \fgit{1.478}{4.029} & \tard{0.019} & \costpp{481451.800}          &                    \\
+      \ql{0.99}   & \cost{53873.500}  & \cost{142398.800} & \cost{164427.200} & \fgiboc{142398.800}{164427.200} & \cost{360699.500}&$^{1}$          & \timem{1.585} & \fgit{1.585}{2.796} & \tard{0.162} & \costpp{360699.500}          &                    \\
+      \rl{0.99}   & \cost{61419.850}  & \cost{181702.200} & \cost{115287.200} & \fgiboc{181702.200}{115287.200} & \cost{358409.250}&                & \timem{1.734} & \fgit{1.734}{3.145} & \tard{0.107} & \costpp{358409.250}          & \rnd{50.787}       \\
+      \rl{1.00}   & \cost{56578.800}  & \cost{199310.000} & \cost{115714.400} & \fgiboc{199310.000}{115714.400} & \cost{371603.200}&                & \timem{1.639} & \fgit{1.639}{3.136} & \tard{0.113} & \costpp{371603.200}          & \rnd{50.575}       \\
+      \BILOne{}   & \cost{48547.100}  & \cost{     0.000} & \cost{776753.600} & \fgiboc{     0.000}{776753.600} & \cost{825300.700}&                & \timem{1.479} & \fgit{1.479}{1.975} & \tard{0.663} & \costpp{825300.700}          &                    \\
+      \BILTwo{}   & \cost{48618.850}  & \cost{ 66927.200} & \cost{248638.400} & \fgiboc{ 66927.200}{248638.400} & \cost{364184.450}&$^{1}$          & \timem{1.480} & \fgit{1.480}{2.312} & \tard{0.220} & \costpp{364184.450}          &                    \\
+      \BILThree{} & \cost{48632.050}  & \cost{222603.200} & \cost{ 74568.800} & \fgiboc{222603.200}{ 74568.800} & \textbf{\cost{345804.050}}&       & \timem{1.481} & \fgit{1.481}{3.094} & \tard{0.063} & \textbf{\costpp{345804.050}} &                    \\
+      \BILFour{}  & \cost{48514.200}  & \cost{409531.200} & \cost{ 23406.400} & \fgiboc{409531.200}{ 23406.400} & \cost{481451.800}       &         & \timem{1.478} & \fgit{1.478}{4.029} & \tard{0.019} & \costpp{481451.800}          &                    \\
       \bottomrule
       \toprule
-      80-\(\Exp\) & WIPC              & FGIC              & BOC               & FGIC+BOC                        & SUM                              & SFTT          & FGIT                & SL(\%)       & Cost.p.P                     & \(\avgrew^{\pol}\) \\
+      80-\(\Exp\) & WIPC              & FGIC              & BOC               & \hide{\fgibocName}                        & SUM                             &\multicolumn{2}{r}{SFTT}          & FGIT                & SL(\%)       & Cost.p.P                     & \(\avgrew^{\pol}\) \\
       \midrule
-      \ql{0.99}   & \cost{87706.700}  & \cost{227485.400} & \cost{225150.400} & \fgiboc{227485.400}{225150.400} & \cost{540342.500}$^{1}$          & \timem{2.043} & \fgit{2.043}{3.541} & \tard{0.166} & \costpp{540342.500}          &                    \\
-      \rl{0.99}   & \cost{89838.950}  & \cost{253476.000} & \cost{185687.200} & \fgiboc{253476.000}{185687.200} & \cost{529002.150}$^{1,2}$        & \timem{2.078} & \fgit{2.078}{3.689} & \tard{0.124} & \costpp{529002.150}          & \rnd{70.527}       \\
-      \rl{1.00}   & \cost{87171.500}  & \cost{242915.600} & \cost{198456.000} & \fgiboc{242915.600}{198456.000} & \cost{528543.100}$^{2,3}$        & \timem{2.034} & \fgit{2.034}{3.599} & \tard{0.134} & \costpp{528543.100}          & \rnd{72.944}       \\
-      \BILOne{}   & \cost{78731.850}  & \cost{     0.000} & \cost{1259709.60} & \fgiboc{     0.000}{1259709.60} & \cost{1338441.45}                & \timem{1.884} & \fgit{1.884}{2.382} & \tard{0.771} & \costpp{1338441.450}         &                    \\
-      \BILTwo{}   & \cost{78203.150}  & \cost{ 52303.200} & \cost{548736.000} & \fgiboc{ 52303.200}{548736.000} & \cost{679242.350}                & \timem{1.875} & \fgit{1.875}{2.602} & \tard{0.350} & \costpp{679242.350}          &                    \\
-      \BILThree{} & \cost{79029.850}  & \cost{200273.200} & \cost{242100.800} & \fgiboc{200273.200}{242100.800} & \textbf{\cost{521403.850}}$^{3}$ & \timem{1.890} & \fgit{1.890}{3.265} & \tard{0.142} & \textbf{\costpp{521403.850}} &                    \\
-      \BILFour{}  & \cost{79384.750}  & \cost{395493.000} & \cost{116943.200} & \fgiboc{395493.000}{116943.200} & \cost{591820.950}                & \timem{1.896} & \fgit{1.896}{4.128} & \tard{0.062} & \costpp{591820.950}          &                    \\
+      \ql{0.99}   & \cost{87706.700}  & \cost{227485.400} & \cost{225150.400} & \fgiboc{227485.400}{225150.400} & \cost{540342.500}&$^{1}$          & \timem{2.043} & \fgit{2.043}{3.541} & \tard{0.166} & \costpp{540342.500}          &                    \\
+      \rl{0.99}   & \cost{89838.950}  & \cost{253476.000} & \cost{185687.200} & \fgiboc{253476.000}{185687.200} & \cost{529002.150}&$^{1,2}$        & \timem{2.078} & \fgit{2.078}{3.689} & \tard{0.124} & \costpp{529002.150}          & \rnd{70.527}       \\
+      \rl{1.00}   & \cost{87171.500}  & \cost{242915.600} & \cost{198456.000} & \fgiboc{242915.600}{198456.000} & \cost{528543.100}&$^{2,3}$        & \timem{2.034} & \fgit{2.034}{3.599} & \tard{0.134} & \costpp{528543.100}          & \rnd{72.944}       \\
+      \BILOne{}   & \cost{78731.850}  & \cost{     0.000} & \cost{1259709.60} & \fgiboc{     0.000}{1259709.60} & \cost{1338441.45}&                & \timem{1.884} & \fgit{1.884}{2.382} & \tard{0.771} & \costpp{1338441.450}         &                    \\
+      \BILTwo{}   & \cost{78203.150}  & \cost{ 52303.200} & \cost{548736.000} & \fgiboc{ 52303.200}{548736.000} & \cost{679242.350}&                & \timem{1.875} & \fgit{1.875}{2.602} & \tard{0.350} & \costpp{679242.350}          &                    \\
+      \BILThree{} & \cost{79029.850}  & \cost{200273.200} & \cost{242100.800} & \fgiboc{200273.200}{242100.800} & \textbf{\cost{521403.850}}&$^{3}$ & \timem{1.890} & \fgit{1.890}{3.265} & \tard{0.142} & \textbf{\costpp{521403.850}} &                    \\
+      \BILFour{}  & \cost{79384.750}  & \cost{395493.000} & \cost{116943.200} & \fgiboc{395493.000}{116943.200} & \cost{591820.950} &               & \timem{1.896} & \fgit{1.896}{4.128} & \tard{0.062} & \costpp{591820.950}          &                    \\
       \bottomrule
       \toprule
-      90-\(\Exp\) & WIPC              & FGIC              & BOC               & FGIC+BOC                        & SUM                              & SFTT          & FGIT                & SL(\%)       & Cost.p.P                     & \(\avgrew^{\pol}\) \\
+      90-\(\Exp\) & WIPC              & FGIC              & BOC               & \hide{\fgibocName}                        & SUM                              & \multicolumn{2}{r}{SFTT}          & FGIT                & SL(\%)       & Cost.p.P                     & \(\avgrew^{\pol}\) \\
       \midrule
-      \ql{0.99}   & \cost{168655.950} & \cost{222145.800} & \cost{788662.400} & \fgiboc{222145.800}{788662.400} & \cost{1179464.150}$^{1}$               & \timem{3.134} & \fgit{3.134}{4.501} & \tard{0.285} & \costpp{1179464.150}         &                    \\
-      \rl{0.99}   & \cost{158034.800} & \cost{212726.200} & \cost{701300.800} & \fgiboc{212726.200}{701300.800} & \cost{1072061.800}$^{2}$               & \timem{2.970} & \fgit{2.970}{4.300} & \tard{0.281} & \costpp{1072061.800}         & \rnd{115.801}      \\
-      \rl{1.00}   & \cost{160453.950} & \cost{286065.800} & \cost{539803.200} & \fgiboc{286065.800}{539803.200} & \textbf{\cost{986322.950}}       & \timem{3.007} & \fgit{3.007}{4.623} & \tard{0.204} & \textbf{\costpp{986322.950}} & \rnd{109.907}      \\
-      \BILOne{}   & \cost{146835.900} & \cost{     0.000} & \cost{2349374.40} & \fgiboc{     0.000}{2349374.40} & \cost{2496210.300}               & \timem{2.796} & \fgit{2.796}{3.295} & \tard{0.867} & \costpp{2496210.300}         &                    \\
-      \BILTwo{}   & \cost{143756.000} & \cost{ 34112.800} & \cost{1412752.00} & \fgiboc{ 34112.800}{1412752.00} & \cost{1590620.800}               & \timem{2.748} & \fgit{2.748}{3.380} & \tard{0.529} & \costpp{1590620.800}         &                    \\
-      \BILThree{} & \cost{144278.250} & \cost{154146.800} & \cost{877450.400} & \fgiboc{154146.800}{877450.400} & \cost{1175875.450}$^{1}$               & \timem{2.756} & \fgit{2.756}{3.857} & \tard{0.294} & \costpp{1175875.450}         &                    \\
-      \BILFour{}  & \cost{147077.800} & \cost{335070.400} & \cost{622141.600} & \fgiboc{335070.400}{622141.600} & \cost{1104289.800}$^{2}$               & \timem{2.800} & \fgit{2.800}{4.608} & \tard{0.175} & \costpp{1104289.800}         &                    \\
+      \ql{0.99}   & \cost{168655.950} & \cost{222145.800} & \cost{788662.400} & \fgiboc{222145.800}{788662.400} & \cost{1179464.150}&$^{1}$               & \timem{3.134} & \fgit{3.134}{4.501} & \tard{0.285} & \costpp{1179464.150}         &                    \\
+      \rl{0.99}   & \cost{158034.800} & \cost{212726.200} & \cost{701300.800} & \fgiboc{212726.200}{701300.800} & \cost{1072061.800}&$^{2}$               & \timem{2.970} & \fgit{2.970}{4.300} & \tard{0.281} & \costpp{1072061.800}         & \rnd{115.801}      \\
+      \rl{1.00}   & \cost{160453.950} & \cost{286065.800} & \cost{539803.200} & \fgiboc{286065.800}{539803.200} & \textbf{\cost{986322.950}}&       & \timem{3.007} & \fgit{3.007}{4.623} & \tard{0.204} & \textbf{\costpp{986322.950}} & \rnd{109.907}      \\
+      \BILOne{}   & \cost{146835.900} & \cost{     0.000} & \cost{2349374.40} & \fgiboc{     0.000}{2349374.40} & \cost{2496210.300}&               & \timem{2.796} & \fgit{2.796}{3.295} & \tard{0.867} & \costpp{2496210.300}         &                    \\
+      \BILTwo{}   & \cost{143756.000} & \cost{ 34112.800} & \cost{1412752.00} & \fgiboc{ 34112.800}{1412752.00} & \cost{1590620.800}&               & \timem{2.748} & \fgit{2.748}{3.380} & \tard{0.529} & \costpp{1590620.800}         &                    \\
+      \BILThree{} & \cost{144278.250} & \cost{154146.800} & \cost{877450.400} & \fgiboc{154146.800}{877450.400} & \cost{1175875.450}&$^{1}$               & \timem{2.756} & \fgit{2.756}{3.857} & \tard{0.294} & \costpp{1175875.450}         &                    \\
+      \BILFour{}  & \cost{147077.800} & \cost{335070.400} & \cost{622141.600} & \fgiboc{335070.400}{622141.600} & \cost{1104289.800}&$^{2}$               & \timem{2.800} & \fgit{2.800}{4.608} & \tard{0.175} & \costpp{1104289.800}         &                    \\
       \bottomrule
@@
     distributed interarrival times of the demand. \(^{1,2,3}\) Costs with the same number are
-    not significant different to each other. }
+    not significant different from each other. }
 \end{table}

### 2021-05-07 paper/paper.tex
updated experiments

@@
 
-\paragraph{Medium Variability.} The results for medium variability in the system are presented in
-Table~\ref{tbl:unif}, where a uniform demand interarrival time is used. We use the tuples
-\(70-\Unif\), \(80-\Unif\), and \(90-\Unif\) to indicate the experimental setup of using uniform
-demand interarrival times with a aimed bottleneck utilisation of \(70\%\), \(80\%\), or \(90\%\)
-respectively. The results state the accumulated costs, averaged over all replications, for the work
-in process (WIPC), the finished goods inventory (FGIC), the backorder costs (BOC), as well as the
-sum of the finished good inventory and backorder costs (\fgibocName{}) and the sum over all three cost
-centers (SUM). For a better insight in the timing performance we also provide the average shopfloor
-throughput time (SFTT), i.e.\@ the average time an order spends in the system after being released
-and until reaching the finished goods inventory, the finished goods inventory time (FGIT), as well
-as the service level (SL). The later is given in percentage and states the percentage of orders that
-were produced on time. Finally, the actual measured costs per period (Cost.p.P) and the by the
-algorithm learned average reward \(\avgrew\) are stated. The tested methods are the deep Q-Learning
-algorithm with discount factor \(0.99\) (\qlt{0.99}), the \ARA{} algorithm with discount factors
-\(0.99\) (\rlt{0.99}) and \(1.00\) (\rlt{1.00}), as well as the backward infinite loading
-technique \BIL{} with fixed lead times between \(1\) and \(4\) as indicated by the number
-($\BILOne$-$\BILFour$).
-
-For the case of medium variability with a utilisation of \(70\%\) (\(70-\Unif\)) we can see that
-\(\BILTwo\) performs best in the sense of accumulated costs. The reinforcement learning algorithm
-follow, where \rlt{0.99} is significantly better then the other two variants. The other \BIL{}
-variants accumulate much more costs. For the timing one can see that \rlt{1.00} performs very good
-with a service level of \(93.2\%\). Furthermore, the learned average reward \(\avgrew\) and actual
-average reward are very similar for both \ARA{} variants. However, with \(70\%\) of utilisation the
-overhead of the reinforcement learning variants are too big to be able to compete with the simple
-rule based release mechanism of \(\BILTwo{}\) as for changing workload the utilisation changes
-rather linearly due to the low variability in the system.
-
-In the case of \(80-\Unif\)
+\paragraph{Medium Demand Variability.} The results for medium variability in the system are
+presented in Table~\ref{tbl:unif}, where a uniform demand interarrival time is used. We use the
+tuples 70-\(\Unif\), 80-\(\Unif\), and 90-\(\Unif\) to indicate the experimental setup of using
+uniform demand interarrival times with a aimed bottleneck utilisation of \(70\%\), \(80\%\), or
+\(90\%\) respectively. The results state the accumulated costs, averaged over all replications, for
+the work in process (WIPC), the finished goods inventory (FGIC), the backorder costs (BOC), as well
+as the sum of the finished good inventory and backorder costs (\fgibocName{}) and the sum over all
+three cost centers (SUM). To ease readability all costs are given in \(1000\)~Dollars. For a better
+insight in the timing performance we also provide the average shopfloor throughput time (SFTT),
+i.e.\@ the average time an order spends in the system after being released and until reaching the
+finished goods inventory, the finished goods inventory time (FGIT), as well as the service level
+(SL). The later is given in percentage and states the percentage of orders that were produced on
+time. Finally, the actual measured costs per period (Cost.p.P) and the by the algorithm learned
+average reward \(\avgrew\) are stated. The tested methods are the deep Q-Learning algorithm with
+discount factor \(0.99\) (\qlt{0.99}), the \ARA{} algorithm with discount factors \(0.99\)
+(\rlt{0.99}) and \(1.00\) (\rlt{1.00}), as well as the backward infinite loading technique \BIL{}
+with fixed lead times between \(1\) and \(4\) as indicated by the number ($\BILOne$-$\BILFour$).
+
+For the case of medium variability with a utilisation of \(70\%\) 70-(\(\Unif\)) we can see that
+\(\BILTwo\) performs best in the sense of accumulated costs (\(220.53\)). The reinforcement learning
+algorithm follow, where \rlt{0.99} is slightly but significantly better then the other two variants
+(\(250.70\), \(254.97\), \(255.06\)). The other \BIL{} variants accumulate much more costs. For the
+timing one can see that \rlt{1.00} performs very good with a service level of \(93.2\%\).
+Furthermore, the learned average reward \(\avgrew\) and actual average reward are very similar for
+both \ARA{} variants. However, with \(70\%\) of utilisation the overhead of the reinforcement
+learning variants are too big to be able to compete with the simple rule based release mechanism of
+\BILTwo{} as for changing workload the utilisation changes rather linearly due to the low
+variability in the system.
+
+In the case of 80-\(\Unif\) the deep Q-Learning variant \qlt{0.99} performs best in the cost related
+measures (\(365.03\) for SUM), followed by \BILThree{} (\(369.97\)) and \rlt{1.00} (\(370.99\)),
+which are not significantly different from each other. One can see the \BILTwo{} (\(377.27\))
+performs slightly worse than \rlt{1.00} and is followed by \rlt{0.99} (\(382.39\)). Again \rlt{1.00}
+achieves a very high service level of \(93.4\%\) with reasonable sum of cost, which is also
+reflected in the low Backorder costs. Furthermore, the measured and learned average reward values
+are quite close. Thus, the results already provide the insight that \ARA{} performs quite well in
+terms of timing for the completion of orders. This pays off when confronted with higher
+probabilistic systems.
+
+For instance, with \(90\%\) utilisation and medium variable demand stream, i.e.\@ 90-(\(\Unif\)), we
+can see that \rlt{0.99} performs best (\(576.48\)), followed by \qlt{0.99} (\(586.92\)) and
+\BILThree{} (\(595.35\)), as well as \rlt{1.00} (\(602.54\)), where the later two are not
+significantly different. In this case \rlt{0.99} also achieves the best service level of \(87.9\%\)
+within this group of best performing algorithms. The WIPC costs, as well as the SFTT, show that the
+reinforcement learning agents have a higher work-in-process, which is clear as they try to smooth
+the workload of demand stream. When looking at the timing cost it becomes even more apparent that
+the agents perform well in timing as \rlt{0.99} performs much better (\(471.13\)) than the other two
+agents (\(488.01\), \(494.66\)) and all the \BIL{} techniques accumulated more \fgibocName{} costs
+(\(507.10\) for the best performing \BIL{}, \BILThree{}).
 
@@
 
+\paragraph{High Demand Variability.} For results of the experiments using exponential demand
+interarrival times, presented in Table~\ref{tbl:exp} we can see a similar pattern as before. In this
+case we use the tuples 70-\(\Exp\), 80-\(\Exp\), and 90-\(\Exp\) to indicate the corresponding
+experimental setups.
+
+For \(70\%\) utilization of the bottleneck, i.e. 80-\(\Exp\), \BILThree{} performs best with costs
+of \(345.80\), followed by \rlt{0.99} with \(358.41\), and then by \qlt{0.99} and \BILTwo{}
+(\(360.70\), \(364.18\)). In terms of timing performance, again \ARA{} outperforms the opponents
+with \fgibocName{} costs of \(296.99\), and \BILThree{} with the second highest costs (\(297.17\)).
+However, the service level of \BILThree{} is higher with \(93.7\%\) due to the lower shop floor
+throughput time (SFTT, \(1.48\) for \BILThree{} versus \(1.73\) for \rlt{0.99}).
+
 % Table~\ref{tbl:res} shows the results, where we evaluate the average reward and discounted
@@
   scheduling + release, etc.}
+
+\MS[t]{Job shop}
 

### 2021-05-08 paper/paper.tex
results + partly conclusion

@@
 %
-Our algorithm, based on the tabular version of \citet{schneckenreither2020average}, does not require
-the exact solution for \(V^{\pol}(s)\), but a solution which is offset suffices. Clearly this
-observation reduces the required iteration steps tremendously as finding the exact solution,
-especially for large discount factors, is tedious. Therefore, we allow to set \(\gamma = 1\), which
-induces \(\X_{\gamma}^{\pol}(s) = V^{\pol}(s) + u\), where \(u\) is for unichain MDPs a scalar value
+Our average reward adjusted learning algorithm (\ARA{}), based on the tabular version of
+\citet{schneckenreither2020average}, does not require the exact solution for \(V^{\pol}(s)\), but a
+solution which is offset suffices. Clearly this observation reduces the required iteration steps
+tremendously as finding the exact solution, especially for large discount factors, is tedious.
+Therefore, we allow to set \(\gamma = 1\), which induces
+\(\X_{\gamma}^{\pol}(s) = V^{\pol}(s) + u\), where \(u\) is for unichain MDPs a scalar value
 independent of \(s\), i.e.\@ equivalent for all states of the MDP~\cite[p.346]{Puterman94}.
@@
 Equation~\ref{eq:constr2} of the average reward RL constraints is used to estimate the average
-reward based on the \MS[t]{reformulate avg RL constraints using s' etc.?} currently inferred state
-values of two consecutive states and the returned reward. % As the average
+reward based on the currently inferred state values of two consecutive states and the returned
+reward. % As the average
 % reward is equal among all states estimating the average reward is based on many observations.
@@
 within this group of best performing algorithms. The WIPC costs, as well as the SFTT, show that the
-reinforcement learning agents have a higher work-in-process, which is clear as they try to smooth
+reinforcement learning agents have a higher work in process, which is clear as they try to smooth
 the workload of demand stream. When looking at the timing cost it becomes even more apparent that
@@
 However, the service level of \BILThree{} is higher with \(93.7\%\) due to the lower shop floor
-throughput time (SFTT, \(1.48\) for \BILThree{} versus \(1.73\) for \rlt{0.99}).
-
-% Table~\ref{tbl:res} shows the results, where we evaluate the average reward and discounted
-% reinforcement learning algorithm by providing either direct reward feedback, that is the reward
-% accumulated in the next period, without keeping track of any orders, as \(\ActShipped\) and
-% \(\ActShippedDQN\) respectively. Further both algorithms are evaluated with keeping track of the
-% order pool orders, such that each action is responsible for all orders in the order pool. These are
-% labelled \(\ActOrdPool\) and \(\ActOrdPoolDQN\). For comparisons reasons we provide
-% \(\BILOne-\BILSix\) and immediate release (interval release).
-
-% The results show that at the current time we are facing issues with scaling the algorithm to the
-% order release problem. Likely this is due to parameterisation issues, which we are investigating
-% currently. However, it can be seen that i) the technique of keeping track of the order pool orders
-% leads to better policies and ii) that despite the parameterisation issues average reward
-% reinforcement learning finds better solutions than its discounted counterpart. However, the least
-% costs are generated by \(\BILThree\), which shows that there is room for improvement for the
-% reinforcement learning policies.
-
-% Interesting is also to see, that all reinforcement learning variants are releasing rather late,
-% leading to low finished goods inventory costs (FGIC), but high backorder costs (BOC). Furthermore,
-% the agents based on the order-pool orders, namely \(\ActOrdPool\) and \(\ActOrdPoolDQN\), are able
-% achieve lower work-in-process costs (WIPC) and also lower tardiness, than the other the agents
-% rewarded directly on the shipped orders.
-
-% \begin{table}[t!]
-%   \centering
-%   \def\arraystretch{1.2}          %  1 is the default, change whatever you need
-%   \setlength\tabcolsep{4pt}
-%   \begin{tabular}{|l||c|c|c|c|c|c|c|c|c|}
-%     % BEGIN RECEIVE ORGTBL res2
-%     \hline
-%     Algorithm/KPI & SUM & BOC & FGIC & WIPC & % \%SL &
-%                                                        TARD & \(\sigma\)TARD & SFTT\\
-%     \hline
-%     % Evaluation 1 & \multicolumn{11}{c|}{} \\
-%     \ActShipped    & 1777.74 & 1322.04 & 94.49 & 361.20       & 2.24 & 1.52 & 3.44\\
-%     \ActOrdPool    & 1593.03 & 1318.65 &14.01 &260.36         & 2.11 & 1.42 & 2.68\\
-%     \ActShippedDQN & 1940.62 & 1570.40 &6.30 &363.92          & 2.35 & 1.59 & 3.95 \\
-%     \ActOrdPoolDQN & 1618.88 & 1306.20 & 40.68& 272.00        & 2.08 & 1.38 & 2.58\\
-%     \BILOne        & 1736.98 & 1510.42  &  0.00  & 226.56     & 2.04 & 1.26 & 2.16 \\
-%     \BILTwo        & 1081.79 & 769.86 & 86.89   & 225.03      & 1.87 & 1.18 & 2.15\\
-%     \BILThree      &  \textbf{922.36} & 362.59 &  334.47  & 225.30     & 1.74 & 1.07 & 2.15\\
-%     \BILFour       & 1062.68 & 155.72  & 681.03   & 225.93    & 1.63 & 0.99 & 2.15\\
-%     \BILFive       & 1380.04 & 65.45  & 1088.94    & 225.64   & 1.67 & 1.02 & 2.15\\
-%     \BILSix        & 1777.77 & 25.04 &  1529.77  & 222.96     & 1.61 & 0.85 & 2.13\\
-%     \imre          & 2209.97 & 12.08 &1974.72 & 223.16        & 1.58 & 0.63 & 2.14 \\
-%     \hline
-%     % END RECEIVE ORGTBL res2
-%   \end{tabular}
-%         %         \egroup
-
-%   \caption{\label{tbl:res}Evaluation results where the monetary units are in \(k\)-values.}
-
-%         %         \end{figure}
-% \end{table}
+throughput time (SFTT, \(1.48\) periods for \BILThree{} versus \(1.73\) periods for \rlt{0.99}).
+
+In case of 80-\(\Exp\) \BILThree{} performs best with costs of \(521.40\), but is not significantly
+better than \rlt{1.00} with \(528.54\). Also \rlt{0.99} performs quite well (\(529.00\)). The
+service level of all three algorithm are good, but the order is reversed: \rlt{0.99} with
+\(87.6\%\), \rlt{1.00} with \(86.6\%\) and \BILThree{} with \(85.8\%\). Again the SFTT of the RL
+agents are slightly higher, i.e. \(0.13\)-\(0.20\) periods, than the ones of the \BIL{} release
+procedures. The other three \BIL{} methods, \BILOne{}, \BILTwo{} and \BILFour{}, are performing
+rather bad (\(1,338.44\), \(679.24\) and \(591.82\) respectively). Interesting to see is that in
+this scenario \rlt{0.99} and \rlt{1.00} handle the due date performance better than \qlt{0.99}, as
+can be seen in the backorder costs (BOC). Here the former two accumulate costs of \(185.69\) and
+\(198.46\), while \qlt{0.99} faces costs of \(225.15\). This results in timing costs (\fgibocName{})
+of \(439.16\) for \rlt{0.99} and \(441.37\) for \rlt{1.00}, which also means a better timing cost
+than \BILThree{} with \(442.37\). For this scenario the actual measured average reward (average
+costs) of about \(75.5\) for both \ARA{} variants and learned average rewards of \(70.53\) for
+\rlt{0.99} and \(72.94\) for \rlt{1.00} are a little bit offset. This probably results from the
+higher variability in the system.
+
+Finally, for the case of 90-\(\Exp\) \rlt{1.00} performs much better than the other variants with
+cots of \(986.32\). It is followed by \rlt{0.99} with costs of \(1,072.06\) and rather far, but not
+significantly, afterwards follows \BILFour{} (\(1,104.29\)). As before the shopfloor throughput time
+(SFTT) is higher for the RL variants, and with that of course also the work in process costs (WIPC).
+The backorder costs of \rlt{1.00} (\(539.80\)) are the lowest of all tested methods. Like for the
+overall costs also in terms of timing costs (\fgibocName{}) \rlt{1.00} performs best (\(825.87\)),
+followed by \rlt{0.99} (\(914.03\)) and \BILFour{} (\(957.21\)).
+%
+\BILThree{} and \ql{0.99} perform quite similar in the sense of accumulated costs, \(1,175.88\) and
+\(1,179.46\) respectively. But the other measures show that \ql{0.99} has a better timing
+performance, thus also a slightly high service level with \(71.5\%\) as compared to \BILThree{} with
+\(70.6\%\). The best service level is achieved by \BILFour{} with \(82.5\%\), before \rlt{1.00}
+\(79.6\%\), but with a high cost as the difference of SUM is \(117.97\).
+%
+Interesting to see is that the measured average reward and the learned average reward are very
+different for both \ARA{} variants (about \(30\)-\(40\)). This is probably due to the very high
+variability in the system, but might also be an indicator that the agents need more time for
+learning. However, due to the high computational complexity we encountered we leave this to future
+work.
+
+Overall we can see that the RL agents, and especially the established \ARA{} algorithm performs very
+well in systems with high variability. In case of the less variable systems the results reveal that
+there is still potential for the improvement of reinforcement learning variants.
 
@@
 % which will consequently lead to better policies.\\
+
+This paper is one of the first papers that use reinforcement learning for setting lead times to
+release orders into the production system. We present a novel average reward adjusted reinforcement
+learning algorithm that uses an artificial neural network to approximate the state-action value
+functions. We add several extensions to this algorithm, including n-step learning, and a special
+experience replay memory that stores its experiences according to the actions takes to prevent
+catastrophic forgetting. Additionally we invent \textit{independent agents} to reduce the action
+space by independently choosing a subset of possible actions, while operating on the same state.
+Additionally we adapt the deep Q-Learning algorithm of \citet{mnih2015human} to be able to present a
+fair comparison between the algorithms, where we also use several backward infinite loading
+techniques.
+
+The results suggest that for high variablity the RL learning agents
 

### 2021-05-09 paper/paper.tex
paper conclusion

@@
 
-The results suggest that for high variablity the RL learning agents
-
-\MS[t]{It remains to be shown that independent agents find the same (or better) policies.}
-
-\MS[t]{Multiple agents on different tasks (hierarchical PP): e.g. Order acceptance + release,
-  scheduling + release, etc.}
-
-\MS[t]{Job shop}
+The results suggest that for high variability the RL learning agents, and here especially the novel
+\ARA{} variant, performs very well. In case of medium variability the achieved results are as good
+as with the \BIL{} variants. And for low variability one can see that the RL agents still have room
+for improvement, as \BIL{} outperforms them. When comparing deep Q-Learning to the established
+\ARA{} algorithm, we can see that for all cases, except one (80-\(\Unif\)), \ARA{} performs better.
+
+Therefore, in the future we plan to develop \ARA{} further to a actor-critic state-of-the-art
+reinforcement learning algorithm. We hypothesise that due to the more correct state value
+estimations of \ARA{} as compared to deep Q-Learning, cf. \citet{schneckenreither2020average}, the
+critic will provide a better feedback of the value function update to the actor. This should lead to
+better policies than with standard actor-critic algorithms, e.g. A3C~\ref{Mnih2016asynchronous}.
+%
+Furthermore, one limitation of this proof-of-concept study is that due to the computational
+complexity we restricted our scope to flow-shop production systems only. In subsequent studies we
+plan to test the algorithms on job-shop queuing systems as well.
+%
+Additionally, we want to exploit hierarchical production
+planning~\citep{schneeweibeta1995hierarchical} by implementing agents on different hierarchical
+levels which can bargain with each other. For instance, one agent might be responsible for medium
+term planning order release decision at the top level, while another agent concentrates on
+short-term scheduling at the base level. At the start of the process a proposed order release plan
+is generated by the first agent. Accordingly the agent from the base level performs a preliminary
+scheduling of these orders and may report very expensive orders as feedback. This information is
+then taken into consideration on the top level which adapts the order release plan accordingly. Here
+the top level agent minimises the overall costs by load leveling on an aggregated model, while at
+the bottom level the agent (mainly) minimises costs backorder costs.
+%
+Finally, although the introduced actions space reduction method, named \textit{independent agents},
+works very well, it remains to be shown that the achievable policies are different to the ones for
+agents without this adaption.
+
+% \MS[t]{It remains to be shown that independent agents find the same (or better) policies.}
+
+% \MS[t]{Multiple agents on different tasks (hierarchical PP): e.g. Order acceptance + release,
+%   scheduling + release, etc.}
+
+% \MS[t]{Job shop}
 

### 2021-05-09 paper/paper.tex
update

@@
 \newcommand\thetitle{Average Reward Adjusted Deep Reinforcement Learning: Near-Blackwell-Optimal
-  Policies for Order Release}
+  Policies applied to the Order Release Problem}
 
@@
   % teilw. von IJPR paper uebernommen
-  \MS[t]{proof-of-concept work}
   An essential task in manufacturing planning and control is to determine when to release orders to
   the shop floor. One key parameter is the lead time which is the planned time that elapses between
-  the release of an order and its completion. Lead times are normally determined based on the actual
-  duration orders previously took to traverse through the production system (flow times).
+  the release of an order and its completion. Lead times are usually determined based on the
+  measured duration orders take to traverse the production system, i.e.\@ the flow times.
   Traditional order release models assume static lead times, although it has been shown that they
-  should be set dynamically to react the dynamic operational characteristics of the system.
-  Therefore, we present an order release model which sets lead times dynamically by using an
-  reinforcement learning approach. Therefore, we provide an in-depth analysis of reinforcement
-  learning to show that average reward reinforcement learning is a better approach for operations
-  research applications as discounted reinforcement learning. Additionally we present an average
-  reward reinforcement learning algorithm which infers near-Blackwell-optimal policies. We use a
-  simulation model of a \MS[t]{todo}two-stage flow-shop to compare the algorithm
-  to % and show that our approach outperforms
-  well-known order release mechanisms. We show that in the current version our proposed model using
-  reinforcement learning outperforms some, but not all other tested approaches.
+  should be set dynamically to be able to react to the dynamic operational characteristics of
+  production system. Therefore, we present an order release model which sets lead times dynamically
+  by using an reinforcement learning approach. This algorithm is designed especially for
+  applications with high variability and periodic feedback to the agent. Therefore, this paper aims
+  at providing a proof-of-concept for the proposed algorithm. Therefore, we compare our method on a
+  three-stage flow-shop simulation model not only to static order release techniques, but also to
+  the commonly used deep Q-Learning algorithm. The results show that the both reinforcement learning
+  agents have room for improvement for queuing systems with low variability, that our proposed
+  method usually outperforms deep Q-Learning and finally for systems with high variability that the
+  proposed reinforcement learning algorithm performs very well.
+  %
+  % We use a  to compare the algorithm
+  % to % and show that our approach outperforms
+  % well-known order release mechanisms. We show that in the current version our proposed model using
+  % reinforcement learning outperforms some, but not all other tested approaches.
   % EVTL: especially for scenarios with...
@@
 ensure scalability we adapt a multi-agent approach, where each agent optimises the lead time
-management of a single product type. The agents do so by assessesing the expected costs for the
+management of a single product type. The agents do so by assessing the expected costs for the
 possible releases imposed by adapting the lead time. According to the learned estimates each agent
@@
 learning with small action-based replay memories and worker agents to infer order release policies
-which outerperform \MS[t]{todo}.
+which, especially in case of high variability, outperform not only the static order release
+mechanisms, but also the commonly applied deep Q-Learning algorithm~\citep[cf.][]{mnih2015human}.
 
@@
 
-\paragraph{Structure.} The rest of the paper is structured as follows. The next section introduced
-average reward reinforcement learning and presents the used algorithm (see also
-\citealt{schneckenreither2020average}). Section~\ref{sec:Experimental_Evaluation} describes the
-simulation model we use to evaluate the approach. \MS[t]{todo}
+\paragraph{Structure.} The rest of the paper is structured as follows. The next section introduces
+average reward reinforcement learning and presents the used algorithm. For a more comprehensive
+discussion of the theoretical foundations we refer to \citet{schneckenreither2020average}.
+Section~\ref{sec:Experimental_Evaluation} describes the simulation model we use to evaluate the
+approach. The results of the computation experiments are presented in
+Section~\ref{sec:Computational}, and Section~\ref{sec:conclusion} concludes the paper.
+
+% \MS[t]{todo}
 % Then~\ref{sec:Preliminary} presents preliminary
@@
 \section{Computational Experiments}
-\label{sec:Preliminary}
+\label{sec:Computational}
 
@@
 critic will provide a better feedback of the value function update to the actor. This should lead to
-better policies than with standard actor-critic algorithms, e.g. A3C~\ref{Mnih2016asynchronous}.
+better policies than with standard actor-critic algorithms, e.g. A3C~\citep{mnih2016asynchronous}.
 %
@@
 
-\MS[t]{todo}
+% \MS[t]{todo}
 

### 2021-05-11 paper/paper.tex
update

@@
   three-stage flow-shop simulation model not only to static order release techniques, but also to
-  the commonly used deep Q-Learning algorithm. The results show that the both reinforcement learning
+  the commonly used deep Q-Learning algorithm. The results show that both reinforcement learning
   agents have room for improvement for queuing systems with low variability, that our proposed
@@
 \(\acts = \size{\States}\), the independent agent configuration, and the number of functions that
-are approximated. For the later, in case of \ARA{} this is \(2\), as we represent both functions
-\(\X^\pol_{\gamma_{0}}\) and \(\X^\pol_{\gamma_{1}}\) with the same neural network, while deep
-Q-Learning \citep[cf.][]{mnih2015human} only approximates the function \(Q_{\gamma}^\pol\).
+are approximated. For the later, in case of \ARA{} this is \(2\), as we represent both state-action
+value functions \(\X^\pol_{\gamma_{0}}\) and \(\X^\pol_{\gamma_{1}}\) with the same neural network,
+while deep Q-Learning \citep[cf.][]{mnih2015human} only approximates the function
+\(Q_{\gamma}^\pol\).
 %

### 2021-05-11 paper/paper.tex
minor

@@
   method usually outperforms deep Q-Learning and finally for systems with high variability that the
-  proposed reinforcement learning algorithm performs very well.
+  proposed reinforcement learning algorithm performs very well, outperforming all other approaches.
   %
@@
 Therefore, in the future we plan to develop \ARA{} further to a actor-critic state-of-the-art
-reinforcement learning algorithm. We hypothesise that due to the more correct state value
+reinforcement learning algorithm. We hypothesise that due to the more correct state-action value
 estimations of \ARA{} as compared to deep Q-Learning, cf. \citet{schneckenreither2020average}, the

### 2021-05-11 paper/paper.tex
minor

@@
 the top level agent minimises the overall costs by load leveling on an aggregated model, while at
-the bottom level the agent (mainly) minimises costs backorder costs.
+the bottom level the agent (mainly) minimises backorder costs.
 %

### 2021-05-11 paper/paper.tex
Updates from Overleaf

@@
   by using an reinforcement learning approach. This algorithm is designed especially for
-  applications with high variability and periodic feedback to the agent. Therefore, this paper aims
+  applications with high process variability and periodic feedback to the agent. Therefore, this paper aims
   at providing a proof-of-concept for the proposed algorithm. Therefore, we compare our method on a
@@
   agents have room for improvement for queuing systems with low variability, that our proposed
-  method usually outperforms deep Q-Learning and finally for systems with high variability that the
+  method usually outperforms deep Q-Learning and finally for systems with high utilisation that the
   proposed reinforcement learning algorithm performs very well, outperforming all other approaches.
@@
 prevent the negative effects of the lead time syndrome. Their model outperforms static and reactive
-lead time management approaches, especially for scenarios with high utilisation and high variability
-in processing times.
+lead time management approaches, especially for scenarios with high utilisation.
 
@@
 learning with small action-based replay memories and worker agents to infer order release policies
-which, especially in case of high variability, outperform not only the static order release
+which, especially in case of high utilisation, outperform not only the static order release
 mechanisms, but also the commonly applied deep Q-Learning algorithm~\citep[cf.][]{mnih2015human}.
@@
 \rlt{0.99} and \(72.94\) for \rlt{1.00} are a little bit offset. This probably results from the
-higher variability in the system.
+higher variability and utilisation in the system.
 
@@
 Interesting to see is that the measured average reward and the learned average reward are very
-different for both \ARA{} variants (about \(30\)-\(40\)). This is probably due to the very high
-variability in the system, but might also be an indicator that the agents need more time for
+different for both \ARA{} variants (about \(30\)-\(40\)). This is probably due to the high 
+variability and utilisation in the system, but might also be an indicator that the agents need more time for
 learning. However, due to the high computational complexity we encountered we leave this to future
@@
 Overall we can see that the RL agents, and especially the established \ARA{} algorithm performs very
-well in systems with high variability. In case of the less variable systems the results reveal that
+well in systems with high utilisation. In case of the less utilised systems the results reveal that
 there is still potential for the improvement of reinforcement learning variants.
@@
 
-The results suggest that for high variability the RL learning agents, and here especially the novel
-\ARA{} variant, performs very well. In case of medium variability the achieved results are as good
-as with the \BIL{} variants. And for low variability one can see that the RL agents still have room
+The results suggest that for high utilisation the RL learning agents, and here especially the novel
+\ARA{} variant, performs very well. In case of medium utilisation the achieved results are as good
+as with the \BIL{} variants. And for low utilisation one can see that the RL agents still have room
 for improvement, as \BIL{} outperforms them. When comparing deep Q-Learning to the established

### 2021-05-14 paper/paper.tex
restructuring backup

@@
 corresponding optimisation problem faces ``sampling issues'' meaning that the flow times depend on
-the lead times. An extreme scenario of this problem is the so called ``lead time syndrome'', which
-describes a vicious cycle where increasing flow times perpetually inflate the lead times which leads
-to worse performance (see e.g., \citealt{Mather1978, knollmann2013control, Selcuk2006}.
+the lead times. Thus, a completely reactive approach can lead to an erratic order release behaviour,
+the Lead Time Syndrome (see also \citealt{Mather1978, knollmann2013control, Selcuk2006}). We
+hypothesis that dynamic lead time management approaches require a predictive design, i.e.\@ the
+method has to incorporate a prediction of the future system state, to be able to compete with
+state-of-the-art order release methods from literature. This is necessary in order to detect issues,
+like the Lead Time Syndrome, in advance and react accordingly \citep{schneckenreither2020order}.
+However, there only exist three papers that propose a predictive lead time management algorithm in
+literature.
+
+One method for which the future system state is by design considered a main driver of decision
+making % that is designed to incorporate the future system state
+is reinforcement learning \citep[e.g.][]{sutton1998introduction}, one of the main three subcategories of
+machine learning. % It stems from dynamic programming the future
+% system state is by design considered as a main driver of decision making in the current period.
+Reinforcement learning is an optimisation technique that stems from dynamic programming, therefore
+also seeks to compute the best stationary policy for a given
+problem. % This policy is usually provided
+% by assessing current and future states (or state-action pairs) of an underlying Markov Decision
+% Process (MDP).
+The advantage of reinforcement learning over dynamic programming is that (i) the problem space is
+explored by an agent and thus only expectantly interesting parts of the problem space need to be
+assessed and (ii) the knowledge (acquisition) of transition probabilities becomes unnecessary as the
+states are evaluated by consecutively observed states solely.
+
+\MS{what is rl}
+
+\MS{why ARAL}
+
+\MS{summary}
+
+%
+%
+
+\section{Related Literature}
+
+This section presents related literature and is split into two parts. First we discuss lead time
+management literature and second the relevant literature of reinforcement learning.
+
+\subsection{Related Lead Time Management Literature}
+
+An extreme scenario of this problem is the so called ``lead time syndrome'', which describes a
+vicious cycle where increasing flow times perpetually inflate the lead times which leads to worse
+performance (see e.g., \citealt{Mather1978, knollmann2013control, Selcuk2006}).
 %
@@
 %
-\citet{schneckenreither2020order} have established
-following categorisation of dynamic lead time management approaches:
+\citet{schneckenreither2020order} have established following categorisation of dynamic lead time
+management approaches:
 
@@
 
-As reinforcement learning stems from dynamic programming the future system state is by design
-considered as a main driver of decision making in the current period.
-% Reinforcement learning is an optimisation technique that stems from dynamic programming and
-Its goal is to find the best stationary policy for a given problem. % This policy is usually provided
-% by assessing current and future states (or state-action pairs) of an underlying Markov Decision
-% Process (MDP).
-The advantage of reinforcement learning over dynamic programming is that (i) the problem
-space is explored by an agent and thus only expectantly interesting parts of the problem space need
-to be assessed and (ii) the knowledge (acquisition) of transition probabilities becomes unnecessary
-as the states are evaluated by consecutively observed states solely.
+\subsection{Related Reinforcement Learning Literature}
 
@@
 %
-This is due to the fact that state value is largely composed of a term defined by the policys'
+This is due to the fact that state value is largely composed of a term defined by the policy's
 average reward value \citep{MillerVeinott1969,Blackwell62} which would otherwise dilute the state

### 2021-05-14 paper/paper.tex
update

@@
 making % that is designed to incorporate the future system state
-is reinforcement learning \citep[e.g.][]{sutton1998introduction}, one of the main three subcategories of
-machine learning. % It stems from dynamic programming the future
+is reinforcement learning \citep[e.g.][]{sutton1998introduction}, one of the main three
+subcategories of machine learning. % It stems from dynamic programming the future
 % system state is by design considered as a main driver of decision making in the current period.
 Reinforcement learning is an optimisation technique that stems from dynamic programming, therefore
-also seeks to compute the best stationary policy for a given
-problem. % This policy is usually provided
-% by assessing current and future states (or state-action pairs) of an underlying Markov Decision
-% Process (MDP).
+also seeks to compute the best stationary policy for a given problem. The idea is to evaluate the
+current system state according to the direct feedback (reward), as well as the previously learned
+state value of the consecutive state. This forms a cyclic constraint problem with as many
+constraints and variables as there are system states. The solution is found through iteration and
+due to the transitivity property of the state value update.
+% as in dynamic programming is that the process and the current state and future states (or
+% state-action pairs) of an underlying Markov Decision Process (MDP).
 The advantage of reinforcement learning over dynamic programming is that (i) the problem space is
@@
 
-\MS{what is rl}
+There are two main types of reinforcement learning algorithms known in literature. The commonly
+applied discounted reinforcement learning method
+\citep[e.g.][]{sutton1998introduction,mnih2015human,mnih2016asynchronous}, in which the future
+system state is discounted by a factor \(0 < \gamma < 1\) to prevent an infinite growth of the state
+evaluations, and average reward reinforcement learning
+\citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults,
+  schwartz1993reinforcement,
+  Mahadevan97_SelfimprovingFactorySimulationUsingContinuoustimeAveragerewardReinforcementLearning}.
+%
+% Like discounted reinforcement learning also average reward reinforcement learning is based on an
+% oracle function, in our case the accumulated costs of a period, to assess the decisions taken by
+% the agent. By repeatedly choosing different actions the agent examines the problem space and rates
+% possible actions for any observed state. The advantage of average reward reinforcement learning
+The advantage of the later % over the widely applied discounted reinforcement learning framework
+is that the underlying optimisation technique is theoretically able to find better policies
+% This yields from the fact that in the standard framework each state is assessed independently and
+% by a single value. To be more precise, average reward reinforcement learning
+by splitting up the evaluation of each state into infinitely many parts. First, the average reward
+per step, second, a bias value that specifies expected reward that is additionally accumulated when
+traversing from the current state to the optimal path, and third, an infinite number of values of
+which each includes the occurrence of the rewards and in which magnitude along the path
+\citep{Howard64,Puterman94,
+  Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults}. % Thus,
+% while in average reward reinforcement learning these terms are learned separately
+To expectantly accumulate the most reward one has to choose the action that leads to the state with
+highest value of the parts, on ties selecting actions following the order above. That is, the
+highest evaluation of the average reward, the bias value, the third component, the fourth component,
+etc. The formulations for computing these values form nested constraints, so for instance the
+current estimate of the average reward is used to compute the bias value. Therefore, a slight
+approximation error may cause harmful damage in subsequent constraints, which makes average reward
+reinforcement computationally very intensive. Worsening the issue, \citet[p.343ff]{Puterman94} shows
+that an additional set of nested constraints have to be solved to gain the correct solution.
+
+In contrast the standard discounted framework estimates only one single scalar value consisting of
+the addition of these subterms, where however, the average reward is scaled by \(1/(1-\gamma)\).
+%
+To collect as much reward as possible, for instance to gain the highest expected profit, the
+discount factor \(\gamma\) has to be very close to one, e.g. \(0.99\).
+%
+% As the discount factor \(\gamma\) is usually set very
+% close to \(1\), e.g. \(0.99\), which
+Such a parameterisation leads to the fact that the subterm including the average reward dominates
+the other two.
+%
+Further, in commonly applied unichain (and thus ergodic) MDPs the average reward per step is equal
+over all states. Thus, independently assessing it is not only computationally unwisely, but also
+problematic when they are iteratively approximated. For this reasons most applications of discounted
+reinforcement learning in literature evaluate their algorithms on processes that have an average
+reward of about zero \citep{schneckenreither2020average}. \citet{schneckenreither2020average} shows
+that tabular based discounted reinforcement learning agents are unable to find even basic policies
+when confronted with processes that have a small average reward per step. Unfortunately, when
+optimising for the highest profit (or the lowest cost), the most intuitive and clear way of reward
+feedback to the agent is presenting it the profit that is accumulated every period.
+
+\MS{todo}
+%
+Therefore, we adapt an algorithm that uses a scalar value for the estimation of the average reward
+over all states, and estimates for every state-action pair that incorporate the bias value and the
+sum, usually referred to as error term, of all other parts. The later values are adjusted by the
+average reward.
+% This, combined with the iterative evaluation and the independently assessing of the average reward
+% for each state lets standard discounted reinforcement learning struggle to find good policies.
+% Thus, this incautious combining of different kinds of state values as done in discounted
+% reinforcement learning leads to the problems that (i) the average reward, which is equal for all
+% states of usually investigated unichain MDPs, is dominating and thus diluting the bias values, and
+% (ii) the state values are deteriorated by the error term that is only imposed due to the
+% discounting technique.
 
@@
 
-Like discounted reinforcement learning also average reward reinforcement learning is based on an
-oracle function, in our case the accumulated costs of a period, to assess the decisions taken by the
-agent. By repeatedly choosing different actions the agent examines the problem space and rates
-possible actions for any observed state. The advantage of average reward reinforcement learning over
-the widely applied discounted reinforcement learning framework is that the underlying optimisation
-technique is able to find better policies. This yields from the fact that in standard discounted
-reinforcement learning method the states are assessed independently and by a single value. To be
-more precise, average reward reinforcement learning splits up the evaluation of the average reward
-per step, a bias value that specifies the amount of collected rewards to reach the optimal path when
-starting in a suboptimal state and an error term which defines the number of steps to reach the
-optimal path \citep{Howard64,Puterman94,
-  Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults}. % Thus,
-% while in average reward reinforcement learning these terms are learned separately
-In contrast to that in the standard discounted framework one single scalar value consisting of the
-addition of these subterms is estimated, where however the average reward is scaled by
-\(1/(1-\gamma)\). The discount factor \(\gamma\) is usually set very close to \(1\), e.g. \(0.99\),
-which leads to the fact that this subterm dominates the other two.
-%
-Further, in commonly applied unichain (and thus ergodic) MDPs the average reward per step is equal
-for all states. Thus, independently assessing it is not only computationally unwisely, but also
-problematic when iteratively annealed as done in reinforcement learning. Therefore, we adapt an
-algorithm that uses a scalar value for the estimation of the average reward over all states, and
-estimates for every state-action pair that incorporate the bias and error term. The later values are
-adjusted by the average reward.
-% This, combined with the iterative evaluation and the independently assessing of the average reward
-% for each state lets standard discounted reinforcement learning struggle to find good policies.
-% Thus, this incautious combining of different kinds of state values as done in discounted
-% reinforcement learning leads to the problems that (i) the average reward, which is equal for all
-% states of usually investigated unichain MDPs, is dominating and thus diluting the bias values, and
-% (ii) the state values are deteriorated by the error term that is only imposed due to the
-% discounting technique.
-
+\MS[]{von hier}
 Thus, as opposed to the commonly applied discounted reinforcement learning algorithm we use an

### 2021-05-15 paper/paper.tex
rel lit

@@
 when confronted with processes that have a small average reward per step. Unfortunately, when
-optimising for the highest profit (or the lowest cost), the most intuitive and clear way of reward
-feedback to the agent is presenting it the profit that is accumulated every period.
-
-\MS{todo}
-%
-Therefore, we adapt an algorithm that uses a scalar value for the estimation of the average reward
-over all states, and estimates for every state-action pair that incorporate the bias value and the
-sum, usually referred to as error term, of all other parts. The later values are adjusted by the
-average reward.
+optimising for the highest profit (or the lowest cost) in the area of operations management, the
+most intuitive and clear way of reward feedback to the agent is presenting it the profit that is
+accumulated every period.
+
+Therefore, we develop, adapt and enhance an algorithm that uses only one scalar value for the
+estimation of the average reward over all states. Further it estimates for every state-action pair
+the bias value and the sum of all other state value parts, usually referred to as error term. The
+later values are adjusted by the average reward.
 % This, combined with the iterative evaluation and the independently assessing of the average reward
@@
 % discounting technique.
-
-\MS{why ARAL}
-
-\MS{summary}
-
+%
+%
+Thus, as opposed to the commonly applied discounted reinforcement learning algorithm we use an
+average reward adjusted reinforcement learning algorithm to dynamically set lead time for order
+release. In contrast to the aforementioned works on average reward reinforcement learning our
+algorithm incorporates the optimisation of not only the average reward over time, but also the bias
+values, which is an important adaption in highly stochastic systems. Only the work by
+\citet{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}
+integrates this second-level refinement optimisation. However, their average reward reinforcement
+learning algorithm requires the selection of a reference state to prevent an infinite increase of
+state values. This results from the lack of feedback in the iterative process of optimising
+different sets of nested constraints. Furthermore, they asses the average reward independently for
+each state.
+
+To achieve a scalable lead time setting algorithm, we propose an extension that reduces the action
+space tremendously. In fact, it transforms the problem size function measured on the number of
+product types from one with an exponential increase to a linear one. The idea is that multiple
+action decision agents are introduced that operate on the same state, but each one is responsible
+for the lead time setting of only one product type.
+% (i) To
+% ensure scalability we adapt a multi-agent approach, where each agent optimises the lead time
+% management of a single product type. The agents do so by assessing the expected costs for the
+% possible releases imposed by adapting the lead time. According to the learned estimates each agent
+% sets a planned lead time for the corresponding product type and with that releases orders into the
+% production system.
+
+Finally, we provide a proof-of-concept experimental section by testing the proposed algorithm on a
+three stage flow shop with six product types, which provides ample evidence of the viability of the
+approach.
+% and are compared to static order release
+% mechanisms. Furthermore, we show that the advantage of average reward reinforcement learning for the
+% problem structures encountered in the context of production and logistics management by comparing
+% the results to the discounted reinforcement learning
+% framework.
+The highly optimised value-iteration algorithm with improved n-step learning, small action-based
+replay memories and concurrent workers infers order release policies which, especially in case of
+high utilisation, outperform not only the static order release mechanisms, but also the commonly
+applied deep Q-Learning algorithm~\citep[cf.][]{mnih2015human}.
+% With regard to practical implications we are confident that decision support
+% tools based on average reward reinforcement learning increase the decision quality of human
+% planners.
+%
+In summary our contributions are as follows:
+\begin{itemize}
+\item We propose a novel \textit{average reward adjusted deep reinforcement learning algorithm} \ARA{}.
+  This is the first average reward adjusted reward learning algorithm that operates using a
+  artificial neural network as function approximation. It includes several known extensions like
+  n-step learning \citep[e.g.][]{mnih2016asynchronous} or experience replay memory
+  \citep[e.g.][]{mnih2015human}, which we advanced to prevent catastrophic forgetting by
+  action-based selection of experiences.
+\item We establish a action reduction method, named \textit{independent agents}, as a subset of the
+  action space is treated as if there were multiple independent agents, although they share the same
+  state. It reduces the growth of the number of possible actions from a exponential growth to a
+  linear one. This is especially useful for product type based lead time management, as each agent
+  is responsible for the lead time of one product type.
+\item Using a three-stage flow shop model we show in an \textit{experimental evaluation} that the
+  proposed algorithm outperforms standard Q-Learning and well established order release methods,
+  especially for systems with high variability.
+\end{itemize}
+
+%
+% \MS[t]{TODO}
+
+\paragraph{Structure.} The rest of the paper is structured as follows. The next section introduces
+average reward reinforcement learning and presents the used algorithm. For a more comprehensive
+discussion of the theoretical foundations we refer to \citet{schneckenreither2020average}.
+Section~\ref{sec:Experimental_Evaluation} describes the simulation model we use to evaluate the
+approach. The results of the computation experiments are presented in
+Section~\ref{sec:Computational}, and Section~\ref{sec:conclusion} concludes the paper.
+
+% \MS[t]{todo}
+% Then~\ref{sec:Preliminary} presents preliminary
+% result data, whereas Section~\ref{sec:conclusion} concludes this working paper.
 %
@@
 
-An extreme scenario of this problem is the so called ``lead time syndrome'', which describes a
-vicious cycle where increasing flow times perpetually inflate the lead times which leads to worse
-performance (see e.g., \citealt{Mather1978, knollmann2013control, Selcuk2006}).
-%
-Thus, setting lead times dynamically harbors optimisation potential \citep{hoyt1978dynamic}, but may
-also substantially degrade the system performance.
+Setting of lead times dynamically is a rather difficult task. Therefore, most state-of-the-art order
+release mechanisms use static or fixed lead times to address the order release problem.
+%
+However, lead times should be set dynamically due to the non-linear relationship between utilisation
+and flow times \citealt{Pahl2007,hoyt1978dynamic}, but this may also substantially degrade the
+system performance. For instance, a solely reactive approach can result in the lead time syndrome.
+%
+The \textit{lead time syndrome} describes a vicious cycle where increasing flow times perpetually
+inflate the lead times which leads to worse performance (see e.g., \citealt{Mather1978,
+  knollmann2013control, Selcuk2006}).
+%
 %
@@
 \begin{itemize}
-\item \textsf{Reactive lead time management} approaches set lead times by reacting on
-  earlier flow times (e.g., \citealt{enns2004work, Selcuk2006}). Note that the forecast is always
-  based on \textit{past} data as the most recent system changes cannot be reflected by flow times
-  until the corresponding orders arrive in the FGI, which might take several
-  periods.
-\item
-  \textsf{Proactive lead time management} may incorporate \textit{past} data in conjunction with
-  the \textit{current} system state to set lead times (e.g., \citealt{Bertrand1983,
-    ChungHuang2002}). Put differently, these methods aim to find a function that provides lead times
-  based on the current state of the production system and possibly information from the past.
-
-\item \textsf{Predictive lead time management} may not only incorporate \textit{past} data and the
-  \textit{current} system state to set lead times, but also utilises the anticipated \textit{future}
-  system state to detect arising issues of future periods and react accordingly (e.g.
-  \citealt{PaterninaArboleda2001,schneckenreither2020order,Schneckenreither2019}). Thus, it extends
-  proactive lead time management from a flow time forecasting or simple lead time setting technique
-  to a lead time management approach that integrates the future behaviour of the system when setting
-  lead times. This allows reasoning of the system dynamics, as for instance triggering the lead time
-  system, and thus such an algorithm can react accordingly to find a more farsighted optimal lead
-  time update.
+\item \textsf{Reactive lead time management} sets lead times by reacting on earlier flow times
+  (e.g., \citealt{enns2004work, Selcuk2006}), i.e.\@ data from the past.
+\item \textsf{Proactive lead time management} may incorporate \textit{past} data and the
+  \textit{current} system state to set lead times (e.g., \citealt{Bertrand1983, ChungHuang2002}).
+\item \textsf{Predictive lead time management} may incorporate \textit{past} data and the
+  \textit{current} system state, and uses an anticipated \textit{future} system state to prevent
+  arising issues of future periods (e.g.
+  \citealt{PaterninaArboleda2001,schneckenreither2020order,Schneckenreither2019}).
 \end{itemize}
@@
 novel value-iteration reinforcement learning agent which exceeds human-level abilities in playing
-many classic Atari 2600 games~\citep{bellemare2012investigating}. Further,
-\citet{mnih2016asynchronous} present improved results with asynchronous actor-critic reinforcement
-learning. Also games like Go~\citep{silver2016mastering} and Chess~\citep{silver2017mastering} have
-been mastered with superhuman performance by \textit{tabula rasa} reinforcement learning agents.
-Furthermore, the method has also been applied in the setting of manufacturing system, e.g.\@ to
-improve the ramp-up process~\citep{doltsinis2012reinforcement}, in locally selecting appropriate
-dispatching rules \citep{zhang1995reinforcement,wang2005application} or scheduling
-\citep{zhang1995reinforcement, waschneck2018optimization}.
+many classic Atari games~\citep{bellemare2012investigating}. Further, \citet{mnih2016asynchronous}
+present improved results with asynchronous actor-critic reinforcement learning. Also games like
+Go~\citep{silver2016mastering} and Chess~\citep{silver2017mastering} have been mastered with
+superhuman performance by \textit{tabula rasa} reinforcement learning agents. Furthermore, the
+method has also been applied in the setting of manufacturing system, e.g.\@ to improve the ramp-up
+process~\citep{doltsinis2012reinforcement}, in locally selecting appropriate dispatching rules
+\citep{zhang1995reinforcement,wang2005application} or scheduling \citep{zhang1995reinforcement,
+  waschneck2018optimization}.
 %
@@
 
-\MS[]{von hier}
-Thus, as opposed to the commonly applied discounted reinforcement learning algorithm we use an
-average reward adjusted reinforcement learning algorithm to adaptively release orders based on the
-assessed state values of the production system. In contrast to the aforementioned works on average
-reward reinforcement learning our algorithm incorporates the optimisation of not only the average
-reward over time, but also the bias values, which is an important adaption in highly stochastic
-systems. Only the work by
-\citet{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}
-integrates this second-level refinement optimisation. However, their algorithm requires the
-selection of a reference state to prevent an infinite increase of state values. This results from
-the lack of feedback in the iterative process of optimising the different refinement optimisation
-levels. Furthermore, they asses the average reward independently for each state.
-
-In summary, we propose a novel average reward adjusted reinforcement learning algorithm to assess
-orders for their release to the shop floor. This is the first average reward adjusted reward
-learning algorithm that operates using a artificial neural network as function approximation. To
-ensure scalability we adapt a multi-agent approach, where each agent optimises the lead time
-management of a single product type. The agents do so by assessing the expected costs for the
-possible releases imposed by adapting the lead time. According to the learned estimates each agent
-sets a planned lead time for the corresponding product type and with that releases orders into the
-production system. The highly optimised value-iteration algorithm uses improved n-step reinforcement
-learning with small action-based replay memories and worker agents to infer order release policies
-which, especially in case of high utilisation, outperform not only the static order release
-mechanisms, but also the commonly applied deep Q-Learning algorithm~\citep[cf.][]{mnih2015human}.
-
-%
-% \MS[t]{TODO} We provide ample evidence of the viability of the approach and compare it to well-known
-% order release mechanisms. Furthermore, we show that the advantage of average reward reinforcement
-% learning for the problem structures encountered in the context of production and logistics by
-% comparing the results to the discounted reinforcement learning framework. The preliminary results
-% show that our approach performs better than the other methods. With regard to practical implications
-% we are confident that decision support tools based on average reward reinforcement learning increase
-% the decision quality of human planners.
-
-\paragraph{Structure.} The rest of the paper is structured as follows. The next section introduces
-average reward reinforcement learning and presents the used algorithm. For a more comprehensive
-discussion of the theoretical foundations we refer to \citet{schneckenreither2020average}.
-Section~\ref{sec:Experimental_Evaluation} describes the simulation model we use to evaluate the
-approach. The results of the computation experiments are presented in
-Section~\ref{sec:Computational}, and Section~\ref{sec:conclusion} concludes the paper.
-
-% \MS[t]{todo}
-% Then~\ref{sec:Preliminary} presents preliminary
-% result data, whereas Section~\ref{sec:conclusion} concludes this working paper.
+% Only the work by
+% \citet{Mahadevan96_AnAveragerewardReinforcementLearningAlgorithmForComputingBiasoptimalPolicies}
+% integrates this second-level refinement optimisation. However, their average reward reinforcement
+% learning algorithm requires the selection of a reference state to prevent an infinite increase of
+% state values. This results from the lack of feedback in the iterative process of optimising the
+% different refinement optimisation levels. Furthermore, they asses the average reward independently
+% for each state.
 

### 2021-05-15 paper/figures/productionsystem.tex
restructing

@@
   \path (M2) edge[->, bend left=10,above,near start ]      node {$\scriptstyle 2$}     (Q5);
-  \path (M2) edge[->, bend left=10,above,near start ]      node[xshift=0.07cm,yshift=-0.12cm] {$\scriptstyle 3$}     (Q6);
-  \path (M3) edge[->, bend right=10,above,very near start] node[xshift=0.22cm,yshift=-0.09cm] {$\scriptstyle 4$}     (Q4);
+  \path (M2) edge[->, bend left=10,above,near start ]      node[xshift=0.17cm,yshift=-0.2cm] {$\scriptstyle 3$}     (Q6);
+960  \path (M3) edge[->, bend right=10,above,very near start] node[xshift=0.32cm,yshift=0.0cm] {$\scriptstyle 4$}     (Q4);
   \path (M3) edge[->, bend right=10,above,very near start]      node[yshift=-0.075cm] {$\scriptstyle 5$}     (Q5);

### 2021-05-15 paper/paper.tex
restructing

@@
 
-\paragraph{Structure.} The rest of the paper is structured as follows. The next section introduces
-average reward reinforcement learning and presents the used algorithm. For a more comprehensive
-discussion of the theoretical foundations we refer to \citet{schneckenreither2020average}.
-Section~\ref{sec:Experimental_Evaluation} describes the simulation model we use to evaluate the
-approach. The results of the computation experiments are presented in
-Section~\ref{sec:Computational}, and Section~\ref{sec:conclusion} concludes the paper.
+\paragraph{Structure.} The rest of the paper is structured as follows. \MS[t]{TODO} The next section
+introduces average reward reinforcement learning and presents the used algorithm. For a more
+comprehensive discussion of the theoretical foundations we refer to
+\citet{schneckenreither2020average}. Section~\ref{sec:Experimental_Evaluation} describes the
+simulation model we use to evaluate the approach. The results of the computation experiments are
+presented in Section~\ref{sec:Computational}, and Section~\ref{sec:conclusion} concludes the paper.
 
@@
 
-We hypothesis that dynamic lead time management approaches need to aim for a predictive design in
+\MS[]{todo: Reactive}
+
+\MS[]{todo: proactive}
+
+We hypothesise that dynamic lead time management approaches need to aim for a predictive design in
 order to be able to compete with state-of-the-art order release methods from literature. However,
@@
 triggering of the lead time syndrome and therefore they introduce an upper lead time bound to
-prevent the negative effects of the lead time syndrome. Their model outperforms static and reactive
-lead time management approaches, especially for scenarios with high utilisation.
-
-\subsection{Related Reinforcement Learning Literature}
+prevent the negative effects of the syndrome. Their model outperforms static and reactive lead time
+management approaches, especially for scenarios with high utilisation.
+
+\subsection{Related Reinforcement Learning Applications}
 
@@
 
-\section{Average Reward Adjusted Reinforcement Learning}
-
-This section briefly reintroduces the most important concepts of average reward adjusted
+\subsection{Reinforcement Learning Preliminaries}
+
+The rest of this section briefly reintroduces the important concepts of average reward adjusted
 reinforcement learning, elaborates on optimality criteria and provides insights of the underlying
@@
 
-Like \cite{MillerVeinott1969} we are considering problems that are observed in a sequence of points
+Like \citet{MillerVeinott1969} we are considering problems that are observed in a sequence of points
 in time labeled \(1,2,\ldots\) and can be modelled using a finite set of states \(\States\),
@@
 
-\paragraph{Average Reward Adjusted Deep RL Algorithm.} %
+\section{Average Reward Adjusted Deep RL Algorithm }
+
 The model-free average reward adjusted RL algorithm is based on the tabular version of
@@
 product type basis and thus is not directly comparable to their order based lead time management
-approach. Nonetheless, this section is largely based on the corresponding section of
-\citet{schneckenreither2020order}.
+approach. % Nonetheless, this section is largely based on the corresponding section of
+% \citet{schneckenreither2020order}.
 
@@
 relationship between flow times and high utilisation levels.
+%%%%%%
+% Calculation
+% (285 / (3 * u)) = interarrival time
+%
+% 285... Proc time bottleneck
+% 3  ... every third product routes of this machine
+% u  ... utilisation, e.g. 0.90
 
@@
 
+\section{Lead Time Setting}
+
 \subsection{Conventional Order Release Rule}
@@
 
-\subsection{Algorithm Setup}
+\subsection{Average Reward Adjusted Deep RL Algorithm}
 \label{subsec:Algorithm_Setup}
@@
 %
-The algorithm is based around The agent first choose an action (steps \(4-5\)), executes the
-selected action and stores the experience in the replay memory (step \(6\)) and then samples a
-mini-batch of experiences each of length \(nstep\) to improve the policy (steps \(7-21\)). Step
-\(17\) uses an exponentially smoothed overestimation (by \(2.5\%\)) of the average reward and rate
-\(2^{-5}\). All average reward updates above this overestimation are rejected.
+The agent first choose an action (steps \(4-5\)), executes the selected action and stores the
+experience in the replay memory (step \(6\)) and then samples a mini-batch of experiences each of
+length \(nstep\) to improve the policy (steps \(7-21\)). Step \(17\) uses an exponentially smoothed
+overestimation (by \(2.5\%\)) of the average reward and rate \(2^{-5}\). All average reward updates
+above this overestimation are rejected.
 

### 2021-05-16 paper/paper.tex
update

@@
   by using an reinforcement learning approach. This algorithm is designed especially for
-  applications with high process variability and periodic feedback to the agent. Therefore, this paper aims
-  at providing a proof-of-concept for the proposed algorithm. Therefore, we compare our method on a
-  three-stage flow-shop simulation model not only to static order release techniques, but also to
-  the commonly used deep Q-Learning algorithm. The results show that both reinforcement learning
-  agents have room for improvement for queuing systems with low variability, that our proposed
-  method usually outperforms deep Q-Learning and finally for systems with high utilisation that the
-  proposed reinforcement learning algorithm performs very well, outperforming all other approaches.
+  applications with high process variability and periodic feedback to the agent. Therefore, this
+  paper aims at providing a proof-of-concept for the proposed algorithm. Therefore, we compare our
+  method on a three-stage flow-shop simulation model not only to static order release techniques,
+  but also to the commonly used deep Q-Learning algorithm. The results show that both reinforcement
+  learning agents have room for improvement for queuing systems with low variability, that our
+  proposed method usually outperforms deep Q-Learning and finally for systems with high utilisation
+  that the proposed reinforcement learning algorithm performs very well, outperforming all other
+  approaches.
   %
@@
 
-\paragraph{Structure.} The rest of the paper is structured as follows. \MS[t]{TODO} The next section
-introduces average reward reinforcement learning and presents the used algorithm. For a more
-comprehensive discussion of the theoretical foundations we refer to
-\citet{schneckenreither2020average}. Section~\ref{sec:Experimental_Evaluation} describes the
-simulation model we use to evaluate the approach. The results of the computation experiments are
-presented in Section~\ref{sec:Computational}, and Section~\ref{sec:conclusion} concludes the paper.
+\paragraph{Structure.} The rest of the paper is structured as follows. The next section introduces
+related literature. Section~\ref{sec:avg} establishes the average reward adjusted deep reinforcement
+learning algorithm and the extension to independent agents.
+Section~\ref{sec:Experimental_Evaluation} describes the simulation model we use to evaluate the
+approach. In Section~\ref{sec:lead} we briefly discuss lead time setting in this paper, while the
+results of the computation experiments are presented in Section~\ref{sec:Computational}. Finally
+Section~\ref{sec:conclusion} concludes the paper.
 
@@
 
-This section presents related literature and is split into two parts. First we discuss lead time
-management literature and second the relevant literature of reinforcement learning.
+This section is split into several parts. First we discuss lead time management literature, then
+reinforcement learning applications to lead time management, third we briefly review the relevant
+notation to reinforcement learning and finally we present the relevant literature of reinforcement
+learning in the rest of the section. The later includes a discussion of discounted reinforcement
+learning, average reward reinforcement learning, the Laurent series expansion of the state values,
+and finally average reward adjusted reinforcement learning.
 
@@
 foundation of our algorithm.
+
+\section{Average Reward Adjusted Deep RL Algorithm }\label{sec:avg}
+
+This section presents the average reward adjusted deep reinforcement learning algorithm \ARA{}.
+First, we present a basic version, which suffices for small problem sizes and introduce the idea of
+independent agents, before presenting an scalable optimised version.
+
+\subsection{Basic Average Reward Adjusted Deep RL Algorithm}
 
@@
 
-\section{Average Reward Adjusted Deep RL Algorithm }
-
 The model-free average reward adjusted RL algorithm is based on the tabular version of
@@
 
-\section{Experimental Evaluation}
-\label{sec:Experimental_Evaluation}
-
-This section introduces the simulation model and the parameterisation of the algorithm. To provide a
-proof-of-concept for the proposed algorithm we adapt the flow shop presented in
-\citet{schneckenreither2020order}. However, our algorithms operates on discrete lead times on a
-product type basis and thus is not directly comparable to their order based lead time management
-approach. % Nonetheless, this section is largely based on the corresponding section of
-% \citet{schneckenreither2020order}.
-
-\subsection{Simulation Model}
-\label{subsec:Simulation_Model}
-
-We adapt the simulation model of a make-to-order (MTO) flow shop with six different
-products and six work-centers of \citet{schneckenreither2020order} and examine the system under
-moderate and high utilisation in combination with moderate or high variability in demand and
-processing times. This section provides the details of the simulation model and the experimental
-setup.
-
-% \subsection{Make-To-Order Flow Shop}
-% \label{subsec:make-to-order_flow_shop}
-% \subsection{Simulation model}
-% \label{subsec:Simulation_model}
-
-The simulation model is depicted in Figure~\ref{fig:ps}, where the customers initiate the process by
-placing orders as indicated by the dashed line.
-%
-\begin{figure*}[t!]
-  \centering
-  \input{figures/productionsystem}                      % Figure production system
-  \caption{Production System of the Simulation Model with routing, processing time distributions and
-    demand interarrival time distributions.} \label{fig:ps}
-\end{figure*}
-%
-%
-The manufacturer produces six different product types, each of which is identified by a natural
-number ${1,2,\ldots,6}$ and has a different routing through the production system. The routing setup
-is provided by corresponding labels of the edges. For instance, product types $1-3$ are routed from
-machine $M1$ to machine $M2$, while the other products are transferred to machine $M3$. The
-production system consists of three production stages with diverging material flow and no return
-visits.
-%
-The incoming orders are uniformly distributed among the product types.
-%
-A period lasts $960$ minutes ($=16$ hours) which represent two \(8\)-hour shifts. At the
-beginning of each period orders can be released into the production system assuming material
-availability for all orders at their release date. Upon release, the orders are
-placed in the buffer at machine $M1$. Buffers and inventories are plotted as filled squares. Thus,
-each workstation is equipped with a buffer which queues the orders until processing is started. All
-buffers use a first-come-first-serve dispatching rule.
-%
-Furthermore, each workstation processes only one order at a time and early deliveries are
-prohibited. Orders completed ahead of their due date remain in the finished goods inventory (\fgi{})
-until they are due and sent to the customer.
-
-\paragraph{Processing Times.} The machine processing time distributions are specified under the
-corresponding node labels of the machines. To simulate high % or moderate
-variance the processing times for all machines are drawn from an exponential % or an uniform
-distribution% , respectively
-. There is one bottleneck machine ($M5$) and therefore we refer to product types $2$ and $5$, those
-routed through machine $M5$, as bottleneck products whereas products $1,3,4,6$ are non-bottleneck
-products.
-
-\paragraph{Demand.} The incoming orders are placed in the order pool with a due date slack of
-\dds{} periods, that is each order arriving within period $t$ will be due at the end of period
-$t+\dds{}$. The due date slack of \(\dds{}\) was chosen to (i) ensure sufficient time between the
-first occurrence of orders in the order pool and the latest possible release of orders and (ii) to
-have a clear cause and effect relationship between the order release decision and the cost
-performance in the analysis.
-%
-To simulate medium and high variability of the demand process the interarrival time between
-consecutive incoming orders is either drawn from an exponential ($\Exp$) or an uniform ($\Unif$)
-distribution, which are adjusted to yield the desired bottleneck utilisation level of either $70\%$
-with $\Exp(135)$ and $\Unif(95,175)$, $80\%$ with $\Exp(118)$ and $\Unif(78,158)$, or $90\%$ with
-$\Exp(105)$ and $\Unif(65,145)$. These values were chosen as they comprise the non-linear
-relationship between flow times and high utilisation levels.
-%%%%%%
-% Calculation
-% (285 / (3 * u)) = interarrival time
-%
-% 285... Proc time bottleneck
-% 3  ... every third product routes of this machine
-% u  ... utilisation, e.g. 0.90
-
-\paragraph{Order Release.}
-%
-Before the order release decision is made, all orders in the order pool are sorted by due date. This
-portrays the implementation of a sequencing rule. According to this sequence, orders are considered
-for release in the beginning of each period starting with the highest priority order.
-
-In our model, orders are released by specifying lead times
-$\lt_{1}, \lt_{2}, \ldots, \lt_{6} \geqslant 1$, where the index corresponds to the product type and
-$\lt_{i} \in \mathbb{N}$.
-%
-%
-By setting a lead time $\lt_{i}$ a \emph{planned release date} is computed for each
-order $j$ of product type $i$ in the order pool given by
-%
-\begin{linenomath*}
-  \begin{equation}
-    \prd{}_j= \dd{}_j - \lt_{i}\ ,
-    \label{eq:Orel}
-  \end{equation}
-\end{linenomath*}
-
-where $\dd_{j}$ denotes the due date of the order.  For dynamic order releases the lead times
-$\lt_{i}$ may vary from period to period, whereas in static order release methods the lead times are
-predetermined and fixed (\citealt{RagatzMabert1988,KimBobrowski1995}).
-%
-For both approaches a job of product type $i$ with lead time $\lt{}_{i}$ in period $t$ and due date
-$\dd_{j}$ is released at the end of period $t$ if and only if $t \geqslant \prd{}_{j}$.
-%
-Thus, in the dynamic setting  the planned
-release date $\prd{}_{j}$ of order $j$ can be updated several times before it is actually released
-to the production system.
-%
-However, once an order is released its release cannot be revoked.
-%
-% If there are several orders with the same due date the order release
-% sequence is randomized.
-%
-Hence, if the set lead time corresponds with the actual flow time the product is finished in the
-same period as it is shipped. If the lead times are set either too long or short,
-the order has to wait in the \fgi{} until it is due or the order is late and backorder (BO) costs
-occur.
-
-\paragraph{Costs.}
-%
-The cost parameters are set by assuming an increase in value from raw material ($1$~Dollar per order
-and period) to the final product ($4$~Dollar per order and period) and the backorder costs are set
-very high ($16$~Dollar per order and period) due to the MTO environment. All costs are assessed
-based on the production system state at the end of each period.
-
-\section{Lead Time Setting}
-
-\subsection{Conventional Order Release Rule}
-
-As external benchmark for comparison we use different parameterized backward infinite loading
-(\BIL{}) techniques \citep{Ackerman1963}. \BIL{} uses Equation~\ref{eq:Orel} with a predetermined
-and fixed lead times \(\lt_{i}\).
-%
-% \begin{equation*}
-%   \rd_{j} = \dd_{j} - \lt_{j} \tcom
-%   \label{eq:BIL}
-% \end{equation*}
-%
-% where the release date \(RD_{j}\) of product type \(j\) is calculated by the difference of the due
-% date \(DD_{j}\) and the lead time \(LT_{j}\).
-%
-% Note that we set lead times for each product group and not for every order as done by
-% \cite{Ackerman1963}.
-
-\subsection{Average Reward Adjusted Deep RL Algorithm}
-\label{subsec:Algorithm_Setup}
-
-\paragraph{Markov Decision Process.} The underlying MDP is unichain and looks as follows. Both,
-state space $\mathcal{S}$ and action space $\mathcal{A}$ are discrete.
-%
-% Recall that we products routed over the bottleneck machine \(M5\) are bottleneck products.
-%
-For the \textit{state space} we adapt the result of \citet{knollmann2013control}, which state that
-the lead times ought to control the WIP level, instead of the due date reliability as changes are
-directly seen. Thus, any state $s \in \mathcal{S}$ of the state space is composed of the following
-information for each product type \(\pt\):
-\begin{itemize}
-
-\item The currently set lead time $LT_{g} \in \{1,2,\ldots,\DDS\}$ (Recall:
-  $\text{Due Date Period} - \text{Lead Time} = \text{Release Period}$). Note that we bound the
-  maximum lead time with due date slack $\DDS$, which is $7$ in our setup.
-
-\item Counters $OP_{g,d}\in \N$ for the number of orders in the order pool divided in time buckets
-  with $d \in \{1,2,\ldots,\DDS \}$, which stands for the number of periods until the due date.
-
-  % \item Flag for each machine whether it is idle or not.
-
-\item Counters $Q_{g,i,d}\in \N$ for the number of orders in the machine and queue \(i \in
-  \{M_{1}, M_{2}, \ldots , M_{6}\}\) divided in
-  time buckets with $d \in \{1,2,\ldots,\DDS \}$, which stands for the number of periods until the
-  due date.
-  % \MS[t]{depend}If a product type is not routed over the corresponding machine, then no entry for this
-  % product type is specified.
-
-\item Counters $FGI_{g,d} \in \N$ of orders in the finished goods inventory divided in time
-  buckets with $d \in \{1, 2,\ldots, \DDS \}$, which stands for the number of periods until the
-  due date. Orders with a due date with more than 3 periods ago are listed in the counter
-  $FGI_{-3}$.
-
-\item Counters $S_{g,d} \in \N$ of shipped orders from the last period divided in time buckets
-  with $d \in\{-3,-2,\ldots, \DDS \}$, which stands for the number of periods until the due date.
-  Orders with a due date with more than 3 periods ago are listed in the counter $S_{-3}$.
-
-\end{itemize}
-
-The algorithm implicitly learns a function which maps the current state of the production
-system to a lead time update and thus a release decision. In an optimal situation it does so by
-multiply exploring every action for each state and assessing its economic viability.
-%
-Orders are released once the due date is within the interval $[t,t+\text{lead time}]$, whereas $t$
-is the current period. Clearly, this yields bulk releases, meaning that either all or no orders with
-the same due date and of the same product type are released.
-
-The \textit{action space} is composed of two decisions, the relative change of the lead times to the
-currently set lead times $LT_{g} \in \{1,2,\ldots,\DDS\}$ for each product type $\pt$.
-%
-We restrict the lead time update for state $s_{t+1}$ according to the lead time $LT_{g}$ from state
-$s_{t}$ by a maximum change of 1. Thus, if $LT_{g}$ is the current lead time of product type $\pt$
-the agent can choose a lead time in $\{1,2,\ldots,\DDS\} \cap \{LT_{g}-1,LT_{g},LT_{g}+1\}$. % Put
-% differently the algorithm can increase or decrease the lead time by 1 or leave it as it is, as long
-% as it acts within the discrete action space given by the set $\{1,2,\ldots,\DDS\}$ for each product
-% type.
-A naive approaches is to define one action for each combination of actions over all product type.
-This however, due to the exponential growth, results in only being able to solve very small
-production systems with only few product types. A common solution in production planning is to
-aggregate the products into product groups, cf.\@ hierarchical production
-planning~\citep{schneeweibeta1995hierarchical}.
-%
-Although applicable in our setup, if we group the products into bottleneck and non-bottleneck
-products as done by \citet{schneckenreither2020order}, also this approach simply does not scale.
+\subsection{Independent Agents}
+
+In our application the agent can update the lead time for each product type by a maximum magnitude
+of \(1\) in regard to the lead time of the current period. A naive approaches is to define one
+action for each combination of possible lead time updates over all product types, e.g.\@ one action
+for a system with six product types could be the vector \(+1, +1, 0, 0, +1, 0\). This however, due
+to the exponential growth, results in only being able to solve very small production systems with
+only few product types. A common solution in production planning is to aggregate the products into
+product groups, cf.\@ hierarchical production planning~\citep{schneeweibeta1995hierarchical}.
+%
+Although applicable in our setup if we group the products into bottleneck and non-bottleneck
+products as done by \citet{schneckenreither2020order}, we refrain from this adaption as it does not
+scale to larger sized production systems.
 
@@
 current state and action taken, the corresponding probabilities for future state and reward change
-during the learning process as the the agent starts to behave differently.
+during the learning process as the agent starts to behave differently.
 %
@@
 
-\paragraph*{Reward.}
-In each period the agents choose actions, which when executed generate a reward while traversing to
-the next period by simulating the production system. The rewards are the accumulated costs at the
-end of the period consisting of the backorders, the WIP orders and the number of orders in the
-FGI\@. The objective is to minimise the returned reward. % Therefore, we adapt the algorithm by
-% swapping all maximisation functions with minimisations.
-
-\paragraph*{Optimised Algorithm.}
-Our implementation includes several optimisations, that help the algorithm in better exploring the
-solution space and stabilise the learning process. These include the above discussed independent
-agents that operate on the same state, \(n\)-step learning \citep{mnih2016asynchronous},
-% overestimating the average reward \(\avgrew^{\pol}\)
-and ensuring to constantly reduce the average reward as searching for policies with higher average
-costs is unnecessary. Algorithm~\ref{alg:full}
-depicts a detailed view on the algorithm including these extensions and with the objective to
-minimise the returned rewards. The idea is the same as in Algorithm~\ref{alg:near}, except that the
-action selection procedure is a loop over all independent agents and the implementation of the
-n-step logic.
-%
-The agent first choose an action (steps \(4-5\)), executes the selected action and stores the
-experience in the replay memory (step \(6\)) and then samples a mini-batch of experiences each of
-length \(nstep\) to improve the policy (steps \(7-21\)). Step \(17\) uses an exponentially smoothed
-overestimation (by \(2.5\%\)) of the average reward and rate \(2^{-5}\). All average reward updates
-above this overestimation are rejected.
-
-% N-Step: 5
-As in \citep{mnih2016asynchronous, mnih2015human} the algorithm uses a \textit{Replay Memory} to
-save experiences and then randomly selects experiences from there for learning. This ensures that
-the independent and identically distributed (i.i.d.) assumption of the samples for the ANN
-backpropagation algorithms are satisfied. We extend this idea by splitting the available size of the
-replay memory into subsets, where each subset represents one action and sample uniformly over these
-subsets first. This prevents overfitting when one action is overrepresented in the policy, e.g.\@ no
-lead time change occurs much more often than the other actions.
-
-\textit{N-step learning} collects a set of consecutive experiences, in our case retrieved from the
-replay memory, and connects the state values starting from the latest observed experience, such that
-the newly computed state value of the corresponding experience propagates to the experience of the
-previous period. This allows a faster convergence, as values bootstrap over multiple periods in
-contrast to just one step. The experiences are stored according to the first taken action in the
-n-step series. We uniformly sample over the actions to ensure each action is constantly trained.
-This adaption reduces catastrophic forgetting of already accumulated knowledge.
-%
-% Overestimating
-% The idea behind overestimating the average reward is to constantly seek for better policies than the
-% currently known. We overestimate the reward by subtracting \(10\%\) of the computed average reward
-% \(\avgrew^{\pol}\) and an exponentially smoothed average reward of the actually returned reward
-% values of greedy actions. Thus, if the computed and actually observed average reward coincide, the
-% overestimation equals \(0\).
-% Rho Minimum
-% Preventing an increase of the average reward with \(\avgrew_{\max}^{\pol}\) reduces the search space
-% and provides a methodology for constantly improving policy.
-%
-% Workers Min Exploration [0.01, 0.02, 0.03, 0.04, 0.05, 0.10, 0.15, 0.20]
-Furthermore, we use \(8\) additional \textsf{worker threads} (parallel RL) that operate on the same
-environment to collect a broader range of experiences~\citep[as e.g. in][]{mnih2016asynchronous}.
-%
-% Shared Rho (Yes)
-The average reward value is shared and computed as average over the main agent and all workers.
-% \MS[t]{shared rho}
-%
-%
-% \MS[t]{worker min exploration, epsilon greedy exploration}
-%
-%
-% Learning rate decay: ExponentialDecay (Just 5e-6) 0.25 100000
-% Learning rate: OptAdam 0.005 0.9 0.999 1e-8 1e-3,
-
-% Algorithms tested:
- %            AlgDQNAvgRewAdjusted 0.8 1.0 ByStateValues
- %          , AlgDQNAvgRewAdjusted 0.8 0.99 ByStateValues
- %          , AlgDQN 0.99 Exact
-
-% Alpha: 0.01, Decay: ExponentialDecay (Just 5e-5) 0.25 50000
-% AlphaRhoMin: 2e-5, Decay: ExponentialDecay (Just 2e-5) 0.25 50000
-% Gamma: 0.01, Decay: ExponentialDecay (Just 1e-3) 0.25 100000
-% Epsilon: 0.25, 0.25, Decay: No decay
-% Exploration: 1.0, Decay: ExponentialDecay (Just 0.005) 0.25 100000
-% LearnRandomAbove: 0.5
-%
-%
-% UNUSED: Delta: 0.005, Decay: ExponentialDecay (Just 5e-4) 0.25 100000
-
-The algorithm initiation strategy is to first fill the replay memory buffer by skipping any learning
-(steps \(7-21\) in Algorithm~\ref{alg:full}) until the period, that corresponds to the replay
-memory size, is reached. Then for \(500\) steps the average reward is set to the last computed
-exponentially smoothed value to prevent divergence due to the untrained network weight
-initialisation values.
-
-% \MS[t]{Extensions: overestimate rho, rho minimum, Multiple Agents (shared state,  shared rho?),
-%   init-phase (0.5 for expSmthRewRate, always adapt rho, same rho for all agents): exp smooth reward
-%   as rho, replay memory actions, n-step (describe replay memory), gradient clipping }
-% \MS[t]{Side info: Replay memory also stores (disallowed) filtered action indices for optimisation purposes and on
-% episodic tasks needs to store a bool indicating an episode end. Uses Int8 for state features values.  }
+\subsection{Scalable Average Reward Adjusted Deep RL Algorithm}
+
+The basic algorithm above suffices only for problem sets of rather small size due to the
+computational complexity and the growth of the action space. In our case, that would either require
+a production system with at most \(3\) product types, or an aggregation of product types into
+product groups. In the sense of this foundational research paper, we rather aim for an enhanced
+version of the algorithm that is scalable.
 
@@
 \end{algorithm}
+
+Therefore, our implementation includes several optimisations, that help the algorithm in better
+exploring the solution space and stabilise the learning process. These include the above discussed
+independent agents that operate on the same state, \(n\)-step learning \citep{mnih2016asynchronous},
+and ensuring to constantly reduce the average reward as searching for policies with higher average
+costs is unnecessary. Algorithm~\ref{alg:full} depicts a detailed view on the algorithm including
+these extensions and with the objective to minimise the returned rewards. The idea is the same as in
+Algorithm~\ref{alg:near}, except that the action selection procedure is a loop over all independent
+agents and the implementation of the n-step logic.
+%
+The agent first chooses an action (steps \(4-5\)), executes the selected action and stores the
+experience in the replay memory (step \(6\)) and then samples a mini-batch of experiences each of
+length \(nstep\) to improve the policy (steps \(7-21\)). Step \(17\) uses an exponentially smoothed
+overestimation (by \(2.5\%\)) of the average reward and rate \(2^{-5}\). All average reward updates
+above this overestimation are rejected.
+
+% N-Step: 5
+As in \citep{mnih2016asynchronous, mnih2015human} the algorithm uses a \textit{Replay Memory} to
+save experiences and then randomly selects experiences from there for learning. This ensures that
+the independent and identically distributed (i.i.d.) assumption of the samples for the ANN
+backpropagation algorithms are satisfied. We extend this idea by splitting the available size of the
+replay memory into subsets, where each subset represents one action and sample uniformly over these
+subsets first. This prevents overfitting when one action is overrepresented in the policy, e.g.\@ no
+lead time change occurs much more often than the other actions.
+
+\textit{N-step learning} collects a set of consecutive experiences, in our case retrieved from the
+replay memory, and connects the state values starting from the latest observed experience, such that
+the newly computed state value of the corresponding experience propagates to the experience of the
+previous period. This allows a faster convergence, as values bootstrap over multiple periods in
+contrast to just one step. The experiences are stored according to the first taken action in the
+n-step series. We uniformly sample over the actions to ensure each action is constantly trained.
+This adaption reduces catastrophic forgetting of already accumulated knowledge.
+%
+% Overestimating
+% The idea behind overestimating the average reward is to constantly seek for better policies than the
+% currently known. We overestimate the reward by subtracting \(10\%\) of the computed average reward
+% \(\avgrew^{\pol}\) and an exponentially smoothed average reward of the actually returned reward
+% values of greedy actions. Thus, if the computed and actually observed average reward coincide, the
+% overestimation equals \(0\).
+% Rho Minimum
+% Preventing an increase of the average reward with \(\avgrew_{\max}^{\pol}\) reduces the search space
+% and provides a methodology for constantly improving policy.
+%
+% Workers Min Exploration [0.01, 0.02, 0.03, 0.04, 0.05, 0.10, 0.15, 0.20]
+Furthermore, we use \(8\) additional \textsf{worker threads} (parallel RL) that operate on the same
+environment to collect a broader range of experiences~\citep[as e.g. in][]{mnih2016asynchronous}.
+%
+% Shared Rho (Yes)
+The average reward value is shared and computed as average over the main agent and all workers.
+% \MS[t]{shared rho}
+%
+%
+% \MS[t]{worker min exploration, epsilon greedy exploration}
+%
+%
+% Learning rate decay: ExponentialDecay (Just 5e-6) 0.25 100000
+% Learning rate: OptAdam 0.005 0.9 0.999 1e-8 1e-3,
+
+% Algorithms tested:
+ %            AlgDQNAvgRewAdjusted 0.8 1.0 ByStateValues
+ %          , AlgDQNAvgRewAdjusted 0.8 0.99 ByStateValues
+ %          , AlgDQN 0.99 Exact
+
+% Alpha: 0.01, Decay: ExponentialDecay (Just 5e-5) 0.25 50000
+% AlphaRhoMin: 2e-5, Decay: ExponentialDecay (Just 2e-5) 0.25 50000
+% Gamma: 0.01, Decay: ExponentialDecay (Just 1e-3) 0.25 100000
+% Epsilon: 0.25, 0.25, Decay: No decay
+% Exploration: 1.0, Decay: ExponentialDecay (Just 0.005) 0.25 100000
+% LearnRandomAbove: 0.5
+%
+%
+% UNUSED: Delta: 0.005, Decay: ExponentialDecay (Just 5e-4) 0.25 100000
+
+The algorithm initiation strategy is to first fill the replay memory buffer by skipping any learning
+(steps \(7-21\) in Algorithm~\ref{alg:full}) until the period, that corresponds to the replay
+memory size, is reached. Then for \(500\) steps the average reward is set to the last computed
+exponentially smoothed value to prevent divergence due to the untrained network weight
+initialisation values.
+
+% \MS[t]{Extensions: overestimate rho, rho minimum, Multiple Agents (shared state,  shared rho?),
+%   init-phase (0.5 for expSmthRewRate, always adapt rho, same rho for all agents): exp smooth reward
+%   as rho, replay memory actions, n-step (describe replay memory), gradient clipping }
+% \MS[t]{Side info: Replay memory also stores (disallowed) filtered action indices for optimisation purposes and on
+% episodic tasks needs to store a bool indicating an episode end. Uses Int8 for state features values.  }
+
+\section{Experimental Evaluation}
+\label{sec:Experimental_Evaluation}
+
+This section introduces the simulation model and the parameterisation of the algorithm. To provide a
+proof-of-concept for the proposed algorithm we adapt the flow shop presented in
+\citet{schneckenreither2020order}. However, our algorithms operates on discrete lead times on a
+product type basis and thus is not directly comparable to their order based lead time management
+approach. % Nonetheless, this section is largely based on the corresponding section of
+% \citet{schneckenreither2020order}.
+
+\subsection{Simulation Model}
+\label{subsec:Simulation_Model}
+
+We adapt the simulation model of a make-to-order (MTO) flow shop with six different
+products and six work-centers of \citet{schneckenreither2020order} and examine the system under
+moderate and high utilisation in combination with moderate or high variability in demand and
+processing times. This section provides the details of the simulation model and the experimental
+setup.
+
+% \subsection{Make-To-Order Flow Shop}
+% \label{subsec:make-to-order_flow_shop}
+% \subsection{Simulation model}
+% \label{subsec:Simulation_model}
+
+The simulation model is depicted in Figure~\ref{fig:ps}, where the customers initiate the process by
+placing orders as indicated by the dashed line.
+%
+\begin{figure*}[t!]
+  \centering
+  \input{figures/productionsystem}                      % Figure production system
+  \caption{Production System of the Simulation Model with routing, processing time distributions and
+    demand interarrival time distributions.} \label{fig:ps}
+\end{figure*}
+%
+%
+The manufacturer produces six different product types, each of which is identified by a natural
+number ${1,2,\ldots,6}$ and has a different routing through the production system. The routing setup
+is provided by corresponding labels of the edges. For instance, product types $1-3$ are routed from
+machine $M1$ to machine $M2$, while the other products are transferred to machine $M3$. The
+production system consists of three production stages with diverging material flow and no return
+visits.
+%
+The incoming orders are uniformly distributed among the product types.
+%
+A period lasts $960$ minutes ($=16$ hours) which represent two \(8\)-hour shifts. At the
+beginning of each period orders can be released into the production system assuming material
+availability for all orders at their release date. Upon release, the orders are
+placed in the buffer at machine $M1$. Buffers and inventories are plotted as filled squares. Thus,
+each workstation is equipped with a buffer which queues the orders until processing is started. All
+buffers use a first-come-first-serve dispatching rule.
+%
+Furthermore, each workstation processes only one order at a time and early deliveries are
+prohibited. Orders completed ahead of their due date remain in the finished goods inventory (\fgi{})
+until they are due and sent to the customer.
+
+\paragraph{Processing Times.} The machine processing time distributions are specified under the
+corresponding node labels of the machines. To simulate high % or moderate
+variance the processing times for all machines are drawn from an exponential % or an uniform
+distribution% , respectively
+. There is one bottleneck machine ($M5$) and therefore we refer to product types $2$ and $5$, those
+routed through machine $M5$, as bottleneck products whereas products $1,3,4,6$ are non-bottleneck
+products.
+
+\paragraph{Demand.} The incoming orders are placed in the order pool with a due date slack of
+\dds{} periods, that is each order arriving within period $t$ will be due at the end of period
+$t+\dds{}$. The due date slack of \(\dds{}\) was chosen to (i) ensure sufficient time between the
+first occurrence of orders in the order pool and the latest possible release of orders and (ii) to
+have a clear cause and effect relationship between the order release decision and the cost
+performance in the analysis.
+%
+To simulate medium and high variability of the demand process the interarrival time between
+consecutive incoming orders is either drawn from an exponential ($\Exp$) or an uniform ($\Unif$)
+distribution, which are adjusted to yield the desired bottleneck utilisation level of either $70\%$
+with $\Exp(135)$ and $\Unif(95,175)$, $80\%$ with $\Exp(118)$ and $\Unif(78,158)$, or $90\%$ with
+$\Exp(105)$ and $\Unif(65,145)$. These values were chosen as they comprise the non-linear
+relationship between flow times and high utilisation levels.
+%%%%%%
+% Calculation
+% (285 / (3 * u)) = interarrival time
+%
+% 285... Proc time bottleneck
+% 3  ... every third product routes of this machine
+% u  ... utilisation, e.g. 0.90
+
+\paragraph{Order Release.}
+%
+Before the order release decision is made, all orders in the order pool are sorted by due date. This
+portrays the implementation of a sequencing rule. According to this sequence, orders are considered
+for release in the beginning of each period starting with the highest priority order.
+
+In our model, orders are released by specifying lead times
+$\lt_{1}, \lt_{2}, \ldots, \lt_{6} \geqslant 1$, where the index corresponds to the product type and
+$\lt_{i} \in \mathbb{N}$.
+%
+%
+By setting a lead time $\lt_{i}$ a \emph{planned release date} is computed for each
+order $j$ of product type $i$ in the order pool given by
+%
+\begin{linenomath*}
+  \begin{equation}
+    \prd{}_j= \dd{}_j - \lt_{i}\ ,
+    \label{eq:Orel}
+  \end{equation}
+\end{linenomath*}
+
+where $\dd_{j}$ denotes the due date of the order.  For dynamic order releases the lead times
+$\lt_{i}$ may vary from period to period, whereas in static order release methods the lead times are
+predetermined and fixed (\citealt{RagatzMabert1988,KimBobrowski1995}).
+%
+For both approaches a job of product type $i$ with lead time $\lt{}_{i}$ in period $t$ and due date
+$\dd_{j}$ is released at the end of period $t$ if and only if $t \geqslant \prd{}_{j}$.
+%
+Thus, in the dynamic setting  the planned
+release date $\prd{}_{j}$ of order $j$ can be updated several times before it is actually released
+to the production system.
+%
+However, once an order is released its release cannot be revoked.
+%
+% If there are several orders with the same due date the order release
+% sequence is randomized.
+%
+Hence, if the set lead time corresponds with the actual flow time the product is finished in the
+same period as it is shipped. If the lead times are set either too long or short,
+the order has to wait in the \fgi{} until it is due or the order is late and backorder (BO) costs
+occur.
+
+\paragraph{Costs.}
+%
+The cost parameters are set by assuming an increase in value from raw material ($1$~Dollar per order
+and period) to the final product ($4$~Dollar per order and period) and the backorder costs are set
+very high ($16$~Dollar per order and period) due to the MTO environment. All costs are assessed
+based on the production system state at the end of each period.
+
+\section{Lead Time Setting}\label{sec:lead}
+
+\subsection{Conventional Order Release Rule}
+
+As external benchmark for comparison we use different parameterized backward infinite loading
+(\BIL{}) techniques \citep{Ackerman1963}. \BIL{} uses Equation~\ref{eq:Orel} with a predetermined
+and fixed lead times \(\lt_{i}\).
+%
+% \begin{equation*}
+%   \rd_{j} = \dd_{j} - \lt_{j} \tcom
+%   \label{eq:BIL}
+% \end{equation*}
+%
+% where the release date \(RD_{j}\) of product type \(j\) is calculated by the difference of the due
+% date \(DD_{j}\) and the lead time \(LT_{j}\).
+%
+% Note that we set lead times for each product group and not for every order as done by
+% \cite{Ackerman1963}.
+
+\subsection{Average Reward Adjusted Deep RL Algorithm}
+\label{subsec:Algorithm_Setup}
+
+\paragraph{Markov Decision Process.} The underlying MDP is unichain and looks as follows. Both,
+state space $\mathcal{S}$ and action space $\mathcal{A}$ are discrete.
+%
+% Recall that we products routed over the bottleneck machine \(M5\) are bottleneck products.
+%
+For the \textit{state space} we adapt the result of \citet{knollmann2013control}, which state that
+the lead times ought to control the WIP level, instead of the due date reliability as changes are
+directly seen. Thus, any state $s \in \mathcal{S}$ of the state space is composed of the following
+information for each product type \(\pt\):
+\begin{itemize}
+
+\item The currently set lead time $LT_{g} \in \{1,2,\ldots,\DDS\}$ (Recall:
+  $\text{Due Date Period} - \text{Lead Time} = \text{Release Period}$). Note that we bound the
+  maximum lead time with due date slack $\DDS$, which is $7$ in our setup.
+
+\item Counters $OP_{g,d}\in \N$ for the number of orders in the order pool divided in time buckets
+  with $d \in \{1,2,\ldots,\DDS \}$, which stands for the number of periods until the due date.
+
+  % \item Flag for each machine whether it is idle or not.
+
+\item Counters $Q_{g,i,d}\in \N$ for the number of orders in the machine and queue \(i \in
+  \{M_{1}, M_{2}, \ldots , M_{6}\}\) divided in
+  time buckets with $d \in \{1,2,\ldots,\DDS \}$, which stands for the number of periods until the
+  due date.
+  % \MS[t]{depend}If a product type is not routed over the corresponding machine, then no entry for this
+  % product type is specified.
+
+\item Counters $FGI_{g,d} \in \N$ of orders in the finished goods inventory divided in time
+  buckets with $d \in \{1, 2,\ldots, \DDS \}$, which stands for the number of periods until the
+  due date. Orders with a due date with more than 3 periods ago are listed in the counter
+  $FGI_{-3}$.
+
+\item Counters $S_{g,d} \in \N$ of shipped orders from the last period divided in time buckets
+  with $d \in\{-3,-2,\ldots, \DDS \}$, which stands for the number of periods until the due date.
+  Orders with a due date with more than 3 periods ago are listed in the counter $S_{-3}$.
+
+\end{itemize}
+
+The algorithm implicitly learns a function which maps the current state of the production
+system to a lead time update and thus a release decision. In an optimal situation it does so by
+multiply exploring every action for each state and assessing its economic viability.
+%
+Orders are released once the due date is within the interval $[t,t+\text{lead time}]$, whereas $t$
+is the current period. Clearly, this yields bulk releases, meaning that either all or no orders with
+the same due date and of the same product type are released.
+
+The \textit{action space} is composed of two decisions, the relative change of the lead times to the
+currently set lead times $LT_{g} \in \{1,2,\ldots,\DDS\}$ for each product type $\pt$.
+%
+We restrict the lead time update for state $s_{t+1}$ according to the lead time $LT_{g}$ from state
+$s_{t}$ by a maximum change of 1. Thus, if $LT_{g}$ is the current lead time of product type $\pt$
+the agent can choose a lead time in $\{1,2,\ldots,\DDS\} \cap \{LT_{g}-1,LT_{g},LT_{g}+1\}$. % Put
+% differently the algorithm can increase or decrease the lead time by 1 or leave it as it is, as long
+% as it acts within the discrete action space given by the set $\{1,2,\ldots,\DDS\}$ for each product
+% type.
+
+\paragraph*{Reward.}
+In each period the agents choose actions, which when executed generate a reward while traversing to
+the next period by simulating the production system. The rewards are the accumulated costs at the
+end of the period consisting of the backorders, the WIP orders and the number of orders in the
+FGI\@. The objective is to minimise the returned reward. % Therefore, we adapt the algorithm by
+% swapping all maximisation functions with minimisations.
 

### 2021-05-17 paper/paper.tex
litrev

@@
   % teilw. von IJPR paper uebernommen
-  An essential task in manufacturing planning and control is to determine when to release orders to
-  the shop floor. One key parameter is the lead time which is the planned time that elapses between
-  the release of an order and its completion. Lead times are usually determined based on the
-  measured duration orders take to traverse the production system, i.e.\@ the flow times.
-  Traditional order release models assume static lead times, although it has been shown that they
-  should be set dynamically to be able to react to the dynamic operational characteristics of
-  production system. Therefore, we present an order release model which sets lead times dynamically
-  by using an reinforcement learning approach. This algorithm is designed especially for
-  applications with high process variability and periodic feedback to the agent. Therefore, this
-  paper aims at providing a proof-of-concept for the proposed algorithm. Therefore, we compare our
-  method on a three-stage flow-shop simulation model not only to static order release techniques,
-  but also to the commonly used deep Q-Learning algorithm. The results show that both reinforcement
-  learning agents have room for improvement for queuing systems with low variability, that our
-  proposed method usually outperforms deep Q-Learning and finally for systems with high utilisation
-  that the proposed reinforcement learning algorithm performs very well, outperforming all other
-  approaches.
+  A major task in hierarchical structured manufacturing planning and control is to determine when to
+  release orders to the shop floor. One key parameter is the lead time which is the planned time
+  that elapses between the release of an order and its completion. Lead times are usually determined
+  based on the flow times, i.e.\@ the measured duration orders take to traverse the production
+  system. Traditional order release models assume static lead times, although it has been shown that
+  they should be set dynamically to be able to react to the dynamic operational characteristics of
+  production system. Therefore, we present a reinforcement learning order release model which sets
+  lead times dynamically. The applied reinforcement learning agent is designed especially for
+  processes with periodic feedback and high variability. This paper aims at providing a
+  proof-of-concept for the proposed algorithm. We compare our method on a three-stage flow-shop
+  simulation model not only to static order release techniques, but also to the commonly used deep
+  Q-Learning agent. The results show that both reinforcement learning agents have room for
+  improvement for queuing systems with low variability, that our proposed method usually outperforms
+  deep Q-Learning and finally for systems with high utilisation that the proposed reinforcement
+  learning algorithm performs very well, outperforming all other approaches.
   %
@@
 
-Setting of lead times dynamically is a rather difficult task. Therefore, most state-of-the-art order
+Setting lead times dynamically is a rather difficult task. Therefore, most state-of-the-art order
 release mechanisms use static or fixed lead times to address the order release problem.
@@
 
-\MS[]{todo: Reactive}
-
-\MS[]{todo: proactive}
-
-We hypothesise that dynamic lead time management approaches need to aim for a predictive design in
-order to be able to compete with state-of-the-art order release methods from literature. However,
-there only exist three papers that propose a predictive lead time management algorithm in
-literature.
-
-The first approach by \citet{PaterninaArboleda2001} introduces a predictive order release model by
-using reinforcement learning in a single product, four-station serial flow line and compare its
-performance (WIP costs) with conventional order release policies (e.g., Kanban and CONWIP).
-%
-% Reinforcement learning is an optimisation technique that stems from dynamic programming and its goal
-% is to find the best stationary policy for a given problem. This policy is usually provided by
-% assessing current and future states (or state-action pairs) of an underlying Markov Decision Process
-% (MDP).
-% The advantage of reinforcement learning over dynamic programming is that (i) the problem
-% space is explored by an agent and thus only expectantly interesting parts of the problem space need
-% to be assessed and (ii) the knowledge (acquisition) of transition probabilities becomes unnecessary
-% as the states are evaluated by consecutively observed states solely.
-%
-The algorithm of \citet{PaterninaArboleda2001} decides on whether or not to release an order after
-each system change, the completion of an operation of any order at any stage or a new order arrival,
-and assumes that any unsatisfied demand is lost. Thus, they use a continuous order release method
-although in practice order release decisions often need to be made on a periodical basis, e.g.\@
-daily (see \citealt{enns2004work,GeldersvanW1982}). They outperform existing control policies with their tabular
-based reinforcement learning agent.
+% \MS[]{todo: Reactive}
+
+In the sequel we review the lead time management (LTM) literature for due date assignment, order
+release and scheduling, with the focus on order release literature.
+
+\paragraph{Reactive LTM.}
+%
+%% Reactive Due Date Ass
+Early due date assignment approaches \citep[e.g.][]{conway1967} set the due dates according to the
+processing time (Total Work Content rule, TWK rule) or the number of operations.
+%
+%% Reactive Order Release
+For order release \citet{enns2004work} investigate the use of a reactive approach with exponential
+smoothing of flow times to set lead times for two-stage distribution requirements planning. Although
+it can be considered an exploratory research paper the main conclusions are that dynamically setting
+planned lead times using exponential smoothing works well, but that the reactions are delayed. The
+later is due to the nature of the flow time measurement only being possible for finished orders.
+%
+Another reactive approach uses exponential smoothing to forecast flow times to set lead times for a
+capacitated multi-stage make-to-stock system \citep{Selcuk2006}. Their approach performs better than
+fixed lead time techniques for most experiments. However, they also exhibit the LTS\@.
+
+\paragraph{Proactive LTM.}
+%
+%% Proactive Due Date Assignment
+The TWK due date assignment rules have been extended by several researchers to proactive methods
+\citep{eilon1976, kanet1986, Enns1995}. As flow times depend on the workload and machine capacity in
+the shop workload dependent \citep{Bertrand1983} due date assignment rules have been developed. The
+basic idea is to estimate the flow time of an order on its arrival by measuring the number of orders
+queued on the corresponding routing, which can be analytically \citep[e.g.][]{ChungHuang2002}, by
+the use of a simulation model \citep[e.g.][]{vig1991}, statistical methods \citep[e.g.][]{Tai2012}
+or using hybrid approaches \citep[e.g.][]{Kaplan1993}. Furthermore several proactive due date
+assignment methods that incorporate neural networks exist \citet{Philipoom1994, Philipoom1997,
+  HsuSha2004,Patil2008,ChangetalFuzzyDDset2008}.
+%
+%% Proactive Order Release
+In the case of order release the analytical papers by \citet{Selcuk2009} and \citet{Selcuk2013} use
+the WIP level to quote lead times. In the former they show that the less frequently updating the
+lead time decreases the nervousness and improves the performance, while in the latter that
+dynamically updating lead times performs much better than static lead times.
+%
+% MIL (Modified infinite loading)
+MIL extends the static BIL procedure to integrate shopfloor congestion by subtracting number of
+queued jobs on the corresponding routing from the due date
+\citep{RagatzMabert1988,wisner1995review}. For non-due-date oriented dispatching rules, their
+approach has a positive effect on total cost performance, while for due-date oriented dispatching
+this holds only if the due dates are loose. For all other cases the release algorithm has no
+negative effect on performance. Variants of this rule-based method exist
+\citep[e.g.][]{park1989job,kim1995evaluating}.
+% Bertrand 2002
+% \citet{bertrand2002investigation} investigate different ways of releasing orders, assigning lead
+% times and operation due dates to assembly orders. In their work a due date is set once the order
+% arrives and then a release time calculated by subtracting the lead time from it. In one approach the
+% lead time calculation is the sum of the operation times plus a factor consisting of the waiting time
+% allowance per operation times the number of operations. They use the average operation waiting time
+% as waiting time allowance factor. In the other approaches the orders are released using the
+% \ac{imre} policy. In terms of order release they find that all work orders of an assembly order
+% should be released at the same time.
+%
+A machine learning approach is presented by \citet{Lee1997}. They combine a genetic algorithm with a
+decision tree to generate sequencing rules in the order pool for release. In the tested job shops
+the CONWIP rule determines the release of the orders and a decision tree to find the order release
+sequence. Their method leads to significant improvements over conventional single release and
+dispatching approaches.
+
+%% Predictive
+\paragraph{Predictive LTM.}
+%
+Due to the complex problem structure we hypothesise that dynamic lead time management approaches
+should aim for a predictive design in order to be able to compete with state-of-the-art order
+release methods from literature. However, there only exist four papers that propose a predictive
+lead time management method.
+
+For scheduling \citet{wang2007reinforcement} present a predictive lead time management approach
+using reinforcement learning. Their algorithm uses multiple agents to forecast the processing time
+of each job for routing decisions. They show on a 10 machine job shop that the RL algorithm performs
+equally or better than tested heuristics.
+%
+In case of order release, \citet{PaterninaArboleda2001} introduce a predictive order release model
+using reinforcement learning in a single product, four-station serial flow line. They compare its
+performance (WIP costs) with conventional order release policies (e.g., Kanban and CONWIP). The
+algorithm decides on whether or not to release an order after each system change, the completion of
+an operation of any order at any stage or a new order arrival, and assumes that any unsatisfied
+demand is lost. Thus, they use a continuous order release method although in practice order release
+decisions often need to be made on a periodical basis, e.g.\@ daily (see
+\citealt{enns2004work,GeldersvanW1982}). They outperform existing control policies with their
+tabular based reinforcement learning agent.
 %
@@
 days. For the same experimental setup the inputs/random variables are the same for each period.
-After learning each order release method was evaluated on \(20\) demand streams, each with \(1000\)
+After learning, each order release method was evaluated on \(20\) demand streams, each with \(1000\)
 periods of warm-up and \(7000\) periods of evaluation phase. Welch’s procedure was applied to
@@
     \begin{minipage}[t]{\dimexpr\textwidth-\leftmargin-\labelsep-\labelwidth-\leftmargin+2pt}
-      With probability \(\frac{\pexp}{n}\) choose action \(a_{t,i}\) randomly or otherwise one of
-      the set defined by
+      With probability \(\frac{\pexp}{n}\) choose action \(a_{t,i}\) randomly or otherwise the one
+      fulfilling
       \(\min_{a \in \mathcal{A}(i)}\X^{\pol}_{\gamma_{1}}(s_{t},a; \netWorker{})\).

### 2021-05-17 paper/paper.tex
minor

@@
   % teilw. von IJPR paper uebernommen
-  A major task in hierarchical structured manufacturing planning and control is to determine when to
-  release orders to the shop floor. One key parameter is the lead time which is the planned time
+  A major task in hierarchically structured manufacturing planning and control is to determine when
+  to release orders to the shop floor. One key parameter is the lead time which is the planned time
   that elapses between the release of an order and its completion. Lead times are usually determined
@@
 % system state is by design considered as a main driver of decision making in the current period.
-Reinforcement learning is an optimisation technique that stems from dynamic programming, therefore
-also seeks to compute the best stationary policy for a given problem. The idea is to evaluate the
-current system state according to the direct feedback (reward), as well as the previously learned
-state value of the consecutive state. This forms a cyclic constraint problem with as many
-constraints and variables as there are system states. The solution is found through iteration and
-due to the transitivity property of the state value update.
+Reinforcement learning (RL) is an optimisation technique that stems from dynamic programming,
+therefore also seeks to compute the best stationary policy for a given problem. The idea is to
+evaluate the current system state according to the direct feedback (reward), as well as the
+previously learned state value of the consecutive state. This forms a cyclic constraint problem with
+as many constraints and variables as there are system states. The solution is found through
+iteration and due to the transitivity property of the state value update.
 % as in dynamic programming is that the process and the current state and future states (or
 % state-action pairs) of an underlying Markov Decision Process (MDP).
-The advantage of reinforcement learning over dynamic programming is that (i) the problem space is
+The advantage of RL over dynamic programming is that (i) the problem space is
 explored by an agent and thus only expectantly interesting parts of the problem space need to be
@@
 
-There are two main types of reinforcement learning algorithms known in literature. The commonly
-applied discounted reinforcement learning method
+There exist two main types of deep RL algorithms in literature. The commonly
+applied discounted RL method
 \citep[e.g.][]{sutton1998introduction,mnih2015human,mnih2016asynchronous}, in which the future
 system state is discounted by a factor \(0 < \gamma < 1\) to prevent an infinite growth of the state
-evaluations, and average reward reinforcement learning
+evaluations, and average reward RL
 \citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults,
@@
 % by a single value. To be more precise, average reward reinforcement learning
-by splitting up the evaluation of each state into infinitely many parts. First, the average reward
-per step, second, a bias value that specifies expected reward that is additionally accumulated when
-traversing from the current state to the optimal path, and third, an infinite number of values of
-which each includes the occurrence of the rewards and in which magnitude along the path
-\citep{Howard64,Puterman94,
+by splitting up the evaluation of each state into infinitely many components. First, the average
+reward per step, second, a bias value that specifies expected reward that is additionally
+accumulated when traversing from the current state to the optimal path, and third, an infinite
+number of values of which each includes the occurrence of the rewards and in which magnitude along
+the path \citep{Howard64,Puterman94,
   Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults}. % Thus,
@@
 To expectantly accumulate the most reward one has to choose the action that leads to the state with
-highest value of the parts, on ties selecting actions following the order above. That is, the
-highest evaluation of the average reward, the bias value, the third component, the fourth component,
-etc. The formulations for computing these values form nested constraints, so for instance the
-current estimate of the average reward is used to compute the bias value. Therefore, a slight
-approximation error may cause harmful damage in subsequent constraints, which makes average reward
-reinforcement computationally very intensive. Worsening the issue, \citet[p.343ff]{Puterman94} shows
-that an additional set of nested constraints have to be solved to gain the correct solution.
+highest value of the components following the order above, on ties
+selecting % within the best actions
+according to the next component. That is, the highest evaluation of the average reward, the bias
+value, the third component, the fourth component, etc. The formulations for computing these values
+form nested constraints, so for instance the current estimate of the average reward is used to
+compute the bias value. Therefore, a slight approximation error may cause harmful damage in
+subsequent constraints, which makes average reward reinforcement computationally very intensive.
+Worsening the issue, \citet[p.343ff]{Puterman94} shows that the constraint problem has to be
+specified to solve one more component as originally intended to gain correct solution.
 
 In contrast the standard discounted framework estimates only one single scalar value consisting of
-the addition of these subterms, where however, the average reward is scaled by \(1/(1-\gamma)\).
-%
-To collect as much reward as possible, for instance to gain the highest expected profit, the
-discount factor \(\gamma\) has to be very close to one, e.g. \(0.99\).
+the addition of these components, where however, the average reward is scaled by \(1/(1-\gamma)\).
+%
+But in order to collect as much reward as possible the discount factor \(\gamma\) has to be very
+close to one, for instance \(0.99\).
 %
@@
 over all states. Thus, independently assessing it is not only computationally unwisely, but also
-problematic when they are iteratively approximated. For this reasons most applications of discounted
-reinforcement learning in literature evaluate their algorithms on processes that have an average
+problematic as they are iteratively approximated. For this reasons most applications of discounted
+RL in literature evaluate their algorithms on processes that have an average
 reward of about zero \citep{schneckenreither2020average}. \citet{schneckenreither2020average} shows
-that tabular based discounted reinforcement learning agents are unable to find even basic policies
+that tabular based discounted RL agents are unable to find even basic policies
 when confronted with processes that have a small average reward per step. Unfortunately, when
@@
 
-Therefore, we develop, adapt and enhance an algorithm that uses only one scalar value for the
-estimation of the average reward over all states. Further it estimates for every state-action pair
-the bias value and the sum of all other state value parts, usually referred to as error term. The
-later values are adjusted by the average reward.
+Therefore, we develop an algorithm that uses only one scalar value for the estimation of the average
+reward over all states. Further it estimates for every state-action pair the bias value and the sum
+of all other state value components, usually referred to as error term. The later values are
+adjusted by the average reward.
 % This, combined with the iterative evaluation and the independently assessing of the average reward
@@
 %
-Thus, as opposed to the commonly applied discounted reinforcement learning algorithm we use an
-average reward adjusted reinforcement learning algorithm to dynamically set lead time for order
-release. In contrast to the aforementioned works on average reward reinforcement learning our
+Thus, as opposed to the commonly applied discounted RL algorithm we use an
+average reward adjusted RL algorithm to dynamically set lead time for order
+release. In contrast to the aforementioned works on average reward RL our
 algorithm incorporates the optimisation of not only the average reward over time, but also the bias
@@
 \begin{itemize}
-\item We propose a novel \textit{average reward adjusted deep reinforcement learning algorithm} \ARA{}.
-  This is the first average reward adjusted reward learning algorithm that operates using a
+\item We propose a novel \textit{average reward adjusted deep reinforcement learning algorithm}
+  \ARA{}. This is the first average reward adjusted reward learning algorithm that operates using a
   artificial neural network as function approximation. It includes several known extensions like
   n-step learning \citep[e.g.][]{mnih2016asynchronous} or experience replay memory
-  \citep[e.g.][]{mnih2015human}, which we advanced to prevent catastrophic forgetting by
+  \citep[e.g.][]{mnih2015human}, which we aditionally advanced to prevent catastrophic forgetting by
   action-based selection of experiences.
@@
 This section is split into several parts. First we discuss lead time management literature, then
-reinforcement learning applications to lead time management, third we briefly review the relevant
-notation to reinforcement learning and finally we present the relevant literature of reinforcement
+RL applications to lead time management, third we briefly review the relevant
+notation to RL and finally we present the relevant literature of reinforcement
 learning in the rest of the section. The later includes a discussion of discounted reinforcement
-learning, average reward reinforcement learning, the Laurent series expansion of the state values,
-and finally average reward adjusted reinforcement learning.
+learning, average reward RL, the Laurent series expansion of the state values,
+and finally average reward adjusted RL.
 
@@
 However, lead times should be set dynamically due to the non-linear relationship between utilisation
-and flow times \citealt{Pahl2007,hoyt1978dynamic}, but this may also substantially degrade the
+and flow times \citep{Pahl2007,hoyt1978dynamic}, but this may also substantially degrade the
 system performance. For instance, a solely reactive approach can result in the lead time syndrome.
@@
 \item \textsf{Reactive lead time management} sets lead times by reacting on earlier flow times
-  (e.g., \citealt{enns2004work, Selcuk2006}), i.e.\@ data from the past.
+  (e.g., \citealt{enns2004work, Selcuk2006}), i.e.\@ data from the \textit{past}.
 \item \textsf{Proactive lead time management} may incorporate \textit{past} data and the
@@
 smoothing of flow times to set lead times for two-stage distribution requirements planning. Although
-it can be considered an exploratory research paper the main conclusions are that dynamically setting
+it is considered an exploratory research paper the main conclusions are that dynamically setting
 planned lead times using exponential smoothing works well, but that the reactions are delayed. The
@@
 \citep{eilon1976, kanet1986, Enns1995}. As flow times depend on the workload and machine capacity in
-the shop workload dependent \citep{Bertrand1983} due date assignment rules have been developed. The
+the shop \citep{Bertrand1983} workload dependent due date assignment rules have been developed. The
 basic idea is to estimate the flow time of an order on its arrival by measuring the number of orders
@@
 or using hybrid approaches \citep[e.g.][]{Kaplan1993}. Furthermore several proactive due date
-assignment methods that incorporate neural networks exist \citet{Philipoom1994, Philipoom1997,
+assignment methods that incorporate neural networks exist \citep{Philipoom1994, Philipoom1997,
   HsuSha2004,Patil2008,ChangetalFuzzyDDset2008}.
@@
 % MIL (Modified infinite loading)
-MIL extends the static BIL procedure to integrate shopfloor congestion by subtracting number of
-queued jobs on the corresponding routing from the due date
-\citep{RagatzMabert1988,wisner1995review}. For non-due-date oriented dispatching rules, their
+Modified infinite loading extends the static Backward Infinte Loading (BIL) procedure to integrate
+shopfloor congestion by subtracting number of queued jobs on the corresponding routing from the due
+date \citep{RagatzMabert1988,wisner1995review}. For non-due-date oriented dispatching rules, their
 approach has a positive effect on total cost performance, while for due-date oriented dispatching
@@
 For scheduling \citet{wang2007reinforcement} present a predictive lead time management approach
-using reinforcement learning. Their algorithm uses multiple agents to forecast the processing time
+using RL. Their algorithm uses multiple agents to forecast the processing time
 of each job for routing decisions. They show on a 10 machine job shop that the RL algorithm performs
@@
 In case of order release, \citet{PaterninaArboleda2001} introduce a predictive order release model
-using reinforcement learning in a single product, four-station serial flow line. They compare its
+using RL in a single product, four-station serial flow line. They compare its
 performance (WIP costs) with conventional order release policies (e.g., Kanban and CONWIP). The
@@
 \citealt{enns2004work,GeldersvanW1982}). They outperform existing control policies with their
-tabular based reinforcement learning agent.
-%
-Then, \citet{Schneckenreither2019} use several different reinforcement learning algorithms to make
+tabular based RL agent.
+%
+Then, \citet{Schneckenreither2019} use several different RL algorithms to make
 periodic order release decisions for a flow shop production system. The algorithm directly sets lead
@@
 
-Over the past decades reinforcement learning has been applied to various problems, for which
+Over the past decades RL has been applied to various problems, for which
 astonishing results have been reported. E.g.\@ only recently \citet{mnih2015human} have presented a
-novel value-iteration reinforcement learning agent which exceeds human-level abilities in playing
+novel value-iteration RL agent which exceeds human-level abilities in playing
 many classic Atari games~\citep{bellemare2012investigating}. Further, \citet{mnih2016asynchronous}
-present improved results with asynchronous actor-critic reinforcement learning. Also games like
+present improved results with asynchronous actor-critic RL. Also games like
 Go~\citep{silver2016mastering} and Chess~\citep{silver2017mastering} have been mastered with
-superhuman performance by \textit{tabula rasa} reinforcement learning agents. Furthermore, the
+superhuman performance by \textit{tabula rasa} RL agents. Furthermore, the
 method has also been applied in the setting of manufacturing system, e.g.\@ to improve the ramp-up
@@
 %
-However, all these applications use discounted reinforcement learning and are either designed to
-investigate a rather simple MDP, e.g.\@ by selecting heuristics instead of optimising the underlying
-problem itself, or by mapping the actual objective in a reward function that is approximately \(0\)
-on average over time.
-%
-This is due to the fact that state value is largely composed of a term defined by the policy's
-average reward value \citep{MillerVeinott1969,Blackwell62} which would otherwise dilute the state
-values and thus decrease the solution quality, as can for instance be observed in
+However, all these applications use discounted RL and are either designed to investigate a rather
+simple MDP, e.g.\@ by selecting heuristics instead of optimising the underlying problem itself, or
+by mapping the actual objective in a reward function that is approximately zero on average over
+time.
+%
+This is due to the fact that state value is largely composed of a term defined by the current
+policys' average reward value \citep{MillerVeinott1969,Blackwell62} which would otherwise dilute the
+state values and thus decrease the solution quality, as can for instance be observed in
 \citet{SchneckenreitherHaeussler2019} and \citet{gijsbrechts2018can}.
@@
 Therefore, most applications that incorporate and directly reflect costs or profit in the reward
-function use average reward reinforcement learning. \citet{aydin2000dynamic} use it in the setting
+function use average reward RL. \citet{aydin2000dynamic} use it in the setting
 of scheduling, while in a series of papers Mahadevan et al.\@ investigated several problem domains
@@
 \citep{Mahadevan96_AverageRewardReinforcementLearningFoundationsAlgorithmsAndEmpiricalResults,Mahadevan96_OptimalityCriteriaInReinforcementLearning}.
-After these foundational works for average reward reinforcement learning they introduced a
-continuous time average reward reinforcement learning algorithm named SMART
+After these foundational works for average reward RL they introduced a
+continuous time average reward RL algorithm named SMART
 \citep{Mahadevan97_SelfimprovingFactorySimulationUsingContinuoustimeAveragerewardReinforcementLearning}.
@@
 The rest of this section briefly reintroduces the important concepts of average reward adjusted
-reinforcement learning, elaborates on optimality criteria and provides insights of the underlying
-algorithm. For a more extensive introduction to average reward reinforcement learning we refer to
+RL, elaborates on optimality criteria and provides insights of the underlying
+algorithm. For a more extensive introduction to average reward RL we refer to
 \citet{schneckenreither2020average}.
@@
 state \(s_{t}\) and chosen action \(a_{t}\). In other words, we expect the system to possess the
-Markov property \citep[p.63]{sutton1998introduction}. Reinforcement learning processes that possess
+Markov property \citep[p.63]{sutton1998introduction}. RL processes that possess
 the Markov property are referred to as Markov decision processes (MDPs)
@@
 factor. That is
+%
 \begin{align*}
@@
 \end{align*}
-
+%
 where \(0 \leqslant \gamma < 1\) is the discount factor and
@@
 policy. For instance, as can be seen in Figure~\ref{fig:printer} setting \(\gamma <0.8027\) the
-printer-loop is preferred over the mail-loop, although the mail-loop accumulates more
-reward over time, i.e.\@ selecting the mail-loop is a sub-optimal choice. Observe that the average
-reward received per step equals \(1\) for the printer-loop and \(2\) for the mail-loop. Thus, the
+printer-loop is preferred over the mail-loop, although the mail-loop accumulates more reward over
+time, i.e.\@ selecting the mail-loop is a sub-optimal choice. Observe that the average reward
+received per step equals \(1\) for the printer-loop and \(2\) for the mail-loop. Thus, the
 (Blackwell-)optimal policy is to choose the mail-loop. However, if
-\(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted reinforcement learning
-chooses the printer loop as going the printer loop has a higher evaluation.
+\(\gamma < 3^{-\frac{1}{5}} \approx 0.8027\) an agent using discounted RL
+chooses the printer loop as doing the printer loop has a higher evaluation.
 %
@@
 % In the sequel we present the Laurent series expansion which not only links discounted reinforcement
-% learning with average reward reinforcement learning, but also connects to \(n\)-discount-optimality.
+% learning with average reward RL, but also connects to \(n\)-discount-optimality.
 %
@@
 state value estimations \citep{MillerVeinott1969} finding a well working parameterisation is
-difficult. Average reward reinforcement learning overcomes these issues by separately assessing the
+difficult. Average reward RL overcomes these issues by separately assessing the
 subterms of the state values directly.
@@
 % Note how the Laurent series expansion dissipates these values in single terms and therefore defines
-% a divide and conquer methodological approach to reinforcement learning. Thus in average reward
-% reinforcement learning these terms are learned separately, while in the discounted framework a
+% a divide and conquer methodological approach to RL. Thus in average reward
+% RL these terms are learned separately, while in the discounted framework a
 % single scalar value for each state with a fixed \(\gamma\)-value is estimated.
@@
 %
-\citet[p.343ff]{Puterman94} shows that due to the given degree of freedom if \(n=-1,0,\ldots,M\)
+\citet[p.343ff]{Puterman94} shows that due to the given degree of freedom, if \(n=-1,0,\ldots,M\)
 constraints are satisfying the above conditions for all states \(s\), then only
 \(\avgrew^{\pol}(s), V^{\pol}(s), W^{\pol}_{1}, \ldots, W^{\pol}_{M-1}\) are unique, whereas
-\(W^{\pol}_{M}\) is offset by the vector \(u\) where for the transition probability matrix \(P\) the
-vector \(u\) is characterised by \((I-P)u = 0\). Note that \(u\) is determined by the number of
+\(W^{\pol}_{M}\) is offset by the vector \(u\), where for the transition probability matrix \(P\)
+the vector \(u\) is characterised by \((I-P)u = 0\). Note that \(u\) is determined by the number of
 closed irreducible classes of \(P\), that is for ergodic MDPs \(u\) is determined by a single
@@
 \(\X_{\gamma}^{\pol}(s) = \E_{\pol}[r_{t} + \gamma \X_{\gamma}^{\pol}(s_{t+1}) - \avgrew^{\pol} \mid
-s_{t} = s]\). The corresponding Bellman optimality equation then is given by
+s_{t} = s]\). The corresponding Bellman optimality equation then is
 \(\X_{\gamma}^{\polopt}(s) = \max_{a} \E_{\polopt}[r_{t} + \gamma \X_{\gamma}^{\polopt}(s_{t+1}) -
-\avgrew^{\pol} \mid s_{t} = s]\). By turning this formula into an update rule it builds the
-foundation of our algorithm.
+\avgrew^{\pol} \mid s_{t} = s]\). Turning this formula into an update rule it builds the foundation
+of our algorithm.
 
@@
 
-This section presents the average reward adjusted deep reinforcement learning algorithm \ARA{}.
+This section presents the average reward adjusted deep RL algorithm (\ARA{}).
 First, we present a basic version, which suffices for small problem sizes and introduce the idea of
@@
 
-\begin{algorithm}[t!]
+\begin{algorithm}[b!]
   \begin{algorithmic}[1]
@@
 \citet{schneckenreither2020average} and depicted in Algorithm~\ref{alg:near}. The algorithm operates
-on a target network \(\netTarget{}\) and a worker network \(\netWorker{}\) as in
-\citet{mnih2015human}. % The target network reduces fluctuations in the state value estimates and
+on a target network \(\netTarget{}\) and a worker network \(\netWorker{}\) as in deep Q-Learning
+\citep{mnih2015human}. % The target network reduces fluctuations in the state value estimates and
 % therefore reduces the
@@
 reward by Equation~\ref{eq:constr2} is superior as compared to exponentially smoothing the actual
-observed rewards. This makes sense as the later rather evaluates past policies.
+observed rewards. This makes sense as the later rather describes policies from the past.
 
@@
 In our application the agent can update the lead time for each product type by a maximum magnitude
-of \(1\) in regard to the lead time of the current period. A naive approaches is to define one
-action for each combination of possible lead time updates over all product types, e.g.\@ one action
-for a system with six product types could be the vector \(+1, +1, 0, 0, +1, 0\). This however, due
-to the exponential growth, results in only being able to solve very small production systems with
-only few product types. A common solution in production planning is to aggregate the products into
-product groups, cf.\@ hierarchical production planning~\citep{schneeweibeta1995hierarchical}.
+of \(1\) in regard to the lead time of the current period. A naive approach is to define one action
+for each combination of possible lead time updates over all product types, e.g.\@ one action for a
+system with six product types could be the vector \((+1, +1, +0, +0, +1, +0)\) for which each
+component defines the update of the corresponding product type lead time from the current lead time.
+This however, due to the exponential growth, results in only being able to solve very small
+production systems with only few product types. A common solution in production planning is to
+aggregate the products into product groups, cf.\@ hierarchical production
+planning~\citep{schneeweibeta1995hierarchical}.
 %
@@
 %
-This allows us to investigate the algorithms behaviour for even larger solution spaces. In
-particular, this adaption enables us to present the algorithm the above defined state space not only
-on the basis of product groups, but directly for each product type \(\pt\). Note that without this
-extension the problem is computationally infeasible due to \(3^{6} = 729\) possible actions.
-Removing the dependency between the lead times of different product types reduces the set of
-possible actions, and thus the number of ANN output layer nodes, to \(3 \cdot 6 = 18\). In
-preliminary experiments on small examples we only observed a very slight increase of required steps
-during training. Therefore, this is a vital adaption to ensure scalability.
 
@@
 a production system with at most \(3\) product types, or an aggregation of product types into
-product groups. In the sense of this foundational research paper, we rather aim for an enhanced
+product groups. In the sense of this foundational research paper, we aim for an enhanced
 version of the algorithm that is scalable.
@@
 costs is unnecessary. Algorithm~\ref{alg:full} depicts a detailed view on the algorithm including
-these extensions and with the objective to minimise the returned rewards. The idea is the same as in
-Algorithm~\ref{alg:near}, except that the action selection procedure is a loop over all independent
-agents and the implementation of the n-step logic.
+these extensions and with the objective to minimise the returned rewards, as we aim for minimal
+costs in our experiments. The idea is the same as in Algorithm~\ref{alg:near}, except that the
+action selection procedure is a loop over all independent agents and the implementation of the
+n-step logic.
 %
@@
 
+In this section we define the tested lead time setting algorithms. This includes a description of
+the MDP used for the reinforcement learning agents.
+
 \subsection{Conventional Order Release Rule}
@@
 % where the release date \(RD_{j}\) of product type \(j\) is calculated by the difference of the due
-% date \(DD_{j}\) and the lead time \(LT_{j}\).
+% date \(DD_{j}\) and the lead time \(\lt_{j}\).
 %
@@
 
-\item The currently set lead time $LT_{g} \in \{1,2,\ldots,\DDS\}$ (Recall:
+\item The currently set lead time $\lt_{g} \in \{1,2,\ldots,\DDS\}$ (Recall:
   $\text{Due Date Period} - \text{Lead Time} = \text{Release Period}$). Note that we bound the
@@
 The \textit{action space} is composed of two decisions, the relative change of the lead times to the
-currently set lead times $LT_{g} \in \{1,2,\ldots,\DDS\}$ for each product type $\pt$.
-%
-We restrict the lead time update for state $s_{t+1}$ according to the lead time $LT_{g}$ from state
-$s_{t}$ by a maximum change of 1. Thus, if $LT_{g}$ is the current lead time of product type $\pt$
-the agent can choose a lead time in $\{1,2,\ldots,\DDS\} \cap \{LT_{g}-1,LT_{g},LT_{g}+1\}$. % Put
+currently set lead times $\lt_{\pt} \in \{1,2,\ldots,\DDS\}$ for each product type $\pt$.
+%
+We restrict the lead time update for state $s_{t+1}$ according to the lead time $\lt_{\pt}$ from state
+$s_{t}$ by a maximum change of 1. Thus, if $\lt_{\pt}$ is the current lead time of product type $\pt$
+the agent can choose a lead time in $\{1,2,\ldots,\DDS\} \cap \{\lt_{\pt}-1,\lt_{\pt},\lt_{\pt}+1\}$. % Put
 % differently the algorithm can increase or decrease the lead time by 1 or leave it as it is, as long
@@
 % type.
+%
+Recall that we implement independent agents to reduce the action space. In particular, this adaption
+enables us to present the algorithm the above defined state space not only on the basis of product
+groups, but directly for each product type. Note that without this extension the problem is
+computationally infeasible due to \(3^{6} = 729\) possible actions. Removing the dependency between
+the lead times of different product types reduces the set of possible actions, and thus the number
+of ANN output layer nodes, to \(3 \cdot 6 = 18\). In preliminary experiments on small examples we
+only observed a very slight increase of required steps during training. Therefore, this is a vital
+adaption to ensure scalability.
 
@@
 The features representing order counts are scaled using min-max scaling with minimum \(0\) and
-maximum \(5\), the lead times with minimum \(0\) and maximum \(7\) respectively.
+maximum \(5\), the lead times with minimum \(0\) and maximum \(7\).
 % Output layer scaling
@@
 mini-batch size each consisting of n-step experiences are selected, there gradients computed and the
-sum of gradients backpropagated though the worker network using Adam \citep{kingma2014adam}, where
+sum of gradients backpropagated trough the worker network using Adam \citep{kingma2014adam}, where
 the weight decay with parameter \(\lambda\) is implemented as presented by
@@
 For the case of medium variability with a utilisation of \(70\%\) 70-(\(\Unif\)) we can see that
-\(\BILTwo\) performs best in the sense of accumulated costs (\(220.53\)). The reinforcement learning
+\(\BILTwo\) performs best in the sense of accumulated costs (\(220.53\)). The RL
 algorithm follow, where \rlt{0.99} is slightly but significantly better then the other two variants
@@
 within this group of best performing algorithms. The WIPC costs, as well as the SFTT, show that the
-reinforcement learning agents have a higher work in process, which is clear as they try to smooth
+RL agents have a higher work in process, which is clear as they try to smooth
 the workload of demand stream. When looking at the timing cost it becomes even more apparent that
@@
 well in systems with high utilisation. In case of the less utilised systems the results reveal that
-there is still potential for the improvement of reinforcement learning variants.
+there is still potential for the improvement of RL variants.
 
@@
 % This would be in line with the observation of overestimated state values of standard discounted
-% reinforcement learning algorithms, which lead to the development of
+% RL algorithms, which lead to the development of
 % extensions~\cite[]{hessel2018rainbow}. We believe that instead of developing extensions to resolve
@@
 
-This paper is one of the first papers that use reinforcement learning for setting lead times to
+This paper is one of the first papers that use RL for setting lead times to
 release orders into the production system. We present a novel average reward adjusted reinforcement
@@
 Therefore, in the future we plan to develop \ARA{} further to a actor-critic state-of-the-art
-reinforcement learning algorithm. We hypothesise that due to the more correct state-action value
+RL algorithm. We hypothesise that due to the more correct state-action value
 estimations of \ARA{} as compared to deep Q-Learning, cf. \citet{schneckenreither2020average}, the
@@
 % the discounted framework. The paper describes an application of an order release model based on
-% reinforcement learning. The performance is tested on a multi-product, two-stage hypothetical flow
+% RL. The performance is tested on a multi-product, two-stage hypothetical flow
 % shop and is measured by cost, delivery and lead time related measures. We show that in the current
@@
 % approaches. In the future we expect to be able to yield better results by investigating the
-% parameterisation and resolving current issues when scaling average reinforcement learning to complex
+% parameterisation and resolving current issues when scaling average RL to complex
 % scenarios.
@@
 
-\bibliographystyle{plainnat}
+\bibliographystyle{tfcadinitials}
+% \bibliographystyle{plainnat}
 \bibliography{references}
